### 2020-12-29 analysis.Rmd
first committ

---
title: "Analysis"
author: "Joseph Bulbulia"
date: "24/12/2020"
output: html_document
---

Notes to Suzanne:
* Question the attention check transformation -- try to get orginal values
* Question the decidion to code factors as numbers (below) -- to to get raw responses
* Question where the scales came from
* Question for NAs, what is actually missing and what is "prefer not to say" and what is "other"

```{r global_options, echo=FALSE, include = FALSE}
knitr::opts_chunk$set(#
  fig.path="Figs/", 
  message=FALSE, 
  warning=FALSE,
  collapse =TRUE,
  echo=FALSE, 
  #results="hide", 
  fig.width= 10,
  fig.height=10,
  tidy = "styler")
```

```{r runanalysis, echo=FALSE, include = FALSE}
source("libs.R") # libraries 
source("funs.R") # custom functions
```

```{r}
# import data
d <- data_read_mymachine()
```

```{r}
#inspect data frame
str(d) # need to create factors, where fractors are needed.
```

Ask Suzanne for raw data. Not sure how they rescaled the ordinal report data. Just a comment. I'd never tranform the factors into numbers scaled between 0-1. This leads to a loss of information.

```{r}
hist(d$rel_1)  # yikes -- factors (church freq was made into a number :(  loss of signal 
hist(d$rel_1)  # yikes -- factors prayer freq was made into a number  :(  loss of signal
```

This is  unfortunate. We are saying that people who are not religious are somewhere between atheists and religious people. Need to back indocator this into a factor.

```{r}
table(d$rel_3,useNA = "ifany") # This is absolutely not good. 
```

Note: here, not all NAs are created alike. 
```{r}
#  need to recode factors -- NAs should be NAs, otherwise, we should describe the missingness
table(d$ethnicity,  useNA = "ifany") # need recode NA as "not identified" 
table(d$denomination,useNA = "ifany") # need recode NA as "no denomination" 
df$wb_socSATSEX
# fix
d$denomination[is.na(d$denomination)] <- "not_reported"
d$ethnicity[is.na(d$ethnicity)] <- "not_reported"
d$wb_soc_3[is.na(d$wb_soc_3)] <- "not_reported" # Sat with sex

wb_soc_mean

```

```{r}
# what are the numbers here? 
table(d$attention_check,useNA = "ifany") 
d$wb_phys_mean
```

```{r}
head(d)
# clean data: use informative names! 
df <- d %>%
  rename(
    rel_CHURCHFREQ  = rel_1, # factor need to recode (eg 7 as 0)// yikes factor transformed into a number
    rel_PRAYERFREQ = rel_2, # factor need to recode (eg. 7 as 0)// yikes factor transformed into a number
    IS_RELIIGOUS = rel_3, # factor 1 = yes, 2 = no, 3 = atheist// yikes factor transformed into a number
    BELONGS_DENOMINATION = rel_4, # relig
    rel_EXTENTBELIEVEGOD = rel_5, # do you belong to a religion or religious denomination? 1 = yes, 2 = no (!)
    rel_EXTENTBELIEVEAFTERLIFE= rel_6,
    rel_EXTENTSPIRITUAL  = rel_7,
    rel_IMPORTANCERELIGIOUSLIFE= rel_8, # 1 to five
    rel_IMPORTANCEBELIEVEGOD= rel_9, # 1 to five
    wb_genLIFEQUALITY = wb_gen_1, # not defined # also remove second underscore, which BRMS doesn't like
    wb_genSATHEALTH = wb_gen_2, # not defined
    wb_phyPAIN = wb_phys_1,
    wb_physNEEDSMEDICALTREAT =  wb_phys_2,
    wb_physHASENERGY =  wb_phys_3,
    wb_physMOBILITY = wb_phys_4,
    wb_physSLEEP = wb_phys_5,
    wb_physCANDODAILYACTIVITIES  = wb_phys_6,
    wb_physSATCAPACITYWORK  = wb_phys_7,
    wb_psychENJOYLIFE = wb_psych_1,
    wb_psychMEANING  = wb_psych_2,
    wb_psychCONCENTRATE  = wb_psych_3,
    wb_psychLIKESAPPEARANCE  = wb_psych_4,
    wb_psychSATSELF  = wb_psych_5,
    wb_psychANXIETY  = wb_psych_6,
    wb_socSATPERSONALRELATIONSHIPS  = wb_soc_1,
    wb_socSATSUPPORTFRIENDS  = wb_soc_2,
    wb_socSATSEX   = wb_soc_3,
    male = gender
  )%>%
  dplyr::mutate(IS_RELIIGOUS = as.factor(IS_RELIIGOUS),
                denomination = as.factor(denomination),
                ethnicity = as.factor(ethnicity),
                country = as.factor(country))%>%
  dplyr::mutate(IS_RELIIGOUS = recode_factor(IS_RELIIGOUS, "0" = "atheist", "0.5" = "not_religious", "1" = "religious"))

```

Note the attention check variable. The data documentation says `1` means "passed" however the questionaire lists options 1 ("not at all") -- 7 "very much."

Will retain the attention check fails for now, however it iwll be important to remove these cases if they are affecting the outcomes. (I have requested further information from Suzanne)

```{r}
# get columns for different indicators 
rel <- get_rel_vars(df)
wbgen <- get_wb_gen(df)
wbphys <- get_wb_phys(df)
wbpsych <- get_wb_psych(df)
wbsoc <- get_wb_soc(df)
head(wbsoc)
head(rel)
```

Exploration Graphs

```{r} 
head(rel)
r1 <- rel %>%
  ggplot(aes(
    x = country,
    y = rel_CHURCHFREQ,
    group = country,
    color = country
  )) +
  geom_boxplot(notch = TRUE) +
  geom_point(position = "jitter", alpha = .1) +
  scale_color_viridis(discrete = TRUE) +
  ggtitle("Church Frequency") +
  ylab("") + 
  theme(axis.text.x = element_text(angle = 30))
r1

r2 <- rel %>%
  ggplot(aes(
    x = country,
    y = rel_PRAYERFREQ,
    group = country,
    color = country
  )) +
  geom_boxplot(notch = TRUE) +
  geom_point(position = "jitter", alpha = .1) +
  scale_color_viridis(discrete = TRUE) +
  ggtitle("Prayer Frequency") +
  ylab("") + 
  theme(axis.text.x = element_text(angle = 30))
r2

r3 <- rel %>%
  ggplot(aes(
    x = country,
    y = rel_EXTENTBELIEVEGOD,
    group = country,
    color = country
  )) +
  geom_boxplot(notch = TRUE) +
  geom_point(position = "jitter", alpha = .1) +
  scale_color_viridis(discrete = TRUE) +
  ggtitle("Extent Believe in God") +
  ylab("") + 
  theme(axis.text.x = element_text(angle = 30))

r3

r4 <- rel %>%
  ggplot(aes(
    x = country,
    y = rel_EXTENTBELIEVEAFTERLIFE,
    group = country,
    color = country
  )) +
  geom_boxplot(notch = TRUE) +
  geom_point(position = "jitter", alpha = .1) +
  scale_color_viridis(discrete = TRUE) +
  ggtitle("Extent Believe in Afterlife") +
  ylab("") + 
  theme(axis.text.x = element_text(angle = 30))
r4

r5 <- rel %>%
  ggplot(aes(
    x = country,
    y = rel_EXTENTSPIRITUAL,
    group = country,
    color = country
  )) +
  geom_boxplot(notch = TRUE) +
  geom_point(position = "jitter", alpha = .1) +
  scale_color_viridis(discrete = TRUE) +
  ggtitle("Extent Spiritual") +
  ylab("") + 
  theme(axis.text.x = element_text(angle = 30))
r5

r6 <- rel %>%
  ggplot(aes(
    x = country,
    y = rel_IMPORTANCERELIGIOUSLIFE,
    group = country,
    color = country
  )) +
  geom_boxplot(notch = TRUE) +
  geom_point(position = "jitter", alpha = .1) +
  scale_color_viridis(discrete = TRUE) +
  ggtitle("Importance of Religious Life") +
  ylab("") + 
  theme(axis.text.x = element_text(angle = 30))
r6

r7 <- rel %>%
  ggplot(aes(
    x = country,
    y = rel_IMPORTANCEBELIEVEGOD,
    group = country,
    color = country
  )) +
  geom_boxplot(notch = TRUE) +
  scale_color_viridis(discrete = TRUE) +
  geom_point(position = "jitter", alpha = .1) +
  ggtitle("Importance Believe in God") +
  ylab("") + 
  theme(axis.text.x = element_text(angle = 30))
r7

(r1 + r2  +  r3 )/  (r4  + r5) / (r6 +  r7)   +  plot_annotation(tag_levels = 'A') + plot_layout(guides = "collect")
```

Correlation plots. Note that we need the to model the country level dependences in any correlation matrix, given strong cultural differences

Religion:

```{r}
rel %>%
  correlation( partial = FALSE, multilevel = TRUE ) %>%
  plot() + theme_blackboard()
```

```{r}
str(rel)
report::report_sample(
  rel,
  group_by = "country",
  # select = NULL,
  # exclude = NULL,
  # weights = NULL, 
  total = TRUE, digits = 2
)
report::report_sample(iris, group_by = "Species")
```

General wellbeing
```{r}
wbgen %>%
  correlation( partial = FALSE, multilevel = TRUE ) %>%
  plot() + theme_blackboard()
```

```{r}
wbphys %>%
  correlation( partial = FALSE, multilevel = TRUE ) %>%
  summary()

wbphys %>%
  correlation( partial = FALSE, multilevel = TRUE ) %>%
  plot() + theme_blackboard()
```

Psychological wellbeing
```{r}
wbpsych %>%
  select(-wb_psychMEANING)%>%
  correlation( partial = FALSE, multilevel = TRUE ) %>%
  summary()

wbpsych %>%
  select(-wb_psychMEANING)%>%
  correlation( partial = FALSE, multilevel = TRUE ) %>%
  plot()  + theme_abyss()
```

```{r}
wbsoc %>%
  dplyr::select(-wb_socSATSEX) %>% # too many missing values
  correlation( partial = FALSE, multilevel = TRUE ) %>%
  summary()

wbsoc %>%
  dplyr::select(-wb_socSATSEX) %>% # too many missing values
  correlation( partial = FALSE, multilevel = TRUE ) %>%
  plot() + theme_abyss()
```

# Specify causal model

```{r}
head(df)
tidy_ggdag <- dagify(
  rel ~ age + country + male,
  wb  ~ rel + country,
  wb ~ soc  + ses,
  ses ~ country + age, 
  exposure =  "rel",
  latent = "country",
  outcome =   "wb") %>%
  tidy_dagitty()
#
# graph adjustment sets
ggdag::ggdag_adjustment_set(tidy_ggdag, node_size = 14) +
  theme(legend.position = "bottom")
ggdag_dseparated(tidy_ggdag)
ggdag_collider(tidy_ggdag)
ggdag_instrumental(tidy_ggdag)å
```

```{r}
head(df)
tidy_ggdag <- dagify(
  rel ~ age + country,
  wb ~ soc  + ses + rel + age,
  ses ~ country + age, 
  exposure =  "rel",
  latent = "country",
  outcome =   "wb") %>%
  tidy_dagitty()
#
# graph adjustment sets
ggdag::ggdag_adjustment_set(tidy_ggdag, node_size = 14) +
  theme(legend.position = "bottom")
ggdag_dseparated(tidy_ggdag)
ggdag_collider(tidy_ggdag)
ggdag_instrumental(tidy_ggdag)å
```

### 2020-12-29 funs.R

@@ -1,0 +1,557 @@
+# funs.R
+# functions used in this analysis
+data_read_mymachine <- function() {
+  out <-
+    read.csv("/Users/jbul176/Documents/GitHub/many_analysts/MARP_data_blinded.csv")
+}
+
+get_rel_vars <- function(x) {
+  df <- x %>% dplyr::select(c(starts_with("rel_"), "country"))
+}
+
+get_wb_gen <- function(x) {
+  df <- x %>% dplyr::select(c(starts_with("wb_gen"), "country"))
+}
+
+get_wb_phys <- function(x) {
+  df <- x %>% dplyr::select(c(starts_with("wb_phys"), "country"))
+}
+
+get_wb_psych <- function(x) {
+  df <- x %>% dplyr::select(c(starts_with("wb_psych"), "country"))
+}
+
+
+get_wb_soc <- function(x) {
+  df <- x %>% dplyr::select(c(starts_with("wb_soc"), "country"))
+}
+
+
+
+
+# this code is for citing packages
+cite_packages <- function() {
+  citation(package = "base", lib.loc = "/Users/josephbulbulia/Dropbox/BIBS")
+  toLatex(sessionInfo(), locale = FALSE)
+  sapply(names(sessionInfo()$otherPkgs), function(x)
+    print(citation(x), style = "Bibtex"))
+  out <- sapply(names(sessionInfo()$otherPkgs),
+                function(x)
+                  print(citation(x), style = "Bibtex"))
+  print(out)
+}
+#
+# corrmat_rel() <- d %>%
+#   dplyr::select(
+#     R_attendance,
+#     R_religiousID,
+#     R_member, # can lose this
+#     R_god,
+#     R_afterlife,
+#     R_prayer)%>%
+#   dplyr::mutate(R_attendance.S = scale(R_attendance),
+#                 R_religiousID.S =scale(R_religiousID),
+#                 R_member.S = scale(R_member), # not used
+#                 R_god.S=scale(R_god),
+#                 R_prayer.S =scale(R_prayer),
+#                 R_afterlife.S =scale(R_afterlife))%>%
+#   rowwise() %>%
+#   mutate(REL= mean(c(R_attendance, R_prayer,R_god,R_religiousID,R_afterlife), na.rm=T))%>% # we don't use this, but my sense is R_member isn't going to be worth including.
+#   mutate(REL.S = scale(REL))%>% # we don't use this
[…]

### 2020-12-29 libs.R

@@ -1,0 +1,45 @@
+# libraries and settings
+library("tidyverse")
+library("ggplot2") #plots
+library("sjstats") #test statistics
+library("sjPlot")
+library("sjmisc")
+library("insight")
+library("see")
+library("ggraph")
+library("parameters")
+library("easystats")
+library("lme4")
+library("psych") ## correlation matrix
+library("table1")
+library("ggeffects")
+library("gghighlight")
+#library("naniar") # missingness
+#library("Amelia")
+#library("magrittr")
+library("patchwork")
+#library("qgraph")
+library("texreg")
+library("viridis")
+library("ggpubr")
+library("ggsci") 
+library("here")
+library("citr")
+library("here")
+#library("statsExpressions")
+library("ggstatsplot")
+library("ggthemes")
+library("dagitty")
+library("ggdag")
+library("papaja")
+library("rstan")
+library("rstanarm")
+library("tidybayes")
+library("bayesplot")
+library("rethinking")
+library("dagitty")
+library("ggdag")
+rstan_options(auto_write=TRUE)
+options(mc.cores=parallel::detectCores ())
+theme_set(theme_few())
+

### 2021-01-14 analysis.Rmd
small fix
initial commit

@@
 
-Ask Suzanne for raw data. Not sure how they rescaled the ordinal report data. Just a comment. I'd never tranform the factors into numbers scaled between 0-1. This leads to a loss of information.
+Ask Suzanne for raw data. Not sure how they rescaled the ordinal report data. Just a comment. I'd never trasnform the factors into numbers scaled between 0-1. This leads to a loss of information.
 
@@
 
-Correlation plots. Note that we need the to model the country level dependences in any correlation matrix, given strong cultural differences
+Correlation plots. Note that we need the to model the country level dependencies in any correlation matrix, given strong cultural differences
 

### 2021-01-14 analysis.Rmd
uploading models

@@
 d$wb_soc_3[is.na(d$wb_soc_3)] <- "not_reported" # Sat with sex
-
-wb_soc_mean
-
 ```
@@
 ```{r}
-head(d)
-# clean data: use informative names! 
-df <- d %>%
+head(d$gender)
+
+# clean data:use informative names !
+df<- d %>%
+  dplyr::mutate(fm1m2 = ifelse( gender != "man", 1 ,2)) %>% # for indexical categorical variable
+  dplyr::mutate(male = factor(ifelse( gender != "man", "male" ,"female"))) %>%
 rename(
-    rel_CHURCHFREQ  = rel_1, # factor need to recode (eg 7 as 0)// yikes factor transformed into a number
-    rel_PRAYERFREQ = rel_2, # factor need to recode (eg. 7 as 0)// yikes factor transformed into a number
-    IS_RELIIGOUS = rel_3, # factor 1 = yes, 2 = no, 3 = atheist// yikes factor transformed into a number
-    BELONGS_DENOMINATION = rel_4, # relig
-    rel_EXTENTBELIEVEGOD = rel_5, # do you belong to a religion or religious denomination? 1 = yes, 2 = no (!)
-    rel_EXTENTBELIEVEAFTERLIFE= rel_6,
+    rel_CHURCHFREQ   = rel_1,
+    # factor need to recode (eg 7 as 0)// yikes factor transformed into a number
+    rel_PRAYERFREQ = rel_2,
+    # factor need to recode (eg. 7 as 0)// yikes factor transformed into a number
+    IS_RELIIGOUS = rel_3,
+    # factor 1 = yes, 2 = no, 3 = atheist// yikes factor transformed into a number
+    BELONGS_DENOMINATION = rel_4,
+    # relig
+    rel_EXTENTBELIEVEGOD = rel_5,
+    # do you belong to a religion or religious denomination? 1 = yes, 2 = no (!)
+    rel_EXTENTBELIEVEAFTERLIFE = rel_6,
     rel_EXTENTSPIRITUAL  = rel_7,
-    rel_IMPORTANCERELIGIOUSLIFE= rel_8, # 1 to five
-    rel_IMPORTANCEBELIEVEGOD= rel_9, # 1 to five
-    wb_genLIFEQUALITY = wb_gen_1, # not defined # also remove second underscore, which BRMS doesn't like
-    wb_genSATHEALTH = wb_gen_2, # not defined
+    rel_IMPORTANCERELIGIOUSLIFE = rel_8,
+    # 1 to five
+    rel_IMPORTANCEBELIEVEGOD = rel_9,
+    # 1 to five
+    wb_genLIFEQUALITY = wb_gen_1,
+    # not defined # also remove second underscore, which BRMS doesn't like
+    wb_genSATHEALTH = wb_gen_2,
+    # not defined
     wb_phyPAIN = wb_phys_1,
@@
     wb_socSATSEX   = wb_soc_3,
-    male = gender
   )%>%
@@
                 ethnicity = as.factor(ethnicity),
-                country = as.factor(country))%>%
+                country = as.factor(country)) %>% 
   dplyr::mutate(IS_RELIIGOUS = recode_factor(IS_RELIIGOUS, "0" = "atheist", "0.5" = "not_religious", "1" = "religious"))
 
+table(df$fm1m2, useNA = "always")
 ```
@@
 
+Suppose we believe that
+
+- religion + ses + male + country + a different unmeasured latent variable (e.g. priviledge) causes well being
+
+- age + male + country + and Unmeasured latent variable (e.g. parent's religion, call this U1) causes Religion
+
+- country + age + another unmeasured latent variable (e.g. priviledge U2) causes ses
+
+- our "exposure" is religion and our outcome is "well-being"
+
+Which variables should I condition on to understand the relationship between religion and well-being?
+
 ```{r}
-head(df)
 tidy_ggdag <- dagify(
-  rel ~ age + country + male,
-  wb  ~ rel + country,
-  wb ~ soc  + ses,
-  ses ~ country + age, 
+  wb  ~ rel + ses + male + country + U2,
+  rel ~ age + male + country + U1,
+  ses ~ country + age + U2,
   exposure =  "rel",
-  latent = "country",
+  outcome =   "wb"
+) %>%
+  tidy_dagitty()
+
+# graph adjustment sets
+ggdag::ggdag_adjustment_set(tidy_ggdag, node_size = 14) +
+  theme(legend.position = "bottom") +
+  theme_dag_blank()
+```
+
+Suppose I was to include all variables in my model, except U2 and U1, which are not measured?
+
+In this case I would introduce colider bias
+
+```{r}
+ggdag::ggdag_collider(tidy_ggdag,
+                      from = "rel",
+                      to = "wb",
+                      control_for(c("ses", "male", "country", "age")))
+
+```
+
+## missingness
+
+Next we need to examine patterns of missingness, which might spoil inference.
+
+```{r}
+library(naniar)
+vis_miss(df) # substantial pattern of missing ness in a few variables, 
+```
+
+There is little missingness so we will just use complete cases.
+Note that it would be better to have a method for handling the uncertainty of any case.
+A short cut is a short cut: missing data should be handled by some probabilitic imputation strategy.
+However because there are many larger problems about measurement error, sampling, and theory, lets set these to questions to the side and work with the complete dataset
+
+## The hard work
+
+We have yet to consider a mechanism by which religion might affect well-being, and specifically which aspects of well-being and which aspects of religion.
+Just as it might be tempting to through all variables into a regression equation, and disasterious to do so, it might be tempting to leap into some form of data reduction, and combine well-being and religious indicators without stopping to think about how causation.
+
+There are many theories we might consider.
+We will focus on two non-trivial theories, one related to church attendance and one related to belief in god.
+If we had more time we could consider many others, but as Catholics know, "timing is everything" -- let's keep to time.
+
+### Church attendance
+
+We propose that church attendance reduces psychological anxiety by increasing perceived support.
+This is a new model so we need to write our dag all over again.
+The dag will resemble the dag above, except we know that people who identify as male have less social support (typically), at least in some countries.
+
+Writing the dag
+```{r}
+tidy_ggdag2 <- dagify(
+  wb  ~ support + ses + male + country + U2,
+  support ~ church + male + country,
+  church ~ age + male + country + U1,
+  ses ~ country + age + U2,
+  exposure =  "church",
+  outcome =   "wb"
+) %>%
+  tidy_dagitty()
+
+# graph adjustment sets
+ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) +
+  theme(legend.position = "bottom") +
+  theme_dag_blank()
+```
+
+The adjustments sets are the same so we proceed with a model.
+Note that if my interest is in the direct effect of church attendance on wellbeing I should *not* include social support because this introduces a collider bias!
+
+```{r}
+ggdag::ggdag_collider(tidy_ggdag2,
+                      from = "church",
+                      to = "wb",
+                      controlling_for(c("male", "support", "country", "age")))
+
+```
+
+## The model
+
+A bayesian workflow should include prior predictive checks and the use of informative priors (McElreath 2020).
+
+For this initial analysis I am going to use the weakly informative priors that BRMS uses; my experience is that large datasets are resistent to choice of priors.
+However it is a mistake to trust this experience.
+For now, though, we must keep to time, and timing is everything.
+
+Additionally, the data have been manipulated in advance of the study to be scaled to a range between 0 and 1.
+This is a very bad idea.
+It would be much better to thinking of church attendance frequence as a monotonic effect.
+I will do that analysis at the next round, if the data are made available in their original form.
+
+Here is the model for church attendance and anxiety
+
+Wrangle data
+```{r}
+# get data into shape
+d1 <- df %>%
+  dplyr::mutate(
+    anx = as.numeric(wb_psychANXIETY),
+    anx_s = scale(wb_psychANXIETY),
+    #put outcome in ST units (see McElreath 2021)  # not necessary
+    age_s = scale(age),
+    ch_s = scale(rel_CHURCHFREQ),
+    mn_s = scale(wb_psychMEANING),
+    bg_s = scale(rel_EXTENTBELIEVEGOD)
+  ) %>%
+  dplyr::select(anx, anx_s, age_s, ch_s, mn_s, fm1m2, country, male, bg_s) %>%
+  dplyr::filter(
+    !is.na(anx_s),
+    !is.na(age_s),
+    !is.na(ch_s),
+    !is.na(mn_s),
+    !is.na(male),
+    !is.na(bg_s)
+  )
+table(d1$country)
+```
+
+Model for church
+
+```{r}
+m1 <- brm(
+  anx ~ ch_s +
+    age_s  +
+    male +
+    (1 + ch_s | country),
+  data = d1,
+  family = "gaussian",
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = "m1"
+)
+```
+
+### model for belief
+
+From causal dag above
+
+```{r}
+m2 <- brm(
+  anx ~
+    bg_s  +
+    age_s  +
+    male +
+    (1 + ch_s | country),
+  data = d1,
+  family = "gaussian",
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = "m2"
+)
+summary(m2)
+```
+
+Graph -- No effect, why? 
+```{r}
+p2 <- brms::conditional_effects(m2,
+                                c("bg_s"),
+                                categorical = F,
+                                #
+                                probs = c(0.11, 0.89))
+plot(p2)
+```
+
+To do -- look at between country effects and relative effects within countries
+
+```{r}
+library("rethinking")
+set.seed(867530)
+
+m_ch <- ulam(
+  alist(
+    anx_s ~ normal(mu , sigma),
+    mu <-  alpha + b1 * ch_s + b2 * age_s  + b3 * male + a_country[country] + b_country[country] * ch_s,
+    alpha ~ normal(0,1),
+    b1 ~ dlnorm(0,1),
+    b2 ~ normal(0,1),
+    b3 ~ normal(0,1),
+    u_country ~ dnormal(0,1),
+    c(a, b)[country] ~ multi_normal(c(a, b) , Rho , sigma_country),
+    a ~ normal(0, 2),
+    b ~ normal(-1, 0.5),
+    sigma_country ~ exponential(1),
+    sigma ~ exponential(1),
+    Rho ~ lkj_corr(2),
+  ) ,
+  data = d1 ,
+  chains = 4 ,
+  cores = 4
+)
+```
+
+Excerpt From: McElreath, Richard;. “Statistical Rethinking.” Apple Books. 
+```
+
+## Belief in God
+
+Lets consider the non-trivial theory that belief in God (bg) affects psychological anxiety (wb) by affording a sense of meaning in life.
+The causal qualities will be the same as those of the previous model, and so too will the threats to inference.
+
+Writing the dag
+```{r}
+tidy_ggdag2 <- dagify(
+  wb  ~ meaning + ses + male + country  +U2,
+  meaning ~ bg +  male + country, 
+  bg ~ age + male + country + U1,
+  ses ~ country + age + U2,
+  exposure =  "bg",
   outcome =   "wb") %>%
   tidy_dagitty()
+
+# graph adjustment sets
+ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) + 
+  theme(legend.position = "bottom") + 
+  theme_dag_blank()
+```
+
+And again we face a collider bias by introducing meaning measure, conditional on assumed causal relationships
+
+```{r}
+ggdag::ggdag_collider(tidy_ggdag2,
+                      from = "church",
+                      to = "wb",
+                      controlling_for(c("male","support","country","age")))
+
+```
+
+### OLD ANALYSIS
+
 #
-# graph adjustment sets
-ggdag::ggdag_adjustment_set(tidy_ggdag, node_size = 14) +
-  theme(legend.position = "bottom")
-ggdag_dseparated(tidy_ggdag)
-ggdag_collider(tidy_ggdag)
-ggdag_instrumental(tidy_ggdag)å
+#setwd("/Users/jbul176/The\ Virtues\ Project\ Dropbox/Joseph\ Bulbulia/00Bulbulia\ Pubs/2020/2020_VAN_ELK") # mine
+
+# read data
+e <- read.csv("/DataStudy3_SourceCredibility_Long.csv", header=T) # old data within/group means
+d <- read.csv('DataLong', header=T) # new data without ingroup means
+#d$statementAndOrder # error but this is not needed
+nrow(d)#20390
+hist(e$religiosity) # group standardized data
+hist(d$religiosity) # note difference
+# check data : note that subject needs to be a factor. also, "Statement" needs to be a factor. for gender there are few "others" we create a "male gender variable (below) for simplicity. 
+#
+# look for patters of missingness ####
+library(naniar)
+vis_miss(d) # substantial pattern of missing ness in a few variables, 
+```
+```{r}
+#plot outcomes: gaussian is reasonable, but we'll come back to this after ppchecks
+str(d)
+# check outcomes
+hist(d$credible) # gaussian ordinal
+hist(d$important)# flat
+# what is the F1 variable?
+hist(d$recallFa ) ##??
+hist(d$recallHit )##??
+```
+```{r}
+#make tables ####
+# and look for anomolies
+library(table1) # useful package
+table1::table1(~Age + Gender + Education + R_religiousID + religiosity + culturalNorms|country, data =d, transpose =T)
+```
+```{r}
+table1::table1(~credible|country, data =d, transpose =TRUE)# outcome variables
 ```
@@
 ```{r}
-head(df)
-tidy_ggdag <- dagify(
-  rel ~ age + country,
-  wb ~ soc  + ses + rel + age,
-  ses ~ country + age, 
-  exposure =  "rel",
-  latent = "country",
-  outcome =   "wb") %>%
-  tidy_dagitty()
+table1::table1( ~R_attendance + R_prayer + R_religiousID + R_member  + R_god   + R_afterlife|country, data =d, transpose =T)
+```
+
+```{r}
+# how correlated are the religion variables? ####
+corrmat<-d%>%
+  dplyr::select(
+    R_attendance,
+    R_religiousID,
+    R_member, # can lose this
+    R_god,
+    R_afterlife,
+    R_prayer)%>%
+  dplyr::mutate(R_attendance.S = scale(R_attendance),
+                R_religiousID.S =scale(R_religiousID),
+                R_member.S = scale(R_member), # not used
+                R_god.S=scale(R_god),
+                R_prayer.S =scale(R_prayer),
+                R_afterlife.S =scale(R_afterlife))%>%
+  rowwise() %>% 
+  mutate(REL= mean(c(R_attendance, R_prayer,R_god,R_religiousID,R_afterlife), na.rm=T))%>% # we don't use this, but my sense is R_member isn't going to be worth including.
+  mutate(REL.S = scale(REL))%>% # we don't use this
+  select(R_attendance.S,R_religiousID.S,R_member.S,R_god.S,R_afterlife.S,R_prayer.S) # don't select religiosity for PCA
+# corrplot
+res <- cor(corrmat, method="pearson")
+corrplot::corrplot(res, method= "color", order = "hclust") # tl.pos = 'n'
+```
+
+```{r}
+## corr.soc  
+sjp.corr(corrmat,
+         title = "Correlation matrix of religion variables",
+         geom.colors = "RdBu")
+```
+```{r}
+# PCA ####
+cpca <- prcomp(corrmat, center=TRUE, scale.=TRUE)
+summary(cpca)
+
+# or this
+pca(corrmat) # God is "good" 
+# get data into shape ####
+# # check the response time log (how I would handle this)
+hist(log(d$RT)) # looks OK
+# create data frame ####
+#  center/standardise for meaningful intercept and for reliable interactions (also for estimation)
+dd<-d%>%
+  dplyr::select(
+    R_attendance,
+    R_religiousID,
+    R_member, # can lose this
+    R_god,
+    R_afterlife,
+    R_prayer,
+    religiosity,
+    credible,source,country,subject,Statement,Gender,Education,SES,culturalNorms,Age,source_first,sourceCorrect,RT,important)%>%
+  dplyr::mutate(subject = factor(subject),
+                Statement = factor(Statement), 
+                Male = factor(ifelse(Gender =="man","male","not_male")),
+                Edu.C = scale(Education,center=T,scale=F),
+                SES.C = scale(SES, center=T,scale=F),
+                culturalNorms.S = scale(culturalNorms,center=T,scale=T),
+                logResp = log(RT),
+                logResp.C = scale(logResp,scale=F,center=TRUE),
+                Age.Decades.C = scale(Age, scale = FALSE, center = TRUE)/10)%>%
+  dplyr::mutate(R_attendance.S = scale(R_attendance),
+                R_religiousID.S =scale(R_religiousID),
+                R_member.S = scale(R_member),
+                R_god.S=scale(R_god),
+                R_prayer.S =scale(R_prayer),
+                R_afterlife.S =scale(R_afterlife),
+                religiosity.S = scale(religiosity))
+
+# Get PCA
+dfp<-data.frame(cpca$x[,1:2])
+ddd<-as.data.frame(cbind(dd,dfp)) # data for PCA comparison
+head(ddd)
+
+```{r}
+# detour ignore#### 
+# table1::table1( ~culturalNorms.S*R_god  |country, data =ddd, transpose =T)
+# ```
+# ```{r}
+# library(sjPlot)
+# ek8<-  plot_xtab(ddd$country,ddd$ R_god, show.n=F,
+#                  show.prc = F , show.total = F,drop.empty = F) +  scale_fill_aaas() +scale_y_continuous(limits=c(0,.30)) + xlab("Country") +  ylab("Frequency God Beliefs") #+
+# #  scale_y_continuous(limits=c()) + #theme(plot.title = element_text(size=09))
+# # theme(axis.text.x = element_text(angle = 0, hjust=1,size =10))
 # 
-# graph adjustment sets
-ggdag::ggdag_adjustment_set(tidy_ggdag, node_size = 14) +
-  theme(legend.position = "bottom")
-ggdag_dseparated(tidy_ggdag)
-ggdag_collider(tidy_ggdag)
-ggdag_instrumental(tidy_ggdag)å
+# ek8
+# ek9<-  plot_xtab(ddd$country,ddd$culturalNorms, show.n=F,
+#                  show.prc = F, show.total = F,drop.empty = F) +  scale_fill_aaas() +
+#   scale_y_continuous(limits=c(0,.30)) +  xlab("Country") +  ylab("Frequency Cultural Norms Perc. Religiosity")
+# 
+# ek9
+# compare_god_norm<-ggarrange(ek8,ek9,nrow=2) # 3 x 1
+# compare_god_norm
 ```
+
+```{r}
+# filthy frequentist model (which we need to check for Variance Inflaction Factors (multicollinearity)) # if our religion variables were not highly correlated we could include all/most of them to understand the dimensionality of religousity on credibility.
+library(lme4)
+# saturated model
+d1<-lmer(credible ~ (R_attendance.S + R_religiousID.S + R_member.S + R_god.S +R_afterlife.S) * source +
+           culturalNorms.S * source +
+           Statement +
+           source_first +
+           sourceCorrect +
+           Age.Decades.C +
+           Edu.C +
+           Male +
+           SES.C +
+           logResp.C+
+           (source|country), data=ddd )
+
+# our model
+d2<-lmer(credible ~religiosity.S * source +  
+           culturalNorms.S * source +
+           Statement +
+           source_first +
+           sourceCorrect +
+           Age.Decades.C +
+           Edu.C +
+           Male +
+           SES.C +
+           logResp.C+
+           (source|country), data=ddd )
+
+# not our model which is just as good as the preferred
+# first and second PC, which predicts 80%+ of the variance without multi-collinearity 
+d3<-lmer(credible ~(PC1 + PC2) * source  + 
+           culturalNorms.S * source +
+           Statement +
+           source_first +
+           sourceCorrect +
+           Age.Decades.C  +
+           Edu.C +
+           SES.C +
+           logResp.C+
+           (source|country), data=ddd )
+
+# god is the strongest religiosity term
+d4<-lmer(credible ~ R_god.S * source  + 
+           culturalNorms.S * source +
+           Statement +
+           source_first +
+           sourceCorrect +
+           Age.Decades.C +
+           Edu.C +
+           Male +
+           SES.C +
+           logResp.C+
+           (source|country), data=ddd )
+
+# simple PC1 model (probably what JB would have done)
+d5<-lmer(credible ~PC1 * source + 
+           culturalNorms.S * source +
+           Statement +
+           source_first +
+           sourceCorrect +
+           Age.Decades.C +
+           Edu.C +
+           Male +
+           SES.C +
+           logResp.C+
+           (source|country), data=ddd )
+# check for multi-colinearity ####
+library(car) 
+vif(d1)# very high, will increase standard errors of regression coeffs.
+
+vif(d2)# OK  our model
+
+vif(d5)# OK my preferred model
+
+library(sjPlot)
+plot_model(d2) # our model (without co-variates)
+plot_model(d2, type= "re") # our model (without co-variates)
+# fem<-plot_model(d2) # our model (without co-variates)
+# plot_model(d2) # our model (without co-variates)
+# ref<-plot_model(d2, type= "re") # our model (without co-variates)
+# plot_model(d3) # same inference
+# plot_model(d4) # same inference
+# plot_model(d5) # same inference
+# ref<-plot_model(d2, type="re")
+# lmer_coefplot<-ggarrange(fem,ref,ncol=1)
+# lmer_coefplot
+```
+
+```{r}
+plot_model(d5) # no diff
+plot_model(d5, type="re") # no diff
+```
+
+```{r}
+# Run this chunk only to satisfy yourself that our religiosity variable is good.
+pl2<-ggpredict(d2, c(terms ="religiosity.S[all]","source"), ci.lvl = 0.95)
+
+pp2<-plot(pl2)+scale_y_continuous(limits=c(1,7)) + xlab("Religiosity (standardised)") + ylab("Credible:1-7")
+
+# pl4<-ggpredict(d4, c(terms ="R_god.S[all]","source"), ci.lvl = 0.95)
+# pp4<-plot(pl4)+scale_y_continuous(limits=c(1,7))+ xlab("Belief in God (standardised)") + ylab("Credible:1-7")
+# 
+# pl5<-ggpredict(d5, c(terms ="PC1[all]","source"), ci.lvl = 0.95)
+# pp5<-plot(pl5)+scale_y_continuous(limits=c(1,7)) + xlab("First Principle Component: Religion Indicators") + ylab("Credible:1-7")
+# 
+# 
+# pl2.1<-ggpredict(d2, c(terms ="culturalNorms.S[all]","source"), ci.lvl = 0.95)
+# 
+# pp2.1<-plot(pl2.1)+scale_y_continuous(limits=c(1,7)) + xlab("Religiosity (standardised), Religiosity Model") + ylab("Credible:1-7")
+# 
+# pl5.1<-ggpredict(d5, c(terms ="culturalNorms.S[all]","source"), ci.lvl = 0.95)
+# pp5.1<-plot(pl5.1)+scale_y_continuous(limits=c(1,7)) + xlab("Cultural Norms (Standardised), PC1 model") + ylab("Credible:1-7")
+# # create plot of the different ways of handling religiosity 
+# lmerplot<-ggarrange(pp2,pp5,pp2.1,pp5.1,labels=c("a","b","c","d"),common.legend = T)
+# lmerplot # 1000 x 1000
+# create plot of our preferred model for the 2 outcomes of interest
+# lmerquick<-ggarrange(ggarrange(lmer_coefplot),ggarrange(pp2,pp2.1,ncol=2),heights=c(2,1),nrow=2)
+# lmerquick # 2000 x 2
+```
+
+```{r}
+#table of our pref model
+library(sjPlot)
+tab_model(d2) # see documentation for options
+```
+
+# bayesian model ####  I'm not running this again, but you can add this to the next chunk and run
+m1 <- brm(credible ~ religiosity.S * source +
+            culturalNorms.S * source +
+            Statement +
+            source_first +
+            sourceCorrect +
+            Age.Decades.C +
+            Edu.C +
+            Male +
+            SES.C +
+            logResp.C+
+            (1+source|country),
+          data=ddd,
+          family = "gaussian",
+          seed=1234,
+          warmup=1500,
+          iter=6500,
+          chains=4)
+
+saveRDS(m1, "m1")
+
+```{r}
+m1<-readRDS("m1") # to avoid running this again in knitr, probably a better way
+#summary(m1) # raw summary
+# compare freq estimates with mcmc
+# tab_model(d2,m1)
+```
+
+```{r}
+# diagnositics
+mcmc_plot(m1,
+          type = "trace")
+# mcmc_plot(m1,
+#           type = "areas")
+# mcmc_plot(m1,
+#           type = "violin")
+# mcmc_plot(m1,
+#           type = "intervals")
+
+# we see the oridinal qualities of our dataset here
+ppg<-pp_check(m1)
+ppg
+```
+
+```{r}
+# for math
+#stancode(m1)
+```
+```{r}
+# stanplot(m1, pars = NA,
+#          type = "intervals",
+#          fixed = FALSE,
+#          prob_outer = 0.95,
+#          exact_match = FALSE)
+```
+```{r}
+#better plot
+# m1fe<-plot_model(m1, title= "Credibility Coefficient Plot")
+# m1fe #1 x 1
+# 
+# m1re<-plot_model(m1, type ="re", sort.est = TRUE)
+# m1re
+# even better plot
+library(see)
+library(parameters)
+par<-parameters::model_parameters(m1,effects = "all", component = "all", test = c("pd"))
+par
+par2<-parameters::model_parameters(m1)
+#param_plot<- plot(par,  sort = TRUE) +theme_lucid() #option
+#param_plot<- plot(par,  sort = TRUE) +theme_abyss() #option
+#param_plot<- plot(par,  sort = TRUE) +theme_radar()#option
+param_plot<- plot(par,  sort = TRUE) +theme_radar()
+#param_plot<- plot(par,  sort = TRUE) +theme_modern()#option
+#param_plot<- plot(par,  sort = TRUE) +theme_blackboard()#option
+param_plot # my pref.
+```
+
+```{r}
+# not used
+# statcon<-make_conditions(dd, vars = c("REL"))
+# statcon
+# str(statcon)
+# cn = data.frame("REL" = c(1,.5,0), "cond__" = c("REL = 1","REL = .5", "REL = 0"))
+
+# Run these lines, I'm just commenting out for the document
+p1<-brms::conditional_effects(m1, c("religiosity.S:source"),
+                               spaghetti = T,
+                              nsamples = 100,
+                              probs = c(0.11, 0.89))
+pp1<-plot(p1, plot = FALSE)[[1]] + scale_y_continuous(limits = c(1,7))  + xlab("Religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Religiosity")  + theme_clean() #theme_radar_dark() + scale_fill_npg()
+pp1 #
+p2<-brms::conditional_effects(m1, c("culturalNorms.S:source"),
+                               spaghetti = T,
+                              nsamples = 100,
+                              probs = c(0.11, 0.89))
+pp2<-plot(p2, plot = FALSE)[[1]] + scale_y_continuous(limits = c(1,7))  + xlab("Cultural Norms of religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Cultural Norms") + theme_clean() #+ theme_radar_dark() + scale_fill_npg()
+pp2 #
+library(ggpubr)
+plotgood<-ggarrange(pp1, pp2, nrow=1, common.legend = T)
+#saveRDS(plotgood,"plotgood")
+#plotgood<-readRDS("plotgood")
+plotgood
+```
+
+# NOTE: an ordinal model would be better, but will require time to run, and it might muddy our inference.
+# 
+m1o <- brm(credible ~ religiosity.S * source +
+             culturalNorms.S * source +
+             Statement +
+             source_first +
+             sourceCorrect +
+             Age.Decades.C +
+             Edu.C +
+             Male +
+             SES.C +
+             logResp.C+
+             (1+source|country),
+           data=ddd,
+           inits = 0,
+           family = cumulative(link ="logit"),
+           seed=1234,
+           warmup=1000,
+           iter=3000,
+           chains=4)
+
+saveRDS(m1o,"m1o")
+
+# table and plots
+```{r}
+
+m1o<-readRDS("m1o")
+ppo<-pp_check(m1o)  ## fit is much better
+ppo
+ordinalmod<-parameters::model_parameters(m1o,effects = "all", component = "all", test = c("pd"))
+ordinalmod
+oplot<-plot(ordinalmod,  sort = TRUE) + theme_radar()
+oplot
+compareordplot<-ggarrange(ggarrange(param_plot,oplot,labels=c("a","b")),ggarrange(ppg, ppo, labels=c("c","d")), nrow=2, heights = c(3,1))
+compareordplot
+```
+## next model
+## have not inclueded this in the knitr call to save time. 
+```{r}
+statcon<-make_conditions(ddd, vars = c("source"))
+statcon
+str(statcon)
+
+p1<-brms::conditional_effects(m1o, c("religiosity.S"),
+                              categorical =T, # 
+                              probs = c(0.11, 0.89),
+                              conditions = statcon)
+pp1<-plot(p1, plot = FALSE)[[1]]   + xlab("Religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Religiosity")  + scale_y_continuous(limits = c(0,.4))  + theme_clean() + scale_color_grey() +scale_fill_grey() # note of you set categorical to true, then scale the y axis limits: c(0,1).
+Same for the plot below
+pp1 #
+p2<-brms::conditional_effects(m1o, c("culturalNorms.S"),
+                               categorical =T,
+                              probs = c(0.11, 0.89),
+                              conditions = statcon)
+pp2<-plot(p2, plot = FALSE)[[1]] + scale_y_continuous(limits = c(0,.4))  + xlab("Cultural Norms of Religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Cultural Norms")  + theme_clean() + scale_color_grey() +scale_fill_grey()
+pp2 #
+library(ggpubr)
+plotgood2.0<-ggarrange(pp1, pp2, labels=c("a","b"), nrow=1, common.legend = T)
+plotgood2.0
+saveRDS(plotgood2.0,"plotgood2.0")
+```
+
+```{r}
+# plot that can be read
+## We can set categorical to false to get more interpretable figures
+p1<-brms::conditional_effects(m1o, c("religiosity.S:source"),
+                              categorical =F, # note set to true to get all the ordinal values
+                              probs = c(0.11, 0.89),
+                                spaghetti = T,
+                              nsamples = 100) # play with this number
+pp1<-plot(p1, plot = FALSE)[[1]]   + xlab("Religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Religiosity")  + scale_y_continuous(limits = c(1,7))  + theme_clean() # note of you set categorical to true, then scale the y axis limits: c(0,1).
+Same for the plot below
+pp1 #
+p2<-brms::conditional_effects(m1o, c("culturalNorms.S:source"),
+                               categorical =F,
+                              probs = c(0.11, 0.89),
+                                spaghetti = T,
+                              nsamples = 100) # play with this number
+pp2<-plot(p2, plot = FALSE)[[1]] + scale_y_continuous(limits = c(1,7))  + xlab("Cultural Norms of Religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Cultural Norms")  + theme_clean()
+pp2 #
+library(ggpubr)
+plotgood2<-ggarrange(pp1, pp2, labels=c("c","d"), nrow=1, common.legend = T)
+plotgood2
+saveRDS(plotgood2,"plotgood2")
+```
+
+```{r}
+# for omni graph
+plotgoodO<-ggarrange(pp1, pp2, nrow=1, common.legend = T)
+plotgoodO
+```
+
+```{r}
+#plotgood2<-readRDS("plotgood2")
+#plotgood2
+# compare ordinal plots
+twotypeordinalplot<-ggarrange(plotgood2.0,plotgood2,nrow=2)
+twotypeordinalplot
+```
+```{r}
+# everything plot
+omniplot<-ggarrange(ggarrange(param_plot,oplot,labels=c("a","b")),ggarrange(ppg, ppo, labels=c("c","d")), ggarrange(plotgood, plotgoodO, labels=c("e","f")), nrow=3, heights = c(3,1,1))
+omniplot
+```
+
+# 
+# 
+# ###### IF WE WANTED BOTH INDICATORS IN THE MODEL THIS IS THE CODe
+# m2 <- brm(mvbind(credible, important) ~ 
+#             religiosity.S * source + 
+#             culturalNorms.S * source +
+#             Statement +
+#             source_first +
+#             sourceCorrect +
+#             Age.Decades.C +
+#             Edu.C +
+#             Male +
+#             SES.C +
+#             logResp.C+
+#             (1+source|country),
+#           data=ddd,
+#           family = "gaussian",
+#           seed=1234,
+#           warmup=1000,
+#           iter=3000,
+#           chains=4)
+
+```

### 2021-01-14 funs.R

@@ -27,8 +27,6 @@ get_wb_soc <- function(x) {
 }
 
 
-
-
 # this code is for citing packages
 cite_packages <- function() {
   citation(package = "base", lib.loc = "/Users/josephbulbulia/Dropbox/BIBS")

### 2021-01-14 libs.R

@@ -18,7 +18,6 @@ library("gghighlight")
 #library("Amelia")
 #library("magrittr")
 library("patchwork")
-#library("qgraph")
 library("texreg")
 library("viridis")
 library("ggpubr")

### 2021-01-14 funs.R
committing functions

@@ -5,6 +5,12 @@ data_read_mymachine <- function() {
     read.csv("/Users/jbul176/Documents/GitHub/many_analysts/MARP_data_blinded.csv")
 }
 
+
+# data_read_mymachine_full <- function() {
+#   out <-
+#     read.csv("DataStudy3_SourceCredibility_Long.csv")
+# }
+
 get_rel_vars <- function(x) {
   df <- x %>% dplyr::select(c(starts_with("rel_"), "country"))
 }

### 2021-01-14 libs.R

@@ -25,6 +25,7 @@ library("ggsci")
 library("here")
 library("citr")
 library("here")
+library("rethinking")
 #library("statsExpressions")
 library("ggstatsplot")
 library("ggthemes")

### 2021-01-14 analysis.Rmd
within and between country analysis

--- 2021-01-14 analysis.Rmd
@@
 
+Suppose we believe that
+
+- religion + ses + male + country + a different unmeasured latent variable (e.g. priviledge) causes well being
+
+- age + male + country + and Unmeasured latent variable (e.g. parent's religion, call this U1) causes Religion
+
+- country + age + another unmeasured latent variable (e.g. priviledge U2) causes ses
+
+- our "exposure" is religion and our outcome is "well-being"
+
+Which variables should I condition on to understand the relationship between religion and well-being?
+
 ```{r}
-head(df)
 tidy_ggdag <- dagify(
-  rel ~ age + country + male,
-  wb  ~ rel + country,
-  wb ~ soc  + ses,
-  ses ~ country + age, 
+  wb  ~ rel + ses + male + country + U2,
+  rel ~ age + male + country + U1,
+  ses ~ country + age + U2,
   exposure =  "rel",
-  latent = "country",
-  outcome =   "wb") %>%
+  outcome =   "wb"
+) %>%
   tidy_dagitty()
-#
+
 # graph adjustment sets
 ggdag::ggdag_adjustment_set(tidy_ggdag, node_size = 14) +
-  theme(legend.position = "bottom")
-ggdag_dseparated(tidy_ggdag)
-ggdag_collider(tidy_ggdag)
-ggdag_instrumental(tidy_ggdag)å
+  theme(legend.position = "bottom") +
+  theme_dag_blank()
+```
+
+Suppose I was to include all variables in my model, except U2 and U1, which are not measured?
+
+In this case I would introduce colider bias
+
+```{r}
+ggdag::ggdag_collider(tidy_ggdag,
+                      from = "rel",
+                      to = "wb",
+                      control_for(c("ses", "male", "country", "age")))
+
+```
+
+## missingness
+
+Next we need to examine patterns of missingness, which might spoil inference.
+
+```{r}
+library(naniar)
+vis_miss(df) # substantial pattern of missing ness in a few variables, 
+```
+
+There is little missingness so we will just use complete cases.
+Note that it would be better to have a method for handling the uncertainty of any case.
+A short cut is a short cut: missing data should be handled by some probabilitic imputation strategy.
+However because there are many larger problems about measurement error, sampling, and theory, lets set these to questions to the side and work with the complete dataset
+
+## The hard work
+
+We have yet to consider a mechanism by which religion might affect well-being, and specifically which aspects of well-being and which aspects of religion.
+Just as it might be tempting to through all variables into a regression equation, and disasterious to do so, it might be tempting to leap into some form of data reduction, and combine well-being and religious indicators without stopping to think about how causation.
+
+There are many theories we might consider.
+We will focus on two non-trivial theories, one related to church attendance and one related to belief in god.
+If we had more time we could consider many others, but as Catholics know, "timing is everything" -- let's keep to time.
+
+### Church attendance
+
+We propose that church attendance reduces psychological anxiety by increasing perceived support.
+This is a new model so we need to write our dag all over again.
+The dag will resemble the dag above, except we know that people who identify as male have less social support (typically), at least in some countries.
+
+Writing the dag
+```{r}
+tidy_ggdag2 <- dagify(
+  wb  ~ support + ses + male + country + U2,
+  support ~ church + male + country,
+  church ~ age + male + country + U1,
+  ses ~ country + age + U2,
+  exposure =  "church",
+  outcome =   "wb"
+) %>%
+  tidy_dagitty()
+
+# graph adjustment sets
+ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) +
+  theme(legend.position = "bottom") +
+  theme_dag_blank()
+```
+
+The adjustments sets are the same so we proceed with a model.
+Note that if my interest is in the direct effect of church attendance on wellbeing I should *not* include social support because this introduces a collider bias!
+
+```{r}
+ggdag::ggdag_collider(tidy_ggdag2,
+                      from = "church",
+                      to = "wb",
+                      controlling_for(c("male", "support", "country", "age")))
+
+```
+
+## The model
+
+A bayesian workflow should include prior predictive checks and the use of informative priors (McElreath 2020).
+
+For this initial analysis I am going to use the weakly informative priors that BRMS uses; my experience is that large datasets are resistent to choice of priors.
+However it is a mistake to trust this experience.
+For now, though, we must keep to time, and timing is everything.
+
+Additionally, the data have been manipulated in advance of the study to be scaled to a range between 0 and 1.
+This is a very bad idea.
+It would be much better to thinking of church attendance frequence as a monotonic effect.
+I will do that analysis at the next round, if the data are made available in their original form.
+
+Here is the model for church attendance and anxiety
+
+Wrangle data
+```{r}
+# get data into shape
+d1 <- df %>%
+  dplyr::filter(
+    !is.na(wb_psychANXIETY),!is.na(age),!is.na(rel_CHURCHFREQ),!is.na(wb_psychMEANING),!is.na(male),!is.na(rel_EXTENTBELIEVEGOD)
+  ) %>%
+  dplyr::mutate(
+    anx = as.numeric(wb_psychANXIETY),
+    anx_s = scale(wb_psychANXIETY),
+    #put outcome in ST units (see McElreath 2021)  # not necessary
+    age_s = scale(age),
+    ch = rel_CHURCHFREQ,
+    mn  = wb_psychMEANING,
+    bg =  rel_EXTENTBELIEVEGOD,
+    ch_s = scale(rel_CHURCHFREQ),
+    mn_s = scale(wb_psychMEANING),
+    bg_s = scale(rel_EXTENTBELIEVEGOD)
+  ) %>%
+  dplyr::select(anx, ch, mn, bg, anx_s, age_s, ch_s, mn_s, fm1m2, country, male, bg_s)
+table(d1$country)
+
+## degroup variables for within and between analysis
+## https://cran.r-project.org/web/packages/parameters/vignettes/demean.html
+d1 <-
+  cbind(d1,
+        demean(d1, select = c("anx", "ch", "bg"), group = "country"))
+head(d1)
+```
+
+We might estimate that effects differ within and between groups, but ignoring within and between country heterogenity and simply pooling effects at the country level
+
+```{r}
+m1 <- brm(
+  anx_s ~ ch_s +
+    age_s  +
+    male +
+    (1 + ch_s | country),
+  data = d1,
+  family = "gaussian",
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = "m1"
+)
+```
+
+very small effect
+
+```{r}
+plot(ggeffects::ggpredict(m1, terms = c("ch_s[minmax]"),type = "fe", ci.lvl = 0.89))  +
+   scale_fill_brewer(palette= "Set2") + labs(
+    title = "Predicted effects of church on anxiety")
+```
+
+Countries
+
+```{r}
+plot(ggeffects::ggpredict(m1, terms = c("country[all]", "ch_s[minmax]"),type = "re", ci.lvl = 0.89))  +
+   scale_fill_brewer(palette= "Set2") + labs(
+    title = "Predicted effects of church on anxiety")
+```
+
+## Within and between
+
+```{r}
+m1_a <- brm(
+  anx_s ~ age_s  + male + ch_within +  ch_between + (1 + ch_within|country + ch_between | country),  
+  data = d1,
+  family = "gaussian",
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = "m1_a.rds")
+summary(m1_a)
+
 ```
@@
 ```{r}
-head(df)
-tidy_ggdag <- dagify(
-  rel ~ age + country,
-  wb ~ soc  + ses + rel + age,
-  ses ~ country + age, 
-  exposure =  "rel",
-  latent = "country",
+plot(ggeffects::ggpredict(m1_a, terms = c("ch_within[minmax]"),type = "re", ci.lvl = 0.89))  +
+   scale_fill_brewer(palette= "Set2") + labs(
+    title = "Predicted effects of church on anxiety")
+```
+
+```{r}
+plot(ggeffects::ggpredict(m1_a, terms = c("ch_between[minmax]"),type = "re", ci.lvl = 0.89))  +
+   scale_fill_brewer(palette= "Set2") + labs(
+    title = "Predicted effects of church on anxiety")
+```
+
+```{r}
+plot(ggeffects::ggpredict(m1_a, terms = c("country[all]", "ch_between[minmax]"),type = "re", ci.lvl = 0.89))  +
+   scale_fill_brewer(palette= "Set2") + labs(
+    title = "Predicted effects of church on anxiety")
+```
+
+## Belief in God
+
+Lets consider the non-trivial theory that belief in God (bg) affects psychological anxiety (wb) by affording a sense of meaning in life.
+The causal qualities will be the same as those of the previous model, and so too will the threats to inference.
+
+Writing the dag
+```{r}
+tidy_ggdag2 <- dagify(
+  wb  ~ meaning + ses + male + country  +U2,
+  meaning ~ bg +  male + country, 
+  bg ~ age + male + country + U1,
+  ses ~ country + age + U2,
+  exposure =  "bg",
   outcome =   "wb") %>%
   tidy_dagitty()
+
+# graph adjustment sets
+ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) + 
+  theme(legend.position = "bottom") + 
+  theme_dag_blank()
+```
+
+And again we face a collider bias by introducing meaning measure, conditional on assumed causal relationships
+
+```{r}
+ggdag::ggdag_collider(tidy_ggdag2,
+                      from = "church",
+                      to = "wb",
+                      controlling_for(c("male","support","country","age")))
+
+```
+
+### model for belief
+
+From causal dag above
+
+```{r}
+
+m2 <- brm(
+  anx_s ~
+    bg_s  +
+    age_s  +
+    male +
+    (1 + ch_s | country),
+  data = d1,
+  family = "gaussian",
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = "m2"
+)
+summary(m2)
+```
+
+Graph -- No effect, why? 
+```{r}
+p2 <- brms::conditional_effects(m2,
+                                c("bg_s"),
+                                categorical = F,
                                 #
-# graph adjustment sets
-ggdag::ggdag_adjustment_set(tidy_ggdag, node_size = 14) +
-  theme(legend.position = "bottom")
-ggdag_dseparated(tidy_ggdag)
-ggdag_collider(tidy_ggdag)
-ggdag_instrumental(tidy_ggdag)å
+                                probs = c(0.11, 0.89))
+plot(p2)
 ```
+
+```{r}
+plot(ggeffects::ggpredict(m2, terms = c("country[all]", "bs_s[minmax]"),type = "re", ci.lvl = 0.89))  +
+   scale_fill_brewer(palette= "Set2") + labs(
+    title = "Predicted effects of church on anxiety")
+```
+
+## Within and between
+
+```{r}
+m2_a <- brm(
+  anx_s ~ age_s  + male + ch_within +  bg_between + (1 + bg_within|country + bg_between | country),  
+  data = d1,
+  family = "gaussian",
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = "m2_a.rds")
+summary(m2_a)
+```
+
+Nothing
+
+```{r}
+plot(ggeffects::ggpredict(m2_a, terms = "bg_within[minmax]",type = "re", ci.lvl = 0.89))  +
+   scale_fill_brewer(palette= "Set2") + labs(
+    title = "Predicted effects of belief on anxiety")
+```
+
+Something
+```{r}
+plot(ggeffects::ggpredict(m2_a, terms = "bg_between[minmax]",type = "re", ci.lvl = 0.89))  +
+   scale_fill_brewer(palette= "Set2") + labs(
+    title = "Predicted effects of belief on anxiety")
+```
+
+```{r}
+plot(ggeffects::ggpredict(m2_a, terms = c("country[all]", "bg_within[minmax]"),type = "re", ci.lvl = 0.89))  +
+   scale_fill_brewer(palette= "Set2") + labs(
+    title = "Predicted effects of belief on anxiety")
+```
+
+```{r}
+plot(ggeffects::ggpredict(m2_a, terms = c("country[all]", "bg_between[minmax]"),type = "re", ci.lvl = 0.89))  +
+   scale_fill_brewer(palette= "Set2") + labs(
+    title = "Predicted effects of belief on anxiety")
+```
+
+<!-- ### OLD ANALYSIS -->
+
+<!-- # -->
+<!-- #setwd("/Users/jbul176/The\ Virtues\ Project\ Dropbox/Joseph\ Bulbulia/00Bulbulia\ Pubs/2020/2020_VAN_ELK") # mine -->
+
+<!-- # read data -->
+<!-- e <- read.csv("/DataStudy3_SourceCredibility_Long.csv", header=T) # old data within/group means -->
+<!-- d <- read.csv('DataLong', header=T) # new data without ingroup means -->
+<!-- #d$statementAndOrder # error but this is not needed -->
+<!-- nrow(d)#20390 -->
+<!-- hist(e$religiosity) # group standardized data -->
+<!-- hist(d$religiosity) # note difference -->
+<!-- # check data : note that subject needs to be a factor. also, "Statement" needs to be a factor. for gender there are few "others" we create a "male gender variable (below) for simplicity.  -->
+<!-- # -->
+<!-- # look for patters of missingness #### -->
+<!-- library(naniar) -->
+<!-- vis_miss(d) # substantial pattern of missing ness in a few variables,  -->
+<!-- ``` -->
+<!-- ```{r} -->
+<!-- #plot outcomes: gaussian is reasonable, but we'll come back to this after ppchecks -->
+<!-- str(d) -->
+<!-- # check outcomes -->
+<!-- hist(d$credible) # gaussian ordinal -->
+<!-- hist(d$important)# flat -->
+<!-- # what is the F1 variable? -->
+<!-- hist(d$recallFa ) ##?? -->
+<!-- hist(d$recallHit )##?? -->
+<!-- ``` -->
+<!-- ```{r} -->
+<!-- #make tables #### -->
+<!-- # and look for anomolies -->
+<!-- library(table1) # useful package -->
+<!-- table1::table1(~Age + Gender + Education + R_religiousID + religiosity + culturalNorms|country, data =d, transpose =T) -->
+<!-- ``` -->
+<!-- ```{r} -->
+<!-- table1::table1(~credible|country, data =d, transpose =TRUE)# outcome variables -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- table1::table1( ~R_attendance + R_prayer + R_religiousID + R_member  + R_god   + R_afterlife|country, data =d, transpose =T) -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- # how correlated are the religion variables? #### -->
+<!-- corrmat<-d%>% -->
+<!--   dplyr::select( -->
+<!--     R_attendance, -->
+<!--     R_religiousID, -->
+<!--     R_member, # can lose this -->
+<!--     R_god, -->
+<!--     R_afterlife, -->
+<!--     R_prayer)%>% -->
+<!--   dplyr::mutate(R_attendance.S = scale(R_attendance), -->
+<!--                 R_religiousID.S =scale(R_religiousID), -->
+<!--                 R_member.S = scale(R_member), # not used -->
+<!--                 R_god.S=scale(R_god), -->
+<!--                 R_prayer.S =scale(R_prayer), -->
+<!--                 R_afterlife.S =scale(R_afterlife))%>% -->
+<!--   rowwise() %>%  -->
+<!--   mutate(REL= mean(c(R_attendance, R_prayer,R_god,R_religiousID,R_afterlife), na.rm=T))%>% # we don't use this, but my sense is R_member isn't going to be worth including. -->
+<!--   mutate(REL.S = scale(REL))%>% # we don't use this -->
+<!--   select(R_attendance.S,R_religiousID.S,R_member.S,R_god.S,R_afterlife.S,R_prayer.S) # don't select religiosity for PCA -->
+<!-- # corrplot -->
+<!-- res <- cor(corrmat, method="pearson") -->
+<!-- corrplot::corrplot(res, method= "color", order = "hclust") # tl.pos = 'n' -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- ## corr.soc   -->
+<!-- sjp.corr(corrmat, -->
+<!--          title = "Correlation matrix of religion variables", -->
+<!--          geom.colors = "RdBu") -->
+<!-- ``` -->
+<!-- ```{r} -->
+<!-- # PCA #### -->
+<!-- cpca <- prcomp(corrmat, center=TRUE, scale.=TRUE) -->
+<!-- summary(cpca) -->
+
+<!-- # or this -->
+<!-- pca(corrmat) # God is "good"  -->
+<!-- # get data into shape #### -->
+<!-- # # check the response time log (how I would handle this) -->
+<!-- hist(log(d$RT)) # looks OK -->
+<!-- # create data frame #### -->
+<!-- #  center/standardise for meaningful intercept and for reliable interactions (also for estimation) -->
+<!-- dd<-d%>% -->
+<!--   dplyr::select( -->
+<!--     R_attendance, -->
+<!--     R_religiousID, -->
+<!--     R_member, # can lose this -->
+<!--     R_god, -->
+<!--     R_afterlife, -->
+<!--     R_prayer, -->
+<!--     religiosity, -->
+<!--     credible,source,country,subject,Statement,Gender,Education,SES,culturalNorms,Age,source_first,sourceCorrect,RT,important)%>% -->
+<!--   dplyr::mutate(subject = factor(subject), -->
+<!--                 Statement = factor(Statement),  -->
+<!--                 Male = factor(ifelse(Gender =="man","male","not_male")), -->
+<!--                 Edu.C = scale(Education,center=T,scale=F), -->
+<!--                 SES.C = scale(SES, center=T,scale=F), -->
+<!--                 culturalNorms.S = scale(culturalNorms,center=T,scale=T), -->
+<!--                 logResp = log(RT), -->
+<!--                 logResp.C = scale(logResp,scale=F,center=TRUE), -->
+<!--                 Age.Decades.C = scale(Age, scale = FALSE, center = TRUE)/10)%>% -->
+<!--   dplyr::mutate(R_attendance.S = scale(R_attendance), -->
+<!--                 R_religiousID.S =scale(R_religiousID), -->
+<!--                 R_member.S = scale(R_member), -->
+<!--                 R_god.S=scale(R_god), -->
+<!--                 R_prayer.S =scale(R_prayer), -->
+<!--                 R_afterlife.S =scale(R_afterlife), -->
+<!--                 religiosity.S = scale(religiosity)) -->
+
+<!-- # Get PCA -->
+<!-- dfp<-data.frame(cpca$x[,1:2]) -->
+<!-- ddd<-as.data.frame(cbind(dd,dfp)) # data for PCA comparison -->
+<!-- head(ddd)  -->
+
+<!-- ```{r} -->
+<!-- # detour ignore####  -->
+<!-- # table1::table1( ~culturalNorms.S*R_god  |country, data =ddd, transpose =T) -->
+<!-- # ``` -->
+<!-- # ```{r} -->
+<!-- # library(sjPlot) -->
+<!-- # ek8<-  plot_xtab(ddd$country,ddd$ R_god, show.n=F, -->
+<!-- #                  show.prc = F , show.total = F,drop.empty = F) +  scale_fill_aaas() +scale_y_continuous(limits=c(0,.30)) + xlab("Country") +  ylab("Frequency God Beliefs") #+ -->
+<!-- # #  scale_y_continuous(limits=c()) + #theme(plot.title = element_text(size=09)) -->
+<!-- # # theme(axis.text.x = element_text(angle = 0, hjust=1,size =10)) -->
+<!-- #  -->
+<!-- # ek8 -->
+<!-- # ek9<-  plot_xtab(ddd$country,ddd$culturalNorms, show.n=F, -->
+<!-- #                  show.prc = F, show.total = F,drop.empty = F) +  scale_fill_aaas() + -->
+<!-- #   scale_y_continuous(limits=c(0,.30)) +  xlab("Country") +  ylab("Frequency Cultural Norms Perc. Religiosity") -->
+<!-- #  -->
+<!-- # ek9 -->
+<!-- # compare_god_norm<-ggarrange(ek8,ek9,nrow=2) # 3 x 1 -->
+<!-- # compare_god_norm -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- # filthy frequentist model (which we need to check for Variance Inflaction Factors (multicollinearity)) # if our religion variables were not highly correlated we could include all/most of them to understand the dimensionality of religousity on credibility.
+-->
+<!-- library(lme4) -->
+<!-- # saturated model -->
+<!-- d1<-lmer(credible ~ (R_attendance.S + R_religiousID.S + R_member.S + R_god.S +R_afterlife.S) * source + -->
+<!--            culturalNorms.S * source + -->
+<!--            Statement + -->
+<!--            source_first + -->
+<!--            sourceCorrect + -->
+<!--            Age.Decades.C + -->
+<!--            Edu.C + -->
+<!--            Male + -->
+<!--            SES.C + -->
+<!--            logResp.C+ -->
+<!--            (source|country), data=ddd )   -->
+
+<!-- # our model -->
+<!-- d2<-lmer(credible ~religiosity.S * source +   -->
+<!--            culturalNorms.S * source + -->
+<!--            Statement + -->
+<!--            source_first + -->
+<!--            sourceCorrect + -->
+<!--            Age.Decades.C + -->
+<!--            Edu.C + -->
+<!--            Male + -->
+<!--            SES.C + -->
+<!--            logResp.C+ -->
+<!--            (source|country), data=ddd )   -->
+
+<!-- # not our model which is just as good as the preferred -->
+<!-- # first and second PC, which predicts 80%+ of the variance without multi-collinearity  -->
+<!-- d3<-lmer(credible ~(PC1 + PC2) * source  +  -->
+<!--            culturalNorms.S * source + -->
+<!--            Statement + -->
+<!--            source_first + -->
+<!--            sourceCorrect + -->
+<!--            Age.Decades.C  + -->
+<!--            Edu.C + -->
+<!--            SES.C + -->
+<!--            logResp.C+ -->
+<!--            (source|country), data=ddd ) -->
+
+<!-- # god is the strongest religiosity term -->
+<!-- d4<-lmer(credible ~ R_god.S * source  +  -->
+<!--            culturalNorms.S * source + -->
+<!--            Statement + -->
+<!--            source_first + -->
+<!--            sourceCorrect + -->
+<!--            Age.Decades.C + -->
+<!--            Edu.C + -->
+<!--            Male + -->
+<!--            SES.C + -->
+<!--            logResp.C+ -->
+<!--            (source|country), data=ddd ) -->
+
+<!-- # simple PC1 model (probably what JB would have done) -->
+<!-- d5<-lmer(credible ~PC1 * source +  -->
+<!--            culturalNorms.S * source + -->
+<!--            Statement + -->
+<!--            source_first + -->
+<!--            sourceCorrect + -->
+<!--            Age.Decades.C + -->
+<!--            Edu.C + -->
+<!--            Male + -->
+<!--            SES.C + -->
+<!--            logResp.C+ -->
+<!--            (source|country), data=ddd ) -->
+<!-- # check for multi-colinearity #### -->
+<!-- library(car)  -->
+<!-- vif(d1)# very high, will increase standard errors of regression coeffs.  -->
+
+<!-- vif(d2)# OK  our model -->
+
+<!-- vif(d5)# OK my preferred model -->
+
+<!-- library(sjPlot) -->
+<!-- plot_model(d2) # our model (without co-variates) -->
+<!-- plot_model(d2, type= "re") # our model (without co-variates) -->
+<!-- # fem<-plot_model(d2) # our model (without co-variates) -->
+<!-- # plot_model(d2) # our model (without co-variates) -->
+<!-- # ref<-plot_model(d2, type= "re") # our model (without co-variates) -->
+<!-- # plot_model(d3) # same inference -->
+<!-- # plot_model(d4) # same inference -->
+<!-- # plot_model(d5) # same inference -->
+<!-- # ref<-plot_model(d2, type="re") -->
+<!-- # lmer_coefplot<-ggarrange(fem,ref,ncol=1) -->
+<!-- # lmer_coefplot -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- plot_model(d5) # no diff -->
+<!-- plot_model(d5, type="re") # no diff -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- # Run this chunk only to satisfy yourself that our religiosity variable is good. -->
+<!-- pl2<-ggpredict(d2, c(terms ="religiosity.S[all]","source"), ci.lvl = 0.95) -->
+
+<!-- pp2<-plot(pl2)+scale_y_continuous(limits=c(1,7)) + xlab("Religiosity (standardised)") + ylab("Credible:1-7") -->
+
+<!-- # pl4<-ggpredict(d4, c(terms ="R_god.S[all]","source"), ci.lvl = 0.95) -->
+<!-- # pp4<-plot(pl4)+scale_y_continuous(limits=c(1,7))+ xlab("Belief in God (standardised)") + ylab("Credible:1-7") -->
+<!-- #  -->
+<!-- # pl5<-ggpredict(d5, c(terms ="PC1[all]","source"), ci.lvl = 0.95) -->
+<!-- # pp5<-plot(pl5)+scale_y_continuous(limits=c(1,7)) + xlab("First Principle Component: Religion Indicators") + ylab("Credible:1-7") -->
+<!-- #  -->
+<!-- #  -->
+<!-- # pl2.1<-ggpredict(d2, c(terms ="culturalNorms.S[all]","source"), ci.lvl = 0.95) -->
+<!-- #  -->
+<!-- # pp2.1<-plot(pl2.1)+scale_y_continuous(limits=c(1,7)) + xlab("Religiosity (standardised), Religiosity Model") + ylab("Credible:1-7") -->
+<!-- #  -->
+<!-- # pl5.1<-ggpredict(d5, c(terms ="culturalNorms.S[all]","source"), ci.lvl = 0.95) -->
+<!-- # pp5.1<-plot(pl5.1)+scale_y_continuous(limits=c(1,7)) + xlab("Cultural Norms (Standardised), PC1 model") + ylab("Credible:1-7") -->
+<!-- # # create plot of the different ways of handling religiosity  -->
+<!-- # lmerplot<-ggarrange(pp2,pp5,pp2.1,pp5.1,labels=c("a","b","c","d"),common.legend = T) -->
+<!-- # lmerplot # 1000 x 1000 -->
+<!-- # create plot of our preferred model for the 2 outcomes of interest -->
+<!-- # lmerquick<-ggarrange(ggarrange(lmer_coefplot),ggarrange(pp2,pp2.1,ncol=2),heights=c(2,1),nrow=2) -->
+<!-- # lmerquick # 2000 x 2 -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- #table of our pref model -->
+<!-- library(sjPlot) -->
+<!-- tab_model(d2) # see documentation for options -->
+<!-- ``` -->
+
+<!-- # bayesian model ####  I'm not running this again, but you can add this to the next chunk and run -->
+<!-- m1 <- brm(credible ~ religiosity.S * source + -->
+<!--             culturalNorms.S * source + -->
+<!--             Statement + -->
+<!--             source_first + -->
+<!--             sourceCorrect + -->
+<!--             Age.Decades.C + -->
+<!--             Edu.C + -->
+<!--             Male + -->
+<!--             SES.C + -->
+<!--             logResp.C+ -->
+<!--             (1+source|country), -->
+<!--           data=ddd, -->
+<!--           family = "gaussian", -->
+<!--           seed=1234, -->
+<!--           warmup=1500, -->
+<!--           iter=6500, -->
+<!--           chains=4) -->
+
+<!-- saveRDS(m1, "m1") -->
+
+<!-- ```{r} -->
+<!-- m1<-readRDS("m1") # to avoid running this again in knitr, probably a better way -->
+<!-- #summary(m1) # raw summary -->
+<!-- # compare freq estimates with mcmc -->
+<!-- # tab_model(d2,m1) -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- # diagnositics -->
+<!-- mcmc_plot(m1, -->
+<!--           type = "trace") -->
+<!-- # mcmc_plot(m1, -->
+<!-- #           type = "areas") -->
+<!-- # mcmc_plot(m1, -->
+<!-- #           type = "violin") -->
+<!-- # mcmc_plot(m1, -->
+<!-- #           type = "intervals") -->
+
+<!-- # we see the oridinal qualities of our dataset here -->
+<!-- ppg<-pp_check(m1) -->
+<!-- ppg -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- # for math -->
+<!-- #stancode(m1) -->
+<!-- ``` -->
+<!-- ```{r} -->
+<!-- # stanplot(m1, pars = NA, -->
+<!-- #          type = "intervals", -->
+<!-- #          fixed = FALSE, -->
+<!-- #          prob_outer = 0.95, -->
+<!-- #          exact_match = FALSE) -->
+<!-- ``` -->
+<!-- ```{r} -->
+<!-- #better plot -->
+<!-- # m1fe<-plot_model(m1, title= "Credibility Coefficient Plot") -->
+<!-- # m1fe #1 x 1 -->
+<!-- #  -->
+<!-- # m1re<-plot_model(m1, type ="re", sort.est = TRUE) -->
+<!-- # m1re -->
+<!-- # even better plot -->
+<!-- library(see) -->
+<!-- library(parameters) -->
+<!-- par<-parameters::model_parameters(m1,effects = "all", component = "all", test = c("pd")) -->
+<!-- par -->
+<!-- par2<-parameters::model_parameters(m1) -->
+<!-- #param_plot<- plot(par,  sort = TRUE) +theme_lucid() #option -->
+<!-- #param_plot<- plot(par,  sort = TRUE) +theme_abyss() #option -->
+<!-- #param_plot<- plot(par,  sort = TRUE) +theme_radar()#option -->
+<!-- param_plot<- plot(par,  sort = TRUE) +theme_radar() -->
+<!-- #param_plot<- plot(par,  sort = TRUE) +theme_modern()#option -->
+<!-- #param_plot<- plot(par,  sort = TRUE) +theme_blackboard()#option -->
+<!-- param_plot # my pref. -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- # not used -->
+<!-- # statcon<-make_conditions(dd, vars = c("REL")) -->
+<!-- # statcon -->
+<!-- # str(statcon) -->
+<!-- # cn = data.frame("REL" = c(1,.5,0), "cond__" = c("REL = 1","REL = .5", "REL = 0")) -->
+
+<!-- # Run these lines, I'm just commenting out for the document -->
+<!-- p1<-brms::conditional_effects(m1, c("religiosity.S:source"), -->
+<!--                                spaghetti = T, -->
+<!--                               nsamples = 100, -->
+<!--                               probs = c(0.11, 0.89)) -->
+<!-- pp1<-plot(p1, plot = FALSE)[[1]] + scale_y_continuous(limits = c(1,7))  + xlab("Religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Religiosity")  + theme_clean() #theme_radar_dark() + scale_fill_npg() -->
+<!-- pp1 # -->
+<!-- p2<-brms::conditional_effects(m1, c("culturalNorms.S:source"), -->
+<!--                                spaghetti = T, -->
+<!--                               nsamples = 100, -->
+<!--                               probs = c(0.11, 0.89)) -->
+<!-- pp2<-plot(p2, plot = FALSE)[[1]] + scale_y_continuous(limits = c(1,7))  + xlab("Cultural Norms of religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Cultural Norms") + theme_clean() #+ theme_radar_dark() + scale_fill_npg() -->
+<!-- pp2 # -->
+<!-- library(ggpubr) -->
+<!-- plotgood<-ggarrange(pp1, pp2, nrow=1, common.legend = T) -->
+<!-- #saveRDS(plotgood,"plotgood") -->
+<!-- #plotgood<-readRDS("plotgood") -->
+<!-- plotgood -->
+<!-- ``` -->
+
+<!-- # NOTE: an ordinal model would be better, but will require time to run, and it might muddy our inference. -->
+<!-- #  -->
+<!-- m1o <- brm(credible ~ religiosity.S * source + -->
+<!--              culturalNorms.S * source + -->
+<!--              Statement + -->
+<!--              source_first + -->
+<!--              sourceCorrect + -->
+<!--              Age.Decades.C + -->
+<!--              Edu.C + -->
+<!--              Male + -->
+<!--              SES.C + -->
+<!--              logResp.C+ -->
+<!--              (1+source|country), -->
+<!--            data=ddd, -->
+<!--            inits = 0, -->
+<!--            family = cumulative(link ="logit"), -->
+<!--            seed=1234, -->
+<!--            warmup=1000, -->
+<!--            iter=3000, -->
+<!--            chains=4) -->
+
+<!-- saveRDS(m1o,"m1o") -->
+
+<!-- # table and plots -->
+<!-- ```{r}  -->
+
+<!-- m1o<-readRDS("m1o") -->
+<!-- ppo<-pp_check(m1o)  ## fit is much better -->
+<!-- ppo -->
+<!-- ordinalmod<-parameters::model_parameters(m1o,effects = "all", component = "all", test = c("pd")) -->
+<!-- ordinalmod -->
+<!-- oplot<-plot(ordinalmod,  sort = TRUE) + theme_radar() -->
+<!-- oplot -->
+<!-- compareordplot<-ggarrange(ggarrange(param_plot,oplot,labels=c("a","b")),ggarrange(ppg, ppo, labels=c("c","d")), nrow=2, heights = c(3,1)) -->
+<!-- compareordplot -->
+<!-- ``` -->
+<!-- ## next model -->
+<!-- ## have not inclueded this in the knitr call to save time.  -->
+<!-- ```{r} -->
+<!-- statcon<-make_conditions(ddd, vars = c("source")) -->
+<!-- statcon -->
+<!-- str(statcon) -->
+
+<!-- p1<-brms::conditional_effects(m1o, c("religiosity.S"), -->
+<!--                               categorical =T, #  -->
+<!--                               probs = c(0.11, 0.89), -->
+<!--                               conditions = statcon) -->
+<!-- pp1<-plot(p1, plot = FALSE)[[1]]   + xlab("Religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Religiosity")  + scale_y_continuous(limits = c(0,.4))  + theme_clean() + scale_color_grey() +scale_fill_grey() # note of you set categorical to true, then scale the y axis limits: c(0,1).
+Same for the plot below -->
+<!-- pp1 # -->
+<!-- p2<-brms::conditional_effects(m1o, c("culturalNorms.S"), -->
+<!--                                categorical =T, -->
+<!--                               probs = c(0.11, 0.89), -->
+<!--                               conditions = statcon) -->
+<!-- pp2<-plot(p2, plot = FALSE)[[1]] + scale_y_continuous(limits = c(0,.4))  + xlab("Cultural Norms of Religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Cultural Norms")  + theme_clean() + scale_color_grey() +scale_fill_grey()  -->
+<!-- pp2 # -->
+<!-- library(ggpubr) -->
+<!-- plotgood2.0<-ggarrange(pp1, pp2, labels=c("a","b"), nrow=1, common.legend = T) -->
+<!-- plotgood2.0 -->
+<!-- saveRDS(plotgood2.0,"plotgood2.0") -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- # plot that can be read -->
+<!-- ## We can set categorical to false to get more interpretable figures -->
+<!-- p1<-brms::conditional_effects(m1o, c("religiosity.S:source"), -->
+<!--                               categorical =F, # note set to true to get all the ordinal values -->
+<!--                               probs = c(0.11, 0.89), -->
+<!--                                 spaghetti = T, -->
+<!--                               nsamples = 100) # play with this number -->
+<!-- pp1<-plot(p1, plot = FALSE)[[1]]   + xlab("Religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Religiosity")  + scale_y_continuous(limits = c(1,7))  + theme_clean() # note of you set categorical to true, then scale the y axis limits: c(0,1).
+Same for the plot below -->
+<!-- pp1 # -->
+<!-- p2<-brms::conditional_effects(m1o, c("culturalNorms.S:source"), -->
+<!--                                categorical =F, -->
+<!--                               probs = c(0.11, 0.89), -->
+<!--                                 spaghetti = T, -->
+<!--                               nsamples = 100) # play with this number -->
+<!-- pp2<-plot(p2, plot = FALSE)[[1]] + scale_y_continuous(limits = c(1,7))  + xlab("Cultural Norms of Religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Cultural Norms")  + theme_clean()  -->
+<!-- pp2 # -->
+<!-- library(ggpubr) -->
+<!-- plotgood2<-ggarrange(pp1, pp2, labels=c("c","d"), nrow=1, common.legend = T) -->
+<!-- plotgood2 -->
+<!-- saveRDS(plotgood2,"plotgood2") -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- # for omni graph -->
+<!-- plotgoodO<-ggarrange(pp1, pp2, nrow=1, common.legend = T) -->
+<!-- plotgoodO -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- #plotgood2<-readRDS("plotgood2") -->
+<!-- #plotgood2 -->
+<!-- # compare ordinal plots -->
+<!-- twotypeordinalplot<-ggarrange(plotgood2.0,plotgood2,nrow=2) -->
+<!-- twotypeordinalplot -->
+<!-- ``` -->
+<!-- ```{r} -->
+<!-- # everything plot -->
+<!-- omniplot<-ggarrange(ggarrange(param_plot,oplot,labels=c("a","b")),ggarrange(ppg, ppo, labels=c("c","d")), ggarrange(plotgood, plotgoodO, labels=c("e","f")), nrow=3, heights = c(3,1,1)) -->
+<!-- omniplot -->
+<!-- ``` -->
+
+# 
+# 
+# ###### IF WE WANTED BOTH INDICATORS IN THE MODEL THIS IS THE CODe
+# m2 <- brm(mvbind(credible, important) ~ 
+#             religiosity.S * source + 
+#             culturalNorms.S * source +
+#             Statement +
+#             source_first +
+#             sourceCorrect +
+#             Age.Decades.C +
+#             Edu.C +
+#             Male +
+#             SES.C +
+#             logResp.C+
+#             (1+source|country),
+#           data=ddd,
+#           family = "gaussian",
+#           seed=1234,
+#           warmup=1000,
+#           iter=3000,
+#           chains=4)
+
+```

### 2021-01-14 libs.R

@@ -33,6 +33,7 @@ library("dagitty")
 library("ggdag")
 library("papaja")
 library("rstan")
+library("brms")
 library("rstanarm")
 library("tidybayes")
 library("bayesplot")

### 2021-01-15 analysis.Rmd
within between
added discussion about within/between

@@
 Note: here, not all NAs are created alike.
+
 ```{r}
@@
 table(d$denomination,useNA = "ifany") # need recode NA as "no denomination"
-df$wb_socSATSEX
+
 # fix
@@
 d$ethnicity[is.na(d$ethnicity)] <- "not_reported"
-d$wb_soc_3[is.na(d$wb_soc_3)] <- "not_reported" # Sat with sex
+#d$wb_soc_3[is.na(d$wb_soc_3)] <- "not_reported" # Sat with sex
 ```
@@
     wb_socSATPERSONALRELATIONSHIPS  = wb_soc_1,
-    wb_socSATSUPPORTFRIENDS  = wb_soc_2,
-    wb_socSATSEX   = wb_soc_3,
+    wb_socSATSUPPORTFRIENDS  = wb_soc_2
+ #   wb_socSATSEX   = wb_soc_3,
   )%>%
@@
 )
-```
-
-very small effect
+parameters::model_parameters(m1)
+parameters::random_parameters(m1)
+```
+
+No reliable effect
 
@@
 ```{r}
-plot(ggeffects::ggpredict(m1, terms = c("country[all]", "ch_s[minmax]"),type = "re", ci.lvl = 0.89))  +
+plot(ggeffects::ggpredict(m1, terms = c("country[all]", "ch_s[minmax]"),type = "re", ci.lvl = 0.89))  +  scale_fill_brewer(palette= "Set2") + labs( title = "Predicted effects of church on anxiety")
+```
+
+But recall big differences within and between counties in Church Attendence
+
+```{r}
+r1 <- rel %>%
+  ggplot(aes(
+    x = country,
+    y = rel_CHURCHFREQ,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  scale_color_viridis(discrete = TRUE) +
+  ggtitle("Church Frequency") +
+  ylab("") + 
+  theme(axis.text.x = element_text(angle = 30))
+r1
+```
+
+We can use the within and between centred indicators for attendence to clarify how such heterogeneity affects inferenence.
+
+- within_church centering = relative church atttendance within a country
+
+- between_church centering = the adjustment for country level differences in churhc attendance
+
+## Within and between
+
+```{r}
+plot(ggeffects::ggpredict(m1a, terms = c("ch_within[minmax]"),type = "re", ci.lvl = 0.89))  +
    scale_fill_brewer(palette= "Set2") + labs(
@@
 
-## Within and between
+And we do not find reliable heterogeneity at the country level.
+```{r}
+plot(ggeffects::ggpredict(m1a, terms = c("country[all]", "ch_within[minmax]"),type = "re", ci.lvl = 0.89))  +
+   scale_fill_brewer(palette= "Set2") + labs(
+    title = "Predicted effects of church on anxiety")
+```
 
@@
   file = "m1_a.rds")
-summary(m1_a)
-
-```
+
+parameters::model_parameters(m1_a)
+parameters::random_parameters(m1_a)
+
+```
+
+Within countries we do not find  effects from different levels of church attendance on anxiety
 
@@
 
+And we do not find reliable heterogeneity at the country level.
+```{r}
+plot(ggeffects::ggpredict(m1_a, terms = c("country[all]", "ch_within[minmax]"),type = "re", ci.lvl = 0.89))  +
+   scale_fill_brewer(palette= "Set2") + labs(
+    title = "Predicted effects of church on anxiety")
+```
+
+Countries who differ in their average levels of church attendence differ present differences in their expected anxiety levels. Those with more attendence are expected to present less anxiety.
 ```{r}
@@
     title = "Predicted effects of church on anxiety")
-```
-
+
+```
+
+And we do find thi separation at the country level. 
 ```{r}
@@
 
+Countries who differ in their average levels of church attendence differ present differences in their expected anxiety levels.
+Those countries greater levels of church attendance may be expected to present less anxiety.
+
+Take homes:
+
+- No evidence for a "relative" church attendence effect. 
+- Evidence for a between country effect, however this  is *only* intereting if you are a sociologist
+- Sociological data, averaged at the country level, is unhelpful for psychological inference at the individual level.
+This point is relevant to big historical database projects that are collecting country and culture level data.
+Although it may be interesting to have sociological interests -- e.g.
+in whether countries where people go to church more often have lower levels of anxiety, we must not confuse this question with an individual level question about the psychological effects of church attendance.
+- A key point to keep in mind: the reasons people go to churh might differ.
+For example church attendance might reduce anxiety and for this reason people attend.
+Without understanding the demand characteristics for church attendance we are likely to get the wrong inference.
+Consider an analogy.
+Imagine we modelled the relationship between taking antiobotics and infection.
+Without knowning why people were taking antiobiotics we might be tempted into inferring inferring from cross sectional data that antiobotics *cause* infection.
+For this reason, it is very important to collect longitudional data on people, and to focus on whether there are demand qualities that lead people to attend church.
+
+Next we turn to the belief in God model.
+
 ## Belief in God
@@
 
-Writing the dag
+We begin by drawing the dag:
+
 ```{r}
@@
   theme_dag_blank()
+
 ```
@@
                       controlling_for(c("male","support","country","age")))
-
+parameters::model_parameters(m2)
+parameters::random_parameters(m2)
 ```
@@
 )
-summary(m2)
+parameters::model_parameters(m2)
+parameters::random_parameters(m2)
+
 ```
@@
 m2_a <- brm(
-  anx_s ~ age_s  + male + ch_within +  bg_between + (1 + bg_within|country + bg_between | country),  
+  anx_s ~ age_s  + male +  bg_between + (1 + bg_within|country + bg_between | country),  
   data = d1,
@@
 summary(m2_a)
+model_parameters(m2_a)
+random_parameters(m2_a)
 ```

### 2021-01-15 analysis.Rmd
commit draft analysis for Suzanne

--- 2021-01-14 analysis.Rmd
@@
 ---
-
-Notes to Suzanne:
-* Question the attention check transformation -- try to get orginal values
-* Question the decidion to code factors as numbers (below) -- to to get raw responses
-* Question where the scales came from
-* Question for NAs, what is actually missing and what is "prefer not to say" and what is "other"
 
@@
 ```
-
 ```{r runanalysis, echo=FALSE, include = FALSE}
@@
 
+Workflow and analysis
+
+First import data (see functions)
+
 ```{r}
@@
 
-```{r}
+Inspect the data, and note that many columns data have been transformed from their raw form.
+
+```{r, echo = FALSE, include = FALSE}
 #inspect data frame
-str(d) # need to create factors, where fractors are needed.
-```
-
-Ask Suzanne for raw data. Not sure how they rescaled the ordinal report data. Just a comment. I'd never trasnform the factors into numbers scaled between 0-1. This leads to a loss of information.
-
-```{r}
+str(d) # need to create factors, where factors are needed.
+```
+
+Ask Suzanne for raw data so that we can, e.g. model church frequency as a monotonic categorical variable, and pick appropriate sampling distributions for the outcome variables.
+
+```{r, Include = FALSE}
 hist(d$rel_1)  # yikes -- factors (church freq was made into a number :(  loss of signal 
@@
 ```
-
-This is  unfortunate. We are saying that people who are not religious are somewhere between atheists and religious people. Need to back indocator this into a factor.
-
-```{r}
+```{r, include = FALSE}
 table(d$rel_3,useNA = "ifany") # This is absolutely not good. 
@@
 
-Note: here, not all NAs are created alike. 
-```{r}
-#  need to recode factors -- NAs should be NAs, otherwise, we should describe the missingness
+```{r include = FALSE}
+#  need to recode factors -- NAs should be NAs, otherwise, we should describe them as missingness, so will avoid these variables
 table(d$ethnicity,  useNA = "ifany") # need recode NA as "not identified" 
 table(d$denomination,useNA = "ifany") # need recode NA as "no denomination" 
-df$wb_socSATSEX
 # fix
@@
 d$ethnicity[is.na(d$ethnicity)] <- "not_reported"
-d$wb_soc_3[is.na(d$wb_soc_3)] <- "not_reported" # Sat with sex
-```
-
-```{r}
+#d$wb_soc_3[is.na(d$wb_soc_3)] <- "not_reported" # Sat with sex
+```
+```{r, include = FALSE}
 # what are the numbers here? 
@@
 
-```{r}
-head(d$gender)
-
+```{r cache = TRUE}
 # clean data:use informative names !
 df<- d %>%
-  dplyr::mutate(fm1m2 = ifelse( gender != "man", 1 ,2)) %>% # for indexical categorical variable
-  dplyr::mutate(male = factor(ifelse( gender != "man", "male" ,"female"))) %>%
+  dplyr::mutate(male2 = ifelse( gender != "man", 1 ,2)) %>% # for index categorical variable, not change of meaning for gender. male and people who do not identify as male. 
+  dplyr::mutate(male = factor(ifelse( gender != "man", "male" ,"notmale"))) %>%
 rename(
@@
 
-table(df$fm1m2, useNA = "always")
-```
-
-Note the attention check variable. The data documentation says `1` means "passed" however the questionaire lists options 1 ("not at all") -- 7 "very much."
-
-Will retain the attention check fails for now, however it iwll be important to remove these cases if they are affecting the outcomes. (I have requested further information from Suzanne)
-
-```{r}
+```
+
+## Missingness
+
+Next we need to examine patterns of missingness, which might spoil inference.
+
+```{r cache = TRUE}
+library("naniar")
+vis_miss(df) # substantial pattern of missing ness in a few variables, 
+```
+
+There is little missingness so we will use only complete cases.
+Note that it would be better to have a method for handling the uncertainty of missingness.
+Missing data should be handled by some probabilistic imputation strategy.
+However, because there are many larger problems about measurement error, sampling, and theory, lets set these to questions to the side and work with the complete dataset
+
+```{r, include = FALSE}
 # get columns for different indicators 
@@
 
-Exploration Graphs
-
-```{r} 
+Exploration Graphs.  Inspect the data.
+
+```{r, cache = TRUE, echo = FALSE} 
 head(rel)
@@
 r7
-
+```
+
+```{r, cache = TRUE, include = FALSE} 
+library("patchwork")
 (r1 + r2  +  r3 )/  (r4  + r5) / (r6 +  r7)   +  plot_annotation(tag_levels = 'A') + plot_layout(guides = "collect")
@@
 
-Correlation plots. Note that we need the to model the country level dependencies in any correlation matrix, given strong cultural differences
-
-Religion:
-
-```{r}
+Next, a quick run of correlations. Note that we need the to model the country level dependencies in any correlation matrix, given strong cultural differences
+
+Religion correlations
+```{r, cache = TRUE}
 rel %>%
@@
 
-```{r}
-str(rel)
-report::report_sample(
-  rel,
-  group_by = "country",
-  # select = NULL,
-  # exclude = NULL,
-  # weights = NULL, 
-  total = TRUE, digits = 2
-)
-report::report_sample(iris, group_by = "Species")
-```
-
-General wellbeing
-```{r}
+General well-being correlations
+
+```{r cache = TRUE}
 wbgen %>%
@@
 
+Physical well-being correlations
+
 ```{r}
@@
 
-Psychological wellbeing
+Psychological well-being correlations
+
 ```{r}
@@
 
-```{r}
-wbsoc %>%
-  dplyr::select(-wb_socSATSEX) %>% # too many missing values
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  summary()
-
-wbsoc %>%
-  dplyr::select(-wb_socSATSEX) %>% # too many missing values
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  plot() + theme_abyss()
-```
-
-# Specify causal model
-
-Suppose we believe that
-
-- religion + ses + male + country + a different unmeasured latent variable (e.g. priviledge) causes well being
-
-- age + male + country + and Unmeasured latent variable (e.g. parent's religion, call this U1) causes Religion
-
-- country + age + another unmeasured latent variable (e.g. priviledge U2) causes ses
+# Specify our causal model
+
+Suppose we believe that:
+
+- religion + ses + male + country + a different unmeasured latent variable (e.g. privilege) causes well being (citations... )
+
+- age + male + country + and Unmeasured latent variable (e.g. parent's religion, call this U1) causes Religion (citations... )
+
+- country + age + another unmeasured latent variable (e.g. privilege U2) causes ses  (citations...)
 
@@
 
-Which variables should I condition on to understand the relationship between religion and well-being?
+Which variables should we condition on to understand the relationship between religion and well-being?
 
@@
 
-Suppose I was to include all variables in my model, except U2 and U1, which are not measured?
-
-In this case I would introduce colider bias
+Suppose we were to include *all* the variables in our model, except U2 and U1 which are not measured?
+
+In this case we  would introduce collider bias!
 
@@
 
-## missingness
-
-Next we need to examine patterns of missingness, which might spoil inference.
-
-```{r}
-library(naniar)
-vis_miss(df) # substantial pattern of missing ness in a few variables, 
-```
-
-There is little missingness so we will just use complete cases.
-Note that it would be better to have a method for handling the uncertainty of any case.
-A short cut is a short cut: missing data should be handled by some probabilitic imputation strategy.
-However because there are many larger problems about measurement error, sampling, and theory, lets set these to questions to the side and work with the complete dataset
-
 ## The hard work
 
-We have yet to consider a mechanism by which religion might affect well-being, and specifically which aspects of well-being and which aspects of religion.
-Just as it might be tempting to through all variables into a regression equation, and disasterious to do so, it might be tempting to leap into some form of data reduction, and combine well-being and religious indicators without stopping to think about how causation.
-
-There are many theories we might consider.
-We will focus on two non-trivial theories, one related to church attendance and one related to belief in god.
-If we had more time we could consider many others, but as Catholics know, "timing is everything" -- let's keep to time.
+We have yet to consider a mechanism by which religion might affect well-being, which will tell us specifically which aspects of well-being and which aspects of religion to focus on.
+Just as it might be tempting to throw all variables into a regression equation, and disastrous to do so, it might be tempting to leap into data reduction combining all variables without pausing to think what we want to know.
+
+There are many theories of how religion affects well-being we might consider (cite Highland et al).
+We will focus on two non-trivial theories -- one related to religious behavior and another related to religious belief.
 
@@
 
-We propose that church attendance reduces psychological anxiety by increasing perceived support.
+We propose that church attendance reduces psychological anxiety by increasing perceived social support (which is measured in this dataset as a well-being variable).
+
 This is a new model so we need to write our dag all over again.
@@
 
-Writing the dag
+We write the dag:
+
 ```{r}
@@
 
-The adjustments sets are the same so we proceed with a model.
-Note that if my interest is in the direct effect of church attendance on wellbeing I should *not* include social support because this introduces a collider bias!
+The two adjustment sets are the same as in the original dag that we considered.
+
+Before proceeding with the model, let's note a curiosity.
+If our  interest is in the direct effects of church attendance on well-being we should *not* include social support because this introduces a collider bias!
 
@@
 
+Let this serve as a warning for us to think about our causation, and what we might want to know before we leap into modeling
+
 ## The model
@@
 
-For this initial analysis I am going to use the weakly informative priors that BRMS uses; my experience is that large datasets are resistent to choice of priors.
+For this initial analysis I am going to use the weakly informative priors that BRMS uses; my experience is that large data sets are resistant to choice of priors.
 However it is a mistake to trust this experience.
@@
 This is a very bad idea.
-It would be much better to thinking of church attendance frequence as a monotonic effect.
+It would be much better to thinking of church attendance frequency as a monotonic effect.
 I will do that analysis at the next round, if the data are made available in their original form.
@@
 Wrangle data
+
 ```{r}
@@
 d1 <- df %>%
+  dplyr::filter(
+    !is.na(wb_psychANXIETY),!is.na(age),!is.na(rel_CHURCHFREQ),!is.na(wb_psychMEANING),!is.na(male),!is.na(rel_EXTENTBELIEVEGOD)
+  ) %>%
   dplyr::mutate(
@@
     age_s = scale(age),
+    ch = rel_CHURCHFREQ,
+    mn  = wb_psychMEANING,
+    bg =  rel_EXTENTBELIEVEGOD,
     ch_s = scale(rel_CHURCHFREQ),
@@
   ) %>%
-  dplyr::select(anx, anx_s, age_s, ch_s, mn_s, fm1m2, country, male, bg_s) %>%
-  dplyr::filter(
-    !is.na(anx_s),
-    !is.na(age_s),
-    !is.na(ch_s),
-    !is.na(mn_s),
-    !is.na(male),
-    !is.na(bg_s)
-  )
-table(d1$country)
-```
-
-Model for church
+  dplyr::select(anx, ch, mn, bg, anx_s, age_s, ch_s, mn_s, male2, country, male, bg_s)
+
+## degroup variables for within and between analysis
+## https://cran.r-project.org/web/packages/parameters/vignettes/demean.html
+d1 <-
+  cbind(d1,
+        demean(d1, select = c("anx", "ch", "bg"), group = "country"))
+head(d1)
+```
+
+We might estimate that effects differ within and between groups, but ignoring within and between country heterogeneity and simply pooling effects at the country level
 
@@
 m1 <- brm(
-  anx ~ ch_s +
+  anx_s ~ ch_s +
     age_s  +
@@
 )
-```
-
-### model for belief
-
-From causal dag above
+parameters::model_parameters(m1, test = "p_direction")
+parameters::random_parameters(m1)
+```
+
+We can see there is no effect of church attendance on anxiety (according to this model). Note that throughout, the plots will sample 300 draws from the posterior distribution.
+
+```{r cache = TRUE}
+p1<-brms::conditional_effects(m1, "ch_s", 
+                              spaghetti = T,
+                              nsamples = 300,
+                              prob = .89,
+                              method = 'fitted')
+
+plot(p1, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1,1))  + xlab("church sd units") + ylab ("anxiety in sd units") +
+  labs(title = "Predicted effect of ch_s")
+```
+
+Before racing ahead, lets do some posterior predictive checks
+
+```{r cache = TRUE}
+brms::mcmc_plot(m1,
+          type = "trace")
+```
+
+When we do a posterior predictive check we find problems in the predicted distribution.
+This we called this a gaussian, but it is not.
+The data have large peaks reflecting the categorical qualities of the outcome variable.
+
+We should work with the original, untransformed values of the data.
+
+```{r cache = TRUE}
+# we see the ordinal qualities of our dataset here
+pp_check(m1) # here we pick up the categorical qualities of the distribution
+```
+
+For now, we press ahead. Let's graph the location effects
+
+```{r cache = TRUE}
+mcmc_plot(m1,
+          type = "intervals",
+          prob_outer = 0.89)
+```
+Note: there is huge country-level variability in the MCMC solutions and uncertainty in the correlations between country level intercepts in slopes. How should we interpret these results:
+
+  - Countries have different levels of anxiety;
+
+  - There is variability in the church-level effects on anxiety by country.
+
+  - Whether the effect of church attendance differs by countries that are high or low in anxiety is uncertain, but the mass of probability is below zero.
+The greater a countries anxiety the more likely church attendance will reduce anxiety...
+however, we should not be too confident that this pattern is general.
+
+  - Finally the standard deviation of the overall expected mean level of anxiety is large, which is to be expected because the gaussian is such a poor choice for our sampling distribution.
+
+We'd want to check the priors in this model, and probably use a more informative prior for the correlation of country anxiety and church attendance.
+
+These are our model priors:
+
+```{r}
+prior_summary(m1)
+```
+
+However we cannot sample from the models priors because some are improper.
+
+```{r}
+samples1 <- brms::prior_samples(m1, "b_ch_s")
+samples1
+```
+
+To see this, run the following and note the warning.
+
+```{r}
+m1_prior <- brm(
+  anx_s ~ ch_s +
+    age_s  +
+    male +
+    (1 + ch_s | country),
+  data = d1,
+  family = "gaussian",
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = "m1_prior",
+  sample_prior = "only")
+```
+
+```{r}
+prior_summary(m1)
+```
+
+A key problem is the use of uninformative priors
+
+```{r}
+bayestestR::describe_prior(m1)
+```
+
+Let's try with an informative prior
+
+```{r cache = TRUE}
+m1_prior <- brm(
+  anx_s ~ ch_s +
+    age_s  +
+    male +
+    (1 + ch_s | country),
+  data = d1,
+  family = "gaussian",
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = "m1_prior",
+  set_prior("normal(0,.5)", class = "b"),
+  sample_prior = "only")
+bayestestR::describe_prior(m1_prior)
+```
+
+This prior allows positive and negative relationships from -6 to + 4 standard deviations, so would not be informative.
+
+<!-- ```{r cache = TRUE} -->
+<!-- plot(ggeffects::ggpredict(m1_prior, terms = c("ch_s[minmax]"),type = "fe", ci.lvl = 0.89))  +  scale_fill_brewer(palette= "Set2") + labs( title = "Predicted effects of church on anxiety") -->
+<!-- ``` -->
+
+or we can simulate and see that our prior is scattered across five standard deviations of anxiety, and we can see our prior is telling us that anything can happen
+
+```{r cache = TRUE}
+p<-brms::conditional_effects(m1_prior, "ch_s", 
+                              spaghetti = T,
+                              nsamples = 1000,
+                              prob = .89,
+                              method = 'fitted')
+plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-5,5))  + xlab("church sd units") + ylab ("anxiety in sd units") + 
+labs(title = "Expected values for our attempted informative prior reveals remains too flat")
+```
+
+We'll return to our priors and sampling distribution at the next iteration of this analysis.
+Let's set this to the side, and focus on the country level variability, looking first at the expectation plots.
+
+Nothing happening at the individual level.
+
+```{r include = FALSE}
+# graph not as nice
+# plot(ggeffects::ggpredict(m1, terms = c("ch_s[minmax]"),type = "fe", ci.lvl = 0.89))  +
+#    scale_fill_brewer(palette= "Set2") + labs(
+#     title = "Predicted effects of church on anxiety")
+```
+
+But lots of variability at the country level intercepts for anxiety.
+We can see that India is over two standard deviations  less anxious than the average culture level anxiety.
+By contrast, Japan, Israel, and France are over .2 standard deviations more anxious than average.
+There looks to be perhaps some covariance between the country intercepts and slopes...
+maybe though all estimates crosses zero.
+We'd either want to use stronger priors or collect data from more countries than are in this sample.
+
+```{r cache = TRUE}
+library(see) 
+library(parameters) 
+par<-parameters::model_parameters(m1,effects = "random" ) 
+plot(par,  sort = TRUE) +theme_lucid() #option -->
+```
+
+```{r cache = TRUE}
+plot(ggeffects::ggpredict(m1, terms = c("country[all]", "ch_s[minmax]"),type = "re", ci.lvl = 0.89))  +  scale_fill_brewer(palette= "Set2")
+```
+
+Let's delving into this question of country level effects, but asking whether people who are relatively more church attending tend to have more anxiety.
+
+- within_church centering = relative church attendance within a country.
+
+- between_church centering = the adjustment for country level differences in church attendance.
+
+## Within and between
+
+Here's a model that includes within and between country centered variables. See:
+https://easystats.github.io/parameters/reference/demean.html
+
+```{r}
+m1a <- brm(
+  anx_s ~ age_s  + male + ch_within +  ch_between + (1 + ch_within|country),  
+  data = d1,
+  family = "gaussian",
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = "m1_a.rds")
+
+parameters::model_parameters(m1a,  
+                             test = "p_direction")
+parameters::random_parameters(m1a)
+
+```
+
+No reliable effect within,
+
+```{r cache = TRUE}
+
+p<-brms::conditional_effects(m1a, "ch_within", 
+                              spaghetti = T,
+                              nsamples = 3000,
+                              prob = .89,
+                              method = 'fitted')
+
+plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1,1))  + xlab("church_within") + ylab ("anxiety in sd units") + 
+  labs(title = "Predicted effects of ch_within")
+
+```
+
+But a between-country effect emerges:
+
+```{r cache = TRUE}
+p<-brms::conditional_effects(m1a, "ch_between", 
+                              spaghetti = T,
+                              nsamples = 3000,
+                              prob = .89,
+                              method = 'fitted')
+
+plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1,1))  + xlab("church_between") + ylab ("anxiety in sd units")+ 
+  labs( title = "Predicted effects of ch_between")
+
+```
+
+Same workflow as before.
+
+Trace plots for the model look OK
+
+```{r}
+ mcmc_plot(m1a,
+          type = "trace")
+```
+
+No surprise, there is the same problem with the sampling distribution.
+
+```{r}
+# we see the ordinal qualities of our dataset here
+pp_check(m1a)
+```
+
+Next the interval plots:
+
+```{r}
+mcmc_plot(m1a,
+          type = "intervals",
+          prob_outer = 0.89)
+```
+
+Which we interpret as follows:
+
+  - Countries have different levels of anxiety;
+
+  - There is no relative deprivation effect for church attendance on anxiety.
+Those whose who participate more in church can be expected to have similar anxiety as those who participate less within a country.
+
+  - This within-country effect is variable at the country level (see "se_country_Intercept"). We interpret his to mean that the within country mean church effect differ between countries.
+
+  - Whether the effect of church attendance differs by countries that are high or low in anxiety is uncertain, but the mass of probability is below zero.
+
+  - Finally the standard deviation of the overall expected mean level of anxiety is large, which is to be expected because the gaussian is such a poor choice for our sampling distribution.
+
+Let's look at the country-level variances more closely
+
+```{r}
+par<-parameters::model_parameters(m1a,effects = "random" ) 
+plot(par,  sort = TRUE) +theme_lucid() #option -->
+```
+
+Next the predictive plots...
+
+```{r cache = TRUE}
+plot(ggeffects::ggpredict(
+  m1a,
+  terms = c("country[all]", "ch_within[minmax]"),
+  type = "re",
+  ci.lvl = 0.89))  +
+  scale_fill_brewer(palette = "Set2") + 
+  labs(title = "Predicted effects of church on anxiety: countries X within") + 
+  scale_y_continuous(limits = c(-1,1))
+```
+
+Countries who differ in their average levels of church attendance differ present differences in their expected anxiety levels. Those with more attendance are expected to present less anxiety.
+
+And here we pick up the country level trend, which is really strong.  Countries that are more churchgoing have lower anxiety than countries that are less churchgoing.
+
+```{r cache = TRUE}
+plot(ggeffects::ggpredict(
+  m1a,
+  terms = c("ch_between[minmax]"),
+  type = "re",
+  ci.lvl = 0.89))  +
+  scale_fill_brewer(palette = "Set2") + labs(title = "Predicted effects of church on anxiety: between")  + scale_y_continuous(limits = c(-1, 1))
+```
+
+Take homes:
+
+    - No evidence for a "relative" church attendence effect.
+
+    - There is evidence for a between country effect.
+Countries high in church attendance are less anxious than countries that are low in church attendance.
+We must not confuse this question with an individual level question about the psychological effects of church attendance!
+
+    - A key point to keep in mind: none of these models track **demand** for church among anxious people.Church attendance might reduce anxiety therefore anxious people attend church.
+Without understanding the demand characteristics for church attendance we are likely to get the wrong inference.
+Consider an analogy.
+Imagine we were to model the relationship between taking antiobotics and infection.
+Without knowning why people were taking antiobiotics we might be tempted into inferring that antiobotics *cause* infection.
+For this reason, it is very important to collect longitudional data on people, and to clarify the demand qualities that might lead people to attend church more frequently.
+
+## Belief in God
+
+Lets consider the non-trivial theory that belief in God (bg) affects psychological anxiety (wb) by affording a sense of meaning in life.
+
+We begin by drawing the dag.
+
+```{r}
+tidy_ggdag2 <- dagify(
+  wb  ~ meaning + ses + male + country  +U2,
+  meaning ~ bg +  male + country, 
+  bg ~ age + male + country + U1,
+  ses ~ country + age + U2,
+  exposure =  "bg",
+  outcome =   "wb") %>%
+  tidy_dagitty()
+
+# graph adjustment sets
+ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) + 
+  theme(legend.position = "bottom") + 
+  theme_dag_blank()
+
+```
+
+And again, we face a collider bias. Our causal model shows us that if we were to include our meaning-making indicator in our statistical we will compromise inference!
+
+```{r}
+ggdag::ggdag_collider(tidy_ggdag2,
+                      from = "church",
+                      to = "wb",
+                      controlling_for(c("male", "support", "country", "age")))
+```
+
+### Model for belief
+
+From causal model we write our statistical model:
 
@@
 m2 <- brm(
-  anx ~
+  anx_s ~
     bg_s  +
@@
 )
-summary(m2)
-```
-
-Graph -- No effect, why? 
-```{r}
-p2 <- brms::conditional_effects(m2,
-                                c("bg_s"),
-                                categorical = F,
-                                #
-                                probs = c(0.11, 0.89))
-plot(p2)
-```
-
-To do -- look at between country effects and relative effects within countries
-
-```{r}
-library("rethinking")
-set.seed(867530)
-
-m_ch <- ulam(
-  alist(
-    anx_s ~ normal(mu , sigma),
-    mu <-  alpha + b1 * ch_s + b2 * age_s  + b3 * male + a_country[country] + b_country[country] * ch_s,
-    alpha ~ normal(0,1),
-    b1 ~ dlnorm(0,1),
-    b2 ~ normal(0,1),
-    b3 ~ normal(0,1),
-    u_country ~ dnormal(0,1),
-    c(a, b)[country] ~ multi_normal(c(a, b) , Rho , sigma_country),
-    a ~ normal(0, 2),
-    b ~ normal(-1, 0.5),
-    sigma_country ~ exponential(1),
-    sigma ~ exponential(1),
-    Rho ~ lkj_corr(2),
-  ) ,
-  data = d1 ,
-  chains = 4 ,
-  cores = 4
-)
-```
-
-Excerpt From: McElreath, Richard;. “Statistical Rethinking.” Apple Books. 
-```
-
-## Belief in God
-
-Lets consider the non-trivial theory that belief in God (bg) affects psychological anxiety (wb) by affording a sense of meaning in life.
-The causal qualities will be the same as those of the previous model, and so too will the threats to inference.
-
-Writing the dag
-```{r}
-tidy_ggdag2 <- dagify(
-  wb  ~ meaning + ses + male + country  +U2,
-  meaning ~ bg +  male + country, 
-  bg ~ age + male + country + U1,
-  ses ~ country + age + U2,
-  exposure =  "bg",
-  outcome =   "wb") %>%
-  tidy_dagitty()
-
-# graph adjustment sets
-ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) + 
-  theme(legend.position = "bottom") + 
-  theme_dag_blank()
-```
-
-And again we face a collider bias by introducing meaning measure, conditional on assumed causal relationships
-
-```{r}
-ggdag::ggdag_collider(tidy_ggdag2,
-                      from = "church",
-                      to = "wb",
-                      controlling_for(c("male","support","country","age")))
-
-```
-
-### OLD ANALYSIS
-
-#
-#setwd("/Users/jbul176/The\ Virtues\ Project\ Dropbox/Joseph\ Bulbulia/00Bulbulia\ Pubs/2020/2020_VAN_ELK") # mine
-
-# read data
-e <- read.csv("/DataStudy3_SourceCredibility_Long.csv", header=T) # old data within/group means
-d <- read.csv('DataLong', header=T) # new data without ingroup means
-#d$statementAndOrder # error but this is not needed
-nrow(d)#20390
-hist(e$religiosity) # group standardized data
-hist(d$religiosity) # note difference
-# check data : note that subject needs to be a factor. also, "Statement" needs to be a factor. for gender there are few "others" we create a "male gender variable (below) for simplicity. 
-#
-# look for patters of missingness ####
-library(naniar)
-vis_miss(d) # substantial pattern of missing ness in a few variables, 
-```
-```{r}
-#plot outcomes: gaussian is reasonable, but we'll come back to this after ppchecks
-str(d)
-# check outcomes
-hist(d$credible) # gaussian ordinal
-hist(d$important)# flat
-# what is the F1 variable?
-hist(d$recallFa ) ##??
-hist(d$recallHit )##??
-```
-```{r}
-#make tables ####
-# and look for anomolies
-library(table1) # useful package
-table1::table1(~Age + Gender + Education + R_religiousID + religiosity + culturalNorms|country, data =d, transpose =T)
-```
-```{r}
-table1::table1(~credible|country, data =d, transpose =TRUE)# outcome variables
-```
-
-```{r}
-table1::table1( ~R_attendance + R_prayer + R_religiousID + R_member  + R_god   + R_afterlife|country, data =d, transpose =T)
-```
-
-```{r}
-# how correlated are the religion variables? ####
-corrmat<-d%>%
-  dplyr::select(
-    R_attendance,
-    R_religiousID,
-    R_member, # can lose this
-    R_god,
-    R_afterlife,
-    R_prayer)%>%
-  dplyr::mutate(R_attendance.S = scale(R_attendance),
-                R_religiousID.S =scale(R_religiousID),
-                R_member.S = scale(R_member), # not used
-                R_god.S=scale(R_god),
-                R_prayer.S =scale(R_prayer),
-                R_afterlife.S =scale(R_afterlife))%>%
-  rowwise() %>% 
-  mutate(REL= mean(c(R_attendance, R_prayer,R_god,R_religiousID,R_afterlife), na.rm=T))%>% # we don't use this, but my sense is R_member isn't going to be worth including.
-  mutate(REL.S = scale(REL))%>% # we don't use this
-  select(R_attendance.S,R_religiousID.S,R_member.S,R_god.S,R_afterlife.S,R_prayer.S) # don't select religiosity for PCA
-# corrplot
-res <- cor(corrmat, method="pearson")
-corrplot::corrplot(res, method= "color", order = "hclust") # tl.pos = 'n'
-```
-
-```{r}
-## corr.soc  
-sjp.corr(corrmat,
-         title = "Correlation matrix of religion variables",
-         geom.colors = "RdBu")
-```
-```{r}
-# PCA ####
-cpca <- prcomp(corrmat, center=TRUE, scale.=TRUE)
-summary(cpca)
-
-# or this
-pca(corrmat) # God is "good" 
-# get data into shape ####
-# # check the response time log (how I would handle this)
-hist(log(d$RT)) # looks OK
-# create data frame ####
-#  center/standardise for meaningful intercept and for reliable interactions (also for estimation)
-dd<-d%>%
-  dplyr::select(
-    R_attendance,
-    R_religiousID,
-    R_member, # can lose this
-    R_god,
-    R_afterlife,
-    R_prayer,
-    religiosity,
-    credible,source,country,subject,Statement,Gender,Education,SES,culturalNorms,Age,source_first,sourceCorrect,RT,important)%>%
-  dplyr::mutate(subject = factor(subject),
-                Statement = factor(Statement), 
-                Male = factor(ifelse(Gender =="man","male","not_male")),
-                Edu.C = scale(Education,center=T,scale=F),
-                SES.C = scale(SES, center=T,scale=F),
-                culturalNorms.S = scale(culturalNorms,center=T,scale=T),
-                logResp = log(RT),
-                logResp.C = scale(logResp,scale=F,center=TRUE),
-                Age.Decades.C = scale(Age, scale = FALSE, center = TRUE)/10)%>%
-  dplyr::mutate(R_attendance.S = scale(R_attendance),
-                R_religiousID.S =scale(R_religiousID),
-                R_member.S = scale(R_member),
-                R_god.S=scale(R_god),
-                R_prayer.S =scale(R_prayer),
-                R_afterlife.S =scale(R_afterlife),
-                religiosity.S = scale(religiosity))
-
-# Get PCA
-dfp<-data.frame(cpca$x[,1:2])
-ddd<-as.data.frame(cbind(dd,dfp)) # data for PCA comparison
-head(ddd)
-
-```{r}
-# detour ignore#### 
-# table1::table1( ~culturalNorms.S*R_god  |country, data =ddd, transpose =T)
-# ```
-# ```{r}
-# library(sjPlot)
-# ek8<-  plot_xtab(ddd$country,ddd$ R_god, show.n=F,
-#                  show.prc = F , show.total = F,drop.empty = F) +  scale_fill_aaas() +scale_y_continuous(limits=c(0,.30)) + xlab("Country") +  ylab("Frequency God Beliefs") #+
-# #  scale_y_continuous(limits=c()) + #theme(plot.title = element_text(size=09))
-# # theme(axis.text.x = element_text(angle = 0, hjust=1,size =10))
-# 
-# ek8
-# ek9<-  plot_xtab(ddd$country,ddd$culturalNorms, show.n=F,
-#                  show.prc = F, show.total = F,drop.empty = F) +  scale_fill_aaas() +
-#   scale_y_continuous(limits=c(0,.30)) +  xlab("Country") +  ylab("Frequency Cultural Norms Perc. Religiosity")
-# 
-# ek9
-# compare_god_norm<-ggarrange(ek8,ek9,nrow=2) # 3 x 1
-# compare_god_norm
-```
-
-```{r}
-# filthy frequentist model (which we need to check for Variance Inflaction Factors (multicollinearity)) # if our religion variables were not highly correlated we could include all/most of them to understand the dimensionality of religousity on credibility.
-library(lme4)
-# saturated model
-d1<-lmer(credible ~ (R_attendance.S + R_religiousID.S + R_member.S + R_god.S +R_afterlife.S) * source +
-           culturalNorms.S * source +
-           Statement +
-           source_first +
-           sourceCorrect +
-           Age.Decades.C +
-           Edu.C +
-           Male +
-           SES.C +
-           logResp.C+
-           (source|country), data=ddd )
-
-# our model
-d2<-lmer(credible ~religiosity.S * source +  
-           culturalNorms.S * source +
-           Statement +
-           source_first +
-           sourceCorrect +
-           Age.Decades.C +
-           Edu.C +
-           Male +
-           SES.C +
-           logResp.C+
-           (source|country), data=ddd )
-
-# not our model which is just as good as the preferred
-# first and second PC, which predicts 80%+ of the variance without multi-collinearity 
-d3<-lmer(credible ~(PC1 + PC2) * source  + 
-           culturalNorms.S * source +
-           Statement +
-           source_first +
-           sourceCorrect +
-           Age.Decades.C  +
-           Edu.C +
-           SES.C +
-           logResp.C+
-           (source|country), data=ddd )
-
-# god is the strongest religiosity term
-d4<-lmer(credible ~ R_god.S * source  + 
-           culturalNorms.S * source +
-           Statement +
-           source_first +
-           sourceCorrect +
-           Age.Decades.C +
-           Edu.C +
-           Male +
-           SES.C +
-           logResp.C+
-           (source|country), data=ddd )
-
-# simple PC1 model (probably what JB would have done)
-d5<-lmer(credible ~PC1 * source + 
-           culturalNorms.S * source +
-           Statement +
-           source_first +
-           sourceCorrect +
-           Age.Decades.C +
-           Edu.C +
-           Male +
-           SES.C +
-           logResp.C+
-           (source|country), data=ddd )
-# check for multi-colinearity ####
-library(car) 
-vif(d1)# very high, will increase standard errors of regression coeffs.
-
-vif(d2)# OK  our model
-
-vif(d5)# OK my preferred model
-
-library(sjPlot)
-plot_model(d2) # our model (without co-variates)
-plot_model(d2, type= "re") # our model (without co-variates)
-# fem<-plot_model(d2) # our model (without co-variates)
-# plot_model(d2) # our model (without co-variates)
-# ref<-plot_model(d2, type= "re") # our model (without co-variates)
-# plot_model(d3) # same inference
-# plot_model(d4) # same inference
-# plot_model(d5) # same inference
-# ref<-plot_model(d2, type="re")
-# lmer_coefplot<-ggarrange(fem,ref,ncol=1)
-# lmer_coefplot
-```
-
-```{r}
-plot_model(d5) # no diff
-plot_model(d5, type="re") # no diff
-```
-
-```{r}
-# Run this chunk only to satisfy yourself that our religiosity variable is good.
-pl2<-ggpredict(d2, c(terms ="religiosity.S[all]","source"), ci.lvl = 0.95)
-
-pp2<-plot(pl2)+scale_y_continuous(limits=c(1,7)) + xlab("Religiosity (standardised)") + ylab("Credible:1-7")
-
-# pl4<-ggpredict(d4, c(terms ="R_god.S[all]","source"), ci.lvl = 0.95)
-# pp4<-plot(pl4)+scale_y_continuous(limits=c(1,7))+ xlab("Belief in God (standardised)") + ylab("Credible:1-7")
-# 
-# pl5<-ggpredict(d5, c(terms ="PC1[all]","source"), ci.lvl = 0.95)
-# pp5<-plot(pl5)+scale_y_continuous(limits=c(1,7)) + xlab("First Principle Component: Religion Indicators") + ylab("Credible:1-7")
-# 
-# 
-# pl2.1<-ggpredict(d2, c(terms ="culturalNorms.S[all]","source"), ci.lvl = 0.95)
-# 
-# pp2.1<-plot(pl2.1)+scale_y_continuous(limits=c(1,7)) + xlab("Religiosity (standardised), Religiosity Model") + ylab("Credible:1-7")
-# 
-# pl5.1<-ggpredict(d5, c(terms ="culturalNorms.S[all]","source"), ci.lvl = 0.95)
-# pp5.1<-plot(pl5.1)+scale_y_continuous(limits=c(1,7)) + xlab("Cultural Norms (Standardised), PC1 model") + ylab("Credible:1-7")
-# # create plot of the different ways of handling religiosity 
-# lmerplot<-ggarrange(pp2,pp5,pp2.1,pp5.1,labels=c("a","b","c","d"),common.legend = T)
-# lmerplot # 1000 x 1000
-# create plot of our preferred model for the 2 outcomes of interest
-# lmerquick<-ggarrange(ggarrange(lmer_coefplot),ggarrange(pp2,pp2.1,ncol=2),heights=c(2,1),nrow=2)
-# lmerquick # 2000 x 2
-```
-
-```{r}
-#table of our pref model
-library(sjPlot)
-tab_model(d2) # see documentation for options
-```
-
-# bayesian model ####  I'm not running this again, but you can add this to the next chunk and run
-m1 <- brm(credible ~ religiosity.S * source +
-            culturalNorms.S * source +
-            Statement +
-            source_first +
-            sourceCorrect +
-            Age.Decades.C +
-            Edu.C +
-            Male +
-            SES.C +
-            logResp.C+
-            (1+source|country),
-          data=ddd,
+parameters::model_parameters(m2, test = "p_direction")
+parameters::random_parameters(m2)
+```
+
+We find no effect for belief in God.
+
+```{r cache = TRUE}
+
+p<-brms::conditional_effects(m2, "bg_s", 
+                              spaghetti = T,
+                              nsamples = 300,
+                              prob = .89,
+                              method = 'fitted')
+
+plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1,1))  + xlab("bg_s") + ylab ("anxiety in sd units") + 
+  labs(title = "Predicted effects of bg_s")
+
+```
+How did the model mix?  Trace plots look OK
+
+```{r}
+ mcmc_plot(m2,
+          type = "trace")
+```
+
+Again, unsurprisingly we find a same problem with the sampling distribution.
+
+```{r}
+# we see the ordinal qualities of our dataset here
+pp_check(m2)
+```
+
+Note: we'd want to check the priors in this model, and use more informative priors. TBA...
+
+Next the interval plots:
+```{r}
+mcmc_plot(m2,
+          type = "intervals",
+          prob_outer = 0.89)
+```
+
+Summary:
+
+    - No individual-level effect of belief in God on anxiety (though recall our warning about demand characteristics)
+
+    - there's lots of variability in the country-level intercepts.
+
+    - there is reliable slope effect variability at the country level.
+
+    - The density of probability in the covariance of intercept and slope trends negative. Those countries with higher anxity tend to have a lower benefit from a belief in God effect.
+
+The country level intercepts follow a similar patter for as for church attendance, which is no surprise because we are estimating the same anxiety parameter as before, just different covariates.
+
+```{r}
+par<-parameters::model_parameters(m1,effects = "random" ) 
+plot(par,  sort = TRUE) +theme_lucid()
+```
+
+Here's the absence of any individual level effect for belief in God on anxiety.
+
+```{r cache = TRUE}
+
+p<-brms::conditional_effects(m2, "bg_s", 
+                              spaghetti = T,
+                              nsamples = 300,
+                              prob = .89,
+                              method = 'fitted')
+
+plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1,1))  + xlab("bg_s") + ylab ("anxiety in sd units") + 
+  labs(title = "Predicted effects of bg_s")
+
+```
+
+There country-level variability is all in anxiety, but not in the effect of belief in God.
+
+```{r cache = TRUE}
+plot(ggeffects::ggpredict(m2, terms = c("country[all]", "bs_s[minmax]"),type = "re", ci.lvl = 0.89))  +
+   scale_fill_brewer(palette= "Set2") + labs(
+    title = "Predicted effects of bs_s on anxiety") + 
+  scale_y_continuous(limits = c(-1,1))
+```
+
+Next we turn to the "within country relative differences" model for belief in God on anxiety.
+
+## Within and between: believe in God
+
+```{r cache = TRUE}
+m2b <- brm(
+  anx_s ~ age_s  + male  +  bg_within + bg_between +  (1  +  bg_within|country),
+  data = d1,
   family = "gaussian",
-          seed=1234,
-          warmup=1500,
-          iter=6500,
-          chains=4)
-
-saveRDS(m1, "m1")
-
-```{r}
-m1<-readRDS("m1") # to avoid running this again in knitr, probably a better way
-#summary(m1) # raw summary
-# compare freq estimates with mcmc
-# tab_model(d2,m1)
-```
-
-```{r}
-# diagnositics
-mcmc_plot(m1,
+  seed = 1234,
+  warmup = 1000,
+  iter = 300,
+  chains = 4,
+  file = "m2_aa.rds")
+model_parameters(m2b, test = "p_direction")
+random_parameters(m2b)
+```
+
+Same workflow for our checks, but we'll want to use more informative priors next time.
+
+Trace plots look OK
+
+```{r}
+mcmc_plot(m2b,
           type = "trace")
-# mcmc_plot(m1,
-#           type = "areas")
-# mcmc_plot(m1,
-#           type = "violin")
-# mcmc_plot(m1,
-#           type = "intervals")
-
-# we see the oridinal qualities of our dataset here
-ppg<-pp_check(m1)
-ppg
-```
-
-```{r}
-# for math
-#stancode(m1)
-```
-```{r}
-# stanplot(m1, pars = NA,
-#          type = "intervals",
-#          fixed = FALSE,
-#          prob_outer = 0.95,
-#          exact_match = FALSE)
-```
-```{r}
-#better plot
-# m1fe<-plot_model(m1, title= "Credibility Coefficient Plot")
-# m1fe #1 x 1
-# 
-# m1re<-plot_model(m1, type ="re", sort.est = TRUE)
-# m1re
-# even better plot
-library(see)
-library(parameters)
-par<-parameters::model_parameters(m1,effects = "all", component = "all", test = c("pd"))
-par
-par2<-parameters::model_parameters(m1)
-#param_plot<- plot(par,  sort = TRUE) +theme_lucid() #option
-#param_plot<- plot(par,  sort = TRUE) +theme_abyss() #option
-#param_plot<- plot(par,  sort = TRUE) +theme_radar()#option
-param_plot<- plot(par,  sort = TRUE) +theme_radar()
-#param_plot<- plot(par,  sort = TRUE) +theme_modern()#option
-#param_plot<- plot(par,  sort = TRUE) +theme_blackboard()#option
-param_plot # my pref.
-```
-
-```{r}
-# not used
-# statcon<-make_conditions(dd, vars = c("REL"))
-# statcon
-# str(statcon)
-# cn = data.frame("REL" = c(1,.5,0), "cond__" = c("REL = 1","REL = .5", "REL = 0"))
-
-# Run these lines, I'm just commenting out for the document
-p1<-brms::conditional_effects(m1, c("religiosity.S:source"),
+```
+
+No surprise, there is the same problem with the sampling distribution.
+
+```{r}
+# we see the ordinal qualities of our dataset here
+pp_check(m2b)
+```
+
+Next the interval plots:
+
+```{r}
+ mcmc_plot(m2b,
+          type = "intervals",
+          prob_outer = 0.89)
+```
+
+Summary:
+
+```{r cache = TRUE}
+
+p<-brms::conditional_effects(m2b, "bg_within", 
                               spaghetti = T,
-                              nsamples = 100,
-                              probs = c(0.11, 0.89))
-pp1<-plot(p1, plot = FALSE)[[1]] + scale_y_continuous(limits = c(1,7))  + xlab("Religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Religiosity")  + theme_clean() #theme_radar_dark() + scale_fill_npg()
-pp1 #
-p2<-brms::conditional_effects(m1, c("culturalNorms.S:source"),
+                              nsamples = 300,
+                              prob = .89,
+                              method = 'fitted')
+
+plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1,1))  + xlab("bg_within") + ylab ("anxiety in sd units") + 
+  labs(title = "Predicted effects of bg_within")
+
+```
+
+And between
+
+```{r cache = TRUE}
+
+p<-brms::conditional_effects(m2b, "bg_between", 
                               spaghetti = T,
-                              nsamples = 100,
-                              probs = c(0.11, 0.89))
-pp2<-plot(p2, plot = FALSE)[[1]] + scale_y_continuous(limits = c(1,7))  + xlab("Cultural Norms of religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Cultural Norms") + theme_clean() #+ theme_radar_dark() + scale_fill_npg()
-pp2 #
-library(ggpubr)
-plotgood<-ggarrange(pp1, pp2, nrow=1, common.legend = T)
-#saveRDS(plotgood,"plotgood")
-#plotgood<-readRDS("plotgood")
-plotgood
-```
-
-# NOTE: an ordinal model would be better, but will require time to run, and it might muddy our inference.
-# 
-m1o <- brm(credible ~ religiosity.S * source +
-             culturalNorms.S * source +
-             Statement +
-             source_first +
-             sourceCorrect +
-             Age.Decades.C +
-             Edu.C +
-             Male +
-             SES.C +
-             logResp.C+
-             (1+source|country),
-           data=ddd,
-           inits = 0,
-           family = cumulative(link ="logit"),
-           seed=1234,
-           warmup=1000,
-           iter=3000,
-           chains=4)
-
-saveRDS(m1o,"m1o")
-
-# table and plots
-```{r}
-
-m1o<-readRDS("m1o")
-ppo<-pp_check(m1o)  ## fit is much better
-ppo
-ordinalmod<-parameters::model_parameters(m1o,effects = "all", component = "all", test = c("pd"))
-ordinalmod
-oplot<-plot(ordinalmod,  sort = TRUE) + theme_radar()
-oplot
-compareordplot<-ggarrange(ggarrange(param_plot,oplot,labels=c("a","b")),ggarrange(ppg, ppo, labels=c("c","d")), nrow=2, heights = c(3,1))
-compareordplot
-```
-## next model
-## have not inclueded this in the knitr call to save time. 
-```{r}
-statcon<-make_conditions(ddd, vars = c("source"))
-statcon
-str(statcon)
-
-p1<-brms::conditional_effects(m1o, c("religiosity.S"),
-                              categorical =T, # 
-                              probs = c(0.11, 0.89),
-                              conditions = statcon)
-pp1<-plot(p1, plot = FALSE)[[1]]   + xlab("Religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Religiosity")  + scale_y_continuous(limits = c(0,.4))  + theme_clean() + scale_color_grey() +scale_fill_grey() # note of you set categorical to true, then scale the y axis limits: c(0,1).
-Same for the plot below
-pp1 #
-p2<-brms::conditional_effects(m1o, c("culturalNorms.S"),
-                               categorical =T,
-                              probs = c(0.11, 0.89),
-                              conditions = statcon)
-pp2<-plot(p2, plot = FALSE)[[1]] + scale_y_continuous(limits = c(0,.4))  + xlab("Cultural Norms of Religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Cultural Norms")  + theme_clean() + scale_color_grey() +scale_fill_grey()
-pp2 #
-library(ggpubr)
-plotgood2.0<-ggarrange(pp1, pp2, labels=c("a","b"), nrow=1, common.legend = T)
-plotgood2.0
-saveRDS(plotgood2.0,"plotgood2.0")
-```
-
-```{r}
-# plot that can be read
-## We can set categorical to false to get more interpretable figures
-p1<-brms::conditional_effects(m1o, c("religiosity.S:source"),
-                              categorical =F, # note set to true to get all the ordinal values
-                              probs = c(0.11, 0.89),
-                                spaghetti = T,
-                              nsamples = 100) # play with this number
-pp1<-plot(p1, plot = FALSE)[[1]]   + xlab("Religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Religiosity")  + scale_y_continuous(limits = c(1,7))  + theme_clean() # note of you set categorical to true, then scale the y axis limits: c(0,1).
-Same for the plot below
-pp1 #
-p2<-brms::conditional_effects(m1o, c("culturalNorms.S:source"),
-                               categorical =F,
-                              probs = c(0.11, 0.89),
-                                spaghetti = T,
-                              nsamples = 100) # play with this number
-pp2<-plot(p2, plot = FALSE)[[1]] + scale_y_continuous(limits = c(1,7))  + xlab("Cultural Norms of Religiosity (standarised)") + ylab("Credibility: 1-7") + ggtitle("Expected Crediblity by Cultural Norms")  + theme_clean()
-pp2 #
-library(ggpubr)
-plotgood2<-ggarrange(pp1, pp2, labels=c("c","d"), nrow=1, common.legend = T)
-plotgood2
-saveRDS(plotgood2,"plotgood2")
-```
-
-```{r}
-# for omni graph
-plotgoodO<-ggarrange(pp1, pp2, nrow=1, common.legend = T)
-plotgoodO
-```
-
-```{r}
-#plotgood2<-readRDS("plotgood2")
-#plotgood2
-# compare ordinal plots
-twotypeordinalplot<-ggarrange(plotgood2.0,plotgood2,nrow=2)
-twotypeordinalplot
-```
-```{r}
-# everything plot
-omniplot<-ggarrange(ggarrange(param_plot,oplot,labels=c("a","b")),ggarrange(ppg, ppo, labels=c("c","d")), ggarrange(plotgood, plotgoodO, labels=c("e","f")), nrow=3, heights = c(3,1,1))
-omniplot
-```
-
-# 
-# 
-# ###### IF WE WANTED BOTH INDICATORS IN THE MODEL THIS IS THE CODe
-# m2 <- brm(mvbind(credible, important) ~ 
-#             religiosity.S * source + 
-#             culturalNorms.S * source +
-#             Statement +
-#             source_first +
-#             sourceCorrect +
-#             Age.Decades.C +
-#             Edu.C +
-#             Male +
-#             SES.C +
-#             logResp.C+
-#             (1+source|country),
-#           data=ddd,
-#           family = "gaussian",
-#           seed=1234,
-#           warmup=1000,
-#           iter=3000,
-#           chains=4)
-
-```
+                              nsamples = 300,
+                              prob = .89,
+                              method = 'fitted')
+
+plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1,1))  + xlab("bg_between") + ylab ("anxiety in sd units") + 
+  labs(title = "Predicted effects of bg_between")
+
+```
+
+Here's the between country graphs:
+
+```{r cache = TRUE}
+plot(ggeffects::ggpredict(m2b, terms = c("country[all]", "bg_within[minmax]"),type = "re", ci.lvl = 0.89))  +
+   scale_fill_brewer(palette= "Set2") + labs(
+    title = "Predicted effects of belief on anxiety")
+```
+
+To do:
+
+  1. Tidy up graphs and check all labels.
+
+  2. Use stronger priors and include stronger prior predictive checks
+
+  3. Sample from an ordinal outcome distribution: obtain un-transformed data from Suzanne
+
+  4. Develop discussion about model assumptions and limitations, such as
+
+      - Is the DAG we use here sensible?
+
+      - Measurement equivalence?
+
+      - India???
+
+      - Demand characteristics and heterogeneous populations.
+
+      - Is the theory that religion causes well-being biologically sensible? Why I don't think so...
+
+      - Connect with the literature.
+
+      - Better describe potential for Simpson's paradoxes: country level data collection is likely to give the wrong inference about religion and well being!
+
+      - Note that sociological-level data is what evolutionary scholars are collecting now -- and that such data do not clarify individual-level psychological processes.
+For individual-level processes we need to measure at the individual level.
+
+      - Make the plug for longitudinal data: we have no understanding of demand characteristics. We get no real inference from the model here because our data are too limited.
+
+      - Make a plug for DAGs, clarify how much trouble we can get into without an explicit causal model.
+
+To be continued...

### 2021-01-15 analysis.Rmd
proofed.

@@
 
-Workflow and analysis
-
-First import data (see functions)
+## Workflow and analysis
+
+First, we import the data (see functions).
 
@@
 
-Inspect the data, and note that many columns data have been transformed from their raw form.
+Next, we inspect the data.
+A key problem is that many columns data have been transformed from their raw form into a number between 0 and 1.
+Later we'll see that these transformations render the modelling less reliable than if the data were retained.
+
+So for the next iteration, we'll need the data as they were recorded. Specially we'll focus on the variables such as
+
+> How often do you have negative feelings such as blue mood, despair, anxiety, depression
+
+We'll investigate church frequency and belief in God as predictors of this dimension of well-being.
 
@@
 ```
-
-Ask Suzanne for raw data so that we can, e.g. model church frequency as a monotonic categorical variable, and pick appropriate sampling distributions for the outcome variables.
 
@@
 wbsoc <- get_wb_soc(df)
-head(wbsoc)
-head(rel)
-```
-
-Exploration Graphs.  Inspect the data.
+
+```
+
+Next some exploration Graphs.  Inspecting the sample responses on religion.
 
@@
 
+India is a big outlier: much more religious than other countries.
+
+Let's look at the anxiety indicator: india is much lower:
+. 
+```{r}
+wb1 <- wbpsych %>%
+  ggplot(aes(
+    x = country,
+    y = wb_psychANXIETY,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  scale_color_viridis(discrete = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  ggtitle("Importance Believe in God") +
+  ylab("") + 
+  theme(axis.text.x = element_text(angle = 30))
+wb1
+
+```
+
+We're interested in  meaning of life as a causal mediator of anxiety
+
+```{r}
+wb2 <- wbpsych %>%
+  ggplot(aes(
+    x = country,
+    y = wb_psychMEANING,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  scale_color_viridis(discrete = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  ggtitle("Importance Believe in God") +
+  ylab("") + 
+  theme(axis.text.x = element_text(angle = 30))
+wb2
+```
+
+India again very high, and France is very low.
+Can we really assume measurement invariance across these countries?
+Are people so really different?
+As with any model, we condition on assumptions, however at another iteration of this study, however for now we will assume measurement invariance.
+
 Next, a quick run of correlations. Note that we need the to model the country level dependencies in any correlation matrix, given strong cultural differences
@@
 
-# Specify our causal model
+## A critical step: the causal model
 
@@
 
-Which variables should we condition on to understand the relationship between religion and well-being?
+### Which variables should we condition on to understand the relationship between religion and well-being?
+
+We have two sets adjustments sets to chose from. We will pick the set with the measured indicators.
 
@@
 
+### Note: there is collider counfounding if we include all variables into our model
+
 Suppose we were to include *all* the variables in our model, except U2 and U1 which are not measured?
@@
 
-## The hard work
+## Specifying the mechanism in relation to this dataset.
 
@@
 We will focus on two non-trivial theories -- one related to religious behavior and another related to religious belief.
+(Citations to follow ...)
 
@@
 
+### Dag shoes two adjustment sets
 ```{r}
@@
 
-Before proceeding with the model, let's note a curiosity.
-If our  interest is in the direct effects of church attendance on well-being we should *not* include social support because this introduces a collider bias!
+### Confounder bias if we include our theoretically proposed causal mediators!
+
+Before proceeding with the model, note a curiosity.
+If our interest is in the direct effects of church attendance on well-being we should *not* include social support because this introduces a collider bias.
 
@@
 
-Let this serve as a warning for us to think about our causation, and what we might want to know before we leap into modeling
-
-## The model
+Let this serve as a warning for us to think about our causation. We must develop our causal model before we leap into our statistical model.
+
+## The statistical model
 
@@
         demean(d1, select = c("anx", "ch", "bg"), group = "country"))
-head(d1)
 ```
@@
 
-We can see there is no effect of church attendance on anxiety (according to this model). Note that throughout, the plots will sample 300 draws from the posterior distribution.
+We can see there is no effect of church attendance on anxiety (according to this model). (Note that throughout, the plots will sample 300 draws from the posterior distribution. )
 
@@
 
-Before racing ahead, lets do some posterior predictive checks
+Before racing ahead, lets do some posterior predictive checks on the model
+
+The trace plots look good.
 
@@
 
-When we do a posterior predictive check we find problems in the predicted distribution.
-This we called this a gaussian, but it is not.
-The data have large peaks reflecting the categorical qualities of the outcome variable.
-
-We should work with the original, untransformed values of the data.
+However the posterior predictive check reveals that we should work with the original untransformed values of the outcome variable data.
 
@@
 
-<!-- ```{r cache = TRUE} -->
-<!-- plot(ggeffects::ggpredict(m1_prior, terms = c("ch_s[minmax]"),type = "fe", ci.lvl = 0.89))  +  scale_fill_brewer(palette= "Set2") + labs( title = "Predicted effects of church on anxiety") -->
-<!-- ``` -->
-
 or we can simulate and see that our prior is scattered across five standard deviations of anxiety, and we can see our prior is telling us that anything can happen
+.
 
@@
 
-Nothing happening at the individual level.
-
-```{r include = FALSE}
-# graph not as nice
-# plot(ggeffects::ggpredict(m1, terms = c("ch_s[minmax]"),type = "fe", ci.lvl = 0.89))  +
-#    scale_fill_brewer(palette= "Set2") + labs(
-#     title = "Predicted effects of church on anxiety")
-```
-
-But lots of variability at the country level intercepts for anxiety.
-We can see that India is over two standard deviations  less anxious than the average culture level anxiety.
-By contrast, Japan, Israel, and France are over .2 standard deviations more anxious than average.
+## The group level effects of country
+
+There is lots of variability at the country level intercepts for anxiety.
+We can see that India is over .5 standard deviations  less anxious than the average culture level anxiety.
+By contrast, Japan, Israel, and France are over .3 standard deviations more anxious than average.
 There looks to be perhaps some covariance between the country intercepts and slopes...
@@
 
-Let's delving into this question of country level effects, but asking whether people who are relatively more church attending tend to have more anxiety.
+Let's delving into this question of country level effects by asking whether people who are *relatively* more church attending within their countries tending to have more anxiety.
 
@@
 
-- between_church centering = the adjustment for country level differences in church attendance.
+- between_church centering = country level differences in church attendance.
 
@@
 https://easystats.github.io/parameters/reference/demean.html
+
+The model.
 
@@
 
-No reliable effect within,
+We find no relation of "ch_within" to anxiety.
 
@@
 
-But a between-country effect emerges:
+However, a strong between-country effect emerges:
 
@@
 
-Same workflow as before.
+Before considering this, let's run some model checks workflow as before.
 

### 2021-01-15 funs.R

@@ -16,7 +16,7 @@ get_rel_vars <- function(x) {
 }
 
 get_wb_gen <- function(x) {
-  df <- x %>% dplyr::select(c(starts_with("wb_gen"), "country"))
+  df <- x %>% dplyr::select(c(starts_with("wb_gen"), "country")) 
 }
 
 get_wb_phys <- function(x) {
@@ -24,7 +24,8 @@ get_wb_phys <- function(x) {
 }
 
 get_wb_psych <- function(x) {
-  df <- x %>% dplyr::select(c(starts_with("wb_psych"), "country"))
+  df <- x %>% dplyr::select(c(starts_with("wb_psych"), "country"))%>%
+    dplyr::select(-wb_psych_mean)
 }
 
 

### 2021-01-15 libs.R

@@ -6,6 +6,7 @@ library("sjPlot")
 library("sjmisc")
 library("insight")
 library("see")
+library("performance")
 library("ggraph")
 library("parameters")
 library("easystats")

### 2021-01-28 analysis.Rmd
ordinal analysis
comments on Scott's study

@@
   ) %>%
-  dplyr::select(anx, ch, mn, bg, anx_s, age_s, ch_s, mn_s, male2, country, male, bg_s)
+  dplyr::select(anx, ch, mn, bg, anx_s, age_s, ch_s, mn_s, male2, country, male, bg_s,wb_psychANXIETY)
 
@@
 
-```{r}
+```{r eval= FALSE}
 prior_summary(m1)
@@
 
-```{r}
+```{r eval = FALSE}
 bayestestR::describe_prior(m1)
@@
 The model.
-
+wb_psychANXIETY
 ```{r}
@@
 
+## TRY ORDINAL MODEL (note church attend should be monotonic -- at next iteration)
+
+```{r}
+m1o <- brm(
+  wb_psychANXIETY ~ ch_s +
+    age_s  +
+    male +
+    (1 + ch_s | country),
+  data = d1,
+  family = cumulative(link ="logit"),
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = "m1o"
+)
+parameters::model_parameters(m1o, test = "p_direction")
+parameters::random_parameters(m1o)
+```
+
+This resolves the problem
+```{r}
+# YES
+pp_check(m1o)
+```
+
+Trace plots:
+
+```{r, cache = TRUE}
+mcmc_plot(m1o,
+          type = "trace")
+```
+
+Model summary
+```{r}
+parameters::model_parameters(m1o)
+```
+
+Plot model
+```{r, cache = TRUE}
+library(viridis)
+p0b<-brms::conditional_effects(m1o, "ch_s", 
+                              spaghetti = T,
+                              nsamples = 200,
+                              categorical = T,
+                              prob = .89,
+                              method = 'fitted')
+
+plot(p0b, plot = FALSE)[[1]]   + xlab("ch_s") + ylab ("anxiety 1-5") +
+  labs(title = "Predicted effects of ch_s")# + scale_y_continuous(limits= c(1,5))# +  scale_color_viridis(discrete=TRUE)
+
+```
+
+<!-- Another plot method -->
+
+<!-- ```{r, cache = TRUE} -->
+<!-- library(viridis) -->
+<!-- p0b1<-brms::conditional_effects(m1o, "ch_s",  -->
+<!--                               spaghetti = T, -->
+<!--                               nsamples = 200, -->
+<!--                              categorical = F, -->
+<!--                               prob = .89, -->
+<!--                               method = 'fitted') -->
+
+<!-- plot(p0b1, plot = FALSE)[[1]]   + xlab("ch_s") + ylab ("anxiety 1-5") + -->
+<!--   labs(title = "Predicted effects of ch_s")+ scale_y_continuous(limits= c(1,5))# +  scale_color_viridis(discrete=TRUE)  -->
+
+<!-- ``` -->
+
+### without demeaning
+
+```{r}
+m1_o <- brm(
+  wb_psychANXIETY ~ age_s  + male + ch_within +  ch_between + (1 + ch_within|country),  
+  data = d1,
+  family = cumulative(link ="logit"),
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = "m1a_o")
+```
+
+### Within /between demeaning
+
+```{r}
+m1a_o <- brm(
+  wb_psychANXIETY ~ age_s  + male + ch_within +  ch_between + (1 + ch_within|country),  
+  data = d1,
+  family = cumulative(link ="logit"),
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = "m1a_o")
+```
+
+This resolves the problem
+```{r}
+# YES
+pp_check(m1a_o)
+```
+
+Trace plots:
+
+```{r, cache = TRUE}
+mcmc_plot(m1a_o,
+          type = "trace")
+```
+
+Model summary
+```{r}
+parameters::model_parameters(m1a_o)
+```
+
+Within: NOTHING
+```{r, cache = TRUE}
+library(viridis)
+pch_w<-brms::conditional_effects(m1a_o, "ch_within", 
+                              spaghetti = T,
+                              nsamples = 200,
+                              categorical = TRUE,
+                              prob = .89,
+                              method = 'fitted')
+
+ppch_w<-plot(pch_w, plot = FALSE)[[1]]    + ylab ("anxiety in sd units") +
+  labs(title = "Predicted effects of Church_within") # +  scale_color_viridis(discrete=TRUE) 
+ppch_w
+
+```
+
+BETWEEN: Yes, and at the highest levels of anxiety.
+
+```{r cache = TRUE}
+
+pch_b<-brms::conditional_effects(m1a_o, "ch_between", 
+                              spaghetti = T,
+                              nsamples = 200,
+                              prob = .89,
+                              categorical = TRUE,
+                              method = 'fitted')
+
+ppch_b<- plot(p, plot = FALSE)[[1]]   + xlab("ch_between") + ylab ("anxiety: 1-5") + 
+  labs(title = "Predicted effects of Church_between")
+ppch_b
+```
+
+<!-- ### Different plot method -->
+
+<!-- Summing over the categories: -->
+
+<!-- ```{r, cache = TRUE} -->
+<!-- library(viridis) -->
+<!-- p0a<-brms::conditional_effects(m1a_o, "ch_within",  -->
+<!--                               spaghetti = T, -->
+<!--                               nsamples = 200, -->
+<!--                              # categorical = TRUE, -->
+<!--                               prob = .89, -->
+<!--                               method = 'fitted') -->
+
+<!-- plot(p0a, plot = FALSE)[[1]]   + xlab("bg_within") + ylab ("anxiety in sd units") + -->
+<!--   labs(title = "Predicted effects of bg_within")  + scale_y_continuous(limits = c(1,5))# +  scale_color_viridis(discrete=TRUE)  -->
+<!-- ``` -->
+
+<!-- ```{r cache = TRUE} -->
+
+<!-- ps<-brms::conditional_effects(m1a_o, "ch_between",  -->
+<!--                               spaghetti = T, -->
+<!--                               nsamples = 200, -->
+<!--                               prob = .89, -->
+<!--                              # categorical = TRUE, -->
+<!--                               method = 'fitted') -->
+
+<!-- plot(ps, plot = FALSE)[[1]]   + xlab("bg_between") + ylab ("anxiety: 1-5") +  -->
+<!--   labs(title = "Predicted effects of bg_between") + scale_y_continuous(limits = c(1,5)) -->
+
+<!-- ``` -->
+
+## Summary of Church attendence Model
 Which we interpret as follows:
@@
 ```{r cache = TRUE}
-plot(ggeffects::ggpredict(
-  m1a,
-  terms = c("country[all]", "ch_within[minmax]"),
-  type = "re",
-  ci.lvl = 0.89))  +
-  scale_fill_brewer(palette = "Set2") + 
-  labs(title = "Predicted effects of church on anxiety: countries X within") + 
-  scale_y_continuous(limits = c(-1,1))
+ppch_w
 ```
@@
 ```{r cache = TRUE}
-plot(ggeffects::ggpredict(
-  m1a,
-  terms = c("ch_between[minmax]"),
-  type = "re",
-  ci.lvl = 0.89))  +
-  scale_fill_brewer(palette = "Set2") + labs(title = "Predicted effects of church on anxiety: between")  + scale_y_continuous(limits = c(-1, 1))
+ppch_b
 ```
@@
 tidy_ggdag2 <- dagify(
-  wb  ~ meaning + ses + male + country  +U2,
+  wb  ~ meaning + ses + male + country  + U2,
   meaning ~ bg +  male + country,
@@
   exposure =  "bg",
-  outcome =   "wb") %>%
+  outcome =   "wb"
+) %>%
   tidy_dagitty()
@@
   theme_dag_blank()
+
+tidy_ggdag2 <- dagify(
+  wb  ~ meaning + ses + male + country  + U2,
+  meaning ~ bg +  male + country,
+  bg ~ age + male + country + U1,
+  ses ~ country + age + U2,
+  exposure =  "bg",
+  outcome =   "wb"
+) %>%
+  tidy_dagitty()
+
+# graph adjustment sets
+ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) +
+  theme(legend.position = "bottom") +
+  theme_dag_blank()
+
+```
+
+```{r}
 
@@
 
-p<-brms::conditional_effects(m2, "bg_s", 
+p <- brms::conditional_effects(
+  m2,
+  "bg_s",
   spaghetti = T,
@@
   prob = .89,
-                              method = 'fitted')
-
-plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1,1))  + xlab("bg_s") + ylab ("anxiety in sd units") + 
+  method = 'fitted'
+)
+
+plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1, 1))  + xlab("bg_s") + ylab ("anxiety in sd units") +
   labs(title = "Predicted effects of bg_s")
@@
 ```{r cache = TRUE}
-
-p<-brms::conditional_effects(m2, "bg_s", 
+p <- brms::conditional_effects(
+  m2,
+  "bg_s",
   spaghetti = T,
@@
 ```{r cache = TRUE}
-plot(ggeffects::ggpredict(m2, terms = c("country[all]", "bs_s[minmax]"),type = "re", ci.lvl = 0.89))  +
-   scale_fill_brewer(palette= "Set2") + labs(
-    title = "Predicted effects of bs_s on anxiety") + 
-  scale_y_continuous(limits = c(-1,1))
+plot(ggeffects::ggpredict(
+  m2,
+  terms = c("country[all]", "bs_s[minmax]"),
+  type = "re",
+  ci.lvl = 0.89
+))  +
+  scale_fill_brewer(palette = "Set2") + labs(title = "Predicted effects of bs_s on anxiety") +
+  scale_y_continuous(limits = c(-1, 1))
 ```
@@
 m2b <- brm(
-  anx_s ~ age_s  + male  +  bg_within + bg_between +  (1  +  bg_within|country),
+  anx_s ~ age_s  + male  +  bg_within + bg_between +  (1  +  bg_within |
+                                                         country),
   data = d1,
@@
   chains = 4,
-  file = "m2_aa.rds")
+  file = "m2_aa.rds"
+)
 model_parameters(m2b, test = "p_direction")
@@
 
+## Checks
+
 Same workflow for our checks, but we'll want to use more informative priors next time.
@@
 No surprise, there is the same problem with the sampling distribution.
-
-```{r}
-# we see the ordinal qualities of our dataset here
-pp_check(m2b)
-```
-
-Next the interval plots:
 
@@
 
+```{r}
+# we see the ordinal qualities of our dataset here
+pp_check(m2b)
+```
+
+Next the interval plots:
+
+```{r}
+ mcmc_plot(m2b,
+          type = "intervals",
+          prob_outer = 0.89)
+```
+
 Summary:
@@
 
-p<-brms::conditional_effects(m2b, "bg_within", 
+p <- brms::conditional_effects(
+  m2b,
+  "bg_within",
   spaghetti = T,
@@
   prob = .89,
-                              method = 'fitted')
-
-plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1,1))  + xlab("bg_within") + ylab ("anxiety in sd units") + 
+  method = 'fitted'
+)
+
+plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1, 1))  + xlab("bg_within") + ylab ("anxiety in sd units") +
   labs(title = "Predicted effects of bg_within")
@@
 ```{r cache = TRUE}
-
 p<-brms::conditional_effects(m2b, "bg_between", 
@@
 
-plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1,1))  + xlab("bg_between") + ylab ("anxiety in sd units") + 
+plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1, 1))  + xlab("bg_between") + ylab ("anxiety 1-5") +
   labs(title = "Predicted effects of bg_between")
@@
 ```{r cache = TRUE}
-plot(ggeffects::ggpredict(m2b, terms = c("country[all]", "bg_within[minmax]"),type = "re", ci.lvl = 0.89))  +
-   scale_fill_brewer(palette= "Set2") + labs(
-    title = "Predicted effects of belief on anxiety")
-```
-
-To do:
+plot(ggeffects::ggpredict(
+  m2b,
+  terms = c("country[all]", "bg_within[minmax]"),
+  type = "re",
+  ci.lvl = 0.89
+))  +
+  scale_fill_brewer(palette = "Set2") + labs(title = "Predicted effects of belief on anxiety")
+```
+
+## TRY ORDINAL MODEL
+
+## Without demeaning
+```{r}
+m2o <- brm(
+  wb_psychANXIETY ~ bg_s +
+    age_s  +
+    male +
+    (1 + ch_s | country),
+  data = d1,
+  family = cumulative(link = "logit"),
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = "m2o"
+)
+parameters::model_parameters(m2o, test = "p_direction")
+parameters::random_parameters(m2o)
+```
+
+This resolves the problem
+```{r}
+# YES
+pp_check(m2o)
+```
+
+Trace plots:
+
+```{r, cache = TRUE}
+mcmc_plot(m2o,
+          type = "trace")
+```
+
+Model summary
+```{r}
+parameters::model_parameters(m2o)
+```
+
+Plot model: nothing
+
+```{r, cache = TRUE}
+library(viridis)
+p0 <- brms::conditional_effects(
+  m2o,
+  "bg_s",
+  spaghetti = T,
+  categorical = TRUE,
+  nsamples = 200,
+  prob = .89,
+  method = 'fitted'
+)
+
+plot(p0, plot = FALSE)[[1]]   + xlab("bg") + ylab ("anxiety 1-5") +
+  labs(title = "Predicted effects of bg") # +  scale_color_viridis(discrete=TRUE) 
+```
+
+###  With demeaning
+
+```{r}
+m2_o <- brm(
+  wb_psychANXIETY ~ age_s  + male + bg_within +  bg_between + (1 + ch_within|country),  
+  data = d1,
+  family = cumulative(link ="logit"),
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = "m2_o")
+```
+
+This resolves the problem
+```{r}
+# YES
+pp_check(m2_o)
+```
+
+Trace plots:
+
+```{r, cache = TRUE}
+mcmc_plot(m2_o,
+          type = "trace")
+```
+
+Model summary
+```{r}
+parameters::model_parameters(m2_o)
+```
+
+Within: NOTHING
+```{r, cache = TRUE}
+library(viridis)
+p0bg <- brms::conditional_effects(
+  m2_o,
+  "bg_within",
+  spaghetti = T,
+  nsamples = 200,
+  categorical = TRUE,
+  prob = .89,
+  method = 'fitted'
+)
+
+ppbg_w<- plot(p0bg, plot = FALSE)[[1]]   + xlab("bg_within") + ylab ("anxiety 1-5") +
+  labs(title = "Predicted effects of BeliefGod_within") # +  scale_color_viridis(discrete=TRUE) 
+ppbg_w
+```
+
+BETWEEN: Yes, and at the highest levels of anxiety.
+
+```{r cache = TRUE}
+pbg <- brms::conditional_effects(
+  m2_o,
+  "bg_between",
+  spaghetti = T,
+  nsamples = 200,
+  prob = .89,
+  categorical = TRUE,
+  method = 'fitted'
+)
+
+ppbg_b<- plot(pbg, plot = FALSE)[[1]]   + xlab("bg_between") + ylab ("anxiety: 1-5") +
+  labs(title = "Predicted effects of BeliefGod_between")
+ppbg_b
+```
+
+## Summary graph
+
+```{r}
+library(patchwork)
+ppch_w +  ppbg_w  + ppch_b + ppbg_b + plot_annotation("What is the relationship between belief church and belief on anxiety?", subtitle = "No evidence for within-person differences, evidence for between country differences at the high end of anxiety", tag_levels = "a") +  plot_layout(guides = 'collect')
+
+#bulbulia_summary_graph_many_analysts
+```
+
+## To do
 
@@
 
-  2. Use stronger priors and include stronger prior predictive checks
-
-  3. Sample from an ordinal outcome distribution: obtain un-transformed data from Suzanne
+  2. Remove clutter.
+
+  3. Use stronger priors and include stronger prior predictive checks
+
+  3. Sample from an ordinal outcome distribution: obtain un-transformed data from Suzanne NOTE: DONE!, but still need to get church frequency as a monotonic predictor!
 
@@
 To be continued...
+
+## APPENDIX

### 2021-01-28 analysis.Rmd
looked at missingness in the denomination variable

@@
 vis_miss(df) # substantial pattern of missing ness in a few variables, 
+gg_miss_upset(df)
 ```
@@
 
-### Dag shoes two adjustment sets
+### Dag shoes two adjustment sets both of which contain denomination
 ```{r}
@@
   wb  ~ support + ses + male + country + U2,
-  support ~ church + male + country,
-  church ~ age + male + country + U1,
+  support ~ church + male + country + den,
+  church ~ age + male + country + den + U1,
   ses ~ country + age + U2,
@@
   dplyr::filter(
-    !is.na(wb_psychANXIETY),!is.na(age),!is.na(rel_CHURCHFREQ),!is.na(wb_psychMEANING),!is.na(male),!is.na(rel_EXTENTBELIEVEGOD)
+    !is.na(wb_psychANXIETY),!is.na(age),!is.na(rel_CHURCHFREQ),!is.na(wb_psychMEANING),!is.na(male),!is.na(rel_EXTENTBELIEVEGOD),
   ) %>%

### 2021-02-09 dags_for_scott.Rmd
fixed material

---
title: "Dags for Scott"
author: "Joseph Bulbulia"
date: "28/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Dag shoes two adjustment sets
```{r}
library(ggdag)
library(dagitty)
library(ggplot2)
library(tidyverse)

# more complex
dg <- dagify(
  ms  ~  mg + time,
  mg ~ time + L,
  L ~ time, 
  exposure =  "mg",
  outcome =   "ms"
) %>%
  tidy_dagitty()

colliders<-ggdag::ggdag_collider(dg,
                      from = "mg",
                      to = "ms",
                      control_for(c("time", "L")))+   theme_dag_blank() + labs(title = "colliders")
colliders
p1<-ggdag(dg)+ theme_dag_blank() + labs(title = "dag")

# graph adjustment sets
p2 <-ggdag::ggdag_adjustment_set(dg) + labs(title = "adjustment set") +
  theme(legend.position = "bottom") + theme_dag_blank() 
p1 + p2 + plot_annotation(title = "To infer path from MG --> MS we only need to condition on time",
                          subtitle = "assume:  
                          ms ~ mg + time, 
                          mg ~ time + L, 
                          L ~ time, 
                          exposure =  mg,
                          outcome =   ms")

ggdag_dseparated(dg,
  from = "mg",
  to = "ms",
  controlling_for = c("time", "L"),
  collider_lines = TRUE
) + theme_dag_blank() + labs(title = "d-relationships")

p1 + plot_annotation(title = "To infer path from MG --> MS we only need to condition on time")

# ggdag_drelationship(dg,
#   from = "mg",
#   to = "ms",
#   controlling_for = c("time","L"),
#   collider_lines = TRUE
# ) + theme_dag_blank()

##play
dg2 <- dagify(
  mg  ~ time + L,
  ms ~  time + L,
  mg ~~ ms,
  L ~ time, 
  exposure =  "ms",
  outcome =   "mg"
) %>%
  tidy_dagitty()

ggdag::ggdag_collider(dg2,
                      from = "ms",
                      to = "mg",
                      control_for(c("time", "L"))) +   theme_dag_blank()

# graph adjustment sets
ggdag::ggdag_adjustment_set(dg2, node_size = 14) +
  theme(legend.position = "bottom") + theme_dag_blank()

## SIMPLE
dg2 <- dagify(
  mgPRE  ~ time,
  ms ~  mgPRE + time + U,
  U ~ time, 
  exposure =  "mgPRE",
  outcome =   "ms"
) %>%
  tidy_dagitty()

jb1<-ggdag::ggdag_collider(dg2,
                      from = "mgPRE",
                      to = "ms",
                      control_for(c("time", "U"))) +   theme_dag_blank() + labs(title = "colliders")
jb1

# graph adjustment sets
jb2<- ggdag::ggdag_adjustment_set(dg2, node_size = 14) +
  theme(legend.position = "bottom") + theme_dag_blank()  + labs(title = "adjustment set")

jb2
library(patchwork)
jb1 / jb2 + plot_annotation(title = "If we think mega-societies and moralising gods co-evolve then we we might not get an obtain an unbaised estimate", tag_levels = "a")

#####

# more complex
tidy_ggdag2 <- dagify(
  wb  ~ rel + age + male + ses + cntry + den,
  rel ~ age + male + cntry + den,
  ses ~ age  + cntry,
  cntry ~ lang + den, 
  exposure =  "rel",
  outcome =   "wb"
) %>%
  tidy_dagitty()

#Here's the trouble with Scott's model
ggdag::ggdag_collider(tidy_ggdag2,
                      from = "rel",
                      to = "wb",
                      control_for(c("ses", "male", "cntry", "age","den","lang"))) +   theme_dag_blank()

# graph adjustment sets
ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) +
  theme(legend.position = "bottom")

ggdag_drelationship(tidy_ggdag2,
  from = "rel",
  to = "wb",
  controlling_for = c("ses", "male", "cntry", "age", "den","lang"),
  collider_lines = TRUE
) + theme_dag_blank()

```

```{r}
tidy_ggdag2 <- dagify(
  wb  ~ rel + age + male + ses + cntry + den,
  rel ~ age + male + cntry + den,
  ses ~ age  + cntry,
  cntry ~ den, 
  exposure =  "rel",
  outcome =   "wb"
) %>%
  tidy_dagitty()

#Here's the trouble with Scott's model
ggdag::ggdag_collider(tidy_ggdag2,
                      from = "rel",
                      to = "wb",
                      control_for(c("ses", "male", "cntry", "age","den","lang"))) +   theme_dag_blank()

# graph adjustment sets
ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) +
  theme(legend.position = "bottom") 
```

```{r eval }
tidy_k6 <- dagify(
  k6  ~ covid + mean + rel + male + income + age + rest,
  mean ~ rel + male  + age + rest,
  rel ~ age  + male,
  income ~ edu, 
  exposure =  "rel",
  outcome =   "k6"
) %>%
  tidy_dagitty()

#Here's the trouble with Scott's model
ggdag::ggdag_collider(tidy_k6,
                      from = "covid",
                      to = "k6",
                      control_for(c("rel", "male", "mean", "age","income"))) +   theme_dag_blank()

# graph adjustment sets
ggdag::ggdag_adjustment_set(tidy_k6, node_size = 14) +
  theme(legend.position = "bottom")

ggdag::ggdag_collider(tidy_k6,
                      from = "covid",
                      to = "k6",
                      controlling_for = c( "male","age")) +   theme_dag_blank()

ggdag_drelationship(tidy_k6,
from = "covid",
                      to = "k6",
                      controlling_for = c( "male","age","covid"), collider_lines = TRUE
) + theme_dag_blank()

```

```{r eval = FALSE}
remotes::install_github("BS1125/CMAverse")
library(CMAverse)
cmdag(outcome = "K6", exposure = "cv", mediator = c("FT", "Work"), 
      basec = c("C1", "C2"), postc = NULL, node = FALSE, text_col = "black")

cmdag(
  outcome = "K6",
  exposure = "COV",
  mediator = "F",
  basec = "male",
  postc = "reld"
)

```

### Confounder bias if we include our theoretically proposed causal mediators!

Before proceeding with the model, note a curiosity. If our interest is in the direct effects of church attendance on well-being we should *not* include social support because this introduces a collider bias.

### 2021-04-25 analysis.Rmd
analysis commit

@@
 source("funs.R") # custom functions
+
+# libraries
+# libraries and settings
+library("tidyverse")
+library("easystats")
+library("ggplot2") #plots
+library("sjstats") #test statistics
+library("sjPlot")
+library("sjmisc")
+library("insight")
+library("see")
+library("performance")
+library("ggraph")
+library("parameters")
+library("lme4")
+library("psych") ## correlation matrix
+library("table1")
+library("ggeffects")
+library("gghighlight")
+#library("naniar") # missingness
+#library("Amelia")
+#library("magrittr")
+library("patchwork")
+library("texreg")
+library("viridis")
+library("ggpubr")
+library("ggsci") 
+library("here")
+library("citr")
+library("rethinking")
+#library("statsExpressions")
+library("ggstatsplot")
+library("ggthemes")
+library("ggdag")
+library("papaja")
+library("rstan")
+library("brms")
+library("tidybayes")
+library("bayesplot")
+library("dagitty")
+library("ggdag")
+rstan_options(auto_write=TRUE)
+options(mc.cores=parallel::detectCores ())
+theme_set(theme_few())
 ```
@@
 # import data
-d <- data_read_mymachine()
+d <- as.data.frame(
+  read.csv(here::here("MARP_data_blinded.csv")
+  )
+)
 ```
@@
 
-Next we need to examine patterns of missingness, which might spoil inference.
+Next we need to examine patterns of missing-ness, which might spoil inference.
 
@@
     (1 + ch_s | country),
+  sample_prior = TRUE,
   data = d1,
@@
 ```{r}
+library(coda)
 samples1 <- brms::prior_samples(m1, "b_ch_s")
-samples1
 ```

### 2021-04-25 funs.R

@@ -2,7 +2,7 @@
 # functions used in this analysis
 data_read_mymachine <- function() {
   out <-
-    read.csv("/Users/jbul176/Documents/GitHub/many_analysts/MARP_data_blinded.csv")
+    read.csv(here::here("MARP_data_blinded.csv"))
 }
 
 

### 2021-04-25 analysis.Rmd
ignore big

@@
 )
+head(m2_aa.rds)
 model_parameters(m2b, test = "p_direction")

### 2021-04-29 analysis.Rmd
update analysis, git ignore the large .rds files for BRMS outputs

--- 2021-04-25 analysis.Rmd
@@
 ---
-title: "Analysis"
-author: "Joseph Bulbulia"
-date: "24/12/2020"
-output: html_document
+title: "Many Analystis Project"
+description: 
+author:
+  - name: Joseph Bulbulia
+    url: https://josephbulbulia.netlify.app
+    affiliation: Victoria University of Wellington
+    affiliation_url: https://www.wgtn.ac.nz
+    orcid_id: 0000-0002-5861-2056
+date: 2021-APRIL-29 
+output:
+  distill::distill_article:
+    self_contained: true
+    toc: true
 ---
@@
 ```
+
 ```{r runanalysis, echo=FALSE, include = FALSE}
@@
 d <- as.data.frame(
-  read.csv(here::here("MARP_data_blinded.csv")
+  read.csv(here::here("MARP_data.csv")
   )
@@
 Next, we inspect the data.
-A key problem is that many columns data have been transformed from their raw form into a number between 0 and 1.
+A challenge is that many columns data have been transformed from their raw form into a number between 0 and 1.
 Later we'll see that these transformations render the modelling less reliable than if the data were retained.
@@
 ```
+
 ```{r, include = FALSE}
@@
 table(d$ethnicity,  useNA = "ifany") # need recode NA as "not identified"
-table(d$denomination,useNA = "ifany") # need recode NA as "no denomination" 
+table(d$denomination, useNA = "ifany") # need recode NA as "no denomination" 
 # fix
@@
 ```
+
 ```{r, include = FALSE}
 # what are the numbers here? 
-table(d$attention_check,useNA = "ifany") 
+table(d$attention_check, useNA = "ifany")
 d$wb_phys_mean
@@
 # clean data:use informative names !
-df<- d %>%
-  dplyr::mutate(male2 = ifelse( gender != "man", 1 ,2)) %>% # for index categorical variable, not change of meaning for gender. male and people who do not identify as male. 
-  dplyr::mutate(male = factor(ifelse( gender != "man", "male" ,"notmale"))) %>%
+df <- d %>%
+  dplyr::mutate(male2 = ifelse(gender != "man", 1 , 2)) %>% # for index categorical variable, not change of meaning for gender. male and people who do not identify as male.
+  dplyr::mutate(male = factor(ifelse(gender != "man", "male" , "notmale"))) %>%
   rename(
@@
     #   wb_socSATSEX   = wb_soc_3,
-  )%>%
-  dplyr::mutate(IS_RELIIGOUS = as.factor(IS_RELIIGOUS),
+  ) %>%
+  dplyr::mutate(
+    IS_RELIIGOUS = as.factor(IS_RELIIGOUS),
     denomination = as.factor(denomination),
     ethnicity = as.factor(ethnicity),
-                country = as.factor(country)) %>% 
-  dplyr::mutate(IS_RELIIGOUS = recode_factor(IS_RELIIGOUS, "0" = "atheist", "0.5" = "not_religious", "1" = "religious"))
+    country = as.factor(country)
+  ) %>%
+  dplyr::mutate(
+    IS_RELIIGOUS = recode_factor(
+      IS_RELIIGOUS,
+      "0" = "atheist",
+      "0.5" = "not_religious",
+      "1" = "religious"
+    )
+  )
 
@@
 library("naniar")
-vis_miss(df) # substantial pattern of missing ness in a few variables, 
+vis_miss(df) # substantial pattern of missingness in a few variables, 
 gg_miss_upset(df)
@@
 library("patchwork")
-(r1 + r2  +  r3 )/  (r4  + r5) / (r6 +  r7)   +  plot_annotation(tag_levels = 'A') + plot_layout(guides = "collect")
-```
-
-India is a big outlier: much more religious than other countries.
-
-Let's look at the anxiety indicator: india is much lower:
-. 
+(
+  r1 + r2  +  r3
+)/  (r4  + r5) / (r6 +  r7)   +  plot_annotation(tag_levels = 'A') + plot_layout(guides = "collect")
+```
+
+Substantial between country variation evident. India = very high
+
+Let's look at the anxiety indicator: India is much lower: .
+
 ```{r}
@@
 Religion correlations
+
 ```{r, cache = TRUE}
@@
 
-## A critical step: the causal model
+## Causal model
 
@@
 
-## Specifying the mechanism in relation to this dataset.
+### Specifying the mechanism in relation to this dataset.
 
@@
 
-### Church attendance
+## Church attendance
 
@@
 
-### Dag shoes two adjustment sets both of which contain denomination
+### Causal Dag: church
+
 ```{r}
@@
 
-### Confounder bias if we include our theoretically proposed causal mediators!
+#### Confounder bias if we include our theoretically proposed causal mediators!
 
@@
 
-## The statistical model
+#### The statistical model
 
@@
   dplyr::filter(
-    !is.na(wb_psychANXIETY),!is.na(age),!is.na(rel_CHURCHFREQ),!is.na(wb_psychMEANING),!is.na(male),!is.na(rel_EXTENTBELIEVEGOD),
+    !is.na(wb_psychANXIETY),
+    !is.na(age),
+    !is.na(rel_CHURCHFREQ),
+    !is.na(wb_psychMEANING),
+    !is.na(male),
+    !is.na(rel_EXTENTBELIEVEGOD),
   ) %>%
@@
   ) %>%
-  dplyr::select(anx, ch, mn, bg, anx_s, age_s, ch_s, mn_s, male2, country, male, bg_s,wb_psychANXIETY)
+  dplyr::select(anx,
+                ch,
+                mn,
+                bg,
+                anx_s,
+                age_s,
+                ch_s,
+                mn_s,
+                male2,
+                country,
+                male,
+                bg_s,
+                wb_psychANXIETY)
 
@@
   chains = 4,
-  file = "m1"
+  file = "m1_real_data"
 )
@@
 
-We can see there is no effect of church attendance on anxiety (according to this model). (Note that throughout, the plots will sample 300 draws from the posterior distribution. )
+Greater church attendance is associate with greater anxiety. There is substantial within-group variance, with a low correlation of country-level slope by intercept variances.
 
 ```{r cache = TRUE}
-p1<-brms::conditional_effects(m1, "ch_s", 
+p1 <- brms::conditional_effects(
+  m1,
+  "ch_s",
   spaghetti = T,
@@
   prob = .89,
-                              method = 'fitted')
-
-plot(p1, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1,1))  + xlab("church sd units") + ylab ("anxiety in sd units") +
+  method = 'fitted'
+)
+
+plot(p1, plot = FALSE)[[1]] + 
+  scale_y_continuous(limits = c(-1, 1))  + 
+  xlab("church sd units") + 
+  ylab("anxiety in sd units") +
   labs(title = "Predicted effect of ch_s")
@@
 ```
+
 Note: there is huge country-level variability in the MCMC solutions and uncertainty in the correlations between country level intercepts in slopes. How should we interpret these results:
@@
 
-  - There is variability in the church-level effects on anxiety by country.
-
-  - Whether the effect of church attendance differs by countries that are high or low in anxiety is uncertain, but the mass of probability is below zero.
-The greater a countries anxiety the more likely church attendance will reduce anxiety...
-however, we should not be too confident that this pattern is general.
-
-  - Finally the standard deviation of the overall expected mean level of anxiety is large, which is to be expected because the gaussian is such a poor choice for our sampling distribution.
+-   There is variability in the church-level effects on anxiety by country, but this is unreliably correlated with the slope prediction of church-attendance
 
@@
 
-```{r}
+```{r eval = FALSE}
 m1_prior <- brm(
@@
   file = "m1_prior",
-  sample_prior = "only")
+  sample_prior = "only"
+)
 ```
@@
 
-or we can simulate and see that our prior is scattered across five standard deviations of anxiety, and we can see our prior is telling us that anything can happen
-.
+or we can simulate and see that our prior is scattered across five standard deviations of anxiety, and we can see our prior is telling us that anything can happen .
 
 ```{r cache = TRUE}
-p<-brms::conditional_effects(m1_prior, "ch_s", 
+p <- brms::conditional_effects(
+  m1_prior,
+  "ch_s",
   spaghetti = T,
@@
   prob = .89,
-                              method = 'fitted')
-plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-5,5))  + xlab("church sd units") + ylab ("anxiety in sd units") + 
+  method = 'fitted'
+)
+plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-5, 5))  + xlab("church sd units") + ylab ("anxiety in sd units") +
   labs(title = "Expected values for our attempted informative prior reveals remains too flat")
@@
 
+Even this prior is diffuse.
+
 We'll return to our priors and sampling distribution at the next iteration of this analysis.
-Let's set this to the side, and focus on the country level variability, looking first at the expectation plots.
-
-## The group level effects of country
+Let's set this to the side for now, and focus on the country level variability, looking first at the expectation plots.
+
+#### The group level effects of country
 
@@
 library(parameters) 
-par<-parameters::model_parameters(m1,effects = "random" ) 
-plot(par,  sort = TRUE) +theme_lucid() #option -->
+par <- parameters::model_parameters(m1, effects = "random")
+plot(par,  sort = TRUE) + theme_lucid() #option -->
 ```
@@
 ```{r cache = TRUE}
-plot(ggeffects::ggpredict(m1, terms = c("country[all]", "ch_s[minmax]"),type = "re", ci.lvl = 0.89))  +  scale_fill_brewer(palette= "Set2")
-```
-
-Let's delving into this question of country level effects by asking whether people who are *relatively* more church attending within their countries tending to have more anxiety.
-
-- within_church centering = relative church attendance within a country.
-
-- between_church centering = country level differences in church attendance.
-
-## Within and between
-
-Here's a model that includes within and between country centered variables. See:
-https://easystats.github.io/parameters/reference/demean.html
-
-The model.
-wb_psychANXIETY
-```{r}
-m1a <- brm(
-  anx_s ~ age_s  + male + ch_within +  ch_between + (1 + ch_within|country),  
-  data = d1,
-  family = "gaussian",
-  seed = 1234,
-  warmup = 1000,
-  iter = 3000,
-  chains = 4,
-  file = "m1_a.rds")
-
-parameters::model_parameters(m1a,  
-                             test = "p_direction")
-parameters::random_parameters(m1a)
-
-```
-
-We find no relation of "ch_within" to anxiety.
-
-```{r cache = TRUE}
-
-p<-brms::conditional_effects(m1a, "ch_within", 
-                              spaghetti = T,
-                              nsamples = 3000,
-                              prob = .89,
-                              method = 'fitted')
-
-plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1,1))  + xlab("church_within") + ylab ("anxiety in sd units") + 
-  labs(title = "Predicted effects of ch_within")
-
-```
-
-However, a strong between-country effect emerges:
-
-```{r cache = TRUE}
-p<-brms::conditional_effects(m1a, "ch_between", 
-                              spaghetti = T,
-                              nsamples = 3000,
-                              prob = .89,
-                              method = 'fitted')
-
-plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1,1))  + xlab("church_between") + ylab ("anxiety in sd units")+ 
-  labs( title = "Predicted effects of ch_between")
-
-```
-
-Before considering this, let's run some model checks workflow as before.
-
-Trace plots for the model look OK
-
-```{r}
- mcmc_plot(m1a,
-          type = "trace")
-```
-
-No surprise, there is the same problem with the sampling distribution.
-
-```{r}
-# we see the ordinal qualities of our dataset here
-pp_check(m1a)
-```
-
-Next the interval plots:
-
-```{r}
-mcmc_plot(m1a,
-          type = "intervals",
-          prob_outer = 0.89)
-```
-
-## TRY ORDINAL MODEL (note church attend should be monotonic -- at next iteration)
+plot(ggeffects::ggpredict(
+  m1,
+  terms = c("country[all]", "ch_s[minmax]"),
+  type = "re",
+  ci.lvl = 0.89
+))  +  scale_fill_brewer(palette = "Set2") +
+  theme_classic()
+```
+
+<!-- Let's delving into this question of country level effects by asking whether people who are *relatively* more church attending within their countries tending to have more anxiety.  -->
+
+<!-- - within_church centering = relative church attendance within a country. -->
+
+<!-- - between_church centering = country level differences in church attendance. -->
+
+<!-- ## Within and between -->
+
+<!-- Here's a model that includes within and between country centered variables. See: -->
+
+<!-- https://easystats.github.io/parameters/reference/demean.html -->
+
+<!-- The model. -->
+
+<!-- wb_psychANXIETY -->
+
+<!-- ```{r} -->
+
+<!-- m1a <- brm( -->
+
+<!--   anx_s ~ age_s  + male + ch_within +  ch_between + (1 + ch_within | country), -->
+
+<!--   data = d1, -->
+
+<!--   family = "gaussian", -->
+
+<!--   seed = 1234, -->
+
+<!--   warmup = 1000, -->
+
+<!--   iter = 3000, -->
+
+<!--   chains = 4, -->
+
+<!--   file = "m1_a.rds_real" -->
+
+<!-- ) -->
+
+<!-- parameters::model_parameters(m1a, -->
+
+<!--                              test = "p_direction") -->
+
+<!-- parameters::random_parameters(m1a) -->
+
+<!-- ``` -->
+
+<!-- We find a positive relation of "ch_within" to anxiety.  -->
+
+<!-- ```{r cache = TRUE} -->
+
+<!-- p <- brms::conditional_effects( -->
+
+<!--   m1a, -->
+
+<!--   "ch_within", -->
+
+<!--   spaghetti = T, -->
+
+<!--   nsamples = 3000, -->
+
+<!--   prob = .89, -->
+
+<!--   method = 'fitted' -->
+
+<!-- ) -->
+
+<!-- plot(p, plot = FALSE)[[1]] +  -->
+
+<!--   scale_y_continuous(limits = c(-1, 1)) +  -->
+
+<!--   xlab("church_within") +  -->
+
+<!--   ylab ("anxiety in sd units") + -->
+
+<!--   labs(title = "Predicted effects of ch_within") -->
+
+<!-- ``` -->
+
+<!-- There is, are somewhat reliable between-country difference, however the effect becomes uncertain at the high end of between-country differences: -->
+
+<!-- ```{r cache = TRUE} -->
+
+<!-- p <- brms::conditional_effects( -->
+
+<!--   m1a, -->
+
+<!--   "ch_between", -->
+
+<!--   spaghetti = T, -->
+
+<!--   nsamples = 3000, -->
+
+<!--   prob = .89, -->
+
+<!--   method = 'fitted' -->
+
+<!-- ) -->
+
+<!-- plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1, 1))  + xlab("church_between") + ylab ("anxiety in sd units") + -->
+
+<!--   labs(title = "Predicted effects of ch_between") -->
+
+<!-- ``` -->
+
+<!-- Before considering this, let's run some model checks on our workflow.  -->
+
+<!-- Trace plots for the model look OK -->
+
+<!-- ```{r} -->
+
+<!--  mcmc_plot(m1a, -->
+
+<!--           type = "trace") -->
+
+<!-- ``` -->
+
+<!-- No surprise, there is the same problem with the sampling distribution. -->
+
+<!-- ```{r} -->
+
+<!-- # we see the ordinal qualities of our dataset here -->
+
+<!-- pp_check(m1a) -->
+
+<!-- ``` -->
+
+<!-- Next the interval plots: -->
+
+<!-- ```{r} -->
+
+<!-- mcmc_plot(m1a, -->
+
+<!--           type = "intervals", -->
+
+<!--           prob_outer = 0.89) -->
+
+<!-- ``` -->
+
+### Ordinal regression: church
 
@@
   chains = 4,
-  file = "m1o"
+  file = "m1o_real"
 )
 parameters::model_parameters(m1o, test = "p_direction")
-parameters::random_parameters(m1o)
+parameters::random_parameters(m1o, component = "conditional")
+summary(m1o)
+```
+
+Focussing on country level variances, we find the slope of the country intercept not reliably correlated with the church attendance slope.
+
+```{r}
+summary(m1o)
 ```
@@
 This resolves the problem
+
 ```{r}
@@
 Model summary
+
 ```{r}
@@
 
-Plot model
+Plot model, here we see that the effects are driven at the high-end of the anxiety scale.
+This evidence is consistent with religion as an attractor for people who have high anxiety (see Bulbulia and Sibley 2012, Faith after the Quake)
+
 ```{r, cache = TRUE}
@@
 plot(p0b, plot = FALSE)[[1]]   + xlab("ch_s") + ylab ("anxiety 1-5") +
-  labs(title = "Predicted effects of ch_s")# + scale_y_continuous(limits= c(1,5))# +  scale_color_viridis(discrete=TRUE)
+  labs(title = "Predicted effects of ch_s")
+# + scale_y_continuous(limits= c(1,5))# +  scale_color_viridis(discrete=TRUE)
 
@@
 <!-- ```{r, cache = TRUE} -->
+
 <!-- library(viridis) -->
+
 <!-- p0b1<-brms::conditional_effects(m1o, "ch_s",  -->
+
 <!--                               spaghetti = T, -->
+
 <!--                               nsamples = 200, -->
+
 <!--                              categorical = F, -->
+
 <!--                               prob = .89, -->
+
 <!--                               method = 'fitted') -->
@@
 <!-- plot(p0b1, plot = FALSE)[[1]]   + xlab("ch_s") + ylab ("anxiety 1-5") + -->
+
 <!--   labs(title = "Predicted effects of ch_s")+ scale_y_continuous(limits= c(1,5))# +  scale_color_viridis(discrete=TRUE)  -->
@@
 
-### without demeaning
-
-```{r}
-m1_o <- brm(
-  wb_psychANXIETY ~ age_s  + male + ch_within +  ch_between + (1 + ch_within|country),  
-  data = d1,
-  family = cumulative(link ="logit"),
-  seed = 1234,
-  warmup = 1000,
-  iter = 3000,
-  chains = 4,
-  file = "m1a_o")
-```
-
-### Within /between demeaning
-
-```{r}
-m1a_o <- brm(
-  wb_psychANXIETY ~ age_s  + male + ch_within +  ch_between + (1 + ch_within|country),  
-  data = d1,
-  family = cumulative(link ="logit"),
-  seed = 1234,
-  warmup = 1000,
-  iter = 3000,
-  chains = 4,
-  file = "m1a_o")
-```
-
-This resolves the problem
-```{r}
-# YES
-pp_check(m1a_o)
-```
-
-Trace plots:
-
-```{r, cache = TRUE}
-mcmc_plot(m1a_o,
-          type = "trace")
-```
-
-Model summary
-```{r}
-parameters::model_parameters(m1a_o)
-```
-
-Within: NOTHING
-```{r, cache = TRUE}
-library(viridis)
-pch_w<-brms::conditional_effects(m1a_o, "ch_within", 
-                              spaghetti = T,
-                              nsamples = 200,
-                              categorical = TRUE,
-                              prob = .89,
-                              method = 'fitted')
-
-ppch_w<-plot(pch_w, plot = FALSE)[[1]]    + ylab ("anxiety in sd units") +
-  labs(title = "Predicted effects of Church_within") # +  scale_color_viridis(discrete=TRUE) 
-ppch_w
-
-```
-
-BETWEEN: Yes, and at the highest levels of anxiety.
-
-```{r cache = TRUE}
-
-pch_b<-brms::conditional_effects(m1a_o, "ch_between", 
-                              spaghetti = T,
-                              nsamples = 200,
-                              prob = .89,
-                              categorical = TRUE,
-                              method = 'fitted')
-
-ppch_b<- plot(p, plot = FALSE)[[1]]   + xlab("ch_between") + ylab ("anxiety: 1-5") + 
-  labs(title = "Predicted effects of Church_between")
-ppch_b
-```
+<!-- ### with demeaning -->
+
+<!-- ```{r} -->
+
+<!-- m1a_o <- brm( -->
+
+<!--   wb_psychANXIETY ~  -->
+
+<!--     age_s  +  -->
+
+<!--     male +  -->
+
+<!--     ch_within +   -->
+
+<!--     ch_between +  -->
+
+<!--     (1 + ch_within|country),   -->
+
+<!--   data = d1, -->
+
+<!--   family = cumulative(link ="logit"), -->
+
+<!--   seed = 1234, -->
+
+<!--   warmup = 1000, -->
+
+<!--   iter = 2000, -->
+
+<!--   chains = 4, -->
+
+<!--   file = "m1a_o_real") -->
+
+<!-- ``` -->
+
+<!-- This resolves the problem -->
+
+<!-- ```{r} -->
+
+<!-- # YES -->
+
+<!-- pp_check(m1a_o) -->
+
+<!-- ``` -->
+
+<!-- Trace plots: -->
+
+<!-- ```{r, cache = TRUE} -->
+
+<!-- mcmc_plot(m1a_o, -->
+
+<!--           type = "trace") -->
+
+<!-- ``` -->
+
+<!-- Model summary -->
+
+<!-- ```{r} -->
+
+<!-- parameters::model_parameters(m1a_o) -->
+
+<!-- ``` -->
+
+<!-- We  find a correlation of within person anxiety intercept by country and within person slope of church attentnce on anxiety, but the correlation is attenuated.  -->
+
+<!-- ```{r, cache = TRUE} -->
+
+<!-- library(viridis) -->
+
+<!-- pch_w <- brms::conditional_effects( -->
+
+<!--   m1a_o, -->
+
+<!--   "ch_within", -->
+
+<!--   spaghetti = T, -->
+
+<!--   nsamples = 200, -->
+
+<!--   categorical = TRUE, -->
+
+<!--   prob = .89, -->
+
+<!--   method = 'fitted' -->
+
+<!-- ) -->
+
+<!-- ppch_w <- -->
+
+<!--   plot(pch_w, plot = FALSE)[[1]]    + ylab ("anxiety in sd units") + -->
+
+<!--   labs(title = "Predicted effects of Church_within") # +  scale_color_viridis(discrete=TRUE) -->
+
+<!-- ppch_w -->
+
+<!-- ``` -->
+
+<!-- ```{r} -->
+
+<!-- par <- parameters::model_parameters(m1a_o, effects = "random") -->
+
+<!-- plot(par,  sort = TRUE) + theme_lucid() #option -->
+
+--\> <!-- ``` -->
+
+<!-- BETWEEN: Yes, and at the highest levels of anxiety.  -->
+
+<!-- ```{r cache = TRUE} -->
+
+<!-- pch_b <- brms::conditional_effects( -->
+
+<!--   m1a_o, -->
+
+<!--   "ch_between", -->
+
+<!--   spaghetti = T, -->
+
+<!--   nsamples = 200, -->
+
+<!--   prob = .89, -->
+
+<!--   categorical = TRUE, -->
+
+<!--   method = 'fitted' -->
+
+<!-- ) -->
+
+<!-- ppch_b <- -->
+
+<!--   plot(p, plot = FALSE)[[1]]   + -->
+
+<!--   xlab("ch_between") + ylab ("anxiety: 1-5") + -->
+
+<!--   labs(title = "Predicted effects of Church_between") -->
+
+<!-- ppch_b -->
+
+<!-- ``` -->
 
@@
 <!-- ```{r, cache = TRUE} -->
+
 <!-- library(viridis) -->
+
 <!-- p0a<-brms::conditional_effects(m1a_o, "ch_within",  -->
+
 <!--                               spaghetti = T, -->
+
 <!--                               nsamples = 200, -->
+
 <!--                              # categorical = TRUE, -->
+
 <!--                               prob = .89, -->
+
 <!--                               method = 'fitted') -->
@@
 <!-- plot(p0a, plot = FALSE)[[1]]   + xlab("bg_within") + ylab ("anxiety in sd units") + -->
+
 <!--   labs(title = "Predicted effects of bg_within")  + scale_y_continuous(limits = c(1,5))# +  scale_color_viridis(discrete=TRUE)  -->
+
 <!-- ``` -->
@@
 <!-- ps<-brms::conditional_effects(m1a_o, "ch_between",  -->
+
 <!--                               spaghetti = T, -->
+
 <!--                               nsamples = 200, -->
+
 <!--                               prob = .89, -->
+
 <!--                              # categorical = TRUE, -->
+
 <!--                               method = 'fitted') -->
@@
 <!-- plot(ps, plot = FALSE)[[1]]   + xlab("bg_between") + ylab ("anxiety: 1-5") +  -->
+
 <!--   labs(title = "Predicted effects of bg_between") + scale_y_continuous(limits = c(1,5)) -->
@@
 
-## Summary of Church attendence Model
+### Results: church
+
 Which we interpret as follows:
@@
 
-  - There is no relative deprivation effect for church attendance on anxiety.
-Those whose who participate more in church can be expected to have similar anxiety as those who participate less within a country.
-
-  - This within-country effect is variable at the country level (see "se_country_Intercept"). We interpret his to mean that the within country mean church effect differ between countries.
-
-  - Whether the effect of church attendance differs by countries that are high or low in anxiety is uncertain, but the mass of probability is below zero.
-
-  - Finally the standard deviation of the overall expected mean level of anxiety is large, which is to be expected because the gaussian is such a poor choice for our sampling distribution.
+-   Within a country, those whose who participate more frequently in church can be expected to have higher anxiety as those who do not participate as frequently in church.
+
+-   Between_country predictions trend in the opposite direction, though not reliably at the high end of the anxiety spectrum this indicates.
 
@@
 ```{r}
-par<-parameters::model_parameters(m1a,effects = "random" ) 
+par<-parameters::model_parameters(m1o,effects = "random" ) 
 plot(par,  sort = TRUE) +theme_lucid() #option -->
@@
 
-Next the predictive plots...
-
-```{r cache = TRUE}
-ppch_w
-```
-
-Countries who differ in their average levels of church attendance differ present differences in their expected anxiety levels. Those with more attendance are expected to present less anxiety.
-
-And here we pick up the country level trend, which is really strong.  Countries that are more churchgoing have lower anxiety than countries that are less churchgoing.
-
-```{r cache = TRUE}
-ppch_b
-```
-
-Take homes:
-
-    - No evidence for a "relative" church attendence effect.
-
-    - There is evidence for a between country effect.
-Countries high in church attendance are less anxious than countries that are low in church attendance.
-We must not confuse this question with an individual level question about the psychological effects of church attendance!
-
-    - A key point to keep in mind: none of these models track **demand** for church among anxious people.Church attendance might reduce anxiety therefore anxious people attend church.
-Without understanding the demand characteristics for church attendance we are likely to get the wrong inference.
-Consider an analogy.
-Imagine we were to model the relationship between taking antiobotics and infection.
-Without knowning why people were taking antiobiotics we might be tempted into inferring that antiobotics *cause* infection.
-For this reason, it is very important to collect longitudional data on people, and to clarify the demand qualities that might lead people to attend church more frequently.
+Countries who differ in their average levels of church attendance differ present differences in their expected anxiety levels, with an unreliable trend in the opposite direction.
 
@@
 
-```{r}
-
-```
-
 And again, we face a collider bias. Our causal model shows us that if we were to include our meaning-making indicator in our statistical we will compromise inference!
@@
 
-### Model for belief
+### Causal DAG: belief
 
@@
     male +
-    (1 + ch_s | country),
+    (1 + bg_s | country),
   data = d1,
@@
   chains = 4,
-  file = "m2"
+  file = "m2_real"
 )
 parameters::model_parameters(m2, test = "p_direction")
-parameters::random_parameters(m2)
-```
-
-We find no effect for belief in God.
-
-```{r cache = TRUE}
-
-p <- brms::conditional_effects(
-  m2,
-  "bg_s",
-  spaghetti = T,
-  nsamples = 300,
-  prob = .89,
-  method = 'fitted'
-)
-
-plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1, 1))  + xlab("bg_s") + ylab ("anxiety in sd units") +
-  labs(title = "Predicted effects of bg_s")
-
-```
-How did the model mix?  Trace plots look OK
-
-```{r}
- mcmc_plot(m2,
-          type = "trace")
-```
-
-Again, unsurprisingly we find a same problem with the sampling distribution.
-
-```{r}
-# we see the ordinal qualities of our dataset here
-pp_check(m2)
-```
-
-Note: we'd want to check the priors in this model, and use more informative priors. TBA...
-
-Next the interval plots:
-```{r}
-mcmc_plot(m2,
-          type = "intervals",
-          prob_outer = 0.89)
-```
-
-Summary:
-
-    - No individual-level effect of belief in God on anxiety (though recall our warning about demand characteristics)
-
-    - there's lots of variability in the country-level intercepts.
-
-    - there is reliable slope effect variability at the country level.
-
-    - The density of probability in the covariance of intercept and slope trends negative. Those countries with higher anxity tend to have a lower benefit from a belief in God effect.
-
-The country level intercepts follow a similar patter for as for church attendance, which is no surprise because we are estimating the same anxiety parameter as before, just different covariates.
-
-```{r}
-par<-parameters::model_parameters(m1,effects = "random" ) 
-plot(par,  sort = TRUE) +theme_lucid()
-```
-
-Here's the absence of any individual level effect for belief in God on anxiety.
+brms::ranef(m2)
+summary(m2)
+```
+
+We find weak effect for belief in God. No correlation of intercept and slope but India is an outlier.
 
@@
   prob = .89,
-  method = 'fitted'
-)
-
-plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1, 1))  + xlab("bg_s") + ylab ("anxiety in sd units") +
+  method = 'fitted')
+
+plot(p, plot = FALSE)[[1]] +
+  scale_y_continuous(limits = c(-1, 1))  + 
+  xlab("bg_s") + 
+  ylab ("anxiety in sd units") +
   labs(title = "Predicted effects of bg_s")
@@
 
-There country-level variability is all in anxiety, but not in the effect of belief in God.
-
-```{r cache = TRUE}
-plot(ggeffects::ggpredict(
-  m2,
-  terms = c("country[all]", "bs_s[minmax]"),
-  type = "re",
-  ci.lvl = 0.89
-))  +
-  scale_fill_brewer(palette = "Set2") + labs(title = "Predicted effects of bs_s on anxiety") +
-  scale_y_continuous(limits = c(-1, 1))
-```
-
-Next we turn to the "within country relative differences" model for belief in God on anxiety.
-
-## Within and between: believe in God
-
-```{r cache = TRUE}
-m2b <- brm(
-  anx_s ~ age_s  + male  +  bg_within + bg_between +  (1  +  bg_within |
-                                                         country),
-  data = d1,
-  family = "gaussian",
-  seed = 1234,
-  warmup = 1000,
-  iter = 300,
-  chains = 4,
-  file = "m2_aa.rds"
-)
-model_parameters(m2b, test = "p_direction")
-random_parameters(m2b)
-```
-
-## Checks
-
-Same workflow for our checks, but we'll want to use more informative priors next time.
-
-Trace plots look OK
-
-```{r}
-mcmc_plot(m2b,
+How did the model mix? Trace plots look OK
+
+```{r}
+ mcmc_plot(m2,
           type = "trace")
@@
 
-No surprise, there is the same problem with the sampling distribution.
-
-```{r}
-mcmc_plot(m2b,
+Again, unsurprisingly we find a same problem with the sampling distribution.
+
+```{r}
+# we see the ordinal qualities of our dataset here
+pp_check(m2)
+```
+
+Note: we'd want to check the priors in this model, and use more informative priors. TBA...
+
+Next the interval plots:
+
+```{r}
+mcmc_plot(m2,
           type = "intervals",
@@
 
-```{r}
-# we see the ordinal qualities of our dataset here
-pp_check(m2b)
-```
-
-Next the interval plots:
-
-```{r}
- mcmc_plot(m2b,
-          type = "intervals",
-          prob_outer = 0.89)
-```
-
 Summary:
 
-```{r cache = TRUE}
-
-p <- brms::conditional_effects(
-  m2b,
-  "bg_within",
-  spaghetti = T,
-  nsamples = 300,
-  prob = .89,
-  method = 'fitted'
-)
-
-plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1, 1))  + xlab("bg_within") + ylab ("anxiety in sd units") +
-  labs(title = "Predicted effects of bg_within")
-
-```
-
-And between
-
-```{r cache = TRUE}
-p<-brms::conditional_effects(m2b, "bg_between", 
-                              spaghetti = T,
-                              nsamples = 300,
-                              prob = .89,
-                              method = 'fitted')
-
-plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1, 1))  + xlab("bg_between") + ylab ("anxiety 1-5") +
-  labs(title = "Predicted effects of bg_between")
-
-```
-
-Here's the between country graphs:
-
-```{r cache = TRUE}
-plot(ggeffects::ggpredict(
-  m2b,
-  terms = c("country[all]", "bg_within[minmax]"),
-  type = "re",
-  ci.lvl = 0.89
-))  +
-  scale_fill_brewer(palette = "Set2") + labs(title = "Predicted effects of belief on anxiety")
-```
-
-## TRY ORDINAL MODEL
-
-## Without demeaning
+-   Small individual-level prediction of belief in God on reduced anxiety (though recall our warning about demand characteristics)
+
+## India is an outlier
+
+```{r}
+par<-parameters::model_parameters(m2,effects = "random" ) 
+plot(par,  sort = TRUE) +theme_lucid()
+```
+
+<!-- ```{r cache = TRUE} -->
+
+<!-- p <- brms::conditional_effects( -->
+
+<!--   m2, -->
+
+<!--   "bg_s", -->
+
+<!--   spaghetti = T, -->
+
+<!--   nsamples = 300, -->
+
+<!--   prob = .89, -->
+
+<!--   method = 'fitted' -->
+
+<!-- ) -->
+
+<!-- plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1, 1))  + xlab("bg_s") + ylab ("anxiety in sd units") + -->
+
+<!--   labs(title = "Predicted effects of bg_s") -->
+
+<!-- ``` -->
+
+<!-- There country-level variability is all in anxiety, but not in the effect of belief in God. -->
+
+<!-- ```{r cache = TRUE} -->
+
+<!-- plot(ggeffects::ggpredict( -->
+
+<!--   m2, -->
+
+<!--   terms = c("country[all]", "bs_s[minmax]"), -->
+
+<!--   type = "re", -->
+
+<!--   ci.lvl = 0.89 -->
+
+<!-- ))  + -->
+
+<!--   scale_fill_brewer(palette = "Set2") + labs(title = "Predicted effects of bs_s on anxiety") + -->
+
+<!--   scale_y_continuous(limits = c(-1, 1)) -->
+
+<!-- ``` -->
+
+<!-- Next we turn to the "within country relative differences" model for belief in God on anxiety. -->
+
+### Statistical model: belief (Ordinal)
+
 ```{r}
@@
   chains = 4,
-  file = "m2o"
+  file = "m2o_real.rds"
 )
@@
   prob = .89,
-  method = 'fitted'
+  method = 'posterior_epred'
 )
 
-plot(p0, plot = FALSE)[[1]]   + xlab("bg") + ylab ("anxiety 1-5") +
+plot(p0, 
+  #  point_args = list(width = .3) ,
+  #  points = TRUE,
+     plot = FALSE)[[1]]   + xlab("bg") + ylab ("anxiety 1-5") +
   labs(title = "Predicted effects of bg") # +  scale_color_viridis(discrete=TRUE) 
@@
 
-###  With demeaning
-
-```{r}
-m2_o <- brm(
-  wb_psychANXIETY ~ age_s  + male + bg_within +  bg_between + (1 + ch_within|country),  
-  data = d1,
-  family = cumulative(link ="logit"),
-  seed = 1234,
-  warmup = 1000,
-  iter = 3000,
-  chains = 4,
-  file = "m2_o")
-```
-
-This resolves the problem
-```{r}
-# YES
-pp_check(m2_o)
-```
-
-Trace plots:
-
-```{r, cache = TRUE}
-mcmc_plot(m2_o,
-          type = "trace")
-```
-
-Model summary
-```{r}
-parameters::model_parameters(m2_o)
-```
-
-Within: NOTHING
-```{r, cache = TRUE}
-library(viridis)
-p0bg <- brms::conditional_effects(
-  m2_o,
-  "bg_within",
-  spaghetti = T,
-  nsamples = 200,
-  categorical = TRUE,
-  prob = .89,
-  method = 'fitted'
-)
-
-ppbg_w<- plot(p0bg, plot = FALSE)[[1]]   + xlab("bg_within") + ylab ("anxiety 1-5") +
-  labs(title = "Predicted effects of BeliefGod_within") # +  scale_color_viridis(discrete=TRUE) 
-ppbg_w
-```
-
-BETWEEN: Yes, and at the highest levels of anxiety.
-
-```{r cache = TRUE}
-pbg <- brms::conditional_effects(
-  m2_o,
-  "bg_between",
-  spaghetti = T,
-  nsamples = 200,
-  prob = .89,
-  categorical = TRUE,
-  method = 'fitted'
-)
-
-ppbg_b<- plot(pbg, plot = FALSE)[[1]]   + xlab("bg_between") + ylab ("anxiety: 1-5") +
-  labs(title = "Predicted effects of BeliefGod_between")
-ppbg_b
-```
-
-## Summary graph
-
-```{r}
-library(patchwork)
-ppch_w +  ppbg_w  + ppch_b + ppbg_b + plot_annotation("What is the relationship between belief church and belief on anxiety?", subtitle = "No evidence for within-person differences, evidence for between country differences at the high end of anxiety", tag_levels = "a") +  plot_layout(guides = 'collect')
-
-#bulbulia_summary_graph_many_analysts
-```
+<!-- ###  With demeaning NOT NEEDED-->
+
+<!-- ```{r} -->
+
+<!-- m2_o <- brm( -->
+
+<!--   wb_psychANXIETY ~ age_s  + male + bg_within +  bg_between + (1 + ch_within|country),   -->
+
+<!--   data = d1, -->
+
+<!--   family = cumulative(link ="logit"), -->
+
+<!--   seed = 1234, -->
+
+<!--   warmup = 1000, -->
+
+<!--   iter = 3000, -->
+
+<!--   chains = 4, -->
+
+<!--   file = "m2_o_real.rds") -->
+
+<!-- ``` -->
+
+<!-- This resolves the problem -->
+
+<!-- ```{r} -->
+
+<!-- # YES -->
+
+<!-- pp_check(m2_o) -->
+
+<!-- ``` -->
+
+<!-- Trace plots: -->
+
+<!-- ```{r, cache = TRUE} -->
+
+<!-- mcmc_plot(m2_o, -->
+
+<!--           type = "trace") -->
+
+<!-- ``` -->
+
+<!-- Model summary -->
+
+<!-- ```{r} -->
+
+<!-- parameters::model_parameters(m2_o) -->
+
+<!-- ``` -->
+
+<!-- Within: NOTHING -->
+
+<!-- ```{r, cache = TRUE} -->
+
+<!-- library(viridis) -->
+
+<!-- p0bg <- brms::conditional_effects( -->
+
+<!--   m2_o, -->
+
+<!--   "bg_within", -->
+
+<!--   spaghetti = T, -->
+
+<!--   nsamples = 200, -->
+
+<!--   categorical = TRUE, -->
+
+<!--   prob = .89, -->
+
+<!--   method = 'fitted' -->
+
+<!-- ) -->
+
+<!-- ppbg_w<- plot(p0bg, plot = FALSE)[[1]]   + xlab("bg_within") + ylab ("anxiety 1-5") + -->
+
+<!--   labs(title = "Predicted effects of BeliefGod_within") # +  scale_color_viridis(discrete=TRUE)  -->
+
+<!-- ppbg_w -->
+
+<!-- ``` -->
+
+<!-- BETWEEN: Yes, and at the highest levels of anxiety.  -->
+
+<!-- ```{r cache = TRUE} -->
+
+<!-- pbg <- brms::conditional_effects( -->
+
+<!--   m2_o, -->
+
+<!--   "bg_between", -->
+
+<!--   spaghetti = T, -->
+
+<!--   nsamples = 200, -->
+
+<!--   prob = .89, -->
+
+<!--   categorical = TRUE, -->
+
+<!--   method = 'fitted' -->
+
+<!-- ) -->
+
+<!-- ppbg_b<- plot(pbg, plot = FALSE)[[1]]   + xlab("bg_between") + ylab ("anxiety: 1-5") + -->
+
+<!--   labs(title = "Predicted effects of BeliefGod_between") -->
+
+<!-- ppbg_b -->
+
+<!-- ``` -->
+
+<!-- ## Summary graph -->
+
+<!-- ```{r} -->
+
+<!-- library(patchwork) -->
+
+<!-- ppch_w +  ppbg_w  + ppch_b + ppbg_b + plot_annotation("What is the relationship between belief church and belief on anxiety?", subtitle = "People with greater belief in God have less anxiety accross countries.", tag_levels = "a") +  plot_layout(guides = 'collect') -->
+
+<!-- #bulbulia_summary_graph_many_analysts -->
+
+<!-- ``` -->
 
@@
 
-  3. Sample from an ordinal outcome distribution: obtain un-transformed data from Suzanne NOTE: DONE!, but still need to get church frequency as a monotonic predictor!
-
-  4. Develop discussion about model assumptions and limitations, such as
+4.  Sample from an ordinal outcome distribution: obtain un-transformed data from Suzanne NOTE: DONE!, but still need to get church frequency as a monotonic predictor!
+
+5.  Develop discussion about model assumptions and limitations, such as
 
@@
 
-## APPENDIX
+## Summary
+
+### Individual-level predictions
+
+Church attendance positively predicts anxiety. This is consistent with theories that anxiety leads to greater demand in Church attendance. However, this question awaits future research.
+
+Belief in God is not reliably associated with greater or lower anxiety.
+
+### Group level-predictions
+
+There is substantial variation in country-level intercepts for anxiety.
+And substantial cultural level variation in religious practice and belief.
+All this needs to be explained.
+Finally India is an outlier

### 2021-04-29 analysis.Rmd
allow analysis to git

@@
 )
-```
-
-Next, we inspect the data.
-A challenge is that many columns data have been transformed from their raw form into a number between 0 and 1.
-Later we'll see that these transformations render the modelling less reliable than if the data were retained.
-
-So for the next iteration, we'll need the data as they were recorded. Specially we'll focus on the variables such as
-
-> How often do you have negative feelings such as blue mood, despair, anxiety, depression
-
-We'll investigate church frequency and belief in God as predictors of this dimension of well-being.
-
-```{r, echo = FALSE, include = FALSE}
-#inspect data frame
-str(d) # need to create factors, where factors are needed.
-```
-
-```{r, Include = FALSE}
-hist(d$rel_1)  # yikes -- factors (church freq was made into a number :(  loss of signal 
-hist(d$rel_1)  # yikes -- factors prayer freq was made into a number  :(  loss of signal
-```
-
-```{r, include = FALSE}
-table(d$rel_3,useNA = "ifany") # This is absolutely not good. 
-```
-
-```{r include = FALSE}
-#  need to recode factors -- NAs should be NAs, otherwise, we should describe them as missingness, so will avoid these variables
-table(d$ethnicity,  useNA = "ifany") # need recode NA as "not identified"
-table(d$denomination, useNA = "ifany") # need recode NA as "no denomination" 
-# fix
-d$denomination[is.na(d$denomination)] <- "not_reported"
-d$ethnicity[is.na(d$ethnicity)] <- "not_reported"
-#d$wb_soc_3[is.na(d$wb_soc_3)] <- "not_reported" # Sat with sex
-```
-
-```{r, include = FALSE}
-# what are the numbers here? 
-table(d$attention_check, useNA = "ifany")
-d$wb_phys_mean
 ```
@@
 
-## Missingness
-
-Next we need to examine patterns of missing-ness, which might spoil inference.
-
-```{r cache = TRUE}
-library("naniar")
-vis_miss(df) # substantial pattern of missingness in a few variables, 
-gg_miss_upset(df)
-```
-
-There is little missingness so we will use only complete cases.
-Note that it would be better to have a method for handling the uncertainty of missingness.
-Missing data should be handled by some probabilistic imputation strategy.
-However, because there are many larger problems about measurement error, sampling, and theory, lets set these to questions to the side and work with the complete dataset
-
-```{r, include = FALSE}
-# get columns for different indicators 
-rel <- get_rel_vars(df)
-wbgen <- get_wb_gen(df)
-wbphys <- get_wb_phys(df)
-wbpsych <- get_wb_psych(df)
-wbsoc <- get_wb_soc(df)
-
-```
-
-Next some exploration Graphs. Inspecting the sample responses on religion.
-
-```{r, cache = TRUE, echo = FALSE}
-head(rel)
-r1 <- rel %>%
-  ggplot(aes(
-    x = country,
-    y = rel_CHURCHFREQ,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  scale_color_viridis(discrete = TRUE) +
-  ggtitle("Church Frequency") +
-  ylab("") +
-  theme(axis.text.x = element_text(angle = 30))
-r1
-
-r2 <- rel %>%
-  ggplot(aes(
-    x = country,
-    y = rel_PRAYERFREQ,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  scale_color_viridis(discrete = TRUE) +
-  ggtitle("Prayer Frequency") +
-  ylab("") +
-  theme(axis.text.x = element_text(angle = 30))
-r2
-
-r3 <- rel %>%
-  ggplot(aes(
-    x = country,
-    y = rel_EXTENTBELIEVEGOD,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  scale_color_viridis(discrete = TRUE) +
-  ggtitle("Extent Believe in God") +
-  ylab("") +
-  theme(axis.text.x = element_text(angle = 30))
-
-r3
-
-r4 <- rel %>%
-  ggplot(aes(
-    x = country,
-    y = rel_EXTENTBELIEVEAFTERLIFE,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  scale_color_viridis(discrete = TRUE) +
-  ggtitle("Extent Believe in Afterlife") +
-  ylab("") +
-  theme(axis.text.x = element_text(angle = 30))
-r4
-
-r5 <- rel %>%
-  ggplot(aes(
-    x = country,
-    y = rel_EXTENTSPIRITUAL,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  scale_color_viridis(discrete = TRUE) +
-  ggtitle("Extent Spiritual") +
-  ylab("") +
-  theme(axis.text.x = element_text(angle = 30))
-r5
-
-r6 <- rel %>%
-  ggplot(aes(
-    x = country,
-    y = rel_IMPORTANCERELIGIOUSLIFE,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  scale_color_viridis(discrete = TRUE) +
-  ggtitle("Importance of Religious Life") +
-  ylab("") +
-  theme(axis.text.x = element_text(angle = 30))
-r6
-
-r7 <- rel %>%
-  ggplot(aes(
-    x = country,
-    y = rel_IMPORTANCEBELIEVEGOD,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  scale_color_viridis(discrete = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  ggtitle("Importance Believe in God") +
-  ylab("") +
-  theme(axis.text.x = element_text(angle = 30))
-r7
-```
-
-```{r, cache = TRUE, include = FALSE}
-library("patchwork")
-(
-  r1 + r2  +  r3
-)/  (r4  + r5) / (r6 +  r7)   +  plot_annotation(tag_levels = 'A') + plot_layout(guides = "collect")
-```
-
-Substantial between country variation evident. India = very high
-
-Let's look at the anxiety indicator: India is much lower: .
-
-```{r}
-wb1 <- wbpsych %>%
-  ggplot(aes(
-    x = country,
-    y = wb_psychANXIETY,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  scale_color_viridis(discrete = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  ggtitle("Importance Believe in God") +
-  ylab("") + 
-  theme(axis.text.x = element_text(angle = 30))
-wb1
-
-```
-
-We're interested in meaning of life as a causal mediator of anxiety
-
-```{r}
-wb2 <- wbpsych %>%
-  ggplot(aes(
-    x = country,
-    y = wb_psychMEANING,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  scale_color_viridis(discrete = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  ggtitle("Importance Believe in God") +
-  ylab("") + 
-  theme(axis.text.x = element_text(angle = 30))
-wb2
-```
-
-India again very high, and France is very low.
-Can we really assume measurement invariance across these countries?
-Are people so really different?
-As with any model, we condition on assumptions, however at another iteration of this study, however for now we will assume measurement invariance.
-
-Next, a quick run of correlations. Note that we need the to model the country level dependencies in any correlation matrix, given strong cultural differences
-
-Religion correlations
-
-```{r, cache = TRUE}
-rel %>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  plot() + theme_blackboard()
-```
-
-General well-being correlations
-
-```{r cache = TRUE}
-wbgen %>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  plot() + theme_blackboard()
-```
-
-Physical well-being correlations
-
-```{r}
-wbphys %>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  summary()
-
-wbphys %>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  plot() + theme_blackboard()
-```
-
-Psychological well-being correlations
-
-```{r}
-wbpsych %>%
-  select(-wb_psychMEANING)%>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  summary()
-
-wbpsych %>%
-  select(-wb_psychMEANING)%>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  plot()  + theme_abyss()
-```
-
-## Causal model
-
-Suppose we believe that:
-
--   religion + ses + male + country + a different unmeasured latent variable (e.g. privilege) causes well being (citations... )
-
--   age + male + country + and Unmeasured latent variable (e.g. parent's religion, call this U1) causes Religion (citations... )
-
--   country + age + another unmeasured latent variable (e.g. privilege U2) causes ses (citations...)
-
--   our "exposure" is religion and our outcome is "well-being"
-
-### Which variables should we condition on to understand the relationship between religion and well-being?
-
-We have two sets adjustments sets to chose from. We will pick the set with the measured indicators.
-
-```{r}
-tidy_ggdag <- dagify(
-  wb  ~ rel + ses + male + country +  U2,
-  rel ~ age + male + country + U1,
-  ses ~ country + age + U2,
-  exposure =  "rel",
-  outcome =   "wb"
-) %>%
-  tidy_dagitty()
-
-# graph adjustment sets
-ggdag::ggdag_adjustment_set(tidy_ggdag, node_size = 14) +
-  theme(legend.position = "bottom") +
-  theme_dag_blank()
-```
-
-### Note: there is collider counfounding if we include all variables into our model
-
-Suppose we were to include *all* the variables in our model, except U2 and U1 which are not measured?
-
-In this case we would introduce collider bias!
-
-```{r}
-ggdag::ggdag_collider(tidy_ggdag,
-                      from = "rel",
-                      to = "wb",
-                      control_for(c("ses", "male", "country", "age")))
-
-```
-
-### Specifying the mechanism in relation to this dataset.
-
-We have yet to consider a mechanism by which religion might affect well-being, which will tell us specifically which aspects of well-being and which aspects of religion to focus on.
-Just as it might be tempting to throw all variables into a regression equation, and disastrous to do so, it might be tempting to leap into data reduction combining all variables without pausing to think what we want to know.
-
-There are many theories of how religion affects well-being we might consider (cite Highland et al).
-We will focus on two non-trivial theories -- one related to religious behavior and another related to religious belief.
-(Citations to follow ...)
-
-## Church attendance
-
-We propose that church attendance reduces psychological anxiety by increasing perceived social support (which is measured in this dataset as a well-being variable).
-
-This is a new model so we need to write our dag all over again.
-The dag will resemble the dag above, except we know that people who identify as male have less social support (typically), at least in some countries.
-
-We write the dag:
-
-### Causal Dag: church
-
-```{r}
-tidy_ggdag2 <- dagify(
-  wb  ~ support + ses + male + country + U2,
-  support ~ church + male + country + den,
-  church ~ age + male + country + den + U1,
-  ses ~ country + age + U2,
-  exposure =  "church",
-  outcome =   "wb"
-) %>%
-  tidy_dagitty()
-
-# graph adjustment sets
-ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) +
-  theme(legend.position = "bottom") +
-  theme_dag_blank()
-```
-
-The two adjustment sets are the same as in the original dag that we considered.
-
-#### Confounder bias if we include our theoretically proposed causal mediators!
-
-Before proceeding with the model, note a curiosity.
-If our interest is in the direct effects of church attendance on well-being we should *not* include social support because this introduces a collider bias.
-
-```{r}
-ggdag::ggdag_collider(tidy_ggdag2,
-                      from = "church",
-                      to = "wb",
-                      controlling_for(c("male", "support", "country", "age")))
-
-```
-
-Let this serve as a warning for us to think about our causation. We must develop our causal model before we leap into our statistical model.
-
-#### The statistical model
-
-A bayesian workflow should include prior predictive checks and the use of informative priors (McElreath 2020).
-
-For this initial analysis I am going to use the weakly informative priors that BRMS uses; my experience is that large data sets are resistant to choice of priors.
-However it is a mistake to trust this experience.
-For now, though, we must keep to time, and timing is everything.
-
-Additionally, the data have been manipulated in advance of the study to be scaled to a range between 0 and 1.
-This is a very bad idea.
-It would be much better to thinking of church attendance frequency as a monotonic effect.
-I will do that analysis at the next round, if the data are made available in their original form.
-
-Here is the model for church attendance and anxiety
-
-Wrangle data
-
 ```{r}
@@
 
+## Results summary: Anxiety ~ church
+
+Does church attendance predict anxiety?
+
+### Model
+
+```{r}
+m1o <- brm(
+  wb_psychANXIETY ~ ch_s +
+    age_s  +
+    male +
+    (1 + ch_s | country),
+  data = d1,
+  family = cumulative(link ="logit"),
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = "m1o_real"
+)
+```
+
+### Results
+
+```{r}
+parameters::model_parameters(m1o, test = "p_direction")
+parameters::random_parameters(m1o, component = "conditional")
+```
+
+Coefficient plots
+
+More church attendance predicts greater anxiety
+
+```{r}
+parF<-parameters::model_parameters(m1o ) 
+plot(parF,  sort = TRUE) +theme_lucid() #option -->
+```
+
+Variability in country intercepts, slope interactions not reliable.
+
+India, however, is clearly an outlier.
+
+```{r}
+parR<-parameters::model_parameters(m1o,effects = "random" ) 
+plot(parR,  sort = TRUE) +theme_lucid() #option -->
+```
+
+Graphing results wee see that the effects are driven entirely at the high end of the anxiety scale.
+This is consistent with a demand characteristic.
+Anxiety draws people to religion (Sibley and Bulbulia 2012)
+
+```{r, cache = TRUE}
+library(viridis)
+p0b<-brms::conditional_effects(m1o, "ch_s", 
+                              spaghetti = T,
+                              nsamples = 200,
+                              categorical = T,
+                              prob = .89,
+                              method = 'fitted')
+
+plot(p0b, plot = FALSE)[[1]]   + xlab("ch_s") + ylab ("anxiety 1-5") +
+  labs(title = "Predicted effects of ch_s")
+```
+
+### Checks:
+
+```{r}
+# YES
+pp_check(m1o)
+```
+
+Trace plots:
+
+```{r, cache = TRUE}
+mcmc_plot(m1o,
+          type = "trace")
+```
+
+## Results summary Belief ~ Anxiety
+
+### Model
+```{r}
+m2o <- brm(
+  wb_psychANXIETY ~ bg_s +
+    age_s  +
+    male +
+    (1 + bg_s | country),
+  data = d1,
+  family = cumulative(link = "logit"),
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = "m2o_real.rds"
+)
+```
+
+### Results
+
+No clear predictive effect of belief on anxiety
+
+```{r}
+parameters::model_parameters(m2o, test = "p_direction")
+parameters::random_parameters(m2o)
+```
+
+Coefficient plots
+
+```{r}
+parF<-parameters::model_parameters( m2o ) 
+plot(parF,  sort = TRUE) +theme_lucid() #option -->
+```
+
+Variability in country intercepts, slope interactions not reliable.
+
+India, however, is clearly an outlier.
+
+```{r}
+parR<-parameters::model_parameters(m2o,effects = "random" ) 
+plot(parR,  sort = TRUE) +theme_lucid() #option -->
+```
+
+```{r, cache = TRUE}
+library(viridis)
+p0 <- brms::conditional_effects(
+  m2o,
+  "bg_s",
+  spaghetti = T,
+  categorical = TRUE,
+  nsamples = 200,
+  prob = .89,
+  method = 'posterior_epred'
+)
+
+plot(p0,
+     #  point_args = list(width = .3) ,
+     #  points = TRUE,
+     plot = FALSE)[[1]]   + xlab("bg") + ylab ("anxiety 1-5") +
+  labs(title = "Predicted effects of bg") # +  scale_color_viridis(discrete=TRUE) 
+```
+
+### Checks
+
+This resolves the problem
+
+```{r}
+# YES
+pp_check(m2o)
+```
+
+Trace plots:
+
+```{r, cache = TRUE}
+mcmc_plot(m2o,
+          type = "trace")
+```
+
+### Overall summary
+
+Church attendance positively predicts anxiety. This is consistent with theories that anxiety leads to greater demand in Church attendance. However, this question awaits future research.
+
+Belief in God is not reliably associated with greater or lower anxiety.
+
+### Group level-predictions
+
+There is substantial variation in country-level intercepts for anxiety.
+And substantial cultural level variation in religious practice and belief.
+All this needs to be explained.
+Finally India is an outlier
+
+## DETAILS OF ANLAYSIS FOR REPLICATION
+
+## Workflow and analysis
+
+First, we import the data (see functions).
+
+```{r}
+# import data
+d <- as.data.frame(
+  read.csv(here::here("MARP_data.csv")
+  )
+)
+
+# get data into shape
+d1 <- df %>%
+  dplyr::filter(
+    !is.na(wb_psychANXIETY),
+    !is.na(age),
+    !is.na(rel_CHURCHFREQ),
+    !is.na(wb_psychMEANING),
+    !is.na(male),
+    !is.na(rel_EXTENTBELIEVEGOD),
+  ) %>%
+  dplyr::mutate(
+    anx = as.numeric(wb_psychANXIETY),
+    anx_s = scale(wb_psychANXIETY),
+    #put outcome in ST units (see McElreath 2021)  # not necessary
+    age_s = scale(age),
+    ch = rel_CHURCHFREQ,
+    mn  = wb_psychMEANING,
+    bg =  rel_EXTENTBELIEVEGOD,
+    ch_s = scale(rel_CHURCHFREQ),
+    mn_s = scale(wb_psychMEANING),
+    bg_s = scale(rel_EXTENTBELIEVEGOD)
+  ) %>%
+  dplyr::select(anx,
+                ch,
+                mn,
+                bg,
+                anx_s,
+                age_s,
+                ch_s,
+                mn_s,
+                male2,
+                country,
+                male,
+                bg_s,
+                wb_psychANXIETY)
+
+## degroup variables for within and between analysis
+## https://cran.r-project.org/web/packages/parameters/vignettes/demean.html
+d1 <-
+  cbind(d1,
+        demean(d1, select = c("anx", "ch", "bg"), group = "country"))
+```
+
+Next, we inspect the data.
+A challenge is that many columns data have been transformed from their raw form into a number between 0 and 1.
+Later we'll see that these transformations render the modelling less reliable than if the data were retained.
+
+So for the next iteration, we'll need the data as they were recorded. Specially we'll focus on the variables such as
+
+> How often do you have negative feelings such as blue mood, despair, anxiety, depression
+
+We'll investigate church frequency and belief in God as predictors of this dimension of well-being.
+
+```{r, echo = FALSE, include = FALSE}
+#inspect data frame
+str(d) # need to create factors, where factors are needed.
+```
+
+```{r, Include = FALSE}
+hist(d$rel_1)  # yikes -- factors (church freq was made into a number :(  loss of signal 
+hist(d$rel_1)  # yikes -- factors prayer freq was made into a number  :(  loss of signal
+```
+
+```{r, include = FALSE}
+table(d$rel_3,useNA = "ifany") # This is absolutely not good. 
+```
+
+```{r include = FALSE}
+#  need to recode factors -- NAs should be NAs, otherwise, we should describe them as missingness, so will avoid these variables
+table(d$ethnicity,  useNA = "ifany") # need recode NA as "not identified"
+table(d$denomination, useNA = "ifany") # need recode NA as "no denomination" 
+# fix
+d$denomination[is.na(d$denomination)] <- "not_reported"
+d$ethnicity[is.na(d$ethnicity)] <- "not_reported"
+#d$wb_soc_3[is.na(d$wb_soc_3)] <- "not_reported" # Sat with sex
+```
+
+```{r, include = FALSE}
+# what are the numbers here? 
+table(d$attention_check, useNA = "ifany")
+d$wb_phys_mean
+```
+
+## Missingness
+
+Next we need to examine patterns of missing-ness, which might spoil inference.
+
+```{r cache = TRUE}
+library("naniar")
+vis_miss(df) # substantial pattern of missingness in a few variables, 
+gg_miss_upset(df)
+```
+
+There is little missingness so we will use only complete cases.
+Note that it would be better to have a method for handling the uncertainty of missingness.
+Missing data should be handled by some probabilistic imputation strategy.
+However, because there are many larger problems about measurement error, sampling, and theory, lets set these to questions to the side and work with the complete dataset
+
+```{r, include = FALSE}
+# get columns for different indicators 
+rel <- get_rel_vars(df)
+wbgen <- get_wb_gen(df)
+wbphys <- get_wb_phys(df)
+wbpsych <- get_wb_psych(df)
+wbsoc <- get_wb_soc(df)
+
+```
+
+Next some exploration Graphs. Inspecting the sample responses on religion.
+
+```{r, cache = TRUE, echo = FALSE}
+head(rel)
+r1 <- rel %>%
+  ggplot(aes(
+    x = country,
+    y = rel_CHURCHFREQ,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  scale_color_viridis(discrete = TRUE) +
+  ggtitle("Church Frequency") +
+  ylab("") +
+  theme(axis.text.x = element_text(angle = 30))
+r1
+
+r2 <- rel %>%
+  ggplot(aes(
+    x = country,
+    y = rel_PRAYERFREQ,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  scale_color_viridis(discrete = TRUE) +
+  ggtitle("Prayer Frequency") +
+  ylab("") +
+  theme(axis.text.x = element_text(angle = 30))
+r2
+
+r3 <- rel %>%
+  ggplot(aes(
+    x = country,
+    y = rel_EXTENTBELIEVEGOD,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  scale_color_viridis(discrete = TRUE) +
+  ggtitle("Extent Believe in God") +
+  ylab("") +
+  theme(axis.text.x = element_text(angle = 30))
+
+r3
+
+r4 <- rel %>%
+  ggplot(aes(
+    x = country,
+    y = rel_EXTENTBELIEVEAFTERLIFE,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  scale_color_viridis(discrete = TRUE) +
+  ggtitle("Extent Believe in Afterlife") +
+  ylab("") +
+  theme(axis.text.x = element_text(angle = 30))
+r4
+
+r5 <- rel %>%
+  ggplot(aes(
+    x = country,
+    y = rel_EXTENTSPIRITUAL,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  scale_color_viridis(discrete = TRUE) +
+  ggtitle("Extent Spiritual") +
+  ylab("") +
+  theme(axis.text.x = element_text(angle = 30))
+r5
+
+r6 <- rel %>%
+  ggplot(aes(
+    x = country,
+    y = rel_IMPORTANCERELIGIOUSLIFE,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  scale_color_viridis(discrete = TRUE) +
+  ggtitle("Importance of Religious Life") +
+  ylab("") +
+  theme(axis.text.x = element_text(angle = 30))
+r6
+
+r7 <- rel %>%
+  ggplot(aes(
+    x = country,
+    y = rel_IMPORTANCEBELIEVEGOD,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  scale_color_viridis(discrete = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  ggtitle("Importance Believe in God") +
+  ylab("") +
+  theme(axis.text.x = element_text(angle = 30))
+r7
+```
+
+```{r, cache = TRUE, include = FALSE}
+library("patchwork")
+(
+  r1 + r2  +  r3
+)/  (r4  + r5) / (r6 +  r7)   +  plot_annotation(tag_levels = 'A') + plot_layout(guides = "collect")
+```
+
+Substantial between country variation evident. India = very high
+
+Let's look at the anxiety indicator: India is much lower: .
+
+```{r}
+wb1 <- wbpsych %>%
+  ggplot(aes(
+    x = country,
+    y = wb_psychANXIETY,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  scale_color_viridis(discrete = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  ggtitle("Importance Believe in God") +
+  ylab("") + 
+  theme(axis.text.x = element_text(angle = 30))
+wb1
+
+```
+
+We're interested in meaning of life as a causal mediator of anxiety
+
+```{r}
+wb2 <- wbpsych %>%
+  ggplot(aes(
+    x = country,
+    y = wb_psychMEANING,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  scale_color_viridis(discrete = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  ggtitle("Importance Believe in God") +
+  ylab("") + 
+  theme(axis.text.x = element_text(angle = 30))
+wb2
+```
+
+India again very high, and France is very low.
+Can we really assume measurement invariance across these countries?
+Are people so really different?
+As with any model, we condition on assumptions, however at another iteration of this study, however for now we will assume measurement invariance.
+
+Next, a quick run of correlations. Note that we need the to model the country level dependencies in any correlation matrix, given strong cultural differences
+
+Religion correlations
+
+```{r, cache = TRUE}
+rel %>%
+  correlation( partial = FALSE, multilevel = TRUE ) %>%
+  plot() + theme_blackboard()
+```
+
+General well-being correlations
+
+```{r cache = TRUE}
+wbgen %>%
+  correlation( partial = FALSE, multilevel = TRUE ) %>%
+  plot() + theme_blackboard()
+```
+
+Physical well-being correlations
+
+```{r}
+wbphys %>%
+  correlation( partial = FALSE, multilevel = TRUE ) %>%
+  summary()
+
+wbphys %>%
+  correlation( partial = FALSE, multilevel = TRUE ) %>%
+  plot() + theme_blackboard()
+```
+
+Psychological well-being correlations
+
+```{r}
+wbpsych %>%
+  select(-wb_psychMEANING)%>%
+  correlation( partial = FALSE, multilevel = TRUE ) %>%
+  summary()
+
+wbpsych %>%
+  select(-wb_psychMEANING)%>%
+  correlation( partial = FALSE, multilevel = TRUE ) %>%
+  plot()  + theme_abyss()
+```
+
+## Causal model
+
+Suppose we believe that:
+
+-   religion + ses + male + country + a different unmeasured latent variable (e.g. privilege) causes well being (citations... )
+
+-   age + male + country + and Unmeasured latent variable (e.g. parent's religion, call this U1) causes Religion (citations... )
+
+-   country + age + another unmeasured latent variable (e.g. privilege U2) causes ses (citations...)
+
+-   our "exposure" is religion and our outcome is "well-being"
+
+### Which variables should we condition on to understand the relationship between religion and well-being?
+
+We have two sets adjustments sets to chose from. We will pick the set with the measured indicators.
+
+```{r}
+tidy_ggdag <- dagify(
+  wb  ~ rel + ses + male + country +  U2,
+  rel ~ age + male + country + U1,
+  ses ~ country + age + U2,
+  exposure =  "rel",
+  outcome =   "wb"
+) %>%
+  tidy_dagitty()
+
+# graph adjustment sets
+ggdag::ggdag_adjustment_set(tidy_ggdag, node_size = 14) +
+  theme(legend.position = "bottom") +
+  theme_dag_blank()
+```
+
+### Note: there is collider counfounding if we include all variables into our model
+
+Suppose we were to include *all* the variables in our model, except U2 and U1 which are not measured?
+
+In this case we would introduce collider bias!
+
+```{r}
+ggdag::ggdag_collider(tidy_ggdag,
+                      from = "rel",
+                      to = "wb",
+                      control_for(c("ses", "male", "country", "age")))
+
+```
+
+### Specifying the mechanism in relation to this dataset.
+
+We have yet to consider a mechanism by which religion might affect well-being, which will tell us specifically which aspects of well-being and which aspects of religion to focus on.
+Just as it might be tempting to throw all variables into a regression equation, and disastrous to do so, it might be tempting to leap into data reduction combining all variables without pausing to think what we want to know.
+
+There are many theories of how religion affects well-being we might consider (cite Highland et al).
+We will focus on two non-trivial theories -- one related to religious behavior and another related to religious belief.
+(Citations to follow ...)
+
+## Church attendance
+
+We propose that church attendance reduces psychological anxiety by increasing perceived social support (which is measured in this dataset as a well-being variable).
+
+This is a new model so we need to write our dag all over again.
+The dag will resemble the dag above, except we know that people who identify as male have less social support (typically), at least in some countries.
+
+We write the dag:
+
+### Causal Dag: church
+
+```{r}
+tidy_ggdag2 <- dagify(
+  wb  ~ support + ses + male + country + U2,
+  support ~ church + male + country + den,
+  church ~ age + male + country + den + U1,
+  ses ~ country + age + U2,
+  exposure =  "church",
+  outcome =   "wb"
+) %>%
+  tidy_dagitty()
+
+# graph adjustment sets
+ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) +
+  theme(legend.position = "bottom") +
+  theme_dag_blank()
+```
+
+The two adjustment sets are the same as in the original dag that we considered.
+
+#### Confounder bias if we include our theoretically proposed causal mediators!
+
+Before proceeding with the model, note a curiosity.
+If our interest is in the direct effects of church attendance on well-being we should *not* include social support because this introduces a collider bias.
+
+```{r}
+ggdag::ggdag_collider(tidy_ggdag2,
+                      from = "church",
+                      to = "wb",
+                      controlling_for(c("male", "support", "country", "age")))
+
+```
+
+Let this serve as a warning for us to think about our causation. We must develop our causal model before we leap into our statistical model.
+
+#### The statistical model
+
+A bayesian workflow should include prior predictive checks and the use of informative priors (McElreath 2020).
+
+For this initial analysis I am going to use the weakly informative priors that BRMS uses; my experience is that large data sets are resistant to choice of priors.
+However it is a mistake to trust this experience.
+For now, though, we must keep to time, and timing is everything.
+
+Additionally, the data have been manipulated in advance of the study to be scaled to a range between 0 and 1.
+This is a very bad idea.
+It would be much better to thinking of church attendance frequency as a monotonic effect.
+I will do that analysis at the next round, if the data are made available in their original form.
+
+Here is the model for church attendance and anxiety
+
+Wrangle data
+
+```{r}
+# get data into shape
+d1 <- df %>%
+  dplyr::filter(
+    !is.na(wb_psychANXIETY),
+    !is.na(age),
+    !is.na(rel_CHURCHFREQ),
+    !is.na(wb_psychMEANING),
+    !is.na(male),
+    !is.na(rel_EXTENTBELIEVEGOD),
+  ) %>%
+  dplyr::mutate(
+    anx = as.numeric(wb_psychANXIETY),
+    anx_s = scale(wb_psychANXIETY),
+    #put outcome in ST units (see McElreath 2021)  # not necessary
+    age_s = scale(age),
+    ch = rel_CHURCHFREQ,
+    mn  = wb_psychMEANING,
+    bg =  rel_EXTENTBELIEVEGOD,
+    ch_s = scale(rel_CHURCHFREQ),
+    mn_s = scale(wb_psychMEANING),
+    bg_s = scale(rel_EXTENTBELIEVEGOD)
+  ) %>%
+  dplyr::select(anx,
+                ch,
+                mn,
+                bg,
+                anx_s,
+                age_s,
+                ch_s,
+                mn_s,
+                male2,
+                country,
+                male,
+                bg_s,
+                wb_psychANXIETY)
+
+## degroup variables for within and between analysis
+## https://cran.r-project.org/web/packages/parameters/vignettes/demean.html
+d1 <-
+  cbind(d1,
+        demean(d1, select = c("anx", "ch", "bg"), group = "country"))
+```
+
 We might estimate that effects differ within and between groups, but ignoring within and between country heterogeneity and simply pooling effects at the country level
@@
 # + scale_y_continuous(limits= c(1,5))# +  scale_color_viridis(discrete=TRUE) 
-
 ```
@@
 
---\> <!-- ``` -->
+<!-- ``` -->
 

### 2021-04-29 analysis.Rmd
Results first, details follow in long appendix

@@
     toc: true
+    code_folding: true
+    highlight: kate
 ---
@@
   collapse =TRUE,
-  echo=FALSE, 
+  echo=TRUE, 
   #results="hide", 
@@
 
-```{r}
+```{r echo}
 m1o <- brm(

### 2021-04-29 analysis.Rmd
updated

@@
 
+No country level differences trending at the high end of the anxiety scale for slope/intercept covariance, but not at the low end of anxiety this pattern is reversed, leading to a canceling out ...
+such that overall the correlation isn't reliable [cor(Intercept,ch_s)  0.24  [-0.38, 0.77].
+But worth investigating...and perhaps a more complex mixture model.
+
+```{r}
+plot(ggeffects::ggpredict(
+  m1o,
+  terms = c("country[all]", "ch_s[minmax]"),
+  type = "re",
+  ci.lvl = 0.89
+)) + scale_color_viridis_d() + 
+  theme_classic()
+```
+
 Graphing results wee see that the effects are driven entirely at the high end of the anxiety scale.
@@
 plot(parR,  sort = TRUE) +theme_lucid() #option -->
+```
+
+No country level predictions in the slope/intercept covariance:
+
+```{r}
+plot(ggeffects::ggpredict(
+  m2o,
+  terms = c("country[all]", "bg_s[minmax]"),
+  type = "re",
+  ci.lvl = 0.89
+)) +
+  scale_color_viridis_d() +
+  theme_classic()
 ```

### 2021-09-28 analysis.Rmd
Update analysis.Rmd

@@
 # clean data:use informative names !
+
+# note from Suzanne: 
+# I do think it's actually a negative relation, as the anxiety item was reverse-coded (following the documentation from the WHO quality of life scale, from which the items were taken).
+This is indicated in the data documentation, but I can totally understand this got buried under all information that was provided.
+We tried to make the dataset as clean as possible since most teams just aggregated all well-being items (and the religiosity items).
+But in retrospect, this recoding did cause some confusion for a couple of teams that used only specific items (and prefered the ordinal scale).
+This is definitely something to take into account for future projects like this one.
+
 df <- d %>%
@@
   dplyr::mutate(
-    anx = as.numeric(wb_psychANXIETY),
+    anx = as.integer(6- wb_psychANXIETY), # reverse coded! rescale 1-5
     anx_s = scale(wb_psychANXIETY),
@@
         demean(d1, select = c("anx", "ch", "bg"), group = "country"))
+
 ```
@@
 ```{r echo}
+glimpse(d1)
 m1o <- brm(
-  wb_psychANXIETY ~ ch_s +
+    anx ~ ch_s +
     age_s  +
@@
   chains = 4,
-  file = "m1o_real"
+  file = here::here("models", "m1o_real")
 )
@@
 
-More church attendance predicts greater anxiety
+More church attendance predicts less anxiety
 
@@
 
-Graphing results wee see that the effects are driven entirely at the high end of the anxiety scale.
-This is consistent with a demand characteristic.
-Anxiety draws people to religion (Sibley and Bulbulia 2012)
+Clear relationship between lower distress and greater religious service attendance
 
@@
                               spaghetti = T,
-                              nsamples = 200,
+                              ndraws = 200,
                               categorical = T,
@@
 m2o <- brm(
-  wb_psychANXIETY ~ bg_s +
+  anx ~ bg_s +
     age_s  +
@@
   chains = 4,
-  file = "m2o_real.rds"
+  file = here::here("models", "m2o_real.rds")
 )
@@
 All this needs to be explained.
-Finally India is an outlier
+Finally India is an outlier.
 
@@
 
-First, we import the data (see functions).
-
-```{r}
-# import data
-d <- as.data.frame(
-  read.csv(here::here("MARP_data.csv")
-  )
-)
-
+First, we inspect the data.
+A challenge is that many columns data have been transformed from their raw form into a number between 0 and 1.
+Later we'll see that these transformations render the modelling less reliable than if the data were retained.
+
+So for the next iteration, we'll need the data as they were recorded. Specially we'll focus on the variables such as
+
+> How often do you have negative feelings such as blue mood, despair, anxiety, depression
+
+We'll investigate church frequency and belief in God as predictors of this dimension of well-being.
+
+```{r, echo = FALSE, include = FALSE}
+#inspect data frame
+str(d) # need to create factors, where factors are needed.
+```
+
+```{r, Include = FALSE}
+hist(d$rel_1)  # yikes -- factors (church freq was made into a number :(  loss of signal 
+hist(d$rel_1)  # yikes -- factors prayer freq was made into a number  :(  loss of signal
+```
+
+```{r, include = FALSE}
+table(d$rel_3,useNA = "ifany") # This is absolutely not good. 
+```
+
+```{r include = FALSE}
+#  need to recode factors -- NAs should be NAs, otherwise, we should describe them as missingness, so will avoid these variables
+table(d$ethnicity,  useNA = "ifany") # need recode NA as "not identified"
+table(d$denomination, useNA = "ifany") # need recode NA as "no denomination" 
+# fix
+d$denomination[is.na(d$denomination)] <- "not_reported"
+d$ethnicity[is.na(d$ethnicity)] <- "not_reported"
+#d$wb_soc_3[is.na(d$wb_soc_3)] <- "not_reported" # Sat with sex
+```
+
+```{r, include = FALSE}
+# what are the numbers here? 
+table(d$attention_check, useNA = "ifany")
+d$wb_phys_mean
+```
+
+## Missingness
+
+Next we need to examine patterns of missing-ness, which might spoil inference.
+
+```{r cache = TRUE}
+library("naniar")
+vis_miss(df) # substantial pattern of missingness in a few variables, 
+gg_miss_upset(df)
+```
+
+There is little missingness so we will use only complete cases.
+Note that it would be better to have a method for handling the uncertainty of missingness.
+Missing data should be handled by some probabilistic imputation strategy.
+However, because there are many larger problems about measurement error, sampling, and theory, lets set these to questions to the side and work with the complete dataset
+
+```{r, include = FALSE}
+# get columns for different indicators 
+rel <- get_rel_vars(df)
+wbgen <- get_wb_gen(df)
+wbphys <- get_wb_phys(df)
+wbpsych <- get_wb_psych(df)
+wbsoc <- get_wb_soc(df)
+
+```
+
+Next some exploration Graphs. Inspecting the sample responses on religion.
+
+```{r, cache = TRUE, echo = FALSE}
+head(rel)
+r1 <- rel %>%
+  ggplot(aes(
+    x = country,
+    y = rel_CHURCHFREQ,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  scale_color_viridis(discrete = TRUE) +
+  ggtitle("Church Frequency") +
+  ylab("") +
+  theme(axis.text.x = element_text(angle = 30))
+r1
+
+r2 <- rel %>%
+  ggplot(aes(
+    x = country,
+    y = rel_PRAYERFREQ,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  scale_color_viridis(discrete = TRUE) +
+  ggtitle("Prayer Frequency") +
+  ylab("") +
+  theme(axis.text.x = element_text(angle = 30))
+r2
+
+r3 <- rel %>%
+  ggplot(aes(
+    x = country,
+    y = rel_EXTENTBELIEVEGOD,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  scale_color_viridis(discrete = TRUE) +
+  ggtitle("Extent Believe in God") +
+  ylab("") +
+  theme(axis.text.x = element_text(angle = 30))
+
+r3
+
+r4 <- rel %>%
+  ggplot(aes(
+    x = country,
+    y = rel_EXTENTBELIEVEAFTERLIFE,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  scale_color_viridis(discrete = TRUE) +
+  ggtitle("Extent Believe in Afterlife") +
+  ylab("") +
+  theme(axis.text.x = element_text(angle = 30))
+r4
+
+r5 <- rel %>%
+  ggplot(aes(
+    x = country,
+    y = rel_EXTENTSPIRITUAL,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  scale_color_viridis(discrete = TRUE) +
+  ggtitle("Extent Spiritual") +
+  ylab("") +
+  theme(axis.text.x = element_text(angle = 30))
+r5
+
+r6 <- rel %>%
+  ggplot(aes(
+    x = country,
+    y = rel_IMPORTANCERELIGIOUSLIFE,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  scale_color_viridis(discrete = TRUE) +
+  ggtitle("Importance of Religious Life") +
+  ylab("") +
+  theme(axis.text.x = element_text(angle = 30))
+r6
+
+r7 <- rel %>%
+  ggplot(aes(
+    x = country,
+    y = rel_IMPORTANCEBELIEVEGOD,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  scale_color_viridis(discrete = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  ggtitle("Importance Believe in God") +
+  ylab("") +
+  theme(axis.text.x = element_text(angle = 30))
+r7
+```
+
+```{r, cache = TRUE, include = FALSE}
+library("patchwork")
+(
+  r1 + r2  +  r3
+)/  (r4  + r5) / (r6 +  r7)   +  plot_annotation(tag_levels = 'A') + plot_layout(guides = "collect")
+```
+
+Substantial between country variation evident. India = very high
+
+Let's look at the anxiety indicator: India is much lower: .
+
+```{r}
+wb1 <- wbpsych %>%
+  ggplot(aes(
+    x = country,
+    y = wb_psychANXIETY,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  scale_color_viridis(discrete = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  ggtitle("Importance Believe in God") +
+  ylab("") + 
+  theme(axis.text.x = element_text(angle = 30))
+wb1
+
+```
+
+We're interested in meaning of life as a causal mediator of anxiety
+
+```{r}
+wb2 <- wbpsych %>%
+  ggplot(aes(
+    x = country,
+    y = wb_psychMEANING,
+    group = country,
+    color = country
+  )) +
+  geom_boxplot(notch = TRUE) +
+  scale_color_viridis(discrete = TRUE) +
+  geom_point(position = "jitter", alpha = .1) +
+  ggtitle("Importance Believe in God") +
+  ylab("") + 
+  theme(axis.text.x = element_text(angle = 30))
+wb2
+```
+
+India again very high, and France is very low.
+Can we really assume measurement invariance across these countries?
+Are people so really different?
+As with any model, we condition on assumptions, however at another iteration of this study, however for now we will assume measurement invariance.
+
+Next, a quick run of correlations. Note that we need the to model the country level dependencies in any correlation matrix, given strong cultural differences
+
+Religion correlations
+
+```{r, cache = TRUE}
+rel %>%
+  correlation( partial = FALSE, multilevel = TRUE ) %>%
+  plot() + theme_blackboard()
+```
+
+General well-being correlations
+
+```{r cache = TRUE}
+wbgen %>%
+  correlation( partial = FALSE, multilevel = TRUE ) %>%
+  plot() + theme_blackboard()
+```
+
+Physical well-being correlations
+
+```{r}
+wbphys %>%
+  correlation( partial = FALSE, multilevel = TRUE ) %>%
+  summary()
+
+wbphys %>%
+  correlation( partial = FALSE, multilevel = TRUE ) %>%
+  plot() + theme_blackboard()
+```
+
+Psychological well-being correlations
+
+```{r}
+wbpsych %>%
+  select(-wb_psychMEANING)%>%
+  correlation( partial = FALSE, multilevel = TRUE ) %>%
+  summary()
+
+wbpsych %>%
+  select(-wb_psychMEANING)%>%
+  correlation( partial = FALSE, multilevel = TRUE ) %>%
+  plot()  + theme_abyss()
+```
+
+## Causal model
+
+Suppose we believe that:
+
+-   religion + ses + male + country + a different unmeasured latent variable (e.g. privilege) causes well being (citations... )
+
+-   age + male + country + and Unmeasured latent variable (e.g. parent's religion, call this U1) causes Religion (citations... )
+
+-   country + age + another unmeasured latent variable (e.g. privilege U2) causes ses (citations...)
+
+-   our "exposure" is religion and our outcome is "well-being"
+
+### Which variables should we condition on to understand the relationship between religion and well-being?
+
+We have two sets adjustments sets to chose from. We will pick the set with the measured indicators.
+
+```{r}
+tidy_ggdag <- dagify(
+  wb  ~ rel + ses + male + country +  U2,
+  rel ~ age + male + country + U1,
+  ses ~ country + age + U2,
+  exposure =  "rel",
+  outcome =   "wb"
+) %>%
+  tidy_dagitty()
+
+# graph adjustment sets
+ggdag::ggdag_adjustment_set(tidy_ggdag, node_size = 14) +
+  theme(legend.position = "bottom") +
+  theme_dag_blank()
+```
+
+### Note: there is collider counfounding if we include all variables into our model
+
+Suppose we were to include *all* the variables in our model, except U2 and U1 which are not measured?
+
+In this case we would introduce collider bias!
+
+```{r}
+ggdag::ggdag_collider(tidy_ggdag,
+                      from = "rel",
+                      to = "wb",
+                      control_for(c("ses", "male", "country", "age")))
+
+```
+
+### Specifying the mechanism in relation to this dataset.
+
+We have yet to consider a mechanism by which religion might affect well-being, which will tell us specifically which aspects of well-being and which aspects of religion to focus on.
+Just as it might be tempting to throw all variables into a regression equation, and disastrous to do so, it might be tempting to leap into data reduction combining all variables without pausing to think what we want to know.
+
+There are many theories of how religion affects well-being we might consider (cite Highland et al).
+We will focus on two non-trivial theories -- one related to religious behavior and another related to religious belief.
+(Citations to follow ...)
+
+## Church attendance
+
+We propose that church attendance reduces psychological anxiety by increasing perceived social support (which is measured in this dataset as a well-being variable).
+
+This is a new model so we need to write our dag all over again.
+The dag will resemble the dag above, except we know that people who identify as male have less social support (typically), at least in some countries.
+
+We write the dag:
+
+### Causal Dag: church
+
+```{r}
+tidy_ggdag2 <- dagify(
+  wb  ~ support + ses + male + country + U2,
+  support ~ church + male + country + den,
+  church ~ age + male + country + den + U1,
+  ses ~ country + age + U2,
+  exposure =  "church",
+  outcome =   "wb"
+) %>%
+  tidy_dagitty()
+
+# graph adjustment sets
+ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) +
+  theme(legend.position = "bottom") +
+  theme_dag_blank()
+```
+
+The two adjustment sets are the same as in the original dag that we considered.
+
+#### Confounder bias if we include our theoretically proposed causal mediators!
+
+Before proceeding with the model, note a curiosity.
+If our interest is in the direct effects of church attendance on well-being we should *not* include social support because this introduces a collider bias.
+
+```{r}
+ggdag::ggdag_collider(tidy_ggdag2,
+                      from = "church",
+                      to = "wb",
+                      controlling_for(c("male", "support", "country", "age")))
+
+```
+
+Let this serve as a warning for us to think about our causation. We must develop our causal model before we leap into our statistical model.
+
+#### The statistical model
+
+A bayesian workflow should include prior predictive checks and the use of informative priors (McElreath 2020).
+
+For this initial analysis I am going to use the weakly informative priors that BRMS uses; my experience is that large data sets are resistant to choice of priors.
+However it is a mistake to trust this experience.
+For now, though, we must keep to time, and timing is everything.
+
+Additionally, the data have been manipulated in advance of the study to be scaled to a range between 0 and 1.
+This is a very bad idea.
+It would be much better to thinking of church attendance frequency as a monotonic effect.
+I will do that analysis at the next round, if the data are made available in their original form.
+
+Here is the model for church attendance and anxiety
+
+Wrangle data
+
+```{r}
 # get data into shape
@@
   dplyr::mutate(
-    anx = as.numeric(wb_psychANXIETY),
+    anx = as.numeric(6- wb_psychANXIETY), # recode
     anx_s = scale(wb_psychANXIETY),
@@
 
-Next, we inspect the data.
-A challenge is that many columns data have been transformed from their raw form into a number between 0 and 1.
-Later we'll see that these transformations render the modelling less reliable than if the data were retained.
-
-So for the next iteration, we'll need the data as they were recorded. Specially we'll focus on the variables such as
-
-> How often do you have negative feelings such as blue mood, despair, anxiety, depression
-
-We'll investigate church frequency and belief in God as predictors of this dimension of well-being.
-
-```{r, echo = FALSE, include = FALSE}
-#inspect data frame
-str(d) # need to create factors, where factors are needed.
-```
-
-```{r, Include = FALSE}
-hist(d$rel_1)  # yikes -- factors (church freq was made into a number :(  loss of signal 
-hist(d$rel_1)  # yikes -- factors prayer freq was made into a number  :(  loss of signal
-```
-
-```{r, include = FALSE}
-table(d$rel_3,useNA = "ifany") # This is absolutely not good. 
-```
-
-```{r include = FALSE}
-#  need to recode factors -- NAs should be NAs, otherwise, we should describe them as missingness, so will avoid these variables
-table(d$ethnicity,  useNA = "ifany") # need recode NA as "not identified"
-table(d$denomination, useNA = "ifany") # need recode NA as "no denomination" 
-# fix
-d$denomination[is.na(d$denomination)] <- "not_reported"
-d$ethnicity[is.na(d$ethnicity)] <- "not_reported"
-#d$wb_soc_3[is.na(d$wb_soc_3)] <- "not_reported" # Sat with sex
-```
-
-```{r, include = FALSE}
-# what are the numbers here? 
-table(d$attention_check, useNA = "ifany")
-d$wb_phys_mean
-```
-
-## Missingness
-
-Next we need to examine patterns of missing-ness, which might spoil inference.
-
-```{r cache = TRUE}
-library("naniar")
-vis_miss(df) # substantial pattern of missingness in a few variables, 
-gg_miss_upset(df)
-```
-
-There is little missingness so we will use only complete cases.
-Note that it would be better to have a method for handling the uncertainty of missingness.
-Missing data should be handled by some probabilistic imputation strategy.
-However, because there are many larger problems about measurement error, sampling, and theory, lets set these to questions to the side and work with the complete dataset
-
-```{r, include = FALSE}
-# get columns for different indicators 
-rel <- get_rel_vars(df)
-wbgen <- get_wb_gen(df)
-wbphys <- get_wb_phys(df)
-wbpsych <- get_wb_psych(df)
-wbsoc <- get_wb_soc(df)
-
-```
-
-Next some exploration Graphs. Inspecting the sample responses on religion.
-
-```{r, cache = TRUE, echo = FALSE}
-head(rel)
-r1 <- rel %>%
-  ggplot(aes(
-    x = country,
-    y = rel_CHURCHFREQ,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  scale_color_viridis(discrete = TRUE) +
-  ggtitle("Church Frequency") +
-  ylab("") +
-  theme(axis.text.x = element_text(angle = 30))
-r1
-
-r2 <- rel %>%
-  ggplot(aes(
-    x = country,
-    y = rel_PRAYERFREQ,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  scale_color_viridis(discrete = TRUE) +
-  ggtitle("Prayer Frequency") +
-  ylab("") +
-  theme(axis.text.x = element_text(angle = 30))
-r2
-
-r3 <- rel %>%
-  ggplot(aes(
-    x = country,
-    y = rel_EXTENTBELIEVEGOD,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  scale_color_viridis(discrete = TRUE) +
-  ggtitle("Extent Believe in God") +
-  ylab("") +
-  theme(axis.text.x = element_text(angle = 30))
-
-r3
-
-r4 <- rel %>%
-  ggplot(aes(
-    x = country,
-    y = rel_EXTENTBELIEVEAFTERLIFE,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  scale_color_viridis(discrete = TRUE) +
-  ggtitle("Extent Believe in Afterlife") +
-  ylab("") +
-  theme(axis.text.x = element_text(angle = 30))
-r4
-
-r5 <- rel %>%
-  ggplot(aes(
-    x = country,
-    y = rel_EXTENTSPIRITUAL,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  scale_color_viridis(discrete = TRUE) +
-  ggtitle("Extent Spiritual") +
-  ylab("") +
-  theme(axis.text.x = element_text(angle = 30))
-r5
-
-r6 <- rel %>%
-  ggplot(aes(
-    x = country,
-    y = rel_IMPORTANCERELIGIOUSLIFE,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  scale_color_viridis(discrete = TRUE) +
-  ggtitle("Importance of Religious Life") +
-  ylab("") +
-  theme(axis.text.x = element_text(angle = 30))
-r6
-
-r7 <- rel %>%
-  ggplot(aes(
-    x = country,
-    y = rel_IMPORTANCEBELIEVEGOD,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  scale_color_viridis(discrete = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  ggtitle("Importance Believe in God") +
-  ylab("") +
-  theme(axis.text.x = element_text(angle = 30))
-r7
-```
-
-```{r, cache = TRUE, include = FALSE}
-library("patchwork")
-(
-  r1 + r2  +  r3
-)/  (r4  + r5) / (r6 +  r7)   +  plot_annotation(tag_levels = 'A') + plot_layout(guides = "collect")
-```
-
-Substantial between country variation evident. India = very high
-
-Let's look at the anxiety indicator: India is much lower: .
-
-```{r}
-wb1 <- wbpsych %>%
-  ggplot(aes(
-    x = country,
-    y = wb_psychANXIETY,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  scale_color_viridis(discrete = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  ggtitle("Importance Believe in God") +
-  ylab("") + 
-  theme(axis.text.x = element_text(angle = 30))
-wb1
-
-```
-
-We're interested in meaning of life as a causal mediator of anxiety
-
-```{r}
-wb2 <- wbpsych %>%
-  ggplot(aes(
-    x = country,
-    y = wb_psychMEANING,
-    group = country,
-    color = country
-  )) +
-  geom_boxplot(notch = TRUE) +
-  scale_color_viridis(discrete = TRUE) +
-  geom_point(position = "jitter", alpha = .1) +
-  ggtitle("Importance Believe in God") +
-  ylab("") + 
-  theme(axis.text.x = element_text(angle = 30))
-wb2
-```
-
-India again very high, and France is very low.
-Can we really assume measurement invariance across these countries?
-Are people so really different?
-As with any model, we condition on assumptions, however at another iteration of this study, however for now we will assume measurement invariance.
-
-Next, a quick run of correlations. Note that we need the to model the country level dependencies in any correlation matrix, given strong cultural differences
-
-Religion correlations
-
-```{r, cache = TRUE}
-rel %>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  plot() + theme_blackboard()
-```
-
-General well-being correlations
-
-```{r cache = TRUE}
-wbgen %>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  plot() + theme_blackboard()
-```
-
-Physical well-being correlations
-
-```{r}
-wbphys %>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  summary()
-
-wbphys %>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  plot() + theme_blackboard()
-```
-
-Psychological well-being correlations
-
-```{r}
-wbpsych %>%
-  select(-wb_psychMEANING)%>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  summary()
-
-wbpsych %>%
-  select(-wb_psychMEANING)%>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  plot()  + theme_abyss()
-```
-
-## Causal model
-
-Suppose we believe that:
-
--   religion + ses + male + country + a different unmeasured latent variable (e.g. privilege) causes well being (citations... )
-
--   age + male + country + and Unmeasured latent variable (e.g. parent's religion, call this U1) causes Religion (citations... )
-
--   country + age + another unmeasured latent variable (e.g. privilege U2) causes ses (citations...)
-
--   our "exposure" is religion and our outcome is "well-being"
-
-### Which variables should we condition on to understand the relationship between religion and well-being?
-
-We have two sets adjustments sets to chose from. We will pick the set with the measured indicators.
-
-```{r}
-tidy_ggdag <- dagify(
-  wb  ~ rel + ses + male + country +  U2,
-  rel ~ age + male + country + U1,
-  ses ~ country + age + U2,
-  exposure =  "rel",
-  outcome =   "wb"
-) %>%
-  tidy_dagitty()
-
-# graph adjustment sets
-ggdag::ggdag_adjustment_set(tidy_ggdag, node_size = 14) +
-  theme(legend.position = "bottom") +
-  theme_dag_blank()
-```
-
-### Note: there is collider counfounding if we include all variables into our model
-
-Suppose we were to include *all* the variables in our model, except U2 and U1 which are not measured?
-
-In this case we would introduce collider bias!
-
-```{r}
-ggdag::ggdag_collider(tidy_ggdag,
-                      from = "rel",
-                      to = "wb",
-                      control_for(c("ses", "male", "country", "age")))
-
-```
-
-### Specifying the mechanism in relation to this dataset.
-
-We have yet to consider a mechanism by which religion might affect well-being, which will tell us specifically which aspects of well-being and which aspects of religion to focus on.
-Just as it might be tempting to throw all variables into a regression equation, and disastrous to do so, it might be tempting to leap into data reduction combining all variables without pausing to think what we want to know.
-
-There are many theories of how religion affects well-being we might consider (cite Highland et al).
-We will focus on two non-trivial theories -- one related to religious behavior and another related to religious belief.
-(Citations to follow ...)
-
-## Church attendance
-
-We propose that church attendance reduces psychological anxiety by increasing perceived social support (which is measured in this dataset as a well-being variable).
-
-This is a new model so we need to write our dag all over again.
-The dag will resemble the dag above, except we know that people who identify as male have less social support (typically), at least in some countries.
-
-We write the dag:
-
-### Causal Dag: church
-
-```{r}
-tidy_ggdag2 <- dagify(
-  wb  ~ support + ses + male + country + U2,
-  support ~ church + male + country + den,
-  church ~ age + male + country + den + U1,
-  ses ~ country + age + U2,
-  exposure =  "church",
-  outcome =   "wb"
-) %>%
-  tidy_dagitty()
-
-# graph adjustment sets
-ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) +
-  theme(legend.position = "bottom") +
-  theme_dag_blank()
-```
-
-The two adjustment sets are the same as in the original dag that we considered.
-
-#### Confounder bias if we include our theoretically proposed causal mediators!
-
-Before proceeding with the model, note a curiosity.
-If our interest is in the direct effects of church attendance on well-being we should *not* include social support because this introduces a collider bias.
-
-```{r}
-ggdag::ggdag_collider(tidy_ggdag2,
-                      from = "church",
-                      to = "wb",
-                      controlling_for(c("male", "support", "country", "age")))
-
-```
-
-Let this serve as a warning for us to think about our causation. We must develop our causal model before we leap into our statistical model.
-
-#### The statistical model
-
-A bayesian workflow should include prior predictive checks and the use of informative priors (McElreath 2020).
-
-For this initial analysis I am going to use the weakly informative priors that BRMS uses; my experience is that large data sets are resistant to choice of priors.
-However it is a mistake to trust this experience.
-For now, though, we must keep to time, and timing is everything.
-
-Additionally, the data have been manipulated in advance of the study to be scaled to a range between 0 and 1.
-This is a very bad idea.
-It would be much better to thinking of church attendance frequency as a monotonic effect.
-I will do that analysis at the next round, if the data are made available in their original form.
-
-Here is the model for church attendance and anxiety
-
-Wrangle data
-
-```{r}
-# get data into shape
-d1 <- df %>%
-  dplyr::filter(
-    !is.na(wb_psychANXIETY),
-    !is.na(age),
-    !is.na(rel_CHURCHFREQ),
-    !is.na(wb_psychMEANING),
-    !is.na(male),
-    !is.na(rel_EXTENTBELIEVEGOD),
-  ) %>%
-  dplyr::mutate(
-    anx = as.numeric(wb_psychANXIETY),
-    anx_s = scale(wb_psychANXIETY),
-    #put outcome in ST units (see McElreath 2021)  # not necessary
-    age_s = scale(age),
-    ch = rel_CHURCHFREQ,
-    mn  = wb_psychMEANING,
-    bg =  rel_EXTENTBELIEVEGOD,
-    ch_s = scale(rel_CHURCHFREQ),
-    mn_s = scale(wb_psychMEANING),
-    bg_s = scale(rel_EXTENTBELIEVEGOD)
-  ) %>%
-  dplyr::select(anx,
-                ch,
-                mn,
-                bg,
-                anx_s,
-                age_s,
-                ch_s,
-                mn_s,
-                male2,
-                country,
-                male,
-                bg_s,
-                wb_psychANXIETY)
-
-## degroup variables for within and between analysis
-## https://cran.r-project.org/web/packages/parameters/vignettes/demean.html
-d1 <-
-  cbind(d1,
-        demean(d1, select = c("anx", "ch", "bg"), group = "country"))
-```
-
 We might estimate that effects differ within and between groups, but ignoring within and between country heterogeneity and simply pooling effects at the country level

### 2021-09-28 analysis.Rmd
Update analysis.Rmd

@@
   categorical = TRUE,
-  nsamples = 200,
+  ndraws = 200,
   prob = .89,
@@
 ```{r}
+wbpsych
 wb1 <- wbpsych %>%
@@
     x = country,
-    y = wb_psychANXIETY,
+    y = 6 - wb_psychANXIETY, # reverse code
     group = country,
@@
 
-Religion correlations
-
-```{r, cache = TRUE}
-rel %>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  plot() + theme_blackboard()
-```
-
-General well-being correlations
-
-```{r cache = TRUE}
-wbgen %>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  plot() + theme_blackboard()
-```
-
-Physical well-being correlations
-
-```{r}
-wbphys %>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  summary()
-
-wbphys %>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  plot() + theme_blackboard()
-```
-
-Psychological well-being correlations
-
-```{r}
-wbpsych %>%
-  select(-wb_psychMEANING)%>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  summary()
-
-wbpsych %>%
-  select(-wb_psychMEANING)%>%
-  correlation( partial = FALSE, multilevel = TRUE ) %>%
-  plot()  + theme_abyss()
-```
+<!-- Religion correlations -->
+
+<!-- ```{r, cache = TRUE} -->
+<!-- rel %>% -->
+<!--   correlation( partial = FALSE, multilevel = TRUE ) %>% -->
+<!--   plot() + theme_blackboard() -->
+<!-- ``` -->
+
+<!-- General well-being correlations -->
+
+<!-- ```{r cache = TRUE} -->
+<!-- wbgen %>% -->
+<!--   correlation( partial = FALSE, multilevel = TRUE ) %>% -->
+<!--   plot() + theme_blackboard() -->
+<!-- ``` -->
+
+<!-- Physical well-being correlations -->
+
+<!-- ```{r} -->
+<!-- wbphys %>% -->
+<!--   correlation( partial = FALSE, multilevel = TRUE ) %>% -->
+<!--   summary() -->
+
+<!-- wbphys %>% -->
+<!--   correlation( partial = FALSE, multilevel = TRUE ) %>% -->
+<!--   plot() + theme_blackboard() -->
+<!-- ``` -->
+
+<!-- Psychological well-being correlations -->
+
+<!-- ```{r} -->
+<!-- wbpsych %>% -->
+<!--   select(-wb_psychMEANING)%>% -->
+<!--   correlation( partial = FALSE, multilevel = TRUE ) %>% -->
+<!--   summary() -->
+
+<!-- wbpsych %>% -->
+<!--   select(-wb_psychMEANING)%>% -->
+<!--   correlation( partial = FALSE, multilevel = TRUE ) %>% -->
+<!--   plot()  + theme_abyss() -->
+<!-- ``` -->
 

### 2021-10-01 analysis.Rmd
Update analysis.Rmd

@@
     !is.na(rel_CHURCHFREQ),
-    !is.na(wb_psychMEANING),
+   # !is.na(wb_psychMEANING),
     !is.na(male),
@@
 
+## Causal Model: belief in God
+
+```{r}
+library("ggdag")
+library("dagitty")
+dg1a <- dagify(
+  anx  ~ belief + church +  ses + male + country + U,
+  belief ~ age + male + country + church + U,
+  church ~ age + male + country + U,
+  ses ~ country + age + edu,
+  edu ~ country + male, 
+  exposure =  "belief",
+  outcome =   "anx",
+  latent = "U",
+  labels = c(
+    "belief" = "belief God",
+    "church" = "religious service attendance",
+    "anx" = "anxiety",
+    "edu" = "education",
+    "ses" = "SES",
+    "male" = "male",
+    "country" = "country",
+    "age" = "age",
+    "U" = "Unmeasured
+    distressing life events"
+  )
+) #%>%
+ # tidy_dagitty()
+
+d0a <-ggdag::ggdag(dg1a, text = FALSE, use_labels = "label")  + theme_dag() + ggtitle("What if we were to assume distressing events affect religion?")
+# graph adjustment sets
+
+d1a<-ggdag::ggdag_adjustment_set(dg1a, node_size = 14, text = FALSE, use_labels = "label") +
+  theme(legend.position = "bottom") +
+  theme_dag() + ggtitle("Then crossectional data could not avoid confounding")
+
+d1a
+
+dg1 <- dagify(
+  anx  ~ belief + church +  ses + male + country + U,
+  belief ~ age + male + country + church,
+  church ~ age + male + country,
+  ses ~ country + age + edu,
+  edu ~ country + male, 
+  exposure =  "belief",
+  outcome =   "anx",
+  latent = "U",
+  labels = c(
+    "belief" = "belief God",
+    "church" = "religious service attendance",
+    "anx" = "anxiety",
+    "edu" = "education",
+    "ses" = "SES",
+    "male" = "male",
+    "country" = "country",
+    "age" = "age",
+    "U" = "Unmeasured
+    distressing life events"
+  )
+) #%>%
+ # tidy_dagitty()
+
+d0 <-ggdag::ggdag(dg1, text = FALSE, use_labels = "label")  + theme_dag() + ggtitle("Assumed Causal Relations")
+# graph adjustment sets
+
+d1<-ggdag::ggdag_adjustment_set(dg1, node_size = 14, text = FALSE, use_labels = "label") +
+  theme(legend.position = "bottom") +
+  theme_dag() + ggtitle("Adjustment sets for Belief in God")
+
+d1
+```
+
+## Causal model: religious service
+
+We condition on the following model:
+
+-   religion + ses + male + country + a different unmeasured latent variable1 (e.g. privilege) affects distress.
+
+-   age + male + country + and unmeasured latent variable2 (e.g. parent's religion, call this U1) affects Religion
+
+-   country + age + an unmeasured latent variable1 causes ses (citations...)
+
+-   our "exposure" is religious service attendance or religious belief and our outcome is "anxiety"
+
+```{r}
+library("ggdag")
+dg2 <- dagify(
+  anx  ~ belief + church +  ses + male + country,
+  belief ~ age + male + country + church + U,
+  church ~ age + male + country,
+  ses ~ country + age + edu,
+  edu ~ country + male, 
+  exposure =  "church",
+  outcome =   "anx",
+  latent = "U",
+  labels = c(
+    "belief" = "belief God",
+    "church" = "religious service attendance",
+    "anx" = "anxiety",
+    "edu" = "education",
+    "ses" = "SES",
+    "male" = "male",
+    "country" = "country",
+    "age" = "age",
+    "U" = "Unmeasured
+    distressing life events"
+  )
+) %>%
+  tidy_dagitty()
+
+ggdag::ggdag(dg2) + theme_dag()
+
+ggdag::ggdag(dg2, text = FALSE, use_labels = "label")  + theme_dag()
+# graph adjustment sets
+d2<-ggdag::ggdag_adjustment_set(dg2, node_size = 14, text = FALSE, use_labels = "label") +
+  theme(legend.position = "bottom") +
+  theme_dag() + ggtitle("Adjustment set for Religious Service")
+
+d2
+```
+
+Graph for paper
+
+```{r}
+library(patchwork)
+( (d0)/(d1 + d2) +   plot_layout(guides = 'collect'))  +((d0a) + (d1a)) + plot_annotation(title = 'If distressing life events affect religious belief and behaviour, longtidinal data are needed for causal inference', tag_levels = "a")
+# 14/12
+```
+
+Graph for confounding
+
+```{r}
+library(patchwork)
+
+```
+
+## Results summary Belief ~ Anxiety
+
+###  Model
+```{r}
+m1 <- brm(
+  anx ~
+    bg_s +
+    ch_s +
+    age_s  +
+    male +
+    (1 + bg_s | country),
+  data = d1,
+  family = cumulative(link = "logit"),
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = here::here("models", "m2o_real")
+)
+
+parameters::model_parameters(m1, test = "p_direction")
+brms::ranef(m1)
+summary(m1)
+```
+
+Coefficient plots
+
+Believe in God predicts less anxiety
+
+```{r}
+parF1 <- parameters::model_parameters(m1 ) 
+plot(parF1,  sort = TRUE) +theme_lucid() #option -->
+```
+
+Variability in country intercepts, slope interactions not reliable.
+
+India, however, is clearly an outlier.
+
+```{r}
+parR1<-parameters::model_parameters(m1,effects = "random" ) 
+plot(parR1,  sort = TRUE) +theme_lucid() #option -->
+```
+
+No country level differences trending at the high end of the anxiety scale for slope/intercept covariance, but not at the low end of anxiety this pattern is reversed, leading to a canceling out ...
+such that overall the correlation isn't reliable [cor(Intercept,ch_s)  0.24  [-0.38, 0.77].
+But worth investigating...and perhaps a more complex mixture model.
+
+```{r}
+plot(ggeffects::ggpredict(
+  m1,
+  terms = c("country[all]", "bg_s[minmax]"),
+  type = "re",
+  ci.lvl = 0.89
+)) + scale_color_viridis_d() +  scale_y_continuous(limits = c(0,1)) + 
+  theme_bw()
+```
+
+Clear relationship between lower distress and greater religious service attendance
+
+```{r, cache = TRUE}
+library(viridis)
+p0<-brms::conditional_effects(m1, "bg_s", 
+                              spaghetti = T,
+                              ndraws = 200,
+                              categorical = T,
+                              prob = .89,
+                              method = 'fitted')
+
+plot(p0, plot = FALSE)[[1]]   + xlab("bg_s") + ylab ("anxiety 1-5") +
+  labs(title = "Predicted effects of bg_s")
+```
+
+### Checks:
+
+```{r}
+# YES
+pp_check(m1)
+```
+
+Trace plots:
+
+```{r, cache = TRUE}
+mcmc_plot(m1,
+          type = "trace")
+```
+
 ## Results summary: Anxiety ~ church
@@
 glimpse(d1)
-m1o <- brm(
+m2 <- brm(
     anx ~ ch_s +
@@
 
-```{r}
-parameters::model_parameters(m1o, test = "p_direction")
-parameters::random_parameters(m1o, component = "conditional")
-```
-
 Coefficient plots
@@
 ```{r}
-parF<-parameters::model_parameters(m1o ) 
+parF<-parameters::model_parameters(m2 ) 
 plot(parF,  sort = TRUE) +theme_lucid() #option -->
@@
 ```{r}
-parR<-parameters::model_parameters(m1o,effects = "random" ) 
+parR<-parameters::model_parameters(m2,effects = "random" ) 
 plot(parR,  sort = TRUE) +theme_lucid() #option -->
@@
 plot(ggeffects::ggpredict(
-  m1o,
+  m2,
   terms = c("country[all]", "ch_s[minmax]"),
@@
 library(viridis)
-p0b<-brms::conditional_effects(m1o, "ch_s", 
+p0b<-brms::conditional_effects(m2, "ch_s", 
                               spaghetti = T,
@@
 # YES
-pp_check(m1o)
+pp_check(m2)
 ```
@@
 ```{r, cache = TRUE}
-mcmc_plot(m1o,
+mcmc_plot(m2,
           type = "trace")
@@
 
-## Results summary Belief ~ Anxiety
-
-### Model
-```{r}
-m2o <- brm(
-  anx ~ bg_s +
-    age_s  +
-    male +
-    (1 + bg_s | country),
-  data = d1,
-  family = cumulative(link = "logit"),
-  seed = 1234,
-  warmup = 1000,
-  iter = 3000,
-  chains = 4,
-  file = here::here("models", "m2o_real.rds")
-)
-```
-
-### Results
-
-No clear predictive effect of belief on anxiety
-
-```{r}
-parameters::model_parameters(m2o, test = "p_direction")
-parameters::random_parameters(m2o)
-```
-
-Coefficient plots
-
-```{r}
-parF<-parameters::model_parameters( m2o ) 
-plot(parF,  sort = TRUE) +theme_lucid() #option -->
-```
-
-Variability in country intercepts, slope interactions not reliable.
-
-India, however, is clearly an outlier.
-
-```{r}
-parR<-parameters::model_parameters(m2o,effects = "random" ) 
-plot(parR,  sort = TRUE) +theme_lucid() #option -->
-```
-
-No country level predictions in the slope/intercept covariance:
-
-```{r}
-plot(ggeffects::ggpredict(
-  m2o,
-  terms = c("country[all]", "bg_s[minmax]"),
-  type = "re",
-  ci.lvl = 0.89
-)) +
-  scale_color_viridis_d() +
-  theme_classic()
-```
-
-```{r, cache = TRUE}
-library(viridis)
-p0 <- brms::conditional_effects(
-  m2o,
-  "bg_s",
-  spaghetti = T,
-  categorical = TRUE,
-  ndraws = 200,
-  prob = .89,
-  method = 'posterior_epred'
-)
-
-plot(p0,
-     #  point_args = list(width = .3) ,
-     #  points = TRUE,
-     plot = FALSE)[[1]]   + xlab("bg") + ylab ("anxiety 1-5") +
-  labs(title = "Predicted effects of bg") # +  scale_color_viridis(discrete=TRUE) 
-```
-
-### Checks
-
-This resolves the problem
-
-```{r}
-# YES
-pp_check(m2o)
-```
-
-Trace plots:
-
-```{r, cache = TRUE}
-mcmc_plot(m2o,
-          type = "trace")
-```
-
 ### Overall summary
 
-Church attendance positively predicts anxiety. This is consistent with theories that anxiety leads to greater demand in Church attendance. However, this question awaits future research.
-
-Belief in God is not reliably associated with greater or lower anxiety.
+Belief in God predicts anxiety. This is consistent with theories that anxiety leads to greater demand for belif in God. However, this question awaits future research.
+
+Church attendance positively predicts lower anxiety.
 
@@
 
-There is substantial variation in country-level intercepts for anxiety.
-And substantial cultural level variation in religious practice and belief.
-All this needs to be explained.
-Finally India is an outlier.
+There is substantial variation in country-level intercepts for anxiety. And substantial cultural level variation in religious practice and belief. All this needs to be explained. India is an outlier.
 
@@
 <!-- ``` -->
-
-## Causal model
-
-Suppose we believe that:
-
--   religion + ses + male + country + a different unmeasured latent variable (e.g. privilege) causes well being (citations... )
-
--   age + male + country + and Unmeasured latent variable (e.g. parent's religion, call this U1) causes Religion (citations... )
-
--   country + age + another unmeasured latent variable (e.g. privilege U2) causes ses (citations...)
-
--   our "exposure" is religion and our outcome is "well-being"
-
-### Which variables should we condition on to understand the relationship between religion and well-being?
-
-We have two sets adjustments sets to chose from. We will pick the set with the measured indicators.
-
-```{r}
-tidy_ggdag <- dagify(
-  wb  ~ rel + ses + male + country +  U2,
-  rel ~ age + male + country + U1,
-  ses ~ country + age + U2,
-  exposure =  "rel",
-  outcome =   "wb"
-) %>%
-  tidy_dagitty()
-
-# graph adjustment sets
-ggdag::ggdag_adjustment_set(tidy_ggdag, node_size = 14) +
-  theme(legend.position = "bottom") +
-  theme_dag_blank()
-```
-
-### Note: there is collider counfounding if we include all variables into our model
-
-Suppose we were to include *all* the variables in our model, except U2 and U1 which are not measured?
-
-In this case we would introduce collider bias!
-
-```{r}
-ggdag::ggdag_collider(tidy_ggdag,
-                      from = "rel",
-                      to = "wb",
-                      control_for(c("ses", "male", "country", "age")))
-
-```
-
-### Specifying the mechanism in relation to this dataset.
-
-We have yet to consider a mechanism by which religion might affect well-being, which will tell us specifically which aspects of well-being and which aspects of religion to focus on.
-Just as it might be tempting to throw all variables into a regression equation, and disastrous to do so, it might be tempting to leap into data reduction combining all variables without pausing to think what we want to know.
-
-There are many theories of how religion affects well-being we might consider (cite Highland et al).
-We will focus on two non-trivial theories -- one related to religious behavior and another related to religious belief.
-(Citations to follow ...)
-
-## Church attendance
-
-We propose that church attendance reduces psychological anxiety by increasing perceived social support (which is measured in this dataset as a well-being variable).
-
-This is a new model so we need to write our dag all over again.
-The dag will resemble the dag above, except we know that people who identify as male have less social support (typically), at least in some countries.
-
-We write the dag:
-
-### Causal Dag: church
-
-```{r}
-tidy_ggdag2 <- dagify(
-  wb  ~ support + ses + male + country + U2,
-  support ~ church + male + country + den,
-  church ~ age + male + country + den + U1,
-  ses ~ country + age + U2,
-  exposure =  "church",
-  outcome =   "wb"
-) %>%
-  tidy_dagitty()
-
-# graph adjustment sets
-ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) +
-  theme(legend.position = "bottom") +
-  theme_dag_blank()
-```
-
-The two adjustment sets are the same as in the original dag that we considered.
-
-#### Confounder bias if we include our theoretically proposed causal mediators!
-
-Before proceeding with the model, note a curiosity.
-If our interest is in the direct effects of church attendance on well-being we should *not* include social support because this introduces a collider bias.
-
-```{r}
-ggdag::ggdag_collider(tidy_ggdag2,
-                      from = "church",
-                      to = "wb",
-                      controlling_for(c("male", "support", "country", "age")))
-
-```
-
-Let this serve as a warning for us to think about our causation. We must develop our causal model before we leap into our statistical model.
 
@@
 
-```{r}
-tidy_ggdag2 <- dagify(
-  wb  ~ meaning + ses + male + country  + U2,
-  meaning ~ bg +  male + country,
-  bg ~ age + male + country + U1,
-  ses ~ country + age + U2,
-  exposure =  "bg",
-  outcome =   "wb"
-) %>%
-  tidy_dagitty()
-
-# graph adjustment sets
-ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) +
-  theme(legend.position = "bottom") +
-  theme_dag_blank()
-
-tidy_ggdag2 <- dagify(
-  wb  ~ meaning + ses + male + country  + U2,
-  meaning ~ bg +  male + country,
-  bg ~ age + male + country + U1,
-  ses ~ country + age + U2,
-  exposure =  "bg",
-  outcome =   "wb"
-) %>%
-  tidy_dagitty()
-
-# graph adjustment sets
-ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) +
-  theme(legend.position = "bottom") +
-  theme_dag_blank()
-
-```
-
 And again, we face a collider bias. Our causal model shows us that if we were to include our meaning-making indicator in our statistical we will compromise inference!
@@
 ```{r}
-ggdag::ggdag_collider(tidy_ggdag2,
+ggdag::ggdag_collider(dg1,
                       from = "church",
@@
                       controlling_for(c("male", "support", "country", "age")))
-```
-
-### Causal DAG: belief
-
-From causal model we write our statistical model:
-
-```{r}
-m2 <- brm(
-  anx_s ~
-    bg_s  +
-    age_s  +
-    male +
-    (1 + bg_s | country),
-  data = d1,
-  family = "gaussian",
-  seed = 1234,
-  warmup = 1000,
-  iter = 3000,
-  chains = 4,
-  file = "m2_real"
-)
-parameters::model_parameters(m2, test = "p_direction")
-brms::ranef(m2)
-summary(m2)
 ```

### 2021-10-05 analysis.Rmd
Update analysis.Rmd
updated graphs + simulation

@@
 library("gghighlight")
+library("texreg")
+library("kableExtra")
 #library("naniar") # missingness
@@
 library("texreg")
-library("viridis")
-library("ggpubr")
-library("ggsci") 
+library("viridisLite")
+library("patchwork")
 library("here")
-library("citr")
-library("rethinking")
+#library("rethinking")
 #library("statsExpressions")
-library("ggstatsplot")
+#library("ggstatsplot")
 library("ggthemes")
 library("ggdag")
-library("papaja")
+#library("papaja")
 library("rstan")
@@
 
-d0 <-ggdag::ggdag(dg1, text = FALSE, use_labels = "label")  + theme_dag() + ggtitle("Assumed Causal Relations")
+dp0 <-ggdag::ggdag(dg1, text = FALSE, use_labels = "label")  + theme_dag() + ggtitle("Assumed Causal Relations")
 # graph adjustment sets
 
-d1<-ggdag::ggdag_adjustment_set(dg1, node_size = 14, text = FALSE, use_labels = "label") +
+dp1<-ggdag::ggdag_adjustment_set(dg1, node_size = 14, text = FALSE, use_labels = "label") +
   theme(legend.position = "bottom") +
@@
 
-d1
+#dp1
 ```
@@
 # graph adjustment sets
-d2<-ggdag::ggdag_adjustment_set(dg2, node_size = 14, text = FALSE, use_labels = "label") +
+dp2<-ggdag::ggdag_adjustment_set(dg2, node_size = 14, text = FALSE, use_labels = "label") +
   theme(legend.position = "bottom") +
@@
 
-d2
+#dp2
 ```
@@
 library(patchwork)
-( (d0)/(d1 + d2) +   plot_layout(guides = 'collect'))  +((d0a) + (d1a)) + plot_annotation(title = 'If distressing life events affect religious belief and behaviour, longtidinal data are needed for causal inference', tag_levels = "a")
+( (dp0)/(dp1 + dp2) +   plot_layout(guides = 'collect'))  +((d0a) + (d1a)) + plot_annotation(title = 'If distressing life events affect religious belief and behaviour, longtidinal data are needed for causal inference', tag_levels = "a")
 # 14/12
@@
 
+```{r}
+library(lme4)
+
+```
+
 Graph for confounding
@@
 library(patchwork)
-
-```
-
-## Results summary Belief ~ Anxiety
+m0 <- lmer(
+    anx ~
+    bg_s +
+  #  ch_s +
+    age_s  +
+    male +
+    ( bg_s | country), 
+  data = d1)
+
+sjPlot::tab_model(m0)
+
+m0 <- brm(
+  anx ~
+    bg_s +
+    #ch_s +
+    age_s  +
+    male +
+    (1 + bg_s | country),
+  data = d1,
+  family = cumulative(link = "logit"),
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = here::here("models", "m2o_real")
+)
+
+```
+
+## Results summary Anxiety ~ Belief
 
@@
 parF1 <- parameters::model_parameters(m1 ) 
+parF1
 plot(parF1,  sort = TRUE) +theme_lucid() #option -->
@@
 parR1<-parameters::model_parameters(m1,effects = "random" ) 
-plot(parR1,  sort = TRUE) +theme_lucid() #option -->
+belief_random <- plot(parR1,  sort = TRUE) +theme_lucid()  + labs(title = "Country level heterogeneity in belief (co)-variances: India is an outlier.") 
 ```
@@
 
-plot(p0, plot = FALSE)[[1]]   + xlab("bg_s") + ylab ("anxiety 1-5") +
-  labs(title = "Predicted effects of bg_s")
+belief_fixed <- plot(p0, plot = FALSE)[[1]]   + xlab("belief (standardised)") + ylab ("anxiety (ordinal 1-5)") +
+  labs(title = "Increasing belief predicts increasing anxiety")
 ```
@@
 ```{r}
-parF<-parameters::model_parameters(m2 ) 
-plot(parF,  sort = TRUE) +theme_lucid() #option -->
+parc<-parameters::model_parameters(m2 ) 
+plot(parc,  sort = TRUE) + theme_lucid() #option -->
 ```
@@
 ```{r}
-parR<-parameters::model_parameters(m2,effects = "random" ) 
-plot(parR,  sort = TRUE) +theme_lucid() #option -->
+parcR<-parameters::model_parameters(m2,effects = "random" ) 
+church_random <- plot(parcR,  sort = TRUE) +theme_lucid() + labs(title = "Country level heterogeneity in religious service (co)-variances: India is again an outlier.") #option -->
 ```
@@
 
-plot(p0b, plot = FALSE)[[1]]   + xlab("ch_s") + ylab ("anxiety 1-5") +
-  labs(title = "Predicted effects of ch_s")
+church_fixed <-plot(p0b, plot = FALSE)[[1]]  + xlab("religious service attendance (standardised)") + ylab ("anxiety (ordinal 1-5)") +
+  labs(title = "Increasing religious service attendence predicts lower anxiety.")
+
+church_fixed
+```
+
+Combined plot
+
+```{r}
+(belief_fixed + church_fixed )/( belief_random + church_random  ) +
+  plot_annotation(tag_levels = "a") + plot_layout(guides = "collect")
+
+# coef_graph 15 x 15 
 ```
@@
 There is substantial variation in country-level intercepts for anxiety. And substantial cultural level variation in religious practice and belief. All this needs to be explained. India is an outlier.
+
+## Tables
+
+### Anxiety ~ Belief
+
+```{r}
+library(kableExtra)
+parameters::parameters(
+  m1,
+  test = "pd",
+  priors = FALSE, 
+  effects = "all",
+  drop = TRUE,
+  group_level = FALSE,
+  diagnostic = "Rhat") %>%
+    mutate_if(is.numeric, ~ round(., 2)) %>%
+    mutate_all(funs(str_replace(., "b_", ""))) %>%
+    kable(booktabs = T, "latex", caption =  "Belief in God is positively related to anxiety") %>%
+    print()
+```
+
+### Anxiety ~ Chruch
+```{r}
+library(kableExtra)
+parameters::parameters(
+  m2,
+  test = "pd",
+  priors = FALSE, 
+  effects = "all",
+  drop = TRUE,
+  group_level = FALSE,
+  diagnostic = "Rhat") %>%
+    mutate_if(is.numeric, ~ round(., 2)) %>%
+    mutate_all(funs(str_replace(., "b_", ""))) %>%
+    kable(booktabs = T, "latex", caption =  "Religious Service is negatively related to anxiety") %>%
+    print()
+```
+
+## Simulation
+
+```{r}
+set.seed(123)
+n <- 1000
+U <- rnorm( n ) # distressing events (unmeasured)
+C <- rnorm( n, U ) # religious service attendance, increase with distress
+B <- rnorm( n , C + U ) )# belief increase with service + distressing events
+Y <- rnorm( n , U - C ) # measured distress caused by distress, reduced by church attendance.
+
+# simulate data
+simdat <- data.frame(
+  Belief = B, # belief
+  ReligiousService = C, # religious service
+  U = U, # distressing events (unmeasured)
+  Anxiety = Y  # outcome
+  )
+
+# summarise results
+summary(
+  mod1 <- lm(Anxiety ~ Belief + ReligiousService, data = simdat)
+)
+```
+
+We can convince ourselves of this result by replicating it many times
+
+```{r}
+sim_fun = function() {
+n <- 1000
+U <- rnorm( n ) # distressing events (unmeasured)
+C <- rnorm( n, U ) # religious service attendance, increase with distress
+B <- rnorm( n , C + U ) # belief increase with service + distressing events
+Y <- rnorm( n , U - C ) # measured distress caused by distress, reduced by church attendance. 
+# simulate data
+simdat <- data.frame(Belief = B, # belief
+  ReligiousService = C, # religious service
+  U = U, # distressing events (unmeasured)
+  Anxiety = Y)
+  sim_s <- lm(Anxiety ~ Belief + ReligiousService, data = simdat)
+  sim_s
+}
+
+# Replication 1000xs
+r_lm = replicate(1000, sim_fun(), simplify = FALSE )
+
+# extract simulations
+tab_sim<-purrr::map_dfr(r_lm, broom::tidy)
+
+# extract belief coefficients
+mout<-tab_sim %>%
+  dplyr::filter(term =="Belief")%>%
+  dplyr::mutate_if(is.numeric, round, 5)
+
+# calculate proportion of estimates greater than 0
+# all positive
+sum(mout$estimate > 0) / length(mout$estimate)
+
+# proportion of belief coefficient p.values < .001
+sum(mout$p.value <.001) / length(mout$p.value)
+
+# extract religious service coefficients
+tout<-tab_sim %>%
+  dplyr::filter(term =="ReligiousService")%>%
+  dplyr::mutate_if(is.numeric, round, 5)
+
+# calculate proportion of estimates less than 0
+# all negative
+sum(tout$estimate < 0) / length(tout$estimate)
+
+# proportion of service coefficient p.values < .001
+sum(tout$estimate > 0) / length(tout$estimate)
+```
+
+```{r}
+parameters::parameters(
+  mod1,
+  test = "pd",
+  priors = FALSE, 
+  effects = "all",
+  drop = TRUE,
+  group_level = FALSE,
+  diagnostic = "Rhat") %>%
+    mutate_if(is.numeric, ~ round(., 2)) %>%
+    mutate_all(funs(str_replace(., "b_", ""))) %>%
+    kable(booktabs = T, "latex", caption =  "Simulation reveals how church attendance and belief can deviate where distress increases religious service attendance and belief in God") %>%
+    print()
+
+```
 

### 2021-10-07 analysis.Rmd
update

@@
 dg1a <- dagify(
-  anx  ~ belief + church +  ses + male + country + U,
+  anx  ~ belief + church + male + country + U,
   belief ~ age + male + country + church + U,
@@
   edu ~ country + male, 
+  U ~ age + ses + country,
   exposure =  "belief",
@@
 
-d0a <-ggdag::ggdag(dg1a, text = FALSE, use_labels = "label")  + theme_dag() + ggtitle("What if we were to assume distressing events affect religion?")
+d0a <-ggdag::ggdag(dg1a, text = FALSE, use_labels = "label")  + theme_dag() + ggtitle("What if we were to assume distressing events increase belief in God?")
 # graph adjustment sets
@@
   theme(legend.position = "bottom") +
-  theme_dag() + ggtitle("Then crossectional data could not avoid confounding")
+  theme_dag() + ggtitle("If we assume distressing events increase Religious Belief, 
+Religious Service as well as anxiety, crossectional data cannot avoid confounding ")
 
@@
 dg1 <- dagify(
-  anx  ~ belief + church +  ses + male + country + U,
-  belief ~ age + male + country + church,
-  church ~ age + male + country,
+  anx  ~ belief + church + male + country + U,
+  belief ~ age + male + country + church + U,
+  church ~ age + male + country + U,
   ses ~ country + age + edu,
   edu ~ country + male, 
+  U ~ age + ses + country,
   exposure =  "belief",
@@
   )
-) #%>%
+) #%>%#%>%
  # tidy_dagitty()       
-
+# 
 dp0 <-ggdag::ggdag(dg1, text = FALSE, use_labels = "label")  + theme_dag() + ggtitle("Assumed Causal Relations")
-# graph adjustment sets
-
-dp1<-ggdag::ggdag_adjustment_set(dg1, node_size = 14, text = FALSE, use_labels = "label") +
-  theme(legend.position = "bottom") +
-  theme_dag() + ggtitle("Adjustment sets for Belief in God")
-
+
+dp0
+# # graph adjustment sets
+# 
+#dp1<-ggdag::ggdag_adjustment_set(dg1, node_size = 14, text = FALSE, use_labels = "label") +
+#   theme(legend.position = "bottom") +
+#   theme_dag() + ggtitle("Adjustment sets for Belief in God")
 #dp1
+dp0
 ```
@@
 dg2 <- dagify(
-  anx  ~ belief + church +  ses + male + country,
+  anx  ~ belief + church + male + country + U,
   belief ~ age + male + country + church + U,
@@
   edu ~ country + male, 
+  U ~ age + ses + country,
   exposure =  "church",
@@
   theme(legend.position = "bottom") +
-  theme_dag() + ggtitle("Adjustment set for Religious Service")
-
-#dp2
+  theme_dag() + ggtitle("Adjustment set for Religious Service
+assuming service does not change from distress")
+
+dp2
+
+dpz <- dagify(
+  anx  ~ belief + church + male + country + U,
+  belief ~ age + male + country + church,
+  church ~ age + male + country,
+  ses ~ country + age + edu,
+  edu ~ country + male, 
+  U ~ age + ses + country,
+  exposure =  "belief",
+  outcome =   "anx",
+  latent = "U",
+  labels = c(
+    "belief" = "belief God",
+    "church" = "religious service attendance",
+    "anx" = "anxiety",
+    "edu" = "education",
+    "ses" = "SES",
+    "male" = "male",
+    "country" = "country",
+    "age" = "age",
+    "U" = "Unmeasured
+    distressing life events"
+  )
+) %>%
+  tidy_dagitty()
+
+ggdag::ggdag(dgz) + theme_dag()
+
+# graph adjustment sets
+dg1<-ggdag::ggdag_adjustment_set(dg1z, node_size = 14, text = FALSE, use_labels = "label") +
+  theme(legend.position = "bottom") +
+  theme_dag() + ggtitle("Adjustment set for Religious Belief
+assuming belief does not change from distress")
+
+dg1
+
 ```
@@
 library(patchwork)
-( (dp0)/(dp1 + dp2) +   plot_layout(guides = 'collect'))  +((d0a) + (d1a)) + plot_annotation(title = 'If distressing life events affect religious belief and behaviour, longtidinal data are needed for causal inference', tag_levels = "a")
-# 14/12
+ (dp0 + d1a)/   (dg1 + dp2)  +   plot_layout(guides = 'collect')  + plot_annotation(title = 'If distressing life events affect religious belief and behaviour, longtidinal data are needed for causal inference', tag_levels = "a")
+# dag_ful 15/15
+```
+
+## correlation
+
+```{r}
+correlation::correlation(d1, select = "ch_s",select2 = "bg_s")
+
+correlation::correlation(d1, select = "bg_s",select2 = "anx")
+
+correlation::correlation(d1, select = "ch_s",select2 = "anx")
+
 ```

### 2021-10-08 analysis.Rmd
Update analysis.Rmd

@@
     anx = as.integer(6- wb_psychANXIETY), # reverse coded! rescale 1-5
-    anx_s = scale(wb_psychANXIETY),
+    anx_s = scale(anx),
     #put outcome in ST units (see McElreath 2021)  # not necessary
@@
 d0a <-ggdag::ggdag(dg1a, text = FALSE, use_labels = "label")  + theme_dag() + ggtitle("What if we were to assume distressing events increase belief in God?")
+d0a
 # graph adjustment sets
@@
   theme(legend.position = "bottom") +
-  theme_dag() + ggtitle("If we assume distressing events increase Religious Belief, 
-Religious Service as well as anxiety, crossectional data cannot avoid confounding ")
+  theme_dag() + ggtitle("If we assume distressing events affect belief and service we cannot estimate causal effects ")
 
@@
 # 
-dp0 <-ggdag::ggdag(dg1, text = FALSE, use_labels = "label")  + theme_dag() + ggtitle("Assumed Causal Relations")
+dp0 <-ggdag::ggdag(dg1, text = FALSE, use_labels = "label")  + theme_dag() + ggtitle("Assumed unmeasured causal effects of distressing experiences")
 
@@
   anx  ~ belief + church + male + country + U,
-  belief ~ age + male + country + church + U,
+  belief ~ age + male + country + church,
   church ~ age + male + country,
@@
 
-ggdag::ggdag(dg2, text = FALSE, use_labels = "label")  + theme_dag()
-# graph adjustment sets
+nocause <- ggdag::ggdag(dg2, text = FALSE, use_labels = "label")  + theme_dag() + labs(title = "Assume unmeasured distressing experience
+does not affect Belief or Service")
+nocause
+
 dp2<-ggdag::ggdag_adjustment_set(dg2, node_size = 14, text = FALSE, use_labels = "label") +
@@
   )
-) %>%
-  tidy_dagitty()
-
-ggdag::ggdag(dgz) + theme_dag()
+)
+
+ggdag::ggdag(dpz) + theme_dag()
 
 # graph adjustment sets
-dg1<-ggdag::ggdag_adjustment_set(dg1z, node_size = 14, text = FALSE, use_labels = "label") +
+dg1<-ggdag::ggdag_adjustment_set(dpz, node_size = 14, text = FALSE, use_labels = "label") +
   theme(legend.position = "bottom") +
@@
 
+## Anxiety causes belief DAG
+
+```{r}
+dpL <- dagify(
+  anx  ~  church + male + country + U,
+  belief ~ age + male + country + church + anx,
+  church ~ age + male + country,
+  ses ~ country + age + edu,
+  edu ~ country + male, 
+  U ~ age + ses + country,
+  exposure =  "anx",
+  outcome =   "belief",
+  latent = "U",
+  labels = c(
+    "belief" = "belief God",
+    "church" = "religious service attendance",
+    "anx" = "anxiety",
+    "edu" = "education",
+    "ses" = "SES",
+    "male" = "male",
+    "country" = "country",
+    "age" = "age",
+    "U" = "Unmeasured
+    distressing life events"
+  )
+) %>%
+  tidy_dagitty()
+
+dl <- ggdag::ggdag(dpL, text = F, use_labels = "label") + theme_dag()
+dl1<- dl + labs(title ="Assume Anxiety causes Belief in God")
+dl1
+
+dl2<-ggdag::ggdag_adjustment_set(dpL, node_size = 14, text = FALSE, use_labels = "label") +
+  theme(legend.position = "bottom") +
+  theme_dag() + ggtitle("Adjustment set for Anxiety causing Belief in God")
+dl2
+
+```
+
+## Anxiety causes church DAG
+
+```{r}
+dpC <- dagify(
+  anx  ~   male + country + U,
+  belief ~ age + male + country + church + anx,
+  church ~ age + male + country,
+  ses ~ country + age + edu,
+  edu ~ country + male, 
+  U ~ age + ses + country,
+  exposure =  "anx",
+  outcome =   "church",
+  latent = "U",
+  labels = c(
+    "belief" = "belief God",
+    "church" = "religious service attendance",
+    "anx" = "anxiety",
+    "edu" = "education",
+    "ses" = "SES",
+    "male" = "male",
+    "country" = "country",
+    "age" = "age",
+    "U" = "Unmeasured
+    distressing life events"
+  )
+) %>%
+  tidy_dagitty()
+
+dC <- ggdag::ggdag(dpC, text = F, use_labels = "label") + theme_dag()
+dlC<- dC + labs(title ="Assume Anxiety causes Religious Service")
+dlC
+
+dl2C<-ggdag::ggdag_adjustment_set(dpC, node_size = 14, text = FALSE, use_labels = "label") +
+  theme(legend.position = "bottom") +
+  theme_dag() + ggtitle("Adjustment set for Anxiety causing Religious Service")
+dl2C
+
+```
+
+## Dag graph
+
 ```{r}
 library(patchwork)
- (dp0 + d1a)/   (dg1 + dp2)  +   plot_layout(guides = 'collect')  + plot_annotation(title = 'If distressing life events affect religious belief and behaviour, longtidinal data are needed for causal inference', tag_levels = "a")
-# dag_ful 15/15
+ (dp0 + d1a)/   (nocause + dg1 + dp2)  /(dl1 + dl2 + dl2C)+   plot_layout(guides = 'collect')  + plot_annotation(title = 'Three causal models', tag_levels = "a") 
+# dag_ful 20/20
 ```
@@
 brms::ranef(m1)
-summary(m1)
+
+parameters::parameters(
+  m1,
+  test = "pd",
+  priors = FALSE, 
+  effects = "all",
+  drop = TRUE,
+  group_level = FALSE,
+  diagnostic = "Rhat") %>%
+    mutate_if(is.numeric, ~ round(., 2)) %>%
+    mutate_all(funs(str_replace(., "b_", ""))) %>%
+    kable(booktabs = T, "latex", caption =  "Parameter estimates for model of Belief on Anxiety") %>%
+    print()
+
 ```
@@
 
-Believe in God predicts less anxiety
+Believe in God predicts more anxiety
 
@@
 parR1<-parameters::model_parameters(m1,effects = "random" ) 
-belief_random <- plot(parR1,  sort = TRUE) +theme_lucid()  + labs(title = "Country level heterogeneity in belief (co)-variances: India is an outlier.") 
+belief_random <- plot(parR1,  sort = TRUE) +theme_lucid()  + labs(title = "Country-level variability in anxiety intercepts are evident; country-level anxiety-on-belief co-variances
+are not reliable: India is an outlier.") 
+belief_random
 ```
@@
                               spaghetti = T,
-                              ndraws = 200,
+                              ndraws = 500,
                               categorical = T,
@@
 
-belief_fixed <- plot(p0, plot = FALSE)[[1]]   + xlab("belief (standardised)") + ylab ("anxiety (ordinal 1-5)") +
-  labs(title = "Increasing belief predicts increasing anxiety")
+belief_fixed <- plot(p0, plot = FALSE)[[1]]   +  xlab("belief (standardised)") + ylab ("anxiety (ordinal 1-5)") + labs(title = "Increasing belief is associated with increasing anxiety")
+belief_fixed
 ```
@@
 )
-```
-
-### Results
-
-Coefficient plots
-
-More church attendance predicts less anxiety
-
-```{r}
-parc<-parameters::model_parameters(m2 ) 
-plot(parc,  sort = TRUE) + theme_lucid() #option -->
-```
-
-Variability in country intercepts, slope interactions not reliable.
-
-India, however, is clearly an outlier.
-
-```{r}
-parcR<-parameters::model_parameters(m2,effects = "random" ) 
-church_random <- plot(parcR,  sort = TRUE) +theme_lucid() + labs(title = "Country level heterogeneity in religious service (co)-variances: India is again an outlier.") #option -->
-```
-
-No country level differences trending at the high end of the anxiety scale for slope/intercept covariance, but not at the low end of anxiety this pattern is reversed, leading to a canceling out ...
-such that overall the correlation isn't reliable [cor(Intercept,ch_s)  0.24  [-0.38, 0.77].
-But worth investigating...and perhaps a more complex mixture model.
-
-```{r}
-plot(ggeffects::ggpredict(
-  m2,
-  terms = c("country[all]", "ch_s[minmax]"),
-  type = "re",
-  ci.lvl = 0.89
-)) + scale_color_viridis_d() + 
-  theme_classic()
-```
-
-Clear relationship between lower distress and greater religious service attendance
-
-```{r, cache = TRUE}
-library(viridis)
-p0b<-brms::conditional_effects(m2, "ch_s", 
-                              spaghetti = T,
-                              ndraws = 200,
-                              categorical = T,
-                              prob = .89,
-                              method = 'fitted')
-
-church_fixed <-plot(p0b, plot = FALSE)[[1]]  + xlab("religious service attendance (standardised)") + ylab ("anxiety (ordinal 1-5)") +
-  labs(title = "Increasing religious service attendence predicts lower anxiety.")
-
-church_fixed
-```
-
-Combined plot
-
-```{r}
-(belief_fixed + church_fixed )/( belief_random + church_random  ) +
-  plot_annotation(tag_levels = "a") + plot_layout(guides = "collect")
-
-# coef_graph 15 x 15 
-```
-
-### Checks:
-
-```{r}
-# YES
-pp_check(m2)
-```
-
-Trace plots:
-
-```{r, cache = TRUE}
-mcmc_plot(m2,
-          type = "trace")
-```
-
-### Overall summary
-
-Belief in God predicts anxiety. This is consistent with theories that anxiety leads to greater demand for belif in God. However, this question awaits future research.
-
-Church attendance positively predicts lower anxiety.
-
-### Group level-predictions
-
-There is substantial variation in country-level intercepts for anxiety. And substantial cultural level variation in religious practice and belief. All this needs to be explained. India is an outlier.
-
-## Tables
-
-### Anxiety ~ Belief
-
-```{r}
-library(kableExtra)
-parameters::parameters(
-  m1,
-  test = "pd",
-  priors = FALSE, 
-  effects = "all",
-  drop = TRUE,
-  group_level = FALSE,
-  diagnostic = "Rhat") %>%
-    mutate_if(is.numeric, ~ round(., 2)) %>%
-    mutate_all(funs(str_replace(., "b_", ""))) %>%
-    kable(booktabs = T, "latex", caption =  "Belief in God is positively related to anxiety") %>%
-    print()
-```
-
-### Anxiety ~ Chruch
-```{r}
-library(kableExtra)
+
 parameters::parameters(
@@
     mutate_all(funs(str_replace(., "b_", ""))) %>%
-    kable(booktabs = T, "latex", caption =  "Religious Service is negatively related to anxiety") %>%
+    kable(booktabs = T, "latex", caption =  "Parameter estimates for model of Religious Service on Anxiety") %>%
     print()
 ```
+
+### Results
+
+Coefficient plots
+
+More church attendance predicts less anxiety
+
+```{r}
+parc<-parameters::model_parameters(m2 ) 
+plot(parc,  sort = TRUE) + theme_lucid() #option -->
+```
+
+Variability in country intercepts, slope interactions not reliable.
+
+India, however, is clearly an outlier.
+
+```{r}
+parcR<-parameters::model_parameters(m2,effects = "random" ) 
+church_random <- plot(parcR,  sort = TRUE) +theme_lucid() + labs(title = "Country-level variability in anxiety intercepts are evident; country-level anxiety-on-service co-variances
+are not reliable: India is again an outlier.")
+
+church_random
+```
+
+No country level differences trending at the high end of the anxiety scale for slope/intercept covariance, but not at the low end of anxiety this pattern is reversed, leading to a canceling out ...
+such that overall the correlation isn't reliable [cor(Intercept,ch_s)  0.24  [-0.38, 0.77].
+But worth investigating...and perhaps a more complex mixture model.
+
+```{r}
+plot(ggeffects::ggpredict(
+  m2,
+  terms = c("country[all]", "ch_s[minmax]"),
+  type = "re",
+  ci.lvl = 0.89
+)) + scale_color_viridis_d() + 
+  theme_classic()
+```
+
+Clear relationship between lower distress and greater religious service attendance
+
+```{r, cache = TRUE}
+library(viridis)
+p0b<-brms::conditional_effects(m2, "ch_s", 
+                              spaghetti = T,
+                              ndraws = 500,
+                              categorical =T,
+                              prob = .89,
+                              method = 'fitted')
+
+church_fixed <-plot(p0b, plot = FALSE)[[1]]  + xlab("religious service attendance (standardised)") + ylab ("anxiety (ordinal 1-5)") +
+  labs(title = "Increasing religious service attendence is associated with lower anxiety")
+
+church_fixed
+```
+
+## Results summary Belief ~ Anxiety
+
+```{r}
+m3<- brm(
+  bg_s ~
+    mo(anx)+
+    ch_s +
+    age_s  +
+    male +
+    (1 + mo(anx)| country),
+  data = d1,
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = here::here("models", "m3o_real")
+)
+
+parameters::model_parameters(m3, test = "p_direction")
+brms::ranef(m3)
+
+parameters::parameters(
+  m3,
+  test = "pd",
+  priors = FALSE, 
+  effects = "all",
+  drop = TRUE,
+  group_level = FALSE,
+  diagnostic = "Rhat") %>%
+    mutate_if(is.numeric, ~ round(., 2)) %>%
+    mutate_all(funs(str_replace(., "b_", ""))) %>%
+    kable(booktabs = T, "latex", caption =  "Parameter estimates for model of Anxiety on Belief") %>%
+    print()
+```
+
+<!-- ```{r} -->
+<!-- this is no better -->
+<!-- m3a<- brm( -->
+<!--   bg_s ~ -->
+<!--     anx_s+ -->
+<!--     ch_s + -->
+<!--     age_s  + -->
+<!--     male + -->
+<!--     (1 + anx_s | country), -->
+<!--   data = d1, -->
+<!--   seed = 1234, -->
+<!--   warmup = 1000, -->
+<!--   iter = 3000, -->
+<!--   chains = 4, -->
+<!--   file = here::here("models", "m3a") -->
+<!-- ) -->
+
+<!-- parameters::model_parameters(m3a, test = "p_direction") -->
+<!-- brms::ranef(m3a) -->
+<!-- summary(m3a) -->
+<!-- hist(d1$bg_s) -->
+<!-- # YES -->
+<!-- pp_check(m3a) -->
+<!-- ``` -->
+
+Coefficient plots
+
+Anxiety predicts belief
+
+```{r}
+parF3 <- parameters::model_parameters(m3 ) 
+parF3
+plot(parF3,  sort = TRUE) +theme_lucid() #option -->
+```
+
+Variability in country intercepts, slope interactions not reliable.
+
+India, however, is clearly an outlier.
+
+```{r}
+parR3<-parameters::model_parameters(m3,effects = "random" ) 
+anx_belief_random <- plot(parR3,  sort = TRUE) +theme_lucid()  + labs(title = "Country-level variabilty in belief intercepts are evident; country-level belief-on-anxiety co-variances
+are not reliable.")
+
+anx_belief_random
+```
+
+No country level differences trending at the high end of the anxiety scale for slope/intercept covariance, but not at the low end of anxiety this pattern is reversed, leading to a canceling out ...
+such that overall the correlation isn't reliable [cor(Intercept,ch_s)  0.24  [-0.38, 0.77].
+But worth investigating...and perhaps a more complex mixture model.
+
+```{r}
+plot(ggeffects::ggpredict(
+  m3,
+  terms = c("country[all]", "anx"),
+  type = "re",
+  ci.lvl = 0.89
+)) + scale_color_viridis_d() +  scale_y_continuous(limits = c(0,1)) + 
+  theme_bw()
+```
+
+```{r, cache = TRUE}
+library(viridis)
+p3<-brms::conditional_effects(m3, "anx", 
+                              spaghetti = T,
+                              ndraws = 500,
+                              categorical = F,
+                              prob = .89,
+                              method = 'fitted')
+
+belief_anxiety_random_fixed <- plot(p3, plot = FALSE)[[1]]   + xlab("Anxiety (monotonic ordinal 1-5)") + ylab ("Belief in God( standardised))") +
+  labs(title = "Increasing anxiety is associated with increasing belief in God")
+
+belief_anxiety_random_fixed
+```
+
+### Checks:
+
+```{r}
+# YES
+pp_check(m3)
+```
+
+Trace plots:
+
+```{r, cache = TRUE}
+mcmc_plot(m3,
+          type = "trace")
+```
+
+## Results summary Church ~ Anxiety
+
+```{r}
+m4<- brm(
+  ch_s ~
+    mo(anx)+
+    age_s  +
+    male +
+    (1 + mo(anx)| country),
+  data = d1,
+  seed = 1234,
+  warmup = 1000,
+  iter = 3000,
+  chains = 4,
+  file = here::here("models", "m4o_real")
+)
+
+parameters::model_parameters(m4, test = "p_direction")
+brms::ranef(m4)
+
+parameters::parameters(
+  m4,
+  test = "pd",
+  priors = FALSE, 
+  effects = "all",
+  drop = TRUE,
+  group_level = FALSE,
+  diagnostic = "Rhat") %>%
+    mutate_if(is.numeric, ~ round(., 2)) %>%
+    mutate_all(funs(str_replace(., "b_", ""))) %>%
+    kable(booktabs = T, "latex", caption =  "Parameter estimates for model of Anxiety on Religious Service") %>%
+    print()
+```
+
+<!-- ```{r} -->
+<!-- this is no better -->
+<!-- m3a<- brm( -->
+<!--   bg_s ~ -->
+<!--     anx_s+ -->
+<!--     ch_s + -->
+<!--     age_s  + -->
+<!--     male + -->
+<!--     (1 + anx_s | country), -->
+<!--   data = d1, -->
+<!--   seed = 1234, -->
+<!--   warmup = 1000, -->
+<!--   iter = 3000, -->
+<!--   chains = 4, -->
+<!--   file = here::here("models", "m3a") -->
+<!-- ) -->
+
+<!-- parameters::model_parameters(m3a, test = "p_direction") -->
+<!-- brms::ranef(m3a) -->
+<!-- summary(m3a) -->
+<!-- hist(d1$bg_s) -->
+<!-- # YES -->
+<!-- pp_check(m3a) -->
+<!-- ``` -->
+
+Coefficient plots
+
+Anxiety predicts belief
+
+```{r}
+parF4 <- parameters::model_parameters(m4 ) 
+parF4
+plot(parF4,  sort = TRUE) +theme_lucid() #option -->
+```
+
+Variability in country intercepts, slope interactions not reliable.
+
+India, however, is clearly an outlier.
+
+```{r}
+par4<-parameters::model_parameters(m4,effects = "random" ) 
+anx_church_random <- plot(par4,  sort = TRUE) + theme_lucid()  + labs(title = "Country-level variabilty in religious service intercepts are evident; country-level service-on-anxiety co-variances
+are not reliable.")
+
+anx_church_random
+```
+
+```{r}
+plot(ggeffects::ggpredict(
+  m4,
+  terms = c("country[all]", "anx"),
+  type = "re",
+  ci.lvl = 0.89
+)) + scale_color_viridis_d() +  scale_y_continuous(limits = c(0,1)) + 
+  theme_bw()
+```
+
+```{r, cache = TRUE}
+library(viridis)
+p4<-brms::conditional_effects(m4, "anx", 
+                              spaghetti = T,
+                              ndraws = 500,
+                              categorical = F,
+                              prob = .89,
+                              method = 'fitted')
+
+church_anxiety_random_fixed <- plot(p4, plot = FALSE)[[1]]   + xlab("Anxiety (monotonic ordinal 1-5)") + ylab ("Religious Service(standardised))") +
+  labs(title = "Increasing anxiety is associated with decreasing church attendance.")
+
+church_anxiety_random_fixed
+```
+
+### Checks:
+
+```{r}
+# YES
+pp_check(m4)
+```
+
+Trace plots:
+
+```{r, cache = TRUE}
+mcmc_plot(m4,
+          type = "trace")
+```
+
+## Combined plot
+
+```{r}
+library(patchwork)
+
+(belief_fixed +  belief_random)/(church_fixed+ church_random  ) / 
+(belief_anxiety_random_fixed + anx_belief_random) / (church_anxiety_random_fixed + anx_church_random) +
+  plot_annotation(tag_levels = "a")# + plot_layout(guides = "collect")
+
+# coef_graph 15 x 25
+```
+
+### Checks:
+
+```{r}
+# YES
+pp_check(m2)
+```
+
+Trace plots:
+
+```{r, cache = TRUE}
+mcmc_plot(m2,
+          type = "trace")
+```
+
+### Overall summary
+
+Belief in God predicts anxiety. This is consistent with theories that anxiety leads to greater demand for belif in God. However, this question awaits future research.
+
+Church attendance positively predicts lower anxiety.
+
+### Group level-predictions
+
+There is substantial variation in country-level intercepts for anxiety. And substantial cultural level variation in religious practice and belief. All this needs to be explained. India is an outlier.
 
@@
 n <- 1000
-U <- rnorm( n ) # distressing events (unmeasured)
-C <- rnorm( n, U ) # religious service attendance, increase with distress
-B <- rnorm( n , C + U ) )# belief increase with service + distressing events
-Y <- rnorm( n , U - C ) # measured distress caused by distress, reduced by church attendance.
+# distressing events (unmeasured)
+U <- rnorm( n ) 
+# religious service attendance increases with distress
+C <- rnorm( n, U ) 
+# anxiety caused by distress experience and
+# buffered by religious service
+Y <- rnorm( n , U - C )
+
+# belief increases with service, anxiety,
+# and experience
+B <- rnorm( n , Y + C + U )
 
@@
 )
+
+# summarise results
+summary(
+  mod1b <- lm(Anxiety ~ ReligiousService, data = simdat)
+)
+
+summary(
+  mod2 <- lm(Belief  ~ Anxiety + ReligiousService, data = simdat)
+)
+
+summary(
+  mod3 <- lm(ReligiousService  ~ Anxiety, data = simdat)
+)
+
+belief_on_anxiety_sim <- plot( ggeffects::ggpredict(mod1, terms = c("Belief")), add.data = TRUE, dot.alpha =.5 ) + labs(title = "Simulation recovers positive relationship between belief on anxiety") + scale_y_continuous(limits = c(-7,7))+ scale_x_continuous(limits = c(-10,10))
+
+service_on_anxiety_sim <- plot( ggeffects::ggpredict(mod1b, terms = c("ReligiousService")), add.data = TRUE, dot.alpha =.5 ) + labs(title = "Simulation recovers negative relationship between religious service on anxiety")+ scale_y_continuous(limits = c(-7,7))+ scale_y_continuous(limits = c(-7,7))+ scale_x_continuous(limits = c(-10,10))
+service_on_anxiety_sim
+
+service_on_anxiety_sim <- plot( ggeffects::ggpredict(mod1, terms = c("ReligiousService")), add.data = TRUE, dot.alpha =.5 ) + labs(title = "Simulation recovers negative relationship between religious service on anxiety")+ scale_y_continuous(limits = c(-7,7))+ scale_x_continuous(limits = c(-10,10))
+service_on_anxiety_sim
+
+anxiety_on_belief_sim <- plot( ggeffects::ggpredict(mod2, terms = c("Anxiety")), add.data = TRUE, dot.alpha =.5 ) + labs(title = "Simulation recovers positive relationship between anxiety on belief")+ scale_y_continuous(limits = c(-7,7))+ scale_x_continuous(limits = c(-10,10))
+anxiety_on_belief_sim
+
+anxiety_on_church_sim <- plot( ggeffects::ggpredict(mod3, terms = c("Anxiety")), add.data = TRUE, dot.alpha =.5 ) + labs(title = "Simulation recovers negative relationship between anxiety on religious service")+ scale_y_continuous(limits = c(-7,7))
+anxiety_on_church_sim
+
+library(patchwork)
+
+(belief_on_anxiety_sim +  service_on_anxiety_sim)/ (anxiety_on_belief_sim + anxiety_on_church_sim)
+
+# sim_graph 15 x 15
+
 ```
@@
 C <- rnorm( n, U ) # religious service attendance, increase with distress
-B <- rnorm( n , C + U ) # belief increase with service + distressing events
 Y <- rnorm( n , U - C ) # measured distress caused by distress, reduced by church attendance. 
+B <- rnorm( n , Y + C + U )# belief increase with service + distressing events + anxiety
+
 # simulate data
-simdat <- data.frame(Belief = B, # belief
+simdat <- data.frame(
+  Belief = B, # belief
   ReligiousService = C, # religious service
   U = U, # distressing events (unmeasured)
-  Anxiety = Y)
+  Anxiety = Y  # outcome
+  )
+
   sim_s <- lm(Anxiety ~ Belief + ReligiousService, data = simdat)
@@
 # proportion of service coefficient p.values < .001
-sum(tout$estimate > 0) / length(tout$estimate)
-```
+sum(tout$p.value <.001) / length(tout$p.value)
+```
+
+## Simulation tables
 
@@
 
+parameters::parameters(
+  mod2,
+  test = "pd",
+  priors = FALSE, 
+  effects = "all",
+  drop = TRUE,
+  group_level = FALSE,
+  diagnostic = "Rhat") %>%
+    mutate_if(is.numeric, ~ round(., 2)) %>%
+    mutate_all(funs(str_replace(., "b_", ""))) %>%
+    kable(booktabs = T, "latex", caption =  "Same simulation tracks anxiety as cause of religious belief") %>%
+    print()
+
+parameters::parameters(
+  mod3,
+  test = "pd",
+  priors = FALSE, 
+  effects = "all",
+  drop = TRUE,
+  group_level = FALSE,
+  diagnostic = "Rhat") %>%
+    mutate_if(is.numeric, ~ round(., 2)) %>%
+    mutate_all(funs(str_replace(., "b_", ""))) %>%
+    kable(booktabs = T, "latex", caption =  "Same simulation tracks anxiety as cause of decling religious service") %>%
+    print()
 ```

### 2022-01-02 analysis.Rmd
small change
Update analysis.Rmd
removed whitespace
Update analysis.Rmd

@@
   #results="hide", 
-  fig.width= 10,
-  fig.height=10,
-  tidy = "styler")
+  fig.width= 16,
+  fig.height= 9)
 ```
@@
 ```{r runanalysis, echo=FALSE, include = FALSE}
-source("libs.R") # libraries 
-source("funs.R") # custom functions
+#source("libs.R") # libraries 
+# source("funs.R") # custom functions
 
@@
 library("ggdag")
+library("geepack")  # marginal structural models
+library("ipw") # inverse probability weighting. 
 rstan_options(auto_write=TRUE)
@@
 
-## correlation
+### correlation
 
@@
 
-Graph for confounding
+### Check LMER Results
 
@@
 library(patchwork)
-m0 <- lmer(
+mlmerCH <- lmer(
+    anx ~
+    #bg_s +
+    ch_s +
+    age_s  +
+    male +
+    ( 1 + ch_s | country), 
+  data = d1)
+
+sjPlot::tab_model(mlmerCH)
+
+mlmerBG <- lmer(
     anx ~
     bg_s +
-  #  ch_s +
+    ch_s +
     age_s  +
     male +
-    ( bg_s | country), 
+    ( 1 + bg_s | country), 
   data = d1)
 
-sjPlot::tab_model(m0)
-
-m0 <- brm(
+sjPlot::tab_model(mlmerBG)
+```
+
+## General Estimating Equations
+
+### IPW Anxiety ~ Belief
+
+```{r}
+library("ipw")
+
+d1ipwB <- ipw::ipwpoint(
+  exposure = bg_s,
+  family = "gaussian",
+  numerator = ~ 1,
+  denominator = ~  ch_s + age_s  + male + country,
+  data = as.data.frame(d1)
+)
+
+summary(d1ipwB$ipw.weights)
+sum(d1ipwB$ipw.weights > 10)
+
+d1pB <- d1 %>% 
+  mutate(ipw = d1ipwB$ipw.weights) %>%
+  filter(ipw < 10)
+```
+
+Model:
+
+```{r}
+mgeeB  <- geeglm(
   anx ~
-    bg_s +
-    #ch_s +
-    age_s  +
-    male +
-    (1 + bg_s | country),
-  data = d1,
-  family = cumulative(link = "logit"),
-  seed = 1234,
-  warmup = 1000,
-  iter = 3000,
-  chains = 4,
-  file = here::here("models", "m2o_real")
-)
-
+    bg_s,# + # already weighted
+   # ch_s +
+   # age_s  +
+   # male,
+  data = d1pB,
+  id = country,
+  weights = ipw
+)
+
+parameters::model_parameters(mgeeB) %>%
+  print_html()
+
+library(ggeffects)
+p0BG <- plot(ggeffects::ggemmeans(mgeeB, terms = "bg_s [all]"), add.data = TRUE, dot.alpha = 0.1)    # 
+p0BG 
+```
+
+### IPW Anxiety ~ Church
+
+First we create Inverse Probability Weights
+
+```{r}
+library("ipw")
+
+d1ipw <- ipw::ipwpoint(
+  exposure = ch_s,
+  family = "gaussian",
+  numerator = ~ 1,
+  denominator = ~  age_s  + male + country,
+  data = as.data.frame(d1)
+)
+
+summary(d1ipw$ipw.weights)
+sum(d1ipw$ipw.weights > 10)
+
+d1p <- d1 %>% 
+  mutate(ipw = d1ipw$ipw.weights)
+```
+
+Model
+
+```{r}
+mgeeCH <- geeglm(
+  anx ~
+    ch_s,# + # already weighted
+    #age_s  +
+  #  male,
+  data = d1p,
+  id = country,
+  weights = ipw
+)
+
+parameters::model_parameters(mgeeCH) %>%
+  print_html()
+
+library(ggeffects)
+pmgeeCH <- plot(ggeffects::ggemmeans(mgeeCH, terms = "ch_s [all]"), add.data = TRUE, dot.alpha = 0.1)    # 
+pmgeeCH
+```
+
+Model with with trimmed IPW (as recommended to avoid leverage from rare cases). Note, the result here is the same.
+
+```{r, eval = FALSE}
+
+# trim IPW to less than 10 
+d1pf <- d1 %>% 
+  mutate(ipw = d1ipw$ipw.weights) %>%
+  filter(ipw < 10)
+
+m0f <- geeglm(
+  anx ~
+    ch_s,# +
+    #age_s  +
+  #  male,
+  data = d1pf,
+  id = country,
+  weights = ipw
+)
+
+parameters::model_parameters(m0f) %>%
+  print_html()
+
+library(ggeffects)
+p0f <- plot(ggeffects::ggemmeans(m0f, terms = "ch_s [all]"), add.data = TRUE, dot.alpha = 0.1)    # 
+p0f
 ```
@@
 ###  Model
+
 ```{r}
@@
 
-## Simulation tables
+˝## Simulation tables
 

### 2022-01-30 analysis.Rmd
updated
Update analysis.Rmd

@@
   fig.height= 9)
+
 ```
@@
 ```{r runanalysis, echo=FALSE, include = FALSE}
-#source("libs.R") # libraries 
+# source("libs.R") # libraries 
 # source("funs.R") # custom functions
-
 # libraries
@@
 library("ggdag")
-#library("papaja")
 library("rstan")
@@
 options(mc.cores=parallel::detectCores ())
-theme_set(theme_few())
+theme_set(theme_few())]
 ```

### 2022-02-02 analysis.Rmd
updated

@@
 date: 2021-APRIL-29 
+bibliography: references.bib
+link-citations: true
 output:
@@
 options(mc.cores=parallel::detectCores ())
-theme_set(theme_few())]
+theme_set(theme_few())
 ```
@@
 
-First, we import the data (see functions).
+First, we import the data (see functions). [@tchetgen2012]
 
@@
 
-### IPW Anxiety ~ Belief
+### IPW Anxiety \~ Belief
 
@@
 
-### IPW Anxiety ~ Church
+### IPW Anxiety \~ Church
 
@@
 
-## Results summary Anxiety ~ Belief
+## Results summary Anxiety \~ Belief
 
@@
 
-## Results summary: Anxiety ~ church
+## Results summary: Anxiety \~ church
 
@@
 
-## Results summary Belief ~ Anxiety
+## Results summary Belief \~ Anxiety
 
@@
 <!-- ```{r} -->
+
 <!-- this is no better -->
+
 <!-- m3a<- brm( -->
+
 <!--   bg_s ~ -->
+
 <!--     anx_s+ -->
+
 <!--     ch_s + -->
+
 <!--     age_s  + -->
+
 <!--     male + -->
+
 <!--     (1 + anx_s | country), -->
+
 <!--   data = d1, -->
+
 <!--   seed = 1234, -->
+
 <!--   warmup = 1000, -->
+
 <!--   iter = 3000, -->
+
 <!--   chains = 4, -->
+
 <!--   file = here::here("models", "m3a") -->
+
 <!-- ) -->
@@
 <!-- parameters::model_parameters(m3a, test = "p_direction") -->
+
 <!-- brms::ranef(m3a) -->
+
 <!-- summary(m3a) -->
+
 <!-- hist(d1$bg_s) -->
+
 <!-- # YES -->
+
 <!-- pp_check(m3a) -->
+
 <!-- ``` -->
@@
 
-## Results summary Church ~ Anxiety
+## Results summary Church \~ Anxiety
 
@@
 
-˝## Simulation tables
+˝#\# Simulation tables
 
@@
 <!-- ```{r, cache = TRUE} -->
+
 <!-- rel %>% -->
+
 <!--   correlation( partial = FALSE, multilevel = TRUE ) %>% -->
+
 <!--   plot() + theme_blackboard() -->
+
 <!-- ``` -->
@@
 <!-- ```{r cache = TRUE} -->
+
 <!-- wbgen %>% -->
+
 <!--   correlation( partial = FALSE, multilevel = TRUE ) %>% -->
+
 <!--   plot() + theme_blackboard() -->
+
 <!-- ``` -->
@@
 <!-- ```{r} -->
+
 <!-- wbphys %>% -->
+
 <!--   correlation( partial = FALSE, multilevel = TRUE ) %>% -->
+
 <!--   summary() -->
@@
 <!-- wbphys %>% -->
+
 <!--   correlation( partial = FALSE, multilevel = TRUE ) %>% -->
+
 <!--   plot() + theme_blackboard() -->
+
 <!-- ``` -->
@@
 <!-- ```{r} -->
+
 <!-- wbpsych %>% -->
+
 <!--   select(-wb_psychMEANING)%>% -->
+
 <!--   correlation( partial = FALSE, multilevel = TRUE ) %>% -->
+
 <!--   summary() -->
@@
 <!-- wbpsych %>% -->
+
 <!--   select(-wb_psychMEANING)%>% -->
+
 <!--   correlation( partial = FALSE, multilevel = TRUE ) %>% -->
+
 <!--   plot()  + theme_abyss() -->
+
 <!-- ``` -->

### 2022-02-02 analysis.Rmd
updated

@@
   fig.height= 9)
-
 ```
@@
 library("patchwork")
-library("here")
+library("here")n
 #library("rethinking")
@@
   cbind(d1,
-        demean(d1, select = c("anx", "ch", "bg"), group = "country"))
-
+        parameters::demean(d1, select = c("anx", "ch", "bg"), group = "country"))
+
+hist(d1$anx)
 ```

### 2022-02-06 response-reviewer.md
reorder repo

Reviewer: 1

We are grateful to the reviewer for taking the time to review our manuscript, which has  substantially improved it. Below we address all reviewer comments.

*The motivation of this article is great, and while its execution is reasonable given the constraints of the commentary format, I would encourage the authors to clean it up in a few key ways.*

- Thanks.

First, I should note that it comes across as hastily put together and lacking consideration for the reader.  For example, we should not have to read what appear to be unmodified xtable tables; we should be able to immediate interpret what the estimates are for in Tables B1-B4.

- Thank you. The tables have been placed in the supplementary materials. We do not provide coefficients to better control for the "table two" fallacy. Rational and citation for the table two fallacy is given.

Also, the figures are unreadable without being able to zoom in. The authors should assume readers will use paper to read their contribution. Other examples abound (see below).

- Thank you. We have increased the figure size to meet the journal's requirements, and attached the figures seperately.

I would also encourage the authors to expand this to a full-length article because it is clear that it requires a little more room to breathe.

- Thank you. We agree. However, this is not possible in for an RBB commentary.

That said, the commentary might benefit for some sharper targeting. The commentary should eventually get published, but it needs to guide the reader better: 1) tighten up the framing; 2) lend a bit more consideration to an audience you're trying to convince; your motivations are clear to me, but I doubt they would be to someone who doesn't already know them; 3) stitch the materials together a bit tighter; make sure each part is explicitly connected.

- Thank you. We agree. We have substantially revised the manuscript for clarity.

Framing: 
The MARP project was an attempt to crowd-source analyses, the authors are doing the project and social science more broadly a service by focusing on causal inference. I’m completely sympathetic with the approach and motivation (in fact, the charge is more applicable to the social sciences in general, the field of the scientific study of religion more specifically). The authors don’t imply that the MARP project is guilty of adopting a “model free” approach, but it *might* come across that way given how the framing is a little negative (“...illustrate how a ‘model free’ approach can mislead...”). This might be a little clearer if the scope of criticism were widened (i.e., to also apply to social science more generally).

- We have given a stronger argument against crowd-sourcing in the absence of expert knowledge for causal inference.

> QUOTE HERE

- We have included a reference to new reference to [Miguel Hernans] who makes this point in a recent publication.

Introduction: 
In a few ways, the introduction needs work (another example of its rushed style). First, it doesn’t obviously follow that “religion affects mental health” because it “remains a core feature of human life”. Second, the intro. enumerates three points that are neither introduced nor obviously related to one another. Initially, I thought it would help to say something along the lines of “Yet[,] inferring how religion affects mental health requires care [for at least three important reasons].” but I’m not sure if the following points are reasons for this care. Third, the second point requires some citations and/or precision. As it reads now, there are a few aphoristic truisms here (e.g., “Causality is dynamic”...”the interplay of religion and mental health depends on broader cultural, political...”) that aren’t helping us get to the point. Too much is going on in this paragraph for the relatively mundane point that things can be dynamic.

- We have rewritten the introduction.

Causal inference and models: 
Paragraphs 3-5 that describe the DAGs need a little attention. It’s a little misleading (perhaps more so to an outsider) to suggest that the MARP data can’t address the backdoor path in Fig. 1b; all it would take is thinking of a variable that would block it. If the MARP data (and our imagination) really can’t do it, some qualification is in order. It reads as unsolvable here for no apparent reason. In other words, the commentary is not presenting the virtues of DAGs very well. It’s tough to distill this large literature into a commentary, but the authors should be able to make things a little clearer, especially since their motivation is to convince readers that causal inference really is the way to go here.

- We think the referee wants us to clarify three points. First, if all common causes (or descendents) of the both the exposure and outcome were to be included in our regression analysis, then it would be possible to obtain an unbiased causal estimate conditional on our causal model. Second, it is likely that all such common causes (or descendents) are captured in the MARP dataset. Third, it is therefore important to suggest why cuasal approach is nevertheless interesting.  We have stated these points clearly in the revised text and furthermore offered a sensitivity analysis to evaluate how large such unmeasured counfounding would need to be to render the causal estimate unreliable at the 95th HPD interval.

As mentioned, Figure 1 is simply unreadable. Even if each panel were larger and more legible, the DAGs are quite messy and the key on the side of the graphs is mysterious; we have 3 circles, one square, and three colors/tones. What do they mean and how are they helping us?

- We agree. We have moved all but the essential

Why are age, SES, education, sex, etc. included here? Yes, there is a large literature that predicts their important (cite them), but since they aren’t discussed here, it seems like an unnecessary and untheorized set of variables that would only confound the estimates even more, given the model.

- First, we have included a reference to VANDERWHEELE 2020 which provides advice on confounder control using these indicators. Second, we note that these indicators are unlikely to measure common effects of the exposures and outcomes; as such inclusion of these indicators is unlikely to introduce collider bias. Third, we note that these indicators are unlikely to stand along the path between the exposures and the outcomes; as such, inclusion of these indicators in our regression models is unlikely to induce mediator bias (or pipe) confounding. Fourth, these indicators are unlikely to capture all potential sources of bias. For example, Vanderwheele 2020 includes a much wider set of potential confounders for confounding control. For this reason, we include a sensitivity analysis, which allows us to state how much coufounding would be necessary to render our causal estimate. The new paragraph reads:

>

The y-axis labels in figure 2 are unreadable. The country-level plots probably aren’t necessary anyway, since that’s not really the focus here.

- Thanks, we agree. We have removed these plots.

Statistical model:   
Justifying the priors here might be useful.

- Thanks, we have included posterior predicted values in the

Appendices: 
code for simulations mod2 and mod3 are identical

- Thanks, fixed.

caption in Figure C1 is probably too preachy for a figure; the figure does not clearly convey the idea that “Causal assumptions are needed for causal inference”. The paper needs a bit more pedagogical flair in text to bring readers to appreciate this point.

- Thanks, figure text fixed.

Details: 
footnote 1: powers[.] Perhaps cite Wallace (1966) or Jensen (2019) for this definition? It’s helpful to be clear about this, but there are many out there (and plenty of folks who have thought hard about this).

- Thanks, we revised the introduction, the citation is no longer needed.

line 40 pg. 3: “makes a start”. colloquial.

-  Thanks, we revised the introduction, no longer relevant.

line 25 pg. 4: “Recall that...” I missed the introduction to this point.

- Thanks, we revised the introduction, no longer relevant.

line 12-13 pg. 5: “these interpretations, we...” something’s not right here.

- Thanks, we revised the introduction, sentence no longer exists.

line 38 pg. 5: “being drawn to God when anxious serves functions”. I get what this sentence means, but this construction is a little awkward. Perhaps rephrase.

- Thanks, agree; again, we revised the introduction. Sentence no longer exists.

### 2022-02-06 DAGS_Scotts_submission/dags_for_scott.Rmd
reorder repo

--- 2021-02-09 dags_for_scott.Rmd
@@
 output: html_document
+editor_options: 
+  chunk_output_type: console
 ---
@@
   tidy_dagitty()
+dg
 

### 2022-02-06 analysis.Rmd
reorder repo

--- 2022-02-02 analysis.Rmd
@@
   theme(legend.position = "bottom") +
-  theme_dag() + ggtitle("Adjustment set for Religious Belief
+  theme_dag() + ggtitle("Adjustment set for religious belief
 assuming belief does not change from distress")

### 2022-02-06 analysis.Rmd
Update analysis.Rmd

@@
 
-nocause <- ggdag::ggdag(dg2, text = FALSE, use_labels = "label")  + theme_dag() + labs(title = "Assume unmeasured distressing experience
-does not affect Belief or Service")
+nocause <- ggdag::ggdag(dg2, text = FALSE, use_labels = "label")  + theme_dag() + labs(title = "Assume unmeasured distressing experience does not affect Belief or Service")
 nocause
 
-dp2<-ggdag::ggdag_adjustment_set(dg2, node_size = 14, text = FALSE, use_labels = "label") +
+dg0<-ggdag::ggdag_adjustment_set(dg2, node_size = 14, text = FALSE, use_labels = "label") +
   theme(legend.position = "bottom") +
-  theme_dag() + ggtitle("Adjustment set for Religious Service
+  theme_dag() + ggtitle("Adjustment set for religious service
 assuming service does not change from distress")
 
-dp2
+dp0
 
@@
 dg1
+```
+
+Save image
+
+```{r dg1, eval=FALSE}
+# save graph
+ggsave(
+  dg1,
+  path = here::here("Figs"),
+  width = 10,
+  height = 5,
+  units = "in",
+  filename = "dag_belief.jpg",
+  device = 'jpeg',
+  limitsize = FALSE,
+  dpi = 1200
+)
+
+```
+
+```{r dp0, eval=FALSE}
+# save graph
+ggsave(
+  dg1,
+  path = here::here("Figs"),
+  width = 10,
+  height = 5,
+  units = "in",
+  filename = "dag_church.jpg",
+  device = 'jpeg',
+  limitsize = FALSE,
+  dpi = 1200
+)
 
@@
   theme(legend.position = "bottom") +
-  theme_dag() + ggtitle("Adjustment set for Anxiety causing Belief in God")
+  theme_dag() + ggtitle("Adjustment set assuming anxiety causes belief in God")
 dl2

### 2022-02-08 analysis.Rmd
Update analysis.Rmd

@@
     highlight: kate
+editor_options: 
+  chunk_output_type: console
 ---
@@
 # source("funs.R") # custom functions
-# libraries
-# libraries and settings
 library("tidyverse")
@@
 library("ipw") # inverse probability weighting. 
+library("WeightIt")# inverse probability weighting. 
 rstan_options(auto_write=TRUE)
@@
 # clean data:use informative names !
-
 # note from Suzanne: 
@@
     "U" = "Unmeasured
-    distressing life events"
-  )
-) #%>%#%>%
+    distressing life events",
+    coords = list(x = c(anx =6 , 
+                        belief = 5, 
+                        church = 4, 
+                        ses = 3,
+                        male = 2, 
+                        age = 2,
+                        country = 2,
+                        U = 1),
+                  y = c(anx = 2, 
+                        belief = 2, 
+                        church = 2, 
+                        ses = 2,
+                        male = 2, 
+                        age = 2,
+                        country = 1,
+                        U = 1)) #%>%#%>%
  # tidy_dagitty()       
@@
 library(lme4)
-
 ```
@@
 
+First check for imbalance using the cobolt package.
+
+```{r}
+# CHECK BALANCE
+# test
+cobalt::bal.tab(bg_s ~ age_s  + male + country,
+        data = d1, estimand = "ATE", m.threshold = .05)
+```
+
+Get inverse probability weights
+
+```{r}
+w_out_bs <- weightit(bg_s ~ age_s  +ch_s +  male + country,
+        data = d1, estimand = "ATE", method = "ebal", stabilize = TRUE)
+w_out_bs
+
+summary(w_out_bs)
+sum(w_out_bs$weights)
+
+cobalt::bal.tab(w_out_bs, m.threshold = .05, disp.v.ratio = TRUE)
+```
+
 ```{r}
 library("ipw")
-
 d1ipwB <- ipw::ipwpoint(
@@
 summary(d1ipwB$ipw.weights)
-sum(d1ipwB$ipw.weights > 10)
+sum(d1ipwB$ipw.weights >= 10)
 
 d1pB <- d1 %>% 
-  mutate(ipw = d1ipwB$ipw.weights) %>%
-  filter(ipw < 10)
+  mutate(ipw = w_out_bs$weights) %>%
+  filter(d1ipwB$ipw.weights < 10)
+
 ```
@@
   anx ~
-    bg_s,# + # already weighted
+    bg_s,# already weighted
    # ch_s +
@@
   id = country,
+  corstr = "exchangeable",
   weights = ipw
 )
+mgeeB
 
@@
 d1p <- d1 %>% 
-  mutate(ipw = d1ipw$ipw.weights)
+  mutate(ipw = d1ipw$ipw.weights) %>%
+  filter(ipw < 10)
 ```
@@
 ```{r, eval = FALSE}
-
 # trim IPW to less than 10 
@@
                               ndraws = 500,
-                              categorical = T,
+                              categorical = F,
                               prob = .89,

### 2022-02-18 analysis-v-2.Rmd
update analysis

--- 2022-02-08 analysis.Rmd
@@
 library("ggplot2") #plots
-library("sjstats") #test statistics
-library("sjPlot")
-library("sjmisc")
-library("insight")
-library("see")
-library("performance")
+#library("sjstats") #test statistics
+#library("sjPlot")
+#library("sjmisc")
+##library("insight")
+#library("see")
+#library("performance")
 library("ggraph")
-library("parameters")
+#library("parameters")
 library("lme4")
-library("psych") ## correlation matrix
+#library("psych") ## correlation matrix
 library("table1")
 library("ggeffects")
-library("gghighlight")
-library("texreg")
+#library("gghighlight")
+#library("texreg")
 library("kableExtra")
 #library("naniar") # missingness
-#library("Amelia")
+library("Amelia")
 #library("magrittr")
@@
 library("ggthemes")
-library("ggdag")
-library("rstan")
+=library("rstan")
 library("brms")
@@
 # get data into shape
+
 d1 <- df %>%
   dplyr::filter(
-    !is.na(wb_psychANXIETY),
-    !is.na(age),
-    !is.na(rel_CHURCHFREQ),
-   # !is.na(wb_psychMEANING),
-    !is.na(male),
-    !is.na(rel_EXTENTBELIEVEGOD),
+   #  !is.na(wb_psychANXIETY),
+   #  !is.na(age),
+   #  !is.na(rel_CHURCHFREQ),
+   # # !is.na(wb_psychMEANING),
+   #  !is.na(male),
+   #  !is.na(rel_EXTENTBELIEVEGOD),
   ) %>%
@@
     anx = as.integer(6- wb_psychANXIETY), # reverse coded! rescale 1-5
-    anx_s = scale(anx),
+    #anx_s = scale(anx),
     #put outcome in ST units (see McElreath 2021)  # not necessary
-    age_s = scale(age),
+    #age_s = scale(age),
     ch = rel_CHURCHFREQ,
-    mn  = wb_psychMEANING,
-    bg =  rel_EXTENTBELIEVEGOD,
-    ch_s = scale(rel_CHURCHFREQ),
-    mn_s = scale(wb_psychMEANING),
-    bg_s = scale(rel_EXTENTBELIEVEGOD)
+  #  mn  = wb_psychMEANING,
+    edu = education,
+    bg =  rel_EXTENTBELIEVEGOD#,
+  #  ch_s = scale(rel_CHURCHFREQ),
+  #  mn_s = scale(wb_psychMEANING),
+   # bg_s = scale(rel_EXTENTBELIEVEGOD)
   ) %>%
@@
                 ch,
-                mn,
+                age,
+                ses,
+                #    mn,
+                edu,
                 bg,
-                anx_s,
-                age_s,
-                ch_s,
-                mn_s,
-                male2,
+                # anx_s,
+                # age_s,
+                #  ch_s,
+                #  mn_s,
+                edu,
                 country,
                 male,
-                bg_s,
+                #   bg_s,
                 wb_psychANXIETY)
 
-## degroup variables for within and between analysis
-## https://cran.r-project.org/web/packages/parameters/vignettes/demean.html
-d1 <-
-  cbind(d1,
-        parameters::demean(d1, select = c("anx", "ch", "bg"), group = "country"))
-
-hist(d1$anx)
 ```
@@
 dg1a <- dagify(
-  anx  ~ belief + church + male + country + U,
-  belief ~ age + male + country + church + U,
-  church ~ age + male + country + U,
+  anx  ~ belief + church  + age + male + ses + country ,
+  belief ~ age + church + male + edu + ses  + church + country,
+  church ~ age + male + edu + ses + country,
   ses ~ country + age + edu,
-  edu ~ country + male, 
-  U ~ age + ses + country,
+  edu ~ country + male + age, 
+ # U ~ age + ses + country,
   exposure =  "belief",
   outcome =   "anx",
-  latent = "U",
+ # latent = "U",
   labels = c(
@@
   )
-) #%>%
- # tidy_dagitty()
-
-d0a <-ggdag::ggdag(dg1a, text = FALSE, use_labels = "label")  + theme_dag() + ggtitle("What if we were to assume distressing events increase belief in God?")
-d0a
-# graph adjustment sets
-
-d1a<-ggdag::ggdag_adjustment_set(dg1a, node_size = 14, text = FALSE, use_labels = "label") +
-  theme(legend.position = "bottom") +
-  theme_dag() + ggtitle("If we assume distressing events affect belief and service we cannot estimate causal effects ")
-
-d1a
-
-dg1 <- dagify(
-  anx  ~ belief + church + male + country + U,
-  belief ~ age + male + country + church + U,
-  church ~ age + male + country + U,
-  ses ~ country + age + edu,
-  edu ~ country + male, 
-  U ~ age + ses + country,
-  exposure =  "belief",
-  outcome =   "anx",
-  latent = "U",
-  labels = c(
-    "belief" = "belief God",
-    "church" = "religious service attendance",
-    "anx" = "anxiety",
-    "edu" = "education",
-    "ses" = "SES",
-    "male" = "male",
-    "country" = "country",
-    "age" = "age",
-    "U" = "Unmeasured
-    distressing life events",
-    coords = list(x = c(anx =6 , 
-                        belief = 5, 
-                        church = 4, 
-                        ses = 3,
-                        male = 2, 
-                        age = 2,
-                        country = 2,
-                        U = 1),
-                  y = c(anx = 2, 
-                        belief = 2, 
-                        church = 2, 
-                        ses = 2,
-                        male = 2, 
-                        age = 2,
-                        country = 1,
-                        U = 1)) #%>%#%>%
- # tidy_dagitty()       
+)     
+dg1a
+print( impliedConditionalIndependencies( dg1a ) )
+print( adjustmentSets( dg1a,  "belief","anx" ) )
+
 # 
-dp0 <-ggdag::ggdag(dg1, text = FALSE, use_labels = "label")  + theme_dag() + ggtitle("Assumed unmeasured causal effects of distressing experiences")
-
-dp0
+# d0a <-ggdag::ggdag(dg1a, text = FALSE, use_labels = "label")  + theme_dag() + ggtitle("What if we were to assume distressing events increase belief in God?")
+# d0a
 # # graph adjustment sets
 # 
-#dp1<-ggdag::ggdag_adjustment_set(dg1, node_size = 14, text = FALSE, use_labels = "label") +
+# d1a<-ggdag::ggdag_adjustment_set(dg1a, node_size = 14, text = FALSE, use_labels = "label") +
 #   theme(legend.position = "bottom") +
-#   theme_dag() + ggtitle("Adjustment sets for Belief in God")
-#dp1
-dp0
-```
+#   theme_dag() + ggtitle("If we assume distressing events affect belief and service we cannot estimate causal effects ")
+# 
+# d1a 
+# 
+# 
+# # 
+# dp0 <-ggdag::ggdag(dg1, text = FALSE, use_labels = "label")  + theme_dag() + ggtitle("Assumed unmeasured causal effects of distressing experiences")
+# 
+# dp0
+# # # graph adjustment sets
+# # 
+# #dp1<-ggdag::ggdag_adjustment_set(dg1, node_size = 14, text = FALSE, use_labels = "label") +
+# #   theme(legend.position = "bottom") +
+# #   theme_dag() + ggtitle("Adjustment sets for Belief in God")
+# #dp1
+# dp0
+```
+'
+We adjust for: { church, country, male, ses } . Because we do not have information about education, we condition on the ladder
 
 ## Causal model: religious service
-
-We condition on the following model:
-
--   religion + ses + male + country + a different unmeasured latent variable1 (e.g. privilege) affects distress.
-
--   age + male + country + and unmeasured latent variable2 (e.g. parent's religion, call this U1) affects Religion
-
--   country + age + an unmeasured latent variable1 causes ses (citations...)
-
--   our "exposure" is religious service attendance or religious belief and our outcome is "anxiety"
 
@@
 dg2 <- dagify( 
-  anx  ~ belief + church + male + country + U,
-  belief ~ age + male + country + church,
-  church ~ age + male + country,
+  anx  ~ belief + church  + age + male + ses + country ,
+  belief ~ age + church + male + edu + ses  + church + country,
+  church ~ age + male + edu + ses + country,
   ses ~ country + age + edu,
-  edu ~ country + male, 
-  U ~ age + ses + country,
+  edu ~ country + male + age, 
   exposure =  "church",
   outcome =   "anx",
-  latent = "U",
+#  latent = "U",
   labels = c(
@@
   )
-) %>%
+)
+
+%>%
   tidy_dagitty()
 
-ggdag::ggdag(dg2) + theme_dag()
-
-nocause <- ggdag::ggdag(dg2, text = FALSE, use_labels = "label")  + theme_dag() + labs(title = "Assume unmeasured distressing experience does not affect Belief or Service")
-nocause
-
-dg0<-ggdag::ggdag_adjustment_set(dg2, node_size = 14, text = FALSE, use_labels = "label") +
-  theme(legend.position = "bottom") +
-  theme_dag() + ggtitle("Adjustment set for religious service
-assuming service does not change from distress")
-
-dp0
-
-dpz <- dagify(
-  anx  ~ belief + church + male + country + U,
-  belief ~ age + male + country + church,
-  church ~ age + male + country,
-  ses ~ country + age + edu,
-  edu ~ country + male, 
-  U ~ age + ses + country,
-  exposure =  "belief",
-  outcome =   "anx",
-  latent = "U",
-  labels = c(
-    "belief" = "belief God",
-    "church" = "religious service attendance",
-    "anx" = "anxiety",
-    "edu" = "education",
-    "ses" = "SES",
-    "male" = "male",
-    "country" = "country",
-    "age" = "age",
-    "U" = "Unmeasured
-    distressing life events"
-  )
-)
-
-ggdag::ggdag(dpz) + theme_dag()
-
-# graph adjustment sets
-dg1<-ggdag::ggdag_adjustment_set(dpz, node_size = 14, text = FALSE, use_labels = "label") +
-  theme(legend.position = "bottom") +
-  theme_dag() + ggtitle("Adjustment set for religious belief
-assuming belief does not change from distress")
-
-dg1
-```
-
-Save image
-
-```{r dg1, eval=FALSE}
-# save graph
-ggsave(
-  dg1,
-  path = here::here("Figs"),
-  width = 10,
-  height = 5,
-  units = "in",
-  filename = "dag_belief.jpg",
-  device = 'jpeg',
-  limitsize = FALSE,
-  dpi = 1200
-)
-
-```
-
-```{r dp0, eval=FALSE}
+print( impliedConditionalIndependencies( dg2 ) )
+print( adjustmentSets( dg1a,  "church", "anx" ) )
+
+# 
+# ggdag::ggdag(dg2) + theme_dag()
+# 
+# nocause <- ggdag::ggdag(dg2, text = FALSE, use_labels = "label")  + theme_dag() + labs(title = "Assume unmeasured distressing experience does not affect Belief or Service")
+# nocause
+# 
+# dg0<-ggdag::ggdag_adjustment_set(dg2, node_size = 14, text = FALSE, use_labels = "label") +
+#   theme(legend.position = "bottom") +
+#   theme_dag() + ggtitle("Adjustment set for religious service
+# assuming service does not change from distress")
+# 
+# dp0
+# 
+# 
+# dpz <- dagify(
+#   anx  ~ belief + church + male + country + U,
+#   belief ~ age + male + country + church,
+#   church ~ age + male + country,
+#   ses ~ country + age + edu,
+#   edu ~ country + male, 
+#   U ~ age + ses + country,
+#   exposure =  "belief",
+#   outcome =   "anx",
+#   latent = "U",
+#   labels = c(
+#     "belief" = "belief God",
+#     "church" = "religious service attendance",
+#     "anx" = "anxiety",
+#     "edu" = "education",
+#     "ses" = "SES",
+#     "male" = "male",
+#     "country" = "country",
+#     "age" = "age",
+#     "U" = "Unmeasured
+#     distressing life events"
+#   )
+# )
+# 
+# ggdag::ggdag(dpz) + theme_dag()
+# 
+# 
+# # graph adjustment sets
+# dg1<-ggdag::ggdag_adjustment_set(dpz, node_size = 14, text = FALSE, use_labels = "label") +
+#   theme(legend.position = "bottom") +
+#   theme_dag() + ggtitle("Adjustment set for religious belief
+# assuming belief does not change from distress")
+# 
+# dg1
+```
+
+We have two adjustment sets: { age, church, country, male, ses } and { church, country, edu, male, ses }
+
+<!-- ```{r dg1, eval=FALSE} -->
+<!-- # save graph -->
+<!-- ggsave( -->
+<!--   dg1, -->
+<!--   path = here::here("Figs"), -->
+<!--   width = 10, -->
+<!--   height = 5, -->
+<!--   units = "in", -->
+<!--   filename = "dag_belief.jpg", -->
+<!--   device = 'jpeg', -->
+<!--   limitsize = FALSE, -->
+<!--   dpi = 1200 -->
+<!-- ) -->
+
+<!-- ``` -->
+
+<!-- ```{r dp0, eval=FALSE} -->
+<!-- # save graph -->
+<!-- ggsave( -->
+<!--   dg1, -->
+<!--   path = here::here("Figs"), -->
+<!--   width = 10, -->
+<!--   height = 5, -->
+<!--   units = "in", -->
+<!--   filename = "dag_church.jpg", -->
+<!--   device = 'jpeg', -->
+<!--   limitsize = FALSE, -->
+<!--   dpi = 1200 -->
+<!-- ) -->
+
+<!-- ``` -->
+
+## Multiple imputation
+
+### Visualise missing data
+
+Next we need to examine patterns of missing-ness, which might spoil inference.
+
+```{r cache = TRUE}
+library("naniar")
+vis_miss(df) # substantial pattern of missingness in a few variables, 
+df
+df$wb_soc_3
+dff<-df%>%
+  select(-wb_soc_3)
+sum(is.na(df$education))
+
+miss_graph<- gg_miss_upset(dff)
+miss_graph
 # save graph
@@
 
-```
-
-Graph for paper
-
-## Anxiety causes belief DAG
-
-```{r}
-dpL <- dagify(
-  anx  ~  church + male + country + U,
-  belief ~ age + male + country + church + anx,
-  church ~ age + male + country,
-  ses ~ country + age + edu,
-  edu ~ country + male, 
-  U ~ age + ses + country,
-  exposure =  "anx",
-  outcome =   "belief",
-  latent = "U",
-  labels = c(
-    "belief" = "belief God",
-    "church" = "religious service attendance",
-    "anx" = "anxiety",
-    "edu" = "education",
-    "ses" = "SES",
-    "male" = "male",
-    "country" = "country",
-    "age" = "age",
-    "U" = "Unmeasured
-    distressing life events"
-  )
-) %>%
-  tidy_dagitty()
-
-dl <- ggdag::ggdag(dpL, text = F, use_labels = "label") + theme_dag()
-dl1<- dl + labs(title ="Assume Anxiety causes Belief in God")
-dl1
-
-dl2<-ggdag::ggdag_adjustment_set(dpL, node_size = 14, text = FALSE, use_labels = "label") +
-  theme(legend.position = "bottom") +
-  theme_dag() + ggtitle("Adjustment set assuming anxiety causes belief in God")
-dl2
-
-```
-
-## Anxiety causes church DAG
-
-```{r}
-dpC <- dagify(
-  anx  ~   male + country + U,
-  belief ~ age + male + country + church + anx,
-  church ~ age + male + country,
-  ses ~ country + age + edu,
-  edu ~ country + male, 
-  U ~ age + ses + country,
-  exposure =  "anx",
-  outcome =   "church",
-  latent = "U",
-  labels = c(
-    "belief" = "belief God",
-    "church" = "religious service attendance",
-    "anx" = "anxiety",
-    "edu" = "education",
-    "ses" = "SES",
-    "male" = "male",
-    "country" = "country",
-    "age" = "age",
-    "U" = "Unmeasured
-    distressing life events"
-  )
-) %>%
-  tidy_dagitty()
-
-dC <- ggdag::ggdag(dpC, text = F, use_labels = "label") + theme_dag()
-dlC<- dC + labs(title ="Assume Anxiety causes Religious Service")
-dlC
-
-dl2C<-ggdag::ggdag_adjustment_set(dpC, node_size = 14, text = FALSE, use_labels = "label") +
-  theme(legend.position = "bottom") +
-  theme_dag() + ggtitle("Adjustment set for Anxiety causing Religious Service")
-dl2C
-
-```
-
-## Dag graph
-
-```{r}
-library(patchwork)
- (dp0 + d1a)/   (nocause + dg1 + dp2)  /(dl1 + dl2 + dl2C)+   plot_layout(guides = 'collect')  + plot_annotation(title = 'Three causal models', tag_levels = "a") 
-# dag_ful 20/20
-```
-
-### correlation
-
-```{r}
-correlation::correlation(d1, select = "ch_s",select2 = "bg_s")
-
-correlation::correlation(d1, select = "bg_s",select2 = "anx")
-
-correlation::correlation(d1, select = "ch_s",select2 = "anx")
-
-```
-
-```{r}
-library(lme4)
-```
-
-### Check LMER Results
-
-```{r}
-library(patchwork)
-mlmerCH <- lmer(
-    anx ~
-    #bg_s +
-    ch_s +
-    age_s  +
-    male +
-    ( 1 + ch_s | country), 
-  data = d1)
-
-sjPlot::tab_model(mlmerCH)
-
-mlmerBG <- lmer(
-    anx ~
+library(Amelia)
+
+  # x is the longitudional dataframe (here 'd_3')
+set.seed(1234)
+prep <- d1 %>% # Remove what we do not need anymore
+    arrange(country) %>%
+    dplyr::select(anx,
+                  ch,
+                  age,
+                  edu,
+                  bg,
+                  edu,
+                  country,
+                  male,
+                  anx, 
+                  ses)
+# impute missing
+prep <- as.data.frame(prep) # tibble won't run in amelia
+set.seed(1234)
+prep2 <- Amelia::amelia(
+  prep,
+  #dataset to impute
+  m = 10,
+  # number of imputations
+  cs = c("country"),
+  #ts = c("years"),
+  noms = c(
+    "male"
+  ))
+  #idvars = c("country")#,
+  # not imputing outcomes
+ # polytime = 3
+)  #https://stackoverflow.com/questions/56218702/missing-data-warning-r
+
+out <- Amelia::transform.amelia(
+  prep2,
+  age_s = scale(age),
+  bg_s = scale(bg),
+  anx_s = scale(anx),
+  age_s = scale(age),
+  ch_s =scale(ch),
+  edu_s = scale(edu),
+  ses_s = scale(ses),
+  male_2 = scale(as.numeric(male))/2) # scale by 2 sds
+
+#saveRDS(out, here::here("mods","out"))
+out <-readRDS(here::here("mods","out"))
+```
+
+We find there is substantial missingness in denomination. There is also missingness in ethnicity. Overall missingness is low in other indicators. However, 
+Because missingness can introduce bias, we multiply impute responses.
+
+## Model for belief
+
+### gee
+
+```{r}
+##
+#{ age, country, edu, male, ses }
+m<-10
+models<-NULL
+for(i in 1:m) {
+  models[[i]]  <- geeglm(
+  anx_s ~ bg_s + ch_s + age_s + edu_s + male_2 + ses_s,
+  data = out$imputations[[i]],  
+  id = country,
+  corstr = "exchangeable"
+)
+}
+
+# saveRDS(models, here::here(
+#   "mods","models.rds")
+# )
+
+parameters::pool_parameters(models)
+
+pred <- ggeffects::ggemmeans(models[[1]], terms = c("bg_s[all]"))
+pred
+
+plot(pred,
+     add.data=T,dot.alpha=.1, 
+     facets = F)  + #scale_y_continuous(limits=c(-.5,.5)) + 
+  theme_classic() + ggtitle("") # title to be suppled
+```
+
+### Marginal structural model Church
+
+```{r}
+library(mice)
+d2<-mice(d1)
+d3<-complete(d2)
+saveRDS(d3,here::here("mods","d3"))
+d3 <-readRDS(here::here("mods","d3"))
+d4<-d3%>%
+  select(anx, ch, bg, age, edu,male,ses,country)%>%
+  drop_na()%>%
+  mutate(anx_s = scale(anx),
+         bg_s = scale(bg),
+         ch_s = scale(ch),
+         age_s = scale(age),
+         edu_s = scale(edu),
+         ses_s = scale(ses),
+         male_2 =scale(as.numeric(male))/2,
+         ses_s = scale(ses),
+         country=as.factor(country))
+
+m<-10
+modelsCH<-NULL
+for(i in 1:m) {
+  modelsCH[[i]]  <- geeglm(
+  anx_s ~  ch_s + age_s + edu_s + male_2 + ses_s,
+  data = out$imputations[[i]],  
+  id = country,
+  corstr = "exchangeable"
+)
+}
+
+saveRDS(modelsCH, here::here(
+  "mods","modelsCH.rds")
+)
+
+pred <- ggeffects::ggemmeans(modelsCH[[1]], terms = c("ch_s[all]"))
+pred
+
+plot(pred,
+    # add.data=T,dot.alpha=.1, 
+     facets = F)  + scale_y_continuous(limits=c(-.5,.5)) + theme_classic() + ggtitle("") # title to be suppled
+parameters::pool_parameters(modelsCH)
+
+parameters::pool_parameters(modelsCH)
+#parameters::pool_parameters(modelschG)
+
+library(ipw)
+
+dbl <- ipw::ipwpoint(
+  exposure = bg_g,
+  family = "gaussian",
+  numerator = ~ 1,
+  denominator = ~ age_s +ch_s + edu_s + male_2 + ses_s + country,
+  trunc = 0.01,
+  data = as.data.frame(d4)
+)
+
+dch <- ipw::ipwpoint(
+  exposure = ch_s,
+  family = "gaussian",
+  numerator = ~ 1,
+  denominator = ~ age_s + edu_s + male_2 + ses_s + country,
+  trunc = 0.01,
+  data = as.data.frame(d4)
+)
+
+d1p <- d4 %>% 
+  mutate(ipwCH = dch$weights.trunc,
+         ipwBG = dbl$weights.trunc)
+
+max(d1p$ipwBG)
+max(d1p$ipwCH)
+```
+
+```{r}
+
+ipb_bg<- geeglm(
+  anx_s ~ bg_s,
+  data = d1p,  
+  id = country,
+  corstr = "exchangeable",
+  weights = ipwBG
+)
+
+ipb_ch<- geeglm(
+  anx_s ~  ch_s,
+  data = d1p,  
+  id = country,
+  corstr = "exchangeable",
+  weights = ipwCH
+)
+
+model_parameters(ipb_bg)
+
+model_parameters(ipb_ch)
+
+library(EValue)
+
+# sensitivity = Church
+evalues.OLS(-0.08, se = 0.01, sd = 1, delta = 1)
+
+# sensitivity
+evalues.OLS(0.04, se = 0.01, sd = 1, delta = 1)
+
+bias_plot(1.3009531, xmax = 6)
+
+# An E-value for unmeasured confounding is minimum strength of association, on the risk ratio scale, that unmeasured confounder(s) would need to have with both the treatment and the outcome to fully explain away a specific treatment–outcome association, conditional on the measured covariates.
+# 
+# The estimate is converted appropriately before the E-value is calculated.
+See conversion functions for more details.
+The point estimate and confidence limits after conversion are returned, as is the E-value for the point estimate and the confidence limit closest to the proposed "true" value (by default, the null value.)
+# 
+# For an OLS() estimate, the E-value is for linear regression with a continuous exposure and outcome.
+Regarding the continuous exposure, the choice of delta defines essentially a dichotomization in the exposure between hypothetical groups of subjects with exposures equal to an arbitrary value c versus to another hypothetical group with exposures equal to c + delta.
+# 
+# For example, if resulting E-value is 2, this means that unmeasured confounder(s) would need to double the probability of a subject's having exposure equal to c + delta instead of c, and would also need to double the probability of being high versus low on the outcome, in which the cutoff for "high" versus "low" is arbitrary subject to some distributional assumptions (Chinn, 2000).
+```
+
+Adjusted GEE no weights
+
+```{r}
+
+ipb_bg_noweights <- geeglm(
+  anx_s ~ bg_s + ch_s + age_s + edu_s + male_2 + ses_s,
+  data = d1p,  
+  id = country,
+  corstr = "exchangeable"
+)
+
+ipb_ch_noweights <- geeglm(
+  anx_s ~ ch_s + age_s + edu_s + male_2 + ses_s,
+  data = d1p,  
+  id = country,
+  corstr = "exchangeable"
+)
+
+model_parameters(ipb_bg_noweights)
+model_parameters(ipb_bg)
+
+model_parameters(ipb_ch_noweights)
+model_parameters(ipb_ch)
+```
+
+We note that the multi-level models give different estimates.
+
+### GlMM TMB BELIEF
+
+```{r}
+library(glmmTMB)
+m<-10
+modelsb<-NULL
+for(i in 1:m) {
+  modelsb[[i]]  <- glmmTMB(anx_s ~ bg_s + ch_s + age_s + edu_s + male_2 + ses_s + 
+                            (1+ bg_s|country), 
+                              data = out$imputations[[i]],
+                              family =)
+}
+
+parameters::pool_parameters(modelsb)
+
+pred <- ggeffects::ggemmeans(modelsb[[1]], terms = c("bg_s[all]"))
+pred
+
+plot(pred,
+    # add.data=T,dot.alpha=.1, 
+     facets = F)  + scale_y_continuous(limits=c(-.2,.2)) + theme_classic() + ggtitle("") # title to be suppled
+```
+
+### GlMM TMB CHURCH
+
+```{r}
+library(glmmTMB)
+m<-10
+modelsC<-NULL
+for(i in 1:m) {
+  modelsC[[i]]  <- glmmTMB(anx_s ~  ch_s + age_s + edu_s + male_2 + ses_s + 
+                            (1+ bg_s|country), 
+                              data = out$imputations[[i]],
+                              family =)
+}
+
+parameters::pool_parameters(modelsC)
+
+pred <- ggeffects::ggemmeans(modelsC[[1]], terms = c("ch_s[all]"))
+pred
+
+plot(pred,
+     add.data=T,dot.alpha=.05, 
+     facets = F)  + theme_classic() + ggtitle("") # title to be suppled
+```
+
+### Test implied conditional independence
+
+```{R}
+m<-10
+models<-NULL
+for(i in 1:m) {
+  models[[i]] <- lmer(anx_s ~ 
                         bg_s + 
@@
                         age_s + 
-    male +
-    ( 1 + bg_s | country), 
-  data = d1)
-
-sjPlot::tab_model(mlmerBG)
-```
-
-## General Estimating Equations
-
-### IPW Anxiety \~ Belief
-
-First check for imbalance using the cobolt package.
-
-```{r}
-# CHECK BALANCE
-# test
-cobalt::bal.tab(bg_s ~ age_s  + male + country,
-        data = d1, estimand = "ATE", m.threshold = .05)
-```
-
-Get inverse probability weights
-
-```{r}
-w_out_bs <- weightit(bg_s ~ age_s  +ch_s +  male + country,
-        data = d1, estimand = "ATE", method = "ebal", stabilize = TRUE)
-w_out_bs
-
-summary(w_out_bs)
-sum(w_out_bs$weights)
-
-cobalt::bal.tab(w_out_bs, m.threshold = .05, disp.v.ratio = TRUE)
-```
-
-```{r}
-library("ipw")
-d1ipwB <- ipw::ipwpoint(
-  exposure = bg_s,
-  family = "gaussian",
-  numerator = ~ 1,
-  denominator = ~  ch_s + age_s  + male + country,
-  data = as.data.frame(d1)
-)
-
-summary(d1ipwB$ipw.weights)
-sum(d1ipwB$ipw.weights >= 10)
-
-d1pB <- d1 %>% 
-  mutate(ipw = w_out_bs$weights) %>%
-  filter(d1ipwB$ipw.weights < 10)
-
-```
-
-Model:
-
-```{r}
-mgeeB  <- geeglm(
-  anx ~
-    bg_s,# already weighted
-   # ch_s +
-   # age_s  +
-   # male,
-  data = d1pB,
-  id = country,
-  corstr = "exchangeable",
-  weights = ipw
-)
-mgeeB
-
-parameters::model_parameters(mgeeB) %>%
-  print_html()
-
-library(ggeffects)
-p0BG <- plot(ggeffects::ggemmeans(mgeeB, terms = "bg_s [all]"), add.data = TRUE, dot.alpha = 0.1)    # 
-p0BG 
-```
-
-### IPW Anxiety \~ Church
-
-First we create Inverse Probability Weights
-
-```{r}
-library("ipw")
-
-d1ipw <- ipw::ipwpoint(
-  exposure = ch_s,
-  family = "gaussian",
-  numerator = ~ 1,
-  denominator = ~  age_s  + male + country,
-  data = as.data.frame(d1)
-)
-
-summary(d1ipw$ipw.weights)
-sum(d1ipw$ipw.weights > 10)
-
-d1p <- d1 %>% 
-  mutate(ipw = d1ipw$ipw.weights) %>%
-  filter(ipw < 10)
-```
-
-Model
-
-```{r}
-mgeeCH <- geeglm(
-  anx ~
-    ch_s,# + # already weighted
-    #age_s  +
-  #  male,
-  data = d1p,
-  id = country,
-  weights = ipw
-)
-
-parameters::model_parameters(mgeeCH) %>%
-  print_html()
-
-library(ggeffects)
-pmgeeCH <- plot(ggeffects::ggemmeans(mgeeCH, terms = "ch_s [all]"), add.data = TRUE, dot.alpha = 0.1)    # 
-pmgeeCH
-```
-
-Model with with trimmed IPW (as recommended to avoid leverage from rare cases). Note, the result here is the same.
-
-```{r, eval = FALSE}
-# trim IPW to less than 10 
-d1pf <- d1 %>% 
-  mutate(ipw = d1ipw$ipw.weights) %>%
-  filter(ipw < 10)
-
-m0f <- geeglm(
-  anx ~
-    ch_s,# +
-    #age_s  +
-  #  male,
-  data = d1pf,
-  id = country,
-  weights = ipw
-)
-
-parameters::model_parameters(m0f) %>%
-  print_html()
-
-library(ggeffects)
-p0f <- plot(ggeffects::ggemmeans(m0f, terms = "ch_s [all]"), add.data = TRUE, dot.alpha = 0.1)    # 
-p0f
-```
-
-## Results summary Anxiety \~ Belief
+                        edu_s + 
+                        male_2 + 
+                        ses_s + 
+                        (1+ bg_s|country ), data=out$imputations[[i]])
+}
+
+parameters::pool_parameters(models,
+                            component = "marginal")
+
+pred <- ggeffects::ggemmeans(models[[1]], terms = c("bg_s[all]"))
+pred
+
+plot(pred,
+    # add.data=T,dot.alpha=.1, 
+     facets = F)  + scale_y_continuous(limits=c(-.2,.2)) + theme_classic() + ggtitle("") # title to be suppled
 
@@
 
-```{r}
-m1 <- brm( 
-  anx ~
+listout <- list(
+  out$imputations$imp1,
+  out$imputations$imp2,
+  out$imputations$imp3,
+  out$imputations$imp4,
+  out$imputations$imp5,
+  out$imputations$imp6,
+  out$imputations$imp7,
+  out$imputations$imp8,
+  out$imputations$imp9,
+  out$imputations$imp10
+)
+#saveRDS(listout, here::here("mods", "listout"))
+listout<- readRDS( here::here("mods", "listout"))
+
+m_bg <- brm_multiple( 
+  anx_s ~
     bg_s +
@@
     age_s +
-    male +
+    edu_s +
+    male_2 +
+    ses_s +
     (1 + bg_s | country),
-  data = d1,
-  family = cumulative(link = "logit"),
+  data = listout,
   seed = 1234,
   warmup = 1000,
-  iter = 3000,
+  iter = 2000,
   chains = 4,
-  file = here::here("models", "m2o_real")
-)
-
-parameters::model_parameters(m1, test = "p_direction")
-brms::ranef(m1)
+  backend = "cmdstanr", 
+  threads = threading(4),
+  file = here::here("mods", "m_bg")
+)
+
+parameters::model_parameters(m_bg, test = "p_direction")
+brms::ranef(m_bg)
 
 parameters::parameters(
-  m1,
+  m_bg,
   test = "pd",
@@
 
+```
+
+### Bayesian Church Attendaance
+
+```{r}
+
+listout <- readRDS(here::here("mods","listout"))
+modmc <- brm_multiple(
+  anx_s ~ ch_s + age_s + edu_s + male_2 +ses_s + (1 + ch_s | country),
+  data = listout,
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  backend = "cmdstanr",
+  threads = threading(4)
+)
+
+#saveRDS(modmc, here::here("mods", "modmc.rds"))
+summary(modmc)
+```
+
+### Summary of selection bias
+
+```{r}
+set.seed(1234)
+y<- rnorm(1000,0,1)
+a<- rnorm(1000,0,1)
+s<- rnorm(1000, a -(2*y),4)
+
+dlbind <-data.frame(y,a,s)
+summary(lm(y ~  a + s, data = dlbind))
 ```

### 2022-02-27 analysis-religious-only.Rmd
updated analysis

--- 2022-02-18 analysis-v-2.Rmd
@@
 library("brms")
-library("tidybayes")
+library("bayesplot")
 library("bayesplot")
@@
 library("WeightIt")# inverse probability weighting. 
+library("emmeans")
+library("splines")
 rstan_options(auto_write=TRUE)
@@
 theme_set(theme_few())
+cite_packages(include_R = TRUE)
+
+citation("daggity")
+toBibtex(citation("ggdag"))
+print(citation("EValue") , bibtex=TRUE)
 ```
@@
     # factor 1 = yes, 2 = no, 3 = atheist// yikes factor transformed into a number
+    # NOOO revers scorred (1= religious, 2= not religious, 3=atheist) [R] So note that 3 = religious!!
     BELONGS_DENOMINATION = rel_4,
@@
   )
-
+max(df$rel_EXTENTBELIEVEGOD)
+max(df$rel_CHURCHFREQ)
+length(unique(df$country))
+table(df$country)
 ```
@@
     anx = as.integer(6- wb_psychANXIETY), # reverse coded! rescale 1-5
-    #anx_s = scale(anx),
+    religious = as.integer(IS_RELIIGOUS), #(1= religious, 2= not religious, 3=atheist) [R]
     #put outcome in ST units (see McElreath 2021)  # not necessary
@@
                 male,
+                religious,
                 #   bg_s,
@@
 #
-# 
-# dpz <- dagify(
-#   anx  ~ belief + church + male + country + U,
-#   belief ~ age + male + country + church,
-#   church ~ age + male + country,
-#   ses ~ country + age + edu,
-#   edu ~ country + male, 
-#   U ~ age + ses + country,
-#   exposure =  "belief",
-#   outcome =   "anx",
-#   latent = "U",
-#   labels = c(
-#     "belief" = "belief God",
-#     "church" = "religious service attendance",
-#     "anx" = "anxiety",
-#     "edu" = "education",
-#     "ses" = "SES",
-#     "male" = "male",
-#     "country" = "country",
-#     "age" = "age",
-#     "U" = "Unmeasured
-#     distressing life events"
-#   )
-# )
-# 
-# ggdag::ggdag(dpz) + theme_dag()
+
+dpz <- dagify(
+  anx  ~ age + belief + church + male + country + ses,
+  belief ~ age + male + country + church + ses + edu,
+  church ~ age + male + country + ses +edu,
+  ses ~ country  + edu,
+  edu ~ country + male + age ,
+  U ~ age + ses + country,
+  exposure =  "church",
+  outcome =   "anx",
+  latent = "U",
+  labels = c(
+    "belief" = "belief God",
+    "church" = "religious service attendance",
+    "anx" = "anxiety",
+    "edu" = "education",
+    "ses" = "SES",
+    "male" = "male",
+    "country" = "country",
+    "age" = "age",
+    "U" = "Unmeasured
+    distressing life events"
+  )
+)
+
+ggdag::ggdag(dpz) + theme_dag()
 # 
@@
 # # graph adjustment sets
-# dg1<-ggdag::ggdag_adjustment_set(dpz, node_size = 14, text = FALSE, use_labels = "label") +
-#   theme(legend.position = "bottom") +
-#   theme_dag() + ggtitle("Adjustment set for religious belief
-# assuming belief does not change from distress")
-# 
-# dg1
+dg1<-ggdag::ggdag_adjustment_set(dpz, node_size = 14, text = FALSE, use_labels = "label") +
+  theme(legend.position = "bottom") +
+  theme_dag() + ggtitle("Adjustment set for religious belief
+assuming belief does not change from distress")
+
+dg1
 ```
@@
 library("naniar")
+library(magick)
 vis_miss(df) # substantial pattern of missingness in a few variables, 
@@
   select(-wb_soc_3)
-sum(is.na(df$education))
 
 miss_graph<- gg_miss_upset(dff)
+miss_graphR <-image_ggplot(miss_graph)
+# save graph
 miss_graph
-# save graph
-ggsave(
-  dg1,
-  path = here::here("Figs"),
-  width = 10,
-  height = 5,
+
+ggsave(miss_graph,
+  path = here::here("figs"),
+  width = 16,
+  height = 9,
   units = "in",
-  filename = "dag_church.jpg",
+  filename = "miss_graph.jpg",
   device = 'jpeg',
@@
                   ses)
-# impute missing
-prep <- as.data.frame(prep) # tibble won't run in amelia
-set.seed(1234)
-prep2 <- Amelia::amelia(
-  prep,
-  #dataset to impute
-  m = 10,
-  # number of imputations
-  cs = c("country"),
-  #ts = c("years"),
-  noms = c(
-    "male"
-  ))
-  #idvars = c("country")#,
-  # not imputing outcomes
- # polytime = 3
-)  #https://stackoverflow.com/questions/56218702/missing-data-warning-r
-
-out <- Amelia::transform.amelia(
-  prep2,
+
+table1::table1(~ ch |religious, data = d1)
+
+```
+
+We find there is substantial missingness in denomination. There is also missingness in ethnicity. Overall missingness is low in other indicators. However, 
+Because missingness can introduce bias, we multiply impute responses.
+
+## Model for belief
+
+### Marginal structural model Church
+
+```{r}
+library(mice)
+library(tidyverse)
+d2<-mice(d1)
+
+#d3<-complete(d2)
+#saveRDS(d3,here::here("mods","d3"))
+#d3 <-readRDS(here::here("mods","d3"))
+d4<-d3%>%
+  select(anx, ch, bg, age, edu,male,ses,country,religious)%>%
+  drop_na()%>%
+  mutate(anx_s = scale(anx),
+         religious = factor(religious),
+         bg_s = scale(bg),
+         ch_s = scale(ch),
          age_s = scale(age),
-  bg_s = scale(bg),
-  anx_s = scale(anx),
-  age_s = scale(age),
-  ch_s =scale(ch),
          edu_s = scale(edu),
          ses_s = scale(ses),
-  male_2 = scale(as.numeric(male))/2) # scale by 2 sds
-
-#saveRDS(out, here::here("mods","out"))
-out <-readRDS(here::here("mods","out"))
-```
-
-We find there is substantial missingness in denomination. There is also missingness in ethnicity. Overall missingness is low in other indicators. However, 
-Because missingness can introduce bias, we multiply impute responses.
-
-## Model for belief
-
-### gee
-
-```{r}
-##
-#{ age, country, edu, male, ses }
-m<-10
-models<-NULL
-for(i in 1:m) {
-  models[[i]]  <- geeglm(
-  anx_s ~ bg_s + ch_s + age_s + edu_s + male_2 + ses_s,
-  data = out$imputations[[i]],  
-  id = country,
-  corstr = "exchangeable"
-)
-}
-
-# saveRDS(models, here::here(
-#   "mods","models.rds")
-# )
-
-parameters::pool_parameters(models)
-
-pred <- ggeffects::ggemmeans(models[[1]], terms = c("bg_s[all]"))
-pred
-
-plot(pred,
-     add.data=T,dot.alpha=.1, 
-     facets = F)  + #scale_y_continuous(limits=c(-.5,.5)) + 
-  theme_classic() + ggtitle("") # title to be suppled
-```
-
-### Marginal structural model Church
-
-```{r}
-library(mice)
-d2<-mice(d1)
-d3<-complete(d2)
-saveRDS(d3,here::here("mods","d3"))
+         male_2 =scale(as.numeric(male))/2,
+         ses_s = scale(ses),
+         country=as.factor(country))
+
+## MICE ONLY RELIGIOUS
+table(d1$religious)
+
+# Check
+d1r<- d1 %>%
+  filter(religious == 3)
+
+hist(d1r$ch)
+
+d2r<-mice(d1r)
+
+#check religious rumbers (1= religious, 2= not religious, 3=atheist) [R] So note that 3 = religious!!
+
+d3r<-complete(d2r)
+
+#saveRDS(d3,here::here("mods","d3r"))
+#saveRDS(d3r,here::here("mods","d3r"))
+
 d3 <-readRDS(here::here("mods","d3"))
+d3r <-readRDS(here::here("mods","d3r"))
+
 d4<-d3%>%
+  select(anx, ch, bg, age, edu,male,ses,country,religious)%>%
+  drop_na()%>%
+  mutate(anx_s = scale(anx),
+         bg_s = scale(bg),
+         ch_s = scale(ch),
+         age_s = scale(age),
+         edu_s = scale(edu),
+         ses_s = scale(ses),
+         male_2 =scale(as.numeric(male))/2,
+         ses_s = scale(ses),
+         country=as.factor(country),
+         religious = as.factor(religious))
+
+d4r<-d3r%>%
   select(anx, ch, bg, age, edu,male,ses,country)%>%
@@
 
-m<-10
-modelsCH<-NULL
-for(i in 1:m) {
-  modelsCH[[i]]  <- geeglm(
-  anx_s ~  ch_s + age_s + edu_s + male_2 + ses_s,
-  data = out$imputations[[i]],  
-  id = country,
-  corstr = "exchangeable"
-)
-}
-
-saveRDS(modelsCH, here::here(
-  "mods","modelsCH.rds")
-)
-
-pred <- ggeffects::ggemmeans(modelsCH[[1]], terms = c("ch_s[all]"))
-pred
-
-plot(pred,
-    # add.data=T,dot.alpha=.1, 
-     facets = F)  + scale_y_continuous(limits=c(-.5,.5)) + theme_classic() + ggtitle("") # title to be suppled
-parameters::pool_parameters(modelsCH)
-
-parameters::pool_parameters(modelsCH)
-#parameters::pool_parameters(modelschG)
-
 library(ipw)
-
-dbl <- ipw::ipwpoint(
-  exposure = bg_g,
-  family = "gaussian",
-  numerator = ~ 1,
-  denominator = ~ age_s +ch_s + edu_s + male_2 + ses_s + country,
-  trunc = 0.01,
-  data = as.data.frame(d4)
-)
 
@@
 d1p <- d4 %>% 
-  mutate(ipwCH = dch$weights.trunc,
-         ipwBG = dbl$weights.trunc)
-
-max(d1p$ipwBG)
+  mutate(ipwCH = dch$weights.trunc)
+
 max(d1p$ipwCH)
-```
-
-```{r}
-
-ipb_bg<- geeglm(
-  anx_s ~ bg_s,
-  data = d1p,  
-  id = country,
-  corstr = "exchangeable",
-  weights = ipwBG
-)
-
-ipb_ch<- geeglm(
-  anx_s ~  ch_s,
+
+# weighted for religious only
+
+dchr <- ipw::ipwpoint(
+  exposure = ch_s,
+  family = "gaussian",
+  numerator = ~ 1,
+  denominator = ~ age_s + edu_s + male_2 + ses_s + country,
+  trunc = 0.01,
+  data = as.data.frame(d4r)
+)
+
+dr <- d4r %>% 
+  mutate(ipwCH = dchr$weights.trunc)
+
+# fine, well below 10
+max(dr$ipwCH)
+
+dr$ch_ord <- factor(dr$ch, ordered=TRUE)
+table(dr$ch_ord)
+
+levels(dr$ch_ord)<-c("never","rarely","1xpy","holydays","1xpm","1xpw",">1xpw")
+
+levels(dr$ch_ord)
+is.ordered(dr$ch_ord)
+
+#saveRDS(dr, here::here("mods","dr"))
+
+# saveRDS(d1p, here::here(
+#   "mods","d1p"
+# ))
+
+b_orR <- brm(
+  anx_s|weights(ipwCH) ~ mo(ch_ord) + (1 + mo(ch_ord) | country),
+  data = dr,
+   prior = c(
+    prior(student_t(3, 0, .5), class = Intercept),
+    prior(normal(-.05,.1), class = "b"),
+    prior(dirichlet(1,1,1,1,1,1), class = "simo", coef = moch_ord1)),
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  backend = "cmdstanr",
+  file = here::here("mods","b_orR")
+)
+
+summary(b_orR)
+
+prior_summary(b_orR)
+
+church_plot_R <-plot(ggeffects::ggemmeans(b_or, terms = "ch_ord")) +   xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + labs(title = "Baysian Multi-level Model",
+subtitle = "Inverse Probability Weighting") + scale_y_continuous(limits = c(-.5,.5))+ 
+  theme_classic()
+church_plot_R
+
+### EEMEANS
+ b_orR %>% 
+  emmeans(~ ch_ord,
+          at = list(ch_ord = c("never",">1xpw"),
+                    country = NA),
+          epred = TRUE) %>% 
+  contrast(method = "revpairwise")
+
+effect_draws_b_orR <-  b_orR %>% 
+  emmeans(~ ch_ord,
+         at = list(ch_ord = c("never",">1xpw"),
+                    country = NA),
+          epred = TRUE) %>% 
+  contrast(method = "revpairwise") %>%
+  gather_emmeans_draws()
+
+effect_draws_b_orR %>% mean_hdi()
+
+#SE = (upper limit – lower limit) / 3.92.
+
+# STANDARD ERROR
+Roa <- effect_draws_b_orR %>% mean_hdi() 
+(Roa$.upper - Roa$.lower)/3.92 #
+
+# SE = .0.06751772
+
+library(EValue)
+evalues.OLS(-0.208, se = 0.06751772, sd = 1, delta = 1)
+#             point    lower     upper
+# RR       0.8275548 0.733842 0.9332348
+# E-values 1.7101772       NA 1.3484170
+bias_plot(0.9332348, xmax = 6)
+
+average_marginal_effectR <- ggplot(effect_draws_b_orR, aes(x = .value)) +
+    stat_halfeye(fill = "dodgerblue") +
+  labs(x = "Average marginal effect of church attenance on anxiety", y = "Density",
+       subtitle = "Marginal effect (never vs. 1xpw): continuous model") +
+  theme_classic() +
+       scale_x_continuous(limits=c(-.6,.3)) +
+  theme(legend.position = "top")
+
+average_marginal_effectR
+
+grand_mean_ameR <- b_orR %>% 
+  emmeans(~ ch_ord,
+          at = list(ch_ord = c("never",">1xpw")),
+          epred = TRUE, re_formula = NA) %>% 
+  contrast(method = "revpairwise") %>% 
+  gather_emmeans_draws()
+
+plot_grand_mean_ameR <- ggplot(grand_mean_ameR, 
+                                   aes(x = .value)) +
+  stat_halfeye(fill = palette_okabe_ito(order = 5)) +
+  labs(x = "Average marginal effect of church attenance on anxiety", y = "Density",
+       subtitle = "Marginal effect (never vs. 1xpw): monotonic model") +
+  theme_classic() +
+    scale_x_continuous(limits=c(-.6,.3)) +
+   theme(legend.position = "bottom",
+         legend.title = element_text(size = 5),
+          legend.text = element_text(size = 2)) 
+plot_grand_mean_ameR
+
+grand_mean_autonomy_distR <- b_orR %>% 
+  epred_draws(newdata = expand_grid(ch_ord = c("never",
+                                               #"rarely","1xpy","holydays","1xpm","1xpw",
+                                               ">1xpw")), 
+              re_formula = NA)
+
+plot_grand_mean_ordR <- ggplot(grand_mean_autonomy_distR, 
+                                   aes(x = .epred, 
+                                       fill = ch_ord)) +
+  scale_fill_okabe_ito() +
+    stat_halfeye(slab_alpha = 0.7) +
+    scale_x_continuous(limits=c(-.6,.4)) +
+  labs(x = "Predicted anxiety (standardised)", y = "Density",
+       fill = "Church frequency",
+       subtitle = "Posterior predictions: monotonic model") +
+  theme_classic() +
+     theme(legend.position = "bottom",
+         legend.title = element_text(size = 12),
+          legend.text = element_text(size = 8))
+
+plot_grand_mean_ameR + plot_grand_mean_ordR + plot_annotation(tag_levels = "a",  title = "Religious only model")
+yola
+
+ggsave(
+  plot_grand_mean_ameR,
+  path = here::here("figs"),
+  width = 12,
+  height = 6,
+  units = "in",
+  filename = "plot_grand_mean_ameR.jpg",
+  device = 'jpeg',
+  limitsize = FALSE,
+  dpi = 1200
+)
+
+```
+
+## include religion
+
+```{r}
+library(splines)
+
+ipb_ch_ln_no_mediation <- geeglm(
+  anx_s ~ ch,
   data = d1p,
@@
 
-model_parameters(ipb_bg)
-
+## Model misspecification 
+ipb_ch_ln_mediation <- geeglm(
+  anx_s ~ ch + religious,
+  data = d1p,
+  id = country,
+  corstr = "exchangeable",
+  weights = ipwCH
+)
+saveRDS(ipb_ch_ln_mediation, here::here("mods","ipb_ch_ln_mediation"))
+saveRDS(ipb_ch_ln_no_mediation, here::here("mods","ipb_ch_ln_no_mediation"))
+
+model_parameters(ipb_ch_ln_mediation)
+model_parameters(ipb_ch_ln_no_mediation)
+
+plot(ggeffects::ggemmeans(ipb_ch_ln_mediation, terms = "ch"), add.data = TRUE)
+
+ipb_ch<- geeglm(
+  anx_s ~  bs(ch),
+  data = d1p,  
+  id = country,
+  corstr = "exchangeable",
+  weights = ipwCH
+)
+saveRDS(ipb_ch, here::here("mods","ipb_ch"))
 model_parameters(ipb_ch)
 
+plot(ggeffects::ggemmeans(ipb_ch, terms = "ch[all]"), add.data = TRUE
+     )
+
+## Rel only
+
+library(splines)
+ipb_ch_lnR<- geeglm(
+  anx_s ~ ch,
+  data = dr,
+  id = country,
+  corstr = "exchangeable",
+  weights = ipwCH
+)
+#saveRDS(ipb_ch_lnR, here::here("mods","ipb_ch_lnR"))
+model_parameters(ipb_ch_lnR)
+
+plot(ggeffects::ggemmeans(ipb_ch_lnR, terms = "ch"), add.data = TRUE
+     )
+
+ipb_chR<- geeglm(
+  anx_s ~  bs(ch),
+  data = dr,  
+  id = country,
+  corstr = "exchangeable",
+  weights = ipwCH
+)
+saveRDS(ipb_chR, here::here("mods","ipb_ch"))
+model_parameters(ipb_chR)
+
+plot(ggeffects::ggemmeans(ipb_chR, terms = "ch[all]"), add.data = TRUE
+     )
+
+# model_parameters(ipb_bg)
+
+abs(min(d1p$ch_s) - max(d1p$ch_s))
+# 3.064032
+
+# sensitivity
+evalues.OLS(0.04, se = 0.01, sd = 1, delta = 1.4)
+
+bias_plot(1.3009531, xmax = 6)
+
+### Religous Only
+
+library(splines)
+r_gee <- geeglm(
+  anx_s ~ ch, # retain scale
+  data = dr,
+  id = country,
+  corstr = "exchangeable",
+  weights = ipwCH
+)
+saveRDS(r_gee, here::here("mods","r_gee"))
+
+hist(dr$ch)
+
+# model_parameters(ipb_bg)
+
+model_parameters(r_gee) # no effect
+
+min(dr$ch_s) + max(dr$ch_s)
+
+# sensitivity
 library(EValue)
-
-# sensitivity = Church
-evalues.OLS(-0.08, se = 0.01, sd = 1, delta = 1)
-
-# sensitivity
-evalues.OLS(0.04, se = 0.01, sd = 1, delta = 1)
-
-bias_plot(1.3009531, xmax = 6)
+evalues.OLS(0.2238435, se = 0.01, sd = 1, delta = 1.4)
+
+#bias_plot(0.2238435, xmax = 6)
 
@@
 ```{r}
-
-ipb_bg_noweights <- geeglm(
-  anx_s ~ bg_s + ch_s + age_s + edu_s + male_2 + ses_s,
+# 
+# ipb_bg_noweights <- geeglm(
+#   anx_s ~ bg_s + ch_s + age_s + edu_s + male_2 + ses_s,
+#   data = d1p,  
+#   id = country,
+#   corstr = "exchangeable"
+# )
+# saveRDS(ipb_bg_noweights, here::here("mods","ipb_bg_noweights"))
+#
+
+ipb_ch_noweights <- geeglm(
+  anx_s ~ bs(ch_s) + age_s + edu_s + male_2 + ses_s,
   data = d1p,  
@@
 )
-
-ipb_ch_noweights <- geeglm(
-  anx_s ~ ch_s + age_s + edu_s + male_2 + ses_s,
+saveRDS(ipb_ch_noweights, here::here("mods","ipb_ch_noweights"))
+
+msm_lmer <- lmer(
+  anx_s ~ bs(ch_s) + age_s + edu_s + male_2 + ses_s +  (1+ ch_s|country),
   data = d1p, 
-  id = country,
-  corstr = "exchangeable"
-)
-
-model_parameters(ipb_bg_noweights)
-model_parameters(ipb_bg)
+)
+saveRDS(msm_lmer, here::here("mods","msm_lmer"))
+
+msm_lmer_wg <- lmer(
+  anx_s ~ bs(ch_s) + (1 + ch_s|country),
+  weights = ipwCH,
+  data = d1p
+)
+saveRDS(msm_lmer_wg, here::here("mods","msm_lmer_wg"))
+
+# model_parameters(ipb_bg_noweights)
+# model_parameters(ipb_bg_noweights)
+# model_parameters(ipb_bg)
+# model_parameters(ipb_bg)
 
@@
 model_parameters(ipb_ch)
-```
-
-We note that the multi-level models give different estimates.
-
-### GlMM TMB BELIEF
-
-```{r}
-library(glmmTMB)
-m<-10
-modelsb<-NULL
-for(i in 1:m) {
-  modelsb[[i]]  <- glmmTMB(anx_s ~ bg_s + ch_s + age_s + edu_s + male_2 + ses_s + 
-                            (1+ bg_s|country), 
-                              data = out$imputations[[i]],
-                              family =)
-}
-
-parameters::pool_parameters(modelsb)
-
-pred <- ggeffects::ggemmeans(modelsb[[1]], terms = c("bg_s[all]"))
+model_parameters(ipb_ch)
+model_parameters(msm_lmer)
+model_parameters(msm_lmer_wg)
+
+pred <- ggeffects::ggemmeans(msm_lmer_wg, terms = c("ch_s[all]"))
 pred
@@
     # add.data=T,dot.alpha=.1, 
-     facets = F)  + scale_y_continuous(limits=c(-.2,.2)) + theme_classic() + ggtitle("") # title to be suppled
-```
-
-### GlMM TMB CHURCH
-
-```{r}
-library(glmmTMB)
-m<-10
-modelsC<-NULL
-for(i in 1:m) {
-  modelsC[[i]]  <- glmmTMB(anx_s ~  ch_s + age_s + edu_s + male_2 + ses_s + 
-                            (1+ bg_s|country), 
-                              data = out$imputations[[i]],
-                              family =)
-}
-
-parameters::pool_parameters(modelsC)
-
-pred <- ggeffects::ggemmeans(modelsC[[1]], terms = c("ch_s[all]"))
-pred
-
-plot(pred,
-     add.data=T,dot.alpha=.05, 
-     facets = F)  + theme_classic() + ggtitle("") # title to be suppled
-```
-
-### Test implied conditional independence
-
-```{R}
-m<-10
-models<-NULL
-for(i in 1:m) {
-  models[[i]] <- lmer(anx_s ~ 
-                        bg_s + 
-                        ch_s + 
-                        age_s + 
-                        edu_s + 
-                        male_2 + 
-                        ses_s + 
-                        (1+ bg_s|country ), data=out$imputations[[i]])
-}
-
-parameters::pool_parameters(models,
-                            component = "marginal")
-
-pred <- ggeffects::ggemmeans(models[[1]], terms = c("bg_s[all]"))
-pred
-
-plot(pred,
-    # add.data=T,dot.alpha=.1, 
-     facets = F)  + scale_y_continuous(limits=c(-.2,.2)) + theme_classic() + ggtitle("") # title to be suppled
-
-### Model
-
-listout <- list(
-  out$imputations$imp1,
-  out$imputations$imp2,
-  out$imputations$imp3,
-  out$imputations$imp4,
-  out$imputations$imp5,
-  out$imputations$imp6,
-  out$imputations$imp7,
-  out$imputations$imp8,
-  out$imputations$imp9,
-  out$imputations$imp10
-)
-#saveRDS(listout, here::here("mods", "listout"))
-listout<- readRDS( here::here("mods", "listout"))
-
-m_bg <- brm_multiple( 
-  anx_s ~
-    bg_s +
-    ch_s +
-    age_s +
-    edu_s +
-    male_2 +
-    ses_s +
-    (1 + bg_s | country),
-  data = listout,
-  seed = 1234,
-  warmup = 1000,
-  iter = 2000,
-  chains = 4,
-  backend = "cmdstanr", 
-  threads = threading(4),
-  file = here::here("mods", "m_bg")
-)
-
-parameters::model_parameters(m_bg, test = "p_direction")
-brms::ranef(m_bg)
-
-parameters::parameters(
-  m_bg,
-  test = "pd",
-  priors = FALSE, 
-  effects = "all",
-  drop = TRUE,
-  group_level = FALSE,
-  diagnostic = "Rhat") %>%
-    mutate_if(is.numeric, ~ round(., 2)) %>%
-    mutate_all(funs(str_replace(., "b_", ""))) %>%
-    kable(booktabs = T, "latex", caption =  "Parameter estimates for model of Belief on Anxiety") %>%
-    print()
-
-```
+     facets = F)  + 
+  scale_y_continuous(limits=c(-.7,.7)) + 
+  theme_classic() + ggtitle("") # title to be suppled
+
+```
+
+We note that the multi-level models give different estimates.
 
@@
 ```{r}
-
 listout <- readRDS(here::here("mods","listout"))
@@
   backend = "cmdstanr",
+  file=  here::here("mods", "modmc.rds"),
   threads = threading(4)
@@
 summary(modmc)
+
+```
+## Bayesian wieghted regression
+
+```{r}
+listout <- readRDS(here::here("mods","listout"))
+modmc_wt_linear <- brm(
+  anx_s|weights(ipwCH) ~ ch_s + (1 + ch_s | country),
+  data = d1p,
+    prior = c(
+    prior(normal(0, .1), class = Intercept),
+    prior(normal(-.1,.1), class = b)),
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  backend = "cmdstanr",
+  file = here::here("mods","modmc_wt_linear")
+)
+#saveRDS(modmc, here::here("mods", "modmc.rds"))
+summary(modmc_wt_linear)
+
+prior_summary(modmc_wt_linear)
+summary(modmc_wt_linear)
+
+library(brms)
+library(splines)
+
+listout <- readRDS(here::here("mods","listout"))
+modmc_wt_priorStrong <- brm(
+  anx_s|weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
+  data = d1p,
+    prior = c(
+    prior(student_t(3, 0,.25), class = Intercept),
+    prior(normal(-.05, .1), class = b)),
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  backend = "cmdstanr",
+  file = here::here("mods","modmc_wt_priorStrong")
+)
+
+# prior summary
+
+prior_summary(modmc_wt_priorStrong)
+
+##### TeST PRIORS
+modmc_wt_priorStrong_TEST <- brm(
+  anx_s|weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
+  data = d1p,
+    prior = c(
+    prior(student_t(3, 0,.25), class = Intercept),
+    prior(normal(-.05, .1), class = b)),
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  sample_prior = "only",
+  backend = "cmdstanr",
+ file = here::here("mods","modmc_wt_priorStrong")
+)
+
+## WEAK PRIOR
+modmc_wt_prior_weak <- brm(
+  anx_s|weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
+  data = d1p,
+  prior = c(prior(normal(0, 1), class = b)),
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  sample_prior = "only",
+  backend = "cmdstanr"#,
+ # file = here::here("mods","modmc_wt_prior")
+)
+```
+
+```{r, cache = TRUE}
+library(viridis)
+
+# make new data
+minchur<- min(d1p$ch_s)
+maxchur<- max(d1p$ch_s)
+
+# difference
+abs(maxchur-minchur)
+
+library(emmeans)
+grand_mean
+
+p0<-brms::conditional_effects(modmc_wt_priorStrong, "ch_s", 
+                              spaghetti = T,
+                              ndraws = 1000,
+                              categorical = F,
+                              prob = .95,
+                              method = 'fitted')
+
+church_plot <- plot(p0, plot = FALSE)[[1]]   +  xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + labs(title = "Baysian Multi-level Model",
+subtitle = "Inverse Probability Weighting") + scale_y_continuous(limits = c(-.5,.5))+ 
+  theme_classic()
+church_plot
+
+ipw_gee <- plot(pred,
+    #add.data=T,dot.alpha=.1, 
+     facets = F)  + 
+  scale_y_continuous(limits=c(-.1,.1)) + xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + 
+  theme_classic() +labs(title = "Generalized Estimating Equations", 
+  subtitle = "Inverse Probability Weighting") + scale_y_continuous(limits = c(-.5,.5))
+ipw_gee
+church_plot
+model_plot <- ipw_gee + church_plot + church_plot_test_weak + church_plot_test_strong
+model_plot <- model_plot + plot_annotation(tag_levels = "a")
+
+model_plot
+
+ggsave(
+  model_plot,
+  path = here::here("figs"),
+  width = 10,
+  height = 5,
+  units = "in",
+  filename = "model_plot.jpg",
+  device = 'jpeg',
+  limitsize = FALSE,
+  dpi = 1200
+)
+
+library(brmsmargins)
+modmc_wt_priorStrong
+
+max(d1p$ch_s)#-0.8321022,2.23193
+
+h <-.001
+ames <- brmsmargins(
+  object = modmc_wt_priorStrong,
+            at = data.frame(ch_s= c(-0.8321022,2.23193)),
+            contrasts = cbind("AME x" = c(0.8321022/ h, 2.23193/ h)),
+            effects = "integrateoutRE")
+                              )
+ames$Summary
+
+ames_st <- brmsmargins(
+  object = modmc_wt_priorStrong, 
+            at = data.frame(ch_s = c(-0.8321022,2.23193)),
+            contrasts = cbind("AME x" = c(-0.8321022/ h, 2.23193/ h)),
+            effects = "integrateoutRE", CI = 0.95)
+ames_st$ContrastSummary
+knitr::kable(ames$Summary, digits = 3, "latex")
+
+ames_ln <- brmsmargins(
+  object = modmc_wt_linear, 
+            at = data.frame(ch_s = c(-0.8321022,2.23193)),
+            contrasts = cbind("AME x" = c(-0.8321022/ h, 2.23193/ h)),
+            effects = "integrateoutRE", CI=.95)
+ames_ln$Summary
+knitr::kable(ames$Summary, digits = 3, "latex")
+
+### EEMEANS
+ modmc_wt_priorStrong %>% 
+  emmeans(~ ch_s,
+          at = list(ch_s = c(-0.8321022,2.23193),
+                    country = NA),
+          epred = TRUE) %>% 
+  contrast(method = "revpairwise")
+
+effect_draws_cn <-  modmc_wt_priorStrong %>% 
+  emmeans(~ ch_s,
+          at = list(ch_s = c(-0.8321022,2.23193),
+                    country = NA),
+          epred = TRUE) %>% 
+  contrast(method = "revpairwise") %>% 
+  gather_emmeans_draws()
+
+# marginal mean 
+effect_draws_cn %>% mean_hdi()
+
+# marginal mean 
+effect_draws_cn %>% mean_hdi() 
+effect_draws_cn
+
+average_marginal_effect <- ggplot(effect_draws_cn, aes(x = .value)) +
+    stat_halfeye(fill = "dodgerblue") +
+  labs(x = "Average marginal effect of church attenance on anxiety", y = "Density",
+       subtitle = "Marginal effect (never vs. 1xpw): continuous model") +
+  theme_classic() +
+       scale_x_continuous(limits=c(-.6,.3)) +
+  theme(legend.position = "top")
+
+average_marginal_effect
+
+grand_mean <- modmc_wt_priorStrong %>% 
+  epred_draws(newdata = expand_grid(ch_s = seq(-0.8321022, 2.23193, by = 0.05)), 
+              re_formula = NA)
+
+  plot_grand_mean <- ggplot(grand_mean, 
+                                 aes(x = ch_s, y = .epred)) +
+  stat_lineribbon() +
+  scale_fill_brewer(palette = "Blues") +
+ # scale_fill_viridis_d(palette = "Blues") +
+  labs(x = "Frequency of religious service attendance (SD)", 
+       y = "Predicted anxiety (SD)",
+       fill = "95 HPD interval",
+       subtitle = "Posterior predictions: continuous model") +
+     theme_classic() +
+  theme(legend.position = "bottom",
+        legend.title = element_text(size = 10),
+          legend.text = element_text(size = 8))
+
+plot_grand_mean
+plot_result <- average_marginal_effect + plot_grand_mean +  plot_annotation(tag_levels = "a")
+plot_result
+
+ggsave(
+  plot_result,
+  path = here::here("figs"),
+  width = 16,
+  height = 9,
+  units = "in",
+  filename = "plot_result.jpg",
+  device = 'jpeg',
+  limitsize = FALSE,
+  dpi = 1200
+)
+
+summary(modmc_wt_priorStrong)
+summary(modmc_wt_linear)
+
+#
+
+library(EValue)
+summary(ipb_ch_ln)
+summary(ipb_ch)
+summary(modmc_wt_linear)
+
+# sensitivity = Church
+#yola
+
+#SE = (upper limit – lower limit) / 3.92.
+
+ta <- effect_draws_cn %>% mean_hdi() 
+ta
+(ta$.upper - ta$.lower)/3.92 #
+
+evalues.OLS(-0.147 , se = 0.05607611, sd = 1, delta = 1.1)
+# 
+#              point     lower     upper
+# RR       0.8631671 0.7734115 0.9633389
+# E-values 1.5870732        NA 1.2368137
+
+# multibias
+multi_evalue(est = RR( 0.8631671), biases = selection())
+
+#
+bias_plot(0.8631671, xmax=5)
+
+multi_bound(multi_bias(confounding()),
+                RRAUc = 2.2, RRUcY = 1.7)
+```
+
+## Scaled 0-1
+
+```{r, cache = TRUE}
+d1p<-readRDS(here::here("mods","d1p"))
+
+dchr_sc <- ipw::ipwpoint(
+  exposure = ch,
+  family = "gaussian",
+  numerator = ~ 1,
+  denominator = ~ age_s + edu_s + male_2 + ses_s + country,
+  trunc = 0.01,
+  data = as.data.frame(d1p)
+)
+
+dp <- d1p %>% 
+  mutate(ipwCH = dchr_sc$weights.trunc)
+
+# fine, well below 10
+max(dp$ipwCH)
+
+```
+
+# Montonic effects
+
+```{r}
+# bayesian monotonic effects
+
+dp$ch_ord <- factor(dp$ch, ordered=TRUE)
+table(dp$ch_ord)
+
+levels(dp$ch_ord)<-c("never","rarely","1xpy","holydays","1xpm","1xpw",">1xpw")
+
+levels(dp$ch_ord)
+is.ordered(dp$ch_ord)
+
+#saveRDS(dp, here::here("mods","dp"))
+dp<-readRDS(here::here("mods","dp"))
+```
+
+## ORDINAL MODEL
+
+```{r}
+## Bayesian wieghted regression 
+b_or <- brm(
+  anx_s|weights(ipwCH) ~ mo(ch_ord) + (1 + mo(ch_ord) | country),
+  data = dp,
+   prior = c(
+    prior(student_t(3, 0, .5), class = Intercept),
+    prior(normal(-.05,.1), class = "b"),
+    prior(dirichlet(1,1,1,1,1,1), class = "simo", coef = moch_ord1)),
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  backend = "cmdstanr",
+  file = here::here("mods","b_or")
+)
+
+summary(b_or)
+
+prior_summary(b_or)
+summary(b_or)
+
+church_plot <-plot(ggeffects::ggemmeans(b_or, terms = "ch_ord")) +   xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + labs(title = "Baysian Multi-level Model",
+subtitle = "Inverse Probability Weighting") + scale_y_continuous(limits = c(-.5,.5))+ 
+  theme_classic()
+church_plot
+
+# pord<-brms::conditional_effects(b_or, "ch_ord", 
+#                               spaghetti = T,
+#                               ndraws = 1000,
+#                               categorical = F,
+#                               prob = .9,
+#                               method = 'fitted')
+# 
+# 
+# church_plot <- plot(pord, plot = FALSE)[[1]]   +  xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + labs(title = "Baysian Multi-level Model",
+# subtitle = "Inverse Probability Weighting") + scale_y_continuous(limits = c(-.5,.5))+ 
+#   theme_classic()
+# church_plot
+library(tidybayes)
+
+summary(b_or)
+### EEMEANS
+ b_or %>% 
+  emmeans(~ ch_ord,
+          at = list(ch_ord = c("never",">1xpw"),
+                    country = NA),
+          epred = TRUE) %>% 
+  contrast(method = "revpairwise")
+
+  effect_draws <-  b_or %>% 
+  emmeans(~ ch_ord,
+         at = list(ch_ord = c("never",">1xpw"),
+                    country = NA),
+          epred = TRUE) %>% 
+  contrast(method = "revpairwise") %>%
+  gather_emmeans_draws()
+
+# linear model, for EVALUES
+effect_draws_or <-  b_or %>% 
+  emmeans(~ ch_ord,
+         at = list(ch_ord = c("never",">1xpw"),
+                    country = NA),
+          epred = TRUE) %>%
+  contrast(method = "revpairwise") %>% 
+  gather_emmeans_draws()
+effect_draws_or
+# marginal mean 
+effect_draws_or %>% mean_hdi()
+
+# calculate standard error
+# 
+# se <- function(x) sqrt(var(x) / length(x))
+# 
+# summary(b_or)
+# se(effect_draws_ln$.value)
+library(tidybayes)
+library(ggokabeito)   # Neat accessible color palette
+# emmeans()-based difference across autonomy
+grand_mean_ame <- b_or %>% 
+  emmeans(~ ch_ord,
+          at = list(ch_ord = c("never",">1xpw")),
+          epred = TRUE, re_formula = NA) %>% 
+  contrast(method = "revpairwise") %>% 
+  gather_emmeans_draws()
+
+plot_grand_mean_ame <- ggplot(grand_mean_ame, 
+                                   aes(x = .value)) +
+  stat_halfeye(fill = palette_okabe_ito(order = 5)) +
+  labs(x = "Average marginal effect of church attenance on anxiety", y = "Density",
+       subtitle = "Marginal effect (never vs. 1xpw): monotonic model") +
+  theme_classic() +
+    scale_x_continuous(limits=c(-.6,.3)) +
+   theme(legend.position = "bottom",
+         legend.title = element_text(size = 5),
+          legend.text = element_text(size = 2))
+
+plot_grand_mean_ame
+
+# plot_grand_mean <- ggplot(grand_mean, 
+#                                  aes( x = .epred, fill=(ch_ord))) +
+#   scale_fill_brewer(palette = "Blues") +
+#     stat_halfeye(slab_alpha = 0.8) +
+# 
+#   labs(x = "Church attendence frequency", 
+#        y = "Marginal Density",
+#        fill = "") +
+#   theme_classic() +
+#   theme(legend.position = "top")
+# 
+# plot_grand_mean
+
+grand_mean_autonomy_dist <- b_or %>% 
+  epred_draws(newdata = expand_grid(ch_ord = c("never","rarely","1xpy","holydays","1xpm","1xpw",">1xpw")), 
+              re_formula = NA)
+
+plot_grand_mean_ord <- ggplot(grand_mean_autonomy_dist, 
+                                   aes(x = .epred, 
+                                       fill = ch_ord)) +
+  scale_fill_okabe_ito() +
+    stat_halfeye(slab_alpha = 0.7) +
+    scale_x_continuous(limits=c(-.6,.4)) +
+  labs(x = "Predicted anxiety (standardised)", y = "Density",
+       fill = "Church frequency",
+       subtitle = "Posterior predictions: monotonic model") +
+  theme_classic() +
+     theme(legend.position = "bottom",
+         legend.title = element_text(size = 12),
+          legend.text = element_text(size = 8))
+
+plot_grand_mean_ord
+
+plot_result_ord <- plot_grand_mean_ame + plot_grand_mean_ord+  plot_annotation(tag_levels = "a")
+plot_result_ord
+
+ggsave(
+  plot_result_ord,
+  path = here::here("figs"),
+  width = 12,
+  height = 6,
+  units = "in",
+  filename = "plot_result_ord.jpg",
+  device = 'jpeg',
+  limitsize = FALSE,
+  dpi = 1200
+)
+
+plot_result_full <- (average_marginal_effect + plot_grand_mean)/
+ (plot_grand_mean_ame + plot_grand_mean_ord) + plot_annotation(tag_levels = "a")
+plot_result_full
+
+ggsave(
+  plot_result_full,
+  path = here::here("figs"),
+  width = 12,
+  height = 6,
+  units = "in",
+  filename = "plot_result_full.jpg",
+  device = 'jpeg',
+  limitsize = FALSE,
+  dpi = 1200
+)
+
+# Obtain evalues
+
+# marginal mean 
+effect_draws_or %>% mean_hdi()
+
+#SE = (upper limit – lower limit) / 3.92.
+
+oa <- effect_draws_or %>% mean_hdi() 
+(oa$.upper - oa$.lower)/3.92 #
+
+# SE = .075
+
+library(EValue)
+evalues.OLS(-0.187 , se = .06733407, sd = 1, delta = 1)
+
+#             point     lower     upper
+# RR       0.8435214 0.7482451 0.9509295
+# E-values 1.6544616        NA 1.2845523
+
+#SE = (upper limit – lower limit) / 3.92.
+
+ta <- effect_draws_cn %>% mean_hdi() 
+ta
+(ta$.upper - ta$.lower)/3.92 #
+
+evalues.OLS(-0.147  , se =  0.05607611, sd = 1, delta = 1.1)
+#              point     lower     upper
+# RR       0.8631671 0.7734115 0.9633389
+# E-values 1.5870732        NA 1.2368137
+```
+
+### Test priors ordinal model
+
+```{r}
+##### TeST PRIORS
+b_ord_test_weak_priors <- brm(
+  anx_s|weights(ipwCH) ~ mo(ch_ord) + (1 + mo(ch_ord) | country),
+  data = dp,
+  prior = c(
+    prior(student_t(3, 0, 2.5), class = Intercept),
+    prior(normal(0,1), class = "b"),
+    prior(dirichlet(1,1,1,1,1,1), class = "simo", coef = moch_ord1)),
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  sample_prior = "only",
+  backend = "cmdstanr"
+)
+
+plot_ord_wk_priors<-brms::conditional_effects(b_ord_test_weak_priors, "ch_ord", 
+                              spaghetti = T,
+                              ndraws = 1000,
+                              categorical = F,
+                              prob = .95,
+                              method = 'fitted')
+
+test_or_wk<- plot(plot_ord_wk_priors, plot = FALSE)[[1]]   +  xlab("Service attendance (monotonic)") + ylab ("Anxiety (standardised)") + labs(title = "Check Priors: Baysian Mixed Effects Marginal Model", subtitle = "Sampling from weak priors only (ordinal) ") + scale_y_continuous(limits = c(-20,15))+
+  theme_classic()
+test_or_wk
+
+d1p<-readRDS(here::here("mods","d1p"))
+dp<-readRDS(here::here("mods","dp"))
+
+b_cont_test_st_priors <- brm(
+  anx_s|weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
+  data = d1p,
+    prior = c(
+    prior(student_t(3, 0,.25), class = Intercept),
+    prior(normal(-.05, .1), class = b)),
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  sample_prior = "only",
+  backend = "cmdstanr"#,
+ # file = here::here("mods","b_cont_test_st_priors")
+)
+
+con_st_priors<-brms::conditional_effects(b_cont_test_st_priors, "ch_s", 
+                              spaghetti = T,
+                              ndraws = 1000,
+                              categorical = F,
+                              prob = .95,
+                              method = 'fitted')
+
+test_con_st<- plot(con_st_priors, plot = FALSE)[[1]]   +  xlab("Service attendance (sd)") + ylab ("Anxiety (standardised)") + labs(title = "Check Priors: Baysian Mixed Effects Marginal Model", subtitle = "Sampling from weak priors only (continuous) ") +   scale_y_continuous(limits = c(-10,10))+
+  theme_classic()
+
+b_cont_test_wk_priors <- brm(
+  anx_s|weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
+  data = d1p,
+   prior = c(
+    prior(normal(0, 2.5), class = Intercept),
+    prior(normal(0, 1), class = b)),
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  sample_prior = "only",
+  backend = "cmdstanr"#
+)
+
+con_wk_priors<-brms::conditional_effects(b_cont_test_wk_priors, "ch_s", 
+                              spaghetti = T,
+                              ndraws = 1000,
+                              categorical = F,
+                              prob = .95,
+                              method = 'fitted')
+
+test_con_wk_priors<- plot(con_wk_priors, plot = FALSE)[[1]]   +  xlab("Service attendance (sd)") + ylab ("Anxiety (standardised)") + labs(title = "Check Priors: Baysian Mixed Effects Marginal Model", subtitle = "Sampling from weak priors only (continuous) ")   + scale_y_continuous(limits = c(-10,10))+
+  theme_classic()
+test_con_wk_priors
+
+## THIS IS IT.  TO MAKE WEAKER INCREASE INTERCEPT 
+b_or_test_st_priors4 <- brm(
+  anx_s|weights(ipwCH) ~ mo(ch_ord) + (1 + mo(ch_ord) | country),
+  data = dp,
+  seed = 1234,
+  prior = c(
+    prior(student_t(3, 0, .5), class = Intercept),
+    prior(normal(-.05,.1), class = "b"),
+    prior(dirichlet(1,1,1,1,1,1), class = "simo", coef = moch_ord1)),
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  sample_prior = "only",
+  backend = "cmdstanr"#,
+ # file = here::here("mods","b_or_test_st_priors4")
+)
+
+prior_summary(b_or_test_st_priors4)
+
+b_or_test_st_priors4<-brms::conditional_effects(b_or_test_st_priors4, "ch_ord", 
+                              spaghetti = T,
+                              ndraws = 1000,
+                              categorical = F,
+                              prob = .95,
+                              method = 'fitted')
+
+or_strong<- plot(b_or_test_st_priors4, plot = FALSE)[[1]]   +  xlab("Service attendance (monotonic)") + ylab ("Anxiety (standardised)") + labs(title = "Check Priors: Baysian Mixed Effects Marginal Model", subtitle = "Sampling from strong priors only (ordinal) ") + scale_y_continuous(limits = c(-20,15))+
+  theme_classic()
+or_strong
+
+#saveRDS(modmc, here::here("mods", "modmc.rds"))
+
+library(patchwork)
+
+test_priors <-(test_or_wk + or_strong )/(test_con_wk_priors+test_con_st) + plot_annotation(tag_levels = "a")
+
+test_priors
+
+ggsave(
+  test_priors,
+  path = here::here("figs"),
+  width = 12,
+  height = 12,
+  units = "in",
+  filename = "test_priors.jpg",
+  device = 'jpeg',
+  limitsize = FALSE,
+  dpi = 1200
+)
+
+```
+
+```{r}
+library(brms)
+library(splines)
+
+listout <- readRDS(here::here("mods","listout"))
+modmc_wt_priorStrong <- brm(
+  anx_s|weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
+  data = d1p,
+  prior = c(
+    prior(normal(0, .1), class = Intercept),
+    prior(normal(-.1,.1), class = b)),
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  backend = "cmdstanr",
+  file = here::here("mods","modmc_wt_priorStrong")
+)
+
+##### TeST PRIORS
+modmc_wt_priorStrong_TEST <- brm(
+  anx_s|weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
+  data = d1p,
+  prior = c(
+    prior(normal(0, .1), class = Intercept),
+    prior(normal(-.1,.1), class = b)),
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  sample_prior = "only",
+  backend = "cmdstanr"#,
+ # file = here::here("mods","modmc_wt_priorStrong")
+)
+#saveRDS(modmc, here::here("mods", "modmc.rds"))
+summary(modmc_wt_prior)
+prior_summary(modmc_wt_prior)
+summary(modmc_wt_priorStrong)
+
+```
+
+### Checks:
+
+```{r}
+# YES
+pp_check(b_ln)
+```
+
+Trace plots:
+
+```{r, cache = TRUE}
+mcmc_plot(b_ln,
+          type = "trace")
 ```
@@
 set.seed(1234)
-y<- rnorm(1000,0,1)
-a<- rnorm(1000,0,1)
-s<- rnorm(1000, a -(2*y),4)
-
-dlbind <-data.frame(y,a,s)
-summary(lm(y ~  a + s, data = dlbind))
+n <- 10000 # n
+a <- rnorm(n, 0, 1) # rel service attendance
+y <- rnorm(n, -a*.2, 1) # rel service affects anxiety 
+bg <- rnorm(n, a, 1) # rel service affects belief 
+s <- rnorm(n, .2*y + .2*bg,1) # selection bias
+
+dbind <-data.frame(y,a,s, bg)
+
+summary(lm(s ~  a + bg, data = dbind))
+summary(lm(s ~  a , data = dbind))
+
+##
+
+set.seed(123)
+sim_fun = function() {
+n <- 10000 # n
+a <- rnorm(n, 0, 1) # rel service attendance
+y <- rnorm(n, -a*.2, 1) # rel service affects anxiety 
+l <- rnorm(n, a, 1) # rel service affects belief 
+s <- rnorm(n, .2*l - 5*y,1) # selection bias
+
+# simulate data
+simdat <-data.frame(y, a, s, l)
+
+  sim_s <- lm(s ~ a + l, data = simdat) # belief + church
+  sim_s  # confounded 
+  summary(sim_s)
+}
+
+# Replication 1000xs
+r_lm = replicate(1000, sim_fun(), simplify = FALSE )
+
+# extract simulations
+tab_sim<-purrr::map_dfr(r_lm, broom::tidy)
+
+# extract belief coefficients
+mout<-tab_sim %>%
+  dplyr::filter(term =="l")%>%
+  dplyr::mutate_if(is.numeric, round, 5)
+
+# calculate proportion of estimates greater than 0
+# all positive
+sum(mout$estimate > 0) / length(mout$estimate)
+
+# proportion of belief coefficient p.values < .001
+sum(mout$p.value <.001) / length(mout$p.value)
+
+# extract religious service coefficients
+tout<-tab_sim %>%
+  dplyr::filter(term =="a")%>%
+  dplyr::mutate_if(is.numeric, round, 5)
+
+# calculate proportion of estimates less than 0
+# all negative
+sum(tout$estimate < 0) / length(tout$estimate)
+
+# proportion of service coefficient p.values < .001
+sum(tout$p.value <.001) / length(tout$p.value)
+```
+
+```{r}
+set.seed(123)
+sim_fun = function() {
+n <- 10000 # n
+a <- rnorm(n, 0, 1) # rel service attendance
+y <- rnorm(n, -a*.2, 1) # rel service affects anxiety 
+bg <- rnorm(n, a, 1) # rel service affects belief 
+s <- rnorm(n, .2*y + .2*bg,1) # selection bias
+
+# simulate data
+simdat <-data.frame(y, a, s, bg)
+
+  sim_s <- lm(s ~  a, data = dbind) # belief + church
+  sim_s  # confounded 
+}
+
+```
+
+```{r}
+library(ggplot2)
+library(magick)
+library(patchwork)
+
+dag1 <- image_ggplot(image_read(here::here("figs","dg_1.tiff")),
+                     interpolate = T)
+dag1
+dag2 <- image_ggplot(image_read(here::here("figs","dg_2.tiff")), interpolate = T)
+dag2
+dag3 <- image_ggplot(image_read(here::here("figs","dg_3.tiff")), interpolate = T)
+dag3
+dag4 <- image_ggplot(image_read(here::here("figs","dg_4.tiff")),
+                     interpolate = T)
+dag4
+library(patchwork)
+daggraph <- dag1 + dag4 + dag3  / ( dag2) + 
+  plot_annotation(title = "Directed Acyclic Graphs (DAGs) clarify assumptions and hazards for causal identification", tag_levels = "a")
+
+ggsave(
+  daggraph,
+  path = here::here("figs"),
+  width = 10,
+  height = 5,
+  units = "in",
+  filename = "daggraph.jpg",
+  device = 'jpeg',
+  limitsize = FALSE,
+  dpi = 1200
+)
 ```

### 2022-03-03 analysis-religious-only.Rmd
updated analysis

@@
                 #   bg_s,
+                denomination,
                 wb_psychANXIETY)
 
+```
+
+table
+
+```{r}
+lt <-table1::table1(~  denomination, data = d1, transpose=F)
+lt
+lt <-table1::table1(~ ch + religious + age + male + edu + ses + denomination |country, data = d1, transpose=F)
+lt
+kable(as.data.frame(lt), booktabs=TRUE, "latex")
+
+nrow(dr)
 ```
@@
 max(dr$ipwCH)
-
+```
+
+## Tables
+
+```{r}
+lt <-table1::table1(~ religious|ch_ord, data = dp)
+
+kable(as.data.frame(lt), booktabs=TRUE, "latex")
+
+nrow(dr)
+```
+
+### Simulation
+
+```{r}
+#\begin{lstlisting}[language=R]
+
+set.seed(123)
+sim_fun_A = function() {
+  n <- 1000
+  U <- rnorm(n, 1) # Unmeasured distress 
+  A <- rnorm(n, 1) # religious service attendance,
+  M <- rnorm(n , A + U)# rel id
+  Y <- rnorm(n , - A*.2 + U + M*.1) #  distress reduced by attendance.
+
+# Religious id, caused by attendance + distress
+
+# simulate data
+simdat_A <- data.frame(
+  M = L, # rel id
+  A = A, # rel service
+  U = U, # distressing events (unmeasured)
+  Y = Y ) # outcome
+
+sim_A <- lm(Y ~ A + M, data = simdat_A)
+sim_A
+}
+
+# Replicate including the mediator
+r_lm_A <- NA
+r_lm_A = replicate(100, sim_fun_A(), simplify = FALSE )
+
+# average over outcomes: effect of religion is attenuated
+parameters::pool_parameters(r_lm_A)
+
+## Repeat but with no mediator in model
+set.seed(123)
+sim_fun_B = function() {
+  n <- 1000
+  U <- rnorm(n, 1) # Unmeasured distress 
+  A <- rnorm(n, 1) # religious service attendance,
+  M <- rnorm(n , A + U)# rel id + distress
+  Y <- rnorm(n , - A *.2 + U + M*.1) #  distress reduced by attendance.
+
+# simulate data
+simdat_B <- data.frame(
+  M = L, # rel id
+  A = A, # rel service
+  U = U, # distressing events (unmeasured)
+  Y = Y ) # outcome
+
+# Only exposure
+sim_B <- lm(Y ~ A, data = simdat_A)
+sim_B
+}
+
+# Replication 1000xs
+r_lm_B <- NA
+r_lm_B = replicate(100, sim_fun_B(), simplify = FALSE )
+
+# average over outcomes: effect of religion is attenuated
+parameters::pool_parameters(r_lm_B)
+
+#\end{lstlisting}
+```
+
+```{r}
 dr$ch_ord <- factor(dr$ch, ordered=TRUE)
@@
 
-plot_grand_mean_ameR + plot_grand_mean_ordR + plot_annotation(tag_levels = "a",  title = "Religious only model")
+graph_r<- plot_grand_mean_ameR + plot_grand_mean_ordR + plot_annotation(tag_levels = "a",  title = "Religious only model")
 yola
@@
 ggsave(
-  plot_grand_mean_ameR,
+  graph_r,
   path = here::here("figs"),
@@
   units = "in",
-  filename = "plot_grand_mean_ameR.jpg",
+  filename = "graph_r.jpg",
   device = 'jpeg',
@@
 )
+
+## Table
+parameters::parameters(
+  modmc_wt_priorStrong,
+  test = "pd",
+  component = "all",
+  priors = FALSE, 
+  ci_method = "hdi",
+#  effects = "fixed",
+  group_level = FALSE,
+  diagnostic = "Rhat") %>%
+    mutate_if(is.numeric, ~ round(., 3)) %>%
+    mutate_all(funs(str_replace(., "b_", ""))) %>%
+    mutate_all(funs(str_replace(., "bs", ""))) %>%
+    kable(booktabs = T, "latex", caption =  "Parameter estimates for continuous (spline) model of religious service on anxiety") %>%
+    print()
 
@@
 church_plot
-
+```
+
+## Tables
+
+```{r}
+# tab_or <- model_parameters(b_or,   verbose = F, ci_method = "hdi",test = c("pd"))
+# kbl(tab_or, "latex",booktabs=T)
+
+# sjstats::tidy_stan(tab_or)
+
+# library(broom.mixed)
+# tidy(
+#   b_or,
+#  # parameters = NA,
+#   effects = c("fixed", "ran_pars"),
+#   robust = FALSE,
+#   conf.int = TRUE,
+#   conf.level = 0.95,
+#   conf.method = c("HPDinterval"),
+#   fix.intercept = TRUE
+# )
+
+parameters::parameters(
+  b_or,
+  test = "pd",
+  component = "all",
+  priors = FALSE, 
+  ci_method = "hdi",
+#  effects = "fixed",
+  group_level = FALSE,
+  diagnostic = "Rhat") %>%
+    mutate_if(is.numeric, ~ round(., 3)) %>%
+    mutate_all(funs(str_replace(., "b_", ""))) %>%
+    mutate_all(funs(str_replace(., "_moch", ""))) %>%
+    kable(booktabs = T, "latex", caption =  "Parameter estimates for model of religious service on anxiety") %>%
+    print()
+```
+
+```{r}
 # pord<-brms::conditional_effects(b_or, "ch_ord", 
@@
 
+grand_mean_autonomy_dist2 <- b_or %>% 
+  epred_draws(newdata = expand_grid(ch_ord = c("never",">1xpw")), 
+              re_formula = NA)
+
 plot_grand_mean_ord <- ggplot(grand_mean_autonomy_dist, 
+                                   aes(x = .epred, 
+                                       fill = ch_ord)) +
+  scale_fill_okabe_ito() +
+    stat_halfeye(slab_alpha = 0.7) +
+    scale_x_continuous(limits=c(-.6,.4)) +
+  labs(x = "Predicted anxiety (standardised)", y = "Density",
+       fill = "Church frequency",
+       subtitle = "Posterior predictions: monotonic model") +
+  theme_classic() +
+     theme(legend.position = "bottom",
+         legend.title = element_text(size = 10),
+          legend.text = element_text(size = 8))
+
+plot_grand_mean_ord
+
+plot_grand_mean_ord2 <- ggplot(grand_mean_autonomy_dist2, 
                                    aes(x = .epred, 
@@
 
-plot_grand_mean_ord
-
-plot_result_ord <- plot_grand_mean_ame + plot_grand_mean_ord+  plot_annotation(tag_levels = "a")
-plot_result_ord
+plot_grand_mean_ord2
 
@@
 
-plot_result_full <- (average_marginal_effect + plot_grand_mean)/
- (plot_grand_mean_ame + plot_grand_mean_ord) + plot_annotation(tag_levels = "a")
-plot_result_full
+plot_result_ord2 <- plot_grand_mean_ame + plot_grand_mean_ord2+  plot_annotation(tag_levels = "a")
+plot_result_ord2
 
 ggsave(
-  plot_result_full,
+  plot_result_ord2,
   path = here::here("figs"),
@@
   units = "in",
+  filename = "plot_result_ord2.jpg",
+  device = 'jpeg',
+  limitsize = FALSE,
+  dpi = 1200
+)
+
+plot_result_full <- (average_marginal_effect + plot_grand_mean)/
+ (plot_grand_mean_ame + plot_grand_mean_ord) + plot_annotation(tag_levels = "a")
+plot_result_full
+
+ggsave(
+  plot_result_full,
+  path = here::here("figs"),
+  width = 10,
+  height =10,
+  units = "in",
   filename = "plot_result_full.jpg",
@@
   limitsize = FALSE,
-  dpi = 1200
+  dpi = 1000
 )
@@
 
-˝#\# Simulation tables
+˝## Simulation tables
 

### 2022-03-03 analysis-v-2.Rmd
updated analysis

@@
 # source("funs.R") # custom functions
+
 library("tidyverse")
@@
 library("WeightIt")# inverse probability weighting. 
+library("mice")
 rstan_options(auto_write=TRUE)
@@
 theme_set(theme_few())
+knitr::write_bib(c(.packages("geepack")))
+
+cite_packages()
 ```
@@
   )
+max(df$rel_EXTENTBELIEVEGOD)
+max(df$rel_CHURCHFREQ)
 
@@
 miss_graph
+miss_graph <-image_ggplot(miss_graph)
 # save graph
-ggsave(
-  dg1,
-  path = here::here("Figs"),
+str(miss_graph)
+
+ggsave(miss_graph,
+  path = here::here("figs"),
   width = 10,
@@
   units = "in",
-  filename = "dag_church.jpg",
+  filename = "miss_graph.jpg",
   device = 'jpeg',
@@
 library(mice)
-d2<-mice(d1)
-d3<-complete(d2)
-saveRDS(d3,here::here("mods","d3"))
+library(tidyverse)
+#d2<-mice(d1)
+#d3<-complete(d2)
+#saveRDS(d3,here::here("mods","d3"))
 d3 <-readRDS(here::here("mods","d3"))
@@
 
+d4
 dbl <- ipw::ipwpoint(
-  exposure = bg_g,
+  exposure = bg_s,
   family = "gaussian",
@@
 )
+#saveRDS(ipb_bg, here::here("mods","ipb_bg"))
 
@@
 )
+saveRDS(ipb_ch, here::here("mods","ipb_ch"))
 
@@
 )
+saveRDS(ipb_bg_noweights, here::here("mods","ipb_bg_noweights"))
 
@@
 )
-
+saveRDS(ipb_ch_noweights, here::here("mods","ipb_ch_noweights"))
+
+msm_lmer <- lmer(
+  anx_s ~ ch_s + age_s + edu_s + male_2 + ses_s +  (1+ ch_s|country),
+  data = d1p, 
+)
+saveRDS(msm_lmer, here::here("mods","msm_lmer"))
+
+msm_lmer_wg <- lmer(
+  anx_s ~ ch_s + (1 + ch_s|country),
+  weights = ipwCH,
+  data = d1p
+)
+saveRDS(msm_lmer_wg, here::here("mods","msm_lmer_wg"))
+
+model_parameters(ipb_bg_noweights)
 model_parameters(ipb_bg_noweights)
 model_parameters(ipb_bg)
+model_parameters(ipb_bg)
 
@@
 model_parameters(ipb_ch)
+model_parameters(ipb_ch)
+model_parameters(msm_lmer)
+model_parameters(msm_lmer_wg)
+
+pred <- ggeffects::ggemmeans(msm_lmer_wg, terms = c("ch_s[all]"))
+pred
+
+plot(pred,
+    # add.data=T,dot.alpha=.1, 
+     facets = F)  + 
+  scale_y_continuous(limits=c(-.7,.7)) + 
+  theme_classic() + ggtitle("") # title to be suppled
+
 ```
@@
 summary(modmc)
+
+```
+## Bayesian wieghted regression
+
+```{r}
+listout <- readRDS(here::here("mods","listout"))
+modmc_wt <- brm(
+  anx_s|weights(ipwCH) ~ ch_s + (1 + ch_s | country),
+  data = d1p,
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  backend = "cmdstanr",
+  file = here::here("mods","modmc_wt")
+)
+#saveRDS(modmc, here::here("mods", "modmc.rds"))
+summary(modmc_wt)
+
+```
+
+```{r, cache = TRUE}
+library(viridis)
+p0<-brms::conditional_effects(modmc_wt, "ch_s", 
+                              spaghetti = T,
+                              ndraws = 1000,
+                              categorical = F,
+                              prob = .95,
+                              method = 'fitted')
+
+church_plot <- plot(p0, plot = FALSE)[[1]]   +  xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + labs(title = "Baysian Multi-level Model",
+subtitle = "Inverse Probability Weighting") + scale_y_continuous(limits = c(-.5,.5))+ 
+  theme_classic()
+church_plot
+
+ipw_gee <- plot(pred,
+    # add.data=T,dot.alpha=.1, 
+     facets = F)  + 
+  scale_y_continuous(limits=c(-.1,.1)) + xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + 
+  theme_classic() +labs(title = "Generalized Estimating Equations", 
+  subtitle = "Inverse Probability Weighting") + scale_y_continuous(limits = c(-.5,.5))
+ipw_gee
+model_plot <- ipw_gee + church_plot
+model_plot + plot_annotation(tag_levels = "a")
+
+ggsave(
+  model_plot,
+  path = here::here("figs"),
+  width = 10,
+  height = 5,
+  units = "in",
+  filename = "model_plot.jpg",
+  device = 'jpeg',
+  limitsize = FALSE,
+  dpi = 1200
+)
+
+```
+
+### Checks:
+
+```{r}
+# YES
+pp_check(m1)
+```
+
+Trace plots:
+
+```{r, cache = TRUE}
+mcmc_plot(m1,
+          type = "trace")
 ```
@@
 summary(lm(y ~  a + s, data = dlbind))
+```
+
+```{r}
+library(ggplot2)
+library(magick)
+library(patchwork)
+
+dag1 <- image_ggplot(image_read(here::here("figs","dg_1.tiff")),
+                     interpolate = T)
+dag1
+dag2 <- image_ggplot(image_read(here::here("figs","dg_2.tiff")), interpolate = T)
+dag2
+dag3 <- image_ggplot(image_read(here::here("figs","dg_3.tiff")), interpolate = T)
+dag3
+dag4 <- image_ggplot(image_read(here::here("figs","dg_4.tiff")),
+                     interpolate = T)
+dag4
+library(patchwork)
+daggraph <- dag1 + dag4 + dag3  / ( dag2) + 
+  plot_annotation(title = "Directed Acyclic Graphs (DAGs) clarify assumptions and hazards for causal identification", tag_levels = "a")
+
+ggsave(
+  daggraph,
+  path = here::here("figs"),
+  width = 10,
+  height = 5,
+  units = "in",
+  filename = "daggraph.jpg",
+  device = 'jpeg',
+  limitsize = FALSE,
+  dpi = 1200
+)
 ```

### 2022-03-12 simulation.R
update

@@ -1,0 +1,57 @@
+
+
+sim_fun_A = function() {
+  n <- 1000
+  U <- rnorm(n, 1) # unmeasured distress 
+  A <- rnorm(n, 1) # religious service attendance,
+  M <- rnorm(n , A + U)# religious identification
+  Y <- rnorm(n , - A *.2 + U + M *.1) #  distress reduced by attendance.
+  
+  # Religious id, caused by attendance + distress
+  
+  # simulate data
+  simdat_A <- data.frame(
+    M = M, # rel identification
+    A = A, # rel service
+    U = U, # distressing events (unmeasured)
+    Y = Y )# outcome
+  
+  sim_A <- lm(Y ~ A + M, data = simdat_A)
+  sim_A
+}
+
+# Replicate including the mediator
+r_lm_A <- NA
+r_lm_A = replicate(100, sim_fun_A(), simplify = FALSE )
+
+
+# mediator attenuates true effect ~ .1 instead of ~.1
+parameters::pool_parameters(r_lm_A)
+
+## Repeat but with no mediator in model
+set.seed(123)
+sim_fun_B = function() {
+  n <- 1000
+  U <- rnorm(n, 1) # unmeasured distress 
+  A <- rnorm(n, 1) # religious service attendance,
+  M <- rnorm(n , A + U)# religious identification
+  Y <- rnorm(n , - A *.2 + U + M *.1) #  distress reduced by attendance.
+  
+  # Simulate data
+  simdat_B <- data.frame(
+    M = L, # rel identification
+    A = A, # rel service attendance
+    U = U, # distressing events (unmeasured)
+    Y = Y ) # outcome
+  
+  # Only model exposure
+  sim_B <- lm(Y ~ A, data = simdat_A)
+  sim_B
+}
+
+# Replication 100xs
+r_lm_B <- NA
+r_lm_B = replicate(100, sim_fun_B(), simplify = FALSE )
+
+# recover true effect ~ .2
+parameters::pool_parameters(r_lm_B)

### 2022-03-12 analysis-religious-final.Rmd
delete figs

--- 2022-03-03 analysis-religious-only.Rmd
@@
 library("easystats")
-library("ggplot2") #plots
-#library("sjstats") #test statistics
-#library("sjPlot")
-#library("sjmisc")
-##library("insight")
-#library("see")
-#library("performance")
 library("ggraph")
-#library("parameters")
 library("lme4")
-#library("psych") ## correlation matrix
 library("table1")
 library("ggeffects")
-#library("gghighlight")
-#library("texreg")
 library("kableExtra")
-#library("naniar") # missingness
+library("naniar") # missingness
 library("Amelia")
-#library("magrittr")
 library("patchwork")
@@
 library("here")
-#library("rethinking")
-#library("statsExpressions")
-#library("ggstatsplot")
 library("ggthemes")
@@
 Finally India is an outlier
+
+```{r}
+sim_fun_A = function() {
+  n <- 1000
+  U <- rnorm(n, 1) # unmeasured distress 
+  A <- rnorm(n, 1) # religious service attendance,
+  M <- rnorm(n , A + .2 * U)# religious identification
+  Y <- rnorm(n , - A *.2 + U *.4) #  distress reduced by attendance.
+
+# Religious id, caused by attendance + distress
+
+# simulate data
+simdat_A <- data.frame(
+  M = M, # rel identification
+  A = A, # rel service
+  U = U, # distressing events (unmeasured)
+  Y = Y )# outcome
+
+sim_A <- lm(Y ~ A + M, data = simdat_A)
+sim_A
+}
+
+# Replicate including the mediator
+r_lm_A <- NA
+r_lm_A = replicate(100, sim_fun_A(), simplify = FALSE )
+
+# mediator attenuates true effect ~ .1 instead of ~.1
+parameters::pool_parameters(r_lm_A)
+
+## Repeat but with no mediator in model
+set.seed(123)
+sim_fun_B = function() {
+  n <- 1000
+  U <- rnorm(n, 1) # unmeasured distress 
+  A <- rnorm(n, 1) # religious service attendance,
+  M <- rnorm(n , A + .2 * U)# religious identification
+  Y <- rnorm(n , - A *.2 + .4 * U) #  distress reduced by attendance.
+
+# Simulate data
+simdat_B <- data.frame(
+  M = M, # rel identification
+  A = A, # rel service attendance
+  U = U, # distressing events (unmeasured)
+  Y = Y ) # outcome
+
+# Only model exposure
+sim_B <- lm(Y ~ A, data = simdat_B)
+sim_B
+}
+
+# Replication 100xs
+r_lm_B <- NA
+r_lm_B = replicate(100, sim_fun_B(), simplify = FALSE )
+
+# diminished recover true effect ~ .2
+parameters::pool_parameters(r_lm_B)
+
+```

### 2022-03-12 History/analysis-v-2.Rmd
order files
create history foldrer

--- 2022-03-03 analysis-v-2.Rmd
@@
 dag2
-dag3 <- image_ggplot(image_read(here::here("figs","dg_3.tiff")), interpolate = T)
+dag3 <- image_ggplot(image_read(here::here("figs","dag_3.tiff")), interpolate = T)
 dag3

### 2022-03-12 History/analysis-v-use.Rmd
order files
create history foldrer

--- 2022-03-03 analysis-religious-only.Rmd
@@
 max(df$rel_CHURCHFREQ)
-length(unique(df$country))
-table(df$country)
 ```
@@
                 #   bg_s,
-                denomination,
                 wb_psychANXIETY)
 
-```
-
-table
-
-```{r}
-lt <-table1::table1(~  denomination, data = d1, transpose=F)
-lt
-lt <-table1::table1(~ ch + religious + age + male + edu + ses + denomination |country, data = d1, transpose=F)
-lt
-kable(as.data.frame(lt), booktabs=TRUE, "latex")
-
-nrow(dr)
 ```
@@
 miss_graph<- gg_miss_upset(dff)
-miss_graphR <-image_ggplot(miss_graph)
+miss_graph <-image_ggplot(miss_graph)
 # save graph
-miss_graph
+str(miss_graph)
 
@@
   path = here::here("figs"),
-  width = 16,
-  height = 9,
+  width = 10,
+  height = 5,
   units = "in",
@@
 
-table1::table1(~ ch |religious, data = d1)
-
 ```
@@
 library(tidyverse)
-d2<-mice(d1)
-
+#d2<-mice(d1)
 #d3<-complete(d2)
 #saveRDS(d3,here::here("mods","d3"))
-#d3 <-readRDS(here::here("mods","d3"))
+d3 <-readRDS(here::here("mods","d3"))
 d4<-d3%>%
-  select(anx, ch, bg, age, edu,male,ses,country,religious)%>%
-  drop_na()%>%
-  mutate(anx_s = scale(anx),
-         religious = factor(religious),
-         bg_s = scale(bg),
-         ch_s = scale(ch),
-         age_s = scale(age),
-         edu_s = scale(edu),
-         ses_s = scale(ses),
-         male_2 =scale(as.numeric(male))/2,
-         ses_s = scale(ses),
-         country=as.factor(country))
-
-## MICE ONLY RELIGIOUS
-table(d1$religious)
-
-# Check
-d1r<- d1 %>%
-  filter(religious == 3)
-
-hist(d1r$ch)
-
-d2r<-mice(d1r)
-
-#check religious rumbers (1= religious, 2= not religious, 3=atheist) [R] So note that 3 = religious!!
-
-d3r<-complete(d2r)
-
-#saveRDS(d3,here::here("mods","d3r"))
-#saveRDS(d3r,here::here("mods","d3r"))
-
-d3 <-readRDS(here::here("mods","d3"))
-d3r <-readRDS(here::here("mods","d3r"))
-
-d4<-d3%>%
-  select(anx, ch, bg, age, edu,male,ses,country,religious)%>%
+  select(anx, ch, bg, age, edu,male,ses,country)%>%
   drop_na()%>%
@@
          ses_s = scale(ses),
-         country=as.factor(country),
-         religious = as.factor(religious))
+         country=as.factor(country))
+
+pred <- ggeffects::ggemmeans(modelsCH[[1]], terms = c("ch_s[all]"))
+pred
+
+plot(pred,
+    # add.data=T,dot.alpha=.1, 
+     facets = F)  + scale_y_continuous(limits=c(-.5,.5)) + theme_classic() + ggtitle("") # title to be suppled
+parameters::pool_parameters(modelsCH)
+
+parameters::pool_parameters(modelsCH)
+#parameters::pool_parameters(modelschG)
+
+## MICE ONLY RELIGIOUS
+table(d1$religious)
+
+# Check
+d1r<- d1 %>%
+  filter(religious == 3)
+
+hist(d1r$ch)
+
+d2r<-mice(d1r)
+
+#check religious rumbers (1= religious, 2= not religious, 3=atheist) [R] So note that 3 = religious!!
+
+d3r<-complete(d2r)
+
+#saveRDS(d3,here::here("mods","d3r"))
+#saveRDS(d3r,here::here("mods","d3r"))
+
+d3 <-readRDS(here::here("mods","d3"))
+d3r <-readRDS(here::here("mods","d3r"))
+
+d4<-d3%>%
+  select(anx, ch, bg, age, edu,male,ses,country)%>%
+  drop_na()%>%
+  mutate(anx_s = scale(anx),
+         bg_s = scale(bg),
+         ch_s = scale(ch),
+         age_s = scale(age),
+         edu_s = scale(edu),
+         ses_s = scale(ses),
+         male_2 =scale(as.numeric(male))/2,
+         ses_s = scale(ses),
+         country=as.factor(country))
 
@@
 
-# weighted for religious only
+# weighted
 
@@
 max(dr$ipwCH)
-```
-
-## Tables
-
-```{r}
-lt <-table1::table1(~ religious|ch_ord, data = dp)
-
-kable(as.data.frame(lt), booktabs=TRUE, "latex")
-
-nrow(dr)
-```
-
-### Simulation
-
-```{r}
-#\begin{lstlisting}[language=R]
-
-set.seed(123)
-sim_fun_A = function() {
-  n <- 1000
-  U <- rnorm(n, 1) # Unmeasured distress 
-  A <- rnorm(n, 1) # religious service attendance,
-  M <- rnorm(n , A + U)# rel id
-  Y <- rnorm(n , - A*.2 + U + M*.1) #  distress reduced by attendance.
-
-# Religious id, caused by attendance + distress
-
-# simulate data
-simdat_A <- data.frame(
-  M = L, # rel id
-  A = A, # rel service
-  U = U, # distressing events (unmeasured)
-  Y = Y ) # outcome
-
-sim_A <- lm(Y ~ A + M, data = simdat_A)
-sim_A
-}
-
-# Replicate including the mediator
-r_lm_A <- NA
-r_lm_A = replicate(100, sim_fun_A(), simplify = FALSE )
-
-# average over outcomes: effect of religion is attenuated
-parameters::pool_parameters(r_lm_A)
-
-## Repeat but with no mediator in model
-set.seed(123)
-sim_fun_B = function() {
-  n <- 1000
-  U <- rnorm(n, 1) # Unmeasured distress 
-  A <- rnorm(n, 1) # religious service attendance,
-  M <- rnorm(n , A + U)# rel id + distress
-  Y <- rnorm(n , - A *.2 + U + M*.1) #  distress reduced by attendance.
-
-# simulate data
-simdat_B <- data.frame(
-  M = L, # rel id
-  A = A, # rel service
-  U = U, # distressing events (unmeasured)
-  Y = Y ) # outcome
-
-# Only exposure
-sim_B <- lm(Y ~ A, data = simdat_A)
-sim_B
-}
-
-# Replication 1000xs
-r_lm_B <- NA
-r_lm_B = replicate(100, sim_fun_B(), simplify = FALSE )
-
-# average over outcomes: effect of religion is attenuated
-parameters::pool_parameters(r_lm_B)
-
-#\end{lstlisting}
-```
-
-```{r}
-dr$ch_ord <- factor(dr$ch, ordered=TRUE)
-table(dr$ch_ord)
-
-levels(dr$ch_ord)<-c("never","rarely","1xpy","holydays","1xpm","1xpw",">1xpw")
-
-levels(dr$ch_ord)
-is.ordered(dr$ch_ord)
-
-#saveRDS(dr, here::here("mods","dr"))
 
@@
 # ))
-
-b_orR <- brm(
-  anx_s|weights(ipwCH) ~ mo(ch_ord) + (1 + mo(ch_ord) | country),
-  data = dr,
-   prior = c(
-    prior(student_t(3, 0, .5), class = Intercept),
-    prior(normal(-.05,.1), class = "b"),
-    prior(dirichlet(1,1,1,1,1,1), class = "simo", coef = moch_ord1)),
-  seed = 1234,
-  warmup = 1000,
-  iter = 2000,
-  chains = 4,
-  backend = "cmdstanr",
-  file = here::here("mods","b_orR")
-)
-
-summary(b_orR)
-
-prior_summary(b_orR)
-
-church_plot_R <-plot(ggeffects::ggemmeans(b_or, terms = "ch_ord")) +   xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + labs(title = "Baysian Multi-level Model",
-subtitle = "Inverse Probability Weighting") + scale_y_continuous(limits = c(-.5,.5))+ 
-  theme_classic()
-church_plot_R
-
-### EEMEANS
- b_orR %>% 
-  emmeans(~ ch_ord,
-          at = list(ch_ord = c("never",">1xpw"),
-                    country = NA),
-          epred = TRUE) %>% 
-  contrast(method = "revpairwise")
-
-effect_draws_b_orR <-  b_orR %>% 
-  emmeans(~ ch_ord,
-         at = list(ch_ord = c("never",">1xpw"),
-                    country = NA),
-          epred = TRUE) %>% 
-  contrast(method = "revpairwise") %>%
-  gather_emmeans_draws()
-
-effect_draws_b_orR %>% mean_hdi()
-
-#SE = (upper limit – lower limit) / 3.92.
-
-# STANDARD ERROR
-Roa <- effect_draws_b_orR %>% mean_hdi() 
-(Roa$.upper - Roa$.lower)/3.92 #
-
-# SE = .0.06751772
-
-library(EValue)
-evalues.OLS(-0.208, se = 0.06751772, sd = 1, delta = 1)
-#             point    lower     upper
-# RR       0.8275548 0.733842 0.9332348
-# E-values 1.7101772       NA 1.3484170
-bias_plot(0.9332348, xmax = 6)
-
-average_marginal_effectR <- ggplot(effect_draws_b_orR, aes(x = .value)) +
-    stat_halfeye(fill = "dodgerblue") +
-  labs(x = "Average marginal effect of church attenance on anxiety", y = "Density",
-       subtitle = "Marginal effect (never vs. 1xpw): continuous model") +
-  theme_classic() +
-       scale_x_continuous(limits=c(-.6,.3)) +
-  theme(legend.position = "top")
-
-average_marginal_effectR
-
-grand_mean_ameR <- b_orR %>% 
-  emmeans(~ ch_ord,
-          at = list(ch_ord = c("never",">1xpw")),
-          epred = TRUE, re_formula = NA) %>% 
-  contrast(method = "revpairwise") %>% 
-  gather_emmeans_draws()
-
-plot_grand_mean_ameR <- ggplot(grand_mean_ameR, 
-                                   aes(x = .value)) +
-  stat_halfeye(fill = palette_okabe_ito(order = 5)) +
-  labs(x = "Average marginal effect of church attenance on anxiety", y = "Density",
-       subtitle = "Marginal effect (never vs. 1xpw): monotonic model") +
-  theme_classic() +
-    scale_x_continuous(limits=c(-.6,.3)) +
-   theme(legend.position = "bottom",
-         legend.title = element_text(size = 5),
-          legend.text = element_text(size = 2)) 
-plot_grand_mean_ameR
-
-grand_mean_autonomy_distR <- b_orR %>% 
-  epred_draws(newdata = expand_grid(ch_ord = c("never",
-                                               #"rarely","1xpy","holydays","1xpm","1xpw",
-                                               ">1xpw")), 
-              re_formula = NA)
-
-plot_grand_mean_ordR <- ggplot(grand_mean_autonomy_distR, 
-                                   aes(x = .epred, 
-                                       fill = ch_ord)) +
-  scale_fill_okabe_ito() +
-    stat_halfeye(slab_alpha = 0.7) +
-    scale_x_continuous(limits=c(-.6,.4)) +
-  labs(x = "Predicted anxiety (standardised)", y = "Density",
-       fill = "Church frequency",
-       subtitle = "Posterior predictions: monotonic model") +
-  theme_classic() +
-     theme(legend.position = "bottom",
-         legend.title = element_text(size = 12),
-          legend.text = element_text(size = 8))
-
-graph_r<- plot_grand_mean_ameR + plot_grand_mean_ordR + plot_annotation(tag_levels = "a",  title = "Religious only model")
-yola
-
-ggsave(
-  graph_r,
-  path = here::here("figs"),
-  width = 12,
-  height = 6,
-  units = "in",
-  filename = "graph_r.jpg",
-  device = 'jpeg',
-  limitsize = FALSE,
-  dpi = 1200
-)
-
-```
-
-## include religion
+```
 
@@
 library(splines)
-
-ipb_ch_ln_no_mediation <- geeglm(
-  anx_s ~ ch,
+ipb_ch_ln<- geeglm(
+  anx_s ~ ch_s,
   data = d1p,
@@
 )
-
-## Model misspecification 
-ipb_ch_ln_mediation <- geeglm(
-  anx_s ~ ch + religious,
+saveRDS(ipb_bg, here::here("mods","ipb_bg_ln"))
+
+ipb_ch<- geeglm(
+  anx_s ~  bs(ch_s),
   data = d1p,  
@@
 )
-saveRDS(ipb_ch_ln_mediation, here::here("mods","ipb_ch_ln_mediation"))
-saveRDS(ipb_ch_ln_no_mediation, here::here("mods","ipb_ch_ln_no_mediation"))
-
-model_parameters(ipb_ch_ln_mediation)
-model_parameters(ipb_ch_ln_no_mediation)
-
-plot(ggeffects::ggemmeans(ipb_ch_ln_mediation, terms = "ch"), add.data = TRUE)
-
-ipb_ch<- geeglm(
-  anx_s ~  bs(ch),
-  data = d1p,  
-  id = country,
-  corstr = "exchangeable",
-  weights = ipwCH
-)
 saveRDS(ipb_ch, here::here("mods","ipb_ch"))
+
+# model_parameters(ipb_bg)
+
 model_parameters(ipb_ch)
 
-plot(ggeffects::ggemmeans(ipb_ch, terms = "ch[all]"), add.data = TRUE
-     )
-
-## Rel only
-
-library(splines)
-ipb_ch_lnR<- geeglm(
-  anx_s ~ ch,
-  data = dr,
-  id = country,
-  corstr = "exchangeable",
-  weights = ipwCH
-)
-#saveRDS(ipb_ch_lnR, here::here("mods","ipb_ch_lnR"))
-model_parameters(ipb_ch_lnR)
-
-plot(ggeffects::ggemmeans(ipb_ch_lnR, terms = "ch"), add.data = TRUE
-     )
-
-ipb_chR<- geeglm(
-  anx_s ~  bs(ch),
-  data = dr,  
-  id = country,
-  corstr = "exchangeable",
-  weights = ipwCH
-)
-saveRDS(ipb_chR, here::here("mods","ipb_ch"))
-model_parameters(ipb_chR)
-
-plot(ggeffects::ggemmeans(ipb_chR, terms = "ch[all]"), add.data = TRUE
-     )
-
-# model_parameters(ipb_bg)
-
-abs(min(d1p$ch_s) - max(d1p$ch_s))
-# 3.064032
+min(d1p$ch_s) + max(d1p$ch_s)
 
@@
 
-library(brms)
-library(splines)
-
-listout <- readRDS(here::here("mods","listout"))
-modmc_wt_priorStrong <- brm(
-  anx_s|weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
-  data = d1p,
-    prior = c(
-    prior(student_t(3, 0,.25), class = Intercept),
-    prior(normal(-.05, .1), class = b)),
-  seed = 1234,
-  warmup = 1000,
-  iter = 2000,
-  chains = 4,
-  backend = "cmdstanr",
-  file = here::here("mods","modmc_wt_priorStrong")
-)
-
-## Table
-parameters::parameters(
-  modmc_wt_priorStrong,
-  test = "pd",
-  component = "all",
-  priors = FALSE, 
-  ci_method = "hdi",
-#  effects = "fixed",
-  group_level = FALSE,
-  diagnostic = "Rhat") %>%
-    mutate_if(is.numeric, ~ round(., 3)) %>%
-    mutate_all(funs(str_replace(., "b_", ""))) %>%
-    mutate_all(funs(str_replace(., "bs", ""))) %>%
-    kable(booktabs = T, "latex", caption =  "Parameter estimates for continuous (spline) model of religious service on anxiety") %>%
-    print()
-
-# prior summary
-
-prior_summary(modmc_wt_priorStrong)
-
-##### TeST PRIORS
-modmc_wt_priorStrong_TEST <- brm(
-  anx_s|weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
-  data = d1p,
-    prior = c(
-    prior(student_t(3, 0,.25), class = Intercept),
-    prior(normal(-.05, .1), class = b)),
-  seed = 1234,
-  warmup = 1000,
-  iter = 2000,
-  chains = 4,
-  sample_prior = "only",
-  backend = "cmdstanr",
- file = here::here("mods","modmc_wt_priorStrong")
-)
-
-## WEAK PRIOR
-modmc_wt_prior_weak <- brm(
-  anx_s|weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
-  data = d1p,
-  prior = c(prior(normal(0, 1), class = b)),
-  seed = 1234,
-  warmup = 1000,
-  iter = 2000,
-  chains = 4,
-  sample_prior = "only",
-  backend = "cmdstanr"#,
- # file = here::here("mods","modmc_wt_prior")
-)
-```
-
-```{r, cache = TRUE}
-library(viridis)
-
-# make new data
-minchur<- min(d1p$ch_s)
-maxchur<- max(d1p$ch_s)
-
-# difference
-abs(maxchur-minchur)
-
-library(emmeans)
-grand_mean
-
-p0<-brms::conditional_effects(modmc_wt_priorStrong, "ch_s", 
-                              spaghetti = T,
-                              ndraws = 1000,
-                              categorical = F,
-                              prob = .95,
-                              method = 'fitted')
-
-church_plot <- plot(p0, plot = FALSE)[[1]]   +  xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + labs(title = "Baysian Multi-level Model",
-subtitle = "Inverse Probability Weighting") + scale_y_continuous(limits = c(-.5,.5))+ 
-  theme_classic()
-church_plot
-
-ipw_gee <- plot(pred,
-    #add.data=T,dot.alpha=.1, 
-     facets = F)  + 
-  scale_y_continuous(limits=c(-.1,.1)) + xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + 
-  theme_classic() +labs(title = "Generalized Estimating Equations", 
-  subtitle = "Inverse Probability Weighting") + scale_y_continuous(limits = c(-.5,.5))
-ipw_gee
-church_plot
-model_plot <- ipw_gee + church_plot + church_plot_test_weak + church_plot_test_strong
-model_plot <- model_plot + plot_annotation(tag_levels = "a")
-
-model_plot
-
-ggsave(
-  model_plot,
-  path = here::here("figs"),
-  width = 10,
-  height = 5,
-  units = "in",
-  filename = "model_plot.jpg",
-  device = 'jpeg',
-  limitsize = FALSE,
-  dpi = 1200
-)
-
-library(brmsmargins)
-modmc_wt_priorStrong
-
-max(d1p$ch_s)#-0.8321022,2.23193
-
-h <-.001
-ames <- brmsmargins(
-  object = modmc_wt_priorStrong,
-            at = data.frame(ch_s= c(-0.8321022,2.23193)),
-            contrasts = cbind("AME x" = c(0.8321022/ h, 2.23193/ h)),
-            effects = "integrateoutRE")
-                              )
-ames$Summary
-
-ames_st <- brmsmargins(
-  object = modmc_wt_priorStrong, 
-            at = data.frame(ch_s = c(-0.8321022,2.23193)),
-            contrasts = cbind("AME x" = c(-0.8321022/ h, 2.23193/ h)),
-            effects = "integrateoutRE", CI = 0.95)
-ames_st$ContrastSummary
-knitr::kable(ames$Summary, digits = 3, "latex")
-
-ames_ln <- brmsmargins(
-  object = modmc_wt_linear, 
-            at = data.frame(ch_s = c(-0.8321022,2.23193)),
-            contrasts = cbind("AME x" = c(-0.8321022/ h, 2.23193/ h)),
-            effects = "integrateoutRE", CI=.95)
-ames_ln$Summary
-knitr::kable(ames$Summary, digits = 3, "latex")
-
-### EEMEANS
- modmc_wt_priorStrong %>% 
-  emmeans(~ ch_s,
-          at = list(ch_s = c(-0.8321022,2.23193),
-                    country = NA),
-          epred = TRUE) %>% 
-  contrast(method = "revpairwise")
-
-effect_draws_cn <-  modmc_wt_priorStrong %>% 
-  emmeans(~ ch_s,
-          at = list(ch_s = c(-0.8321022,2.23193),
-                    country = NA),
-          epred = TRUE) %>% 
-  contrast(method = "revpairwise") %>% 
-  gather_emmeans_draws()
-
-# marginal mean 
-effect_draws_cn %>% mean_hdi()
-
-# marginal mean 
-effect_draws_cn %>% mean_hdi() 
-effect_draws_cn
-
-average_marginal_effect <- ggplot(effect_draws_cn, aes(x = .value)) +
-    stat_halfeye(fill = "dodgerblue") +
-  labs(x = "Average marginal effect of church attenance on anxiety", y = "Density",
-       subtitle = "Marginal effect (never vs. 1xpw): continuous model") +
-  theme_classic() +
-       scale_x_continuous(limits=c(-.6,.3)) +
-  theme(legend.position = "top")
-
-average_marginal_effect
-
-grand_mean <- modmc_wt_priorStrong %>% 
-  epred_draws(newdata = expand_grid(ch_s = seq(-0.8321022, 2.23193, by = 0.05)), 
-              re_formula = NA)
-
-  plot_grand_mean <- ggplot(grand_mean, 
-                                 aes(x = ch_s, y = .epred)) +
-  stat_lineribbon() +
-  scale_fill_brewer(palette = "Blues") +
- # scale_fill_viridis_d(palette = "Blues") +
-  labs(x = "Frequency of religious service attendance (SD)", 
-       y = "Predicted anxiety (SD)",
-       fill = "95 HPD interval",
-       subtitle = "Posterior predictions: continuous model") +
-     theme_classic() +
-  theme(legend.position = "bottom",
-        legend.title = element_text(size = 10),
-          legend.text = element_text(size = 8))
-
-plot_grand_mean
-plot_result <- average_marginal_effect + plot_grand_mean +  plot_annotation(tag_levels = "a")
-plot_result
-
-ggsave(
-  plot_result,
-  path = here::here("figs"),
-  width = 16,
-  height = 9,
-  units = "in",
-  filename = "plot_result.jpg",
-  device = 'jpeg',
-  limitsize = FALSE,
-  dpi = 1200
-)
-
-summary(modmc_wt_priorStrong)
-summary(modmc_wt_linear)
-
-#
-
-library(EValue)
-summary(ipb_ch_ln)
-summary(ipb_ch)
-summary(modmc_wt_linear)
-
-# sensitivity = Church
-#yola
-
-#SE = (upper limit – lower limit) / 3.92.
-
-ta <- effect_draws_cn %>% mean_hdi() 
-ta
-(ta$.upper - ta$.lower)/3.92 #
-
-evalues.OLS(-0.147 , se = 0.05607611, sd = 1, delta = 1.1)
-# 
-#              point     lower     upper
-# RR       0.8631671 0.7734115 0.9633389
-# E-values 1.5870732        NA 1.2368137
-
-# multibias
-multi_evalue(est = RR( 0.8631671), biases = selection())
-
-#
-bias_plot(0.8631671, xmax=5)
-
-multi_bound(multi_bias(confounding()),
-                RRAUc = 2.2, RRUcY = 1.7)
-```
-
-## Scaled 0-1
-
-```{r, cache = TRUE}
-d1p<-readRDS(here::here("mods","d1p"))
-
-dchr_sc <- ipw::ipwpoint(
-  exposure = ch,
-  family = "gaussian",
-  numerator = ~ 1,
-  denominator = ~ age_s + edu_s + male_2 + ses_s + country,
-  trunc = 0.01,
-  data = as.data.frame(d1p)
-)
-
-dp <- d1p %>% 
-  mutate(ipwCH = dchr_sc$weights.trunc)
-
-# fine, well below 10
-max(dp$ipwCH)
-
-```
-
-# Montonic effects
-
-```{r}
-# bayesian monotonic effects
-
-dp$ch_ord <- factor(dp$ch, ordered=TRUE)
-table(dp$ch_ord)
-
-levels(dp$ch_ord)<-c("never","rarely","1xpy","holydays","1xpm","1xpw",">1xpw")
-
-levels(dp$ch_ord)
-is.ordered(dp$ch_ord)
-
-#saveRDS(dp, here::here("mods","dp"))
-dp<-readRDS(here::here("mods","dp"))
-```
-
-## ORDINAL MODEL
-
-```{r}
-## Bayesian wieghted regression 
-b_or <- brm(
-  anx_s|weights(ipwCH) ~ mo(ch_ord) + (1 + mo(ch_ord) | country),
-  data = dp,
-   prior = c(
-    prior(student_t(3, 0, .5), class = Intercept),
-    prior(normal(-.05,.1), class = "b"),
-    prior(dirichlet(1,1,1,1,1,1), class = "simo", coef = moch_ord1)),
-  seed = 1234,
-  warmup = 1000,
-  iter = 2000,
-  chains = 4,
-  backend = "cmdstanr",
-  file = here::here("mods","b_or")
-)
-
-summary(b_or)
-
-prior_summary(b_or)
-summary(b_or)
-
-church_plot <-plot(ggeffects::ggemmeans(b_or, terms = "ch_ord")) +   xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + labs(title = "Baysian Multi-level Model",
-subtitle = "Inverse Probability Weighting") + scale_y_continuous(limits = c(-.5,.5))+ 
-  theme_classic()
-church_plot
-```
-
-## Tables
-
-```{r}
-# tab_or <- model_parameters(b_or,   verbose = F, ci_method = "hdi",test = c("pd"))
-# kbl(tab_or, "latex",booktabs=T)
-
-# sjstats::tidy_stan(tab_or)
-
-# library(broom.mixed)
-# tidy(
-#   b_or,
-#  # parameters = NA,
-#   effects = c("fixed", "ran_pars"),
-#   robust = FALSE,
-#   conf.int = TRUE,
-#   conf.level = 0.95,
-#   conf.method = c("HPDinterval"),
-#   fix.intercept = TRUE
-# )
-
-parameters::parameters(
-  b_or,
-  test = "pd",
-  component = "all",
-  priors = FALSE, 
-  ci_method = "hdi",
-#  effects = "fixed",
-  group_level = FALSE,
-  diagnostic = "Rhat") %>%
-    mutate_if(is.numeric, ~ round(., 3)) %>%
-    mutate_all(funs(str_replace(., "b_", ""))) %>%
-    mutate_all(funs(str_replace(., "_moch", ""))) %>%
-    kable(booktabs = T, "latex", caption =  "Parameter estimates for model of religious service on anxiety") %>%
-    print()
-```
-
-```{r}
-# pord<-brms::conditional_effects(b_or, "ch_ord", 
-#                               spaghetti = T,
-#                               ndraws = 1000,
-#                               categorical = F,
-#                               prob = .9,
-#                               method = 'fitted')
-# 
-# 
-# church_plot <- plot(pord, plot = FALSE)[[1]]   +  xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + labs(title = "Baysian Multi-level Model",
-# subtitle = "Inverse Probability Weighting") + scale_y_continuous(limits = c(-.5,.5))+ 
-#   theme_classic()
-# church_plot
-library(tidybayes)
-
-summary(b_or)
-### EEMEANS
- b_or %>% 
-  emmeans(~ ch_ord,
-          at = list(ch_ord = c("never",">1xpw"),
-                    country = NA),
-          epred = TRUE) %>% 
-  contrast(method = "revpairwise")
-
-  effect_draws <-  b_or %>% 
-  emmeans(~ ch_ord,
-         at = list(ch_ord = c("never",">1xpw"),
-                    country = NA),
-          epred = TRUE) %>% 
-  contrast(method = "revpairwise") %>%
-  gather_emmeans_draws()
-
-# linear model, for EVALUES
-effect_draws_or <-  b_or %>% 
-  emmeans(~ ch_ord,
-         at = list(ch_ord = c("never",">1xpw"),
-                    country = NA),
-          epred = TRUE) %>%
-  contrast(method = "revpairwise") %>% 
-  gather_emmeans_draws()
-effect_draws_or
-# marginal mean 
-effect_draws_or %>% mean_hdi()
-
-# calculate standard error
-# 
-# se <- function(x) sqrt(var(x) / length(x))
-# 
-# summary(b_or)
-# se(effect_draws_ln$.value)
-library(tidybayes)
-library(ggokabeito)   # Neat accessible color palette
-# emmeans()-based difference across autonomy
-grand_mean_ame <- b_or %>% 
-  emmeans(~ ch_ord,
-          at = list(ch_ord = c("never",">1xpw")),
-          epred = TRUE, re_formula = NA) %>% 
-  contrast(method = "revpairwise") %>% 
-  gather_emmeans_draws()
-
-plot_grand_mean_ame <- ggplot(grand_mean_ame, 
-                                   aes(x = .value)) +
-  stat_halfeye(fill = palette_okabe_ito(order = 5)) +
-  labs(x = "Average marginal effect of church attenance on anxiety", y = "Density",
-       subtitle = "Marginal effect (never vs. 1xpw): monotonic model") +
-  theme_classic() +
-    scale_x_continuous(limits=c(-.6,.3)) +
-   theme(legend.position = "bottom",
-         legend.title = element_text(size = 5),
-          legend.text = element_text(size = 2))
-
-plot_grand_mean_ame
-
-# plot_grand_mean <- ggplot(grand_mean, 
-#                                  aes( x = .epred, fill=(ch_ord))) +
-#   scale_fill_brewer(palette = "Blues") +
-#     stat_halfeye(slab_alpha = 0.8) +
-# 
-#   labs(x = "Church attendence frequency", 
-#        y = "Marginal Density",
-#        fill = "") +
-#   theme_classic() +
-#   theme(legend.position = "top")
-# 
-# plot_grand_mean
-
-grand_mean_autonomy_dist <- b_or %>% 
-  epred_draws(newdata = expand_grid(ch_ord = c("never","rarely","1xpy","holydays","1xpm","1xpw",">1xpw")), 
-              re_formula = NA)
-
-grand_mean_autonomy_dist2 <- b_or %>% 
-  epred_draws(newdata = expand_grid(ch_ord = c("never",">1xpw")), 
-              re_formula = NA)
-
-plot_grand_mean_ord <- ggplot(grand_mean_autonomy_dist, 
-                                   aes(x = .epred, 
-                                       fill = ch_ord)) +
-  scale_fill_okabe_ito() +
-    stat_halfeye(slab_alpha = 0.7) +
-    scale_x_continuous(limits=c(-.6,.4)) +
-  labs(x = "Predicted anxiety (standardised)", y = "Density",
-       fill = "Church frequency",
-       subtitle = "Posterior predictions: monotonic model") +
-  theme_classic() +
-     theme(legend.position = "bottom",
-         legend.title = element_text(size = 10),
-          legend.text = element_text(size = 8))
-
-plot_grand_mean_ord
-
-plot_grand_mean_ord2 <- ggplot(grand_mean_autonomy_dist2, 
-                                   aes(x = .epred, 
-                                       fill = ch_ord)) +
-  scale_fill_okabe_ito() +
-    stat_halfeye(slab_alpha = 0.7) +
-    scale_x_continuous(limits=c(-.6,.4)) +
-  labs(x = "Predicted anxiety (standardised)", y = "Density",
-       fill = "Church frequency",
-       subtitle = "Posterior predictions: monotonic model") +
-  theme_classic() +
-     theme(legend.position = "bottom",
-         legend.title = element_text(size = 12),
-          legend.text = element_text(size = 8))
-
-plot_grand_mean_ord2
-
-ggsave(
-  plot_result_ord,
-  path = here::here("figs"),
-  width = 12,
-  height = 6,
-  units = "in",
-  filename = "plot_result_ord.jpg",
-  device = 'jpeg',
-  limitsize = FALSE,
-  dpi = 1200
-)
-
-plot_result_ord2 <- plot_grand_mean_ame + plot_grand_mean_ord2+  plot_annotation(tag_levels = "a")
-plot_result_ord2
-
-ggsave(
-  plot_result_ord2,
-  path = here::here("figs"),
-  width = 12,
-  height = 6,
-  units = "in",
-  filename = "plot_result_ord2.jpg",
-  device = 'jpeg',
-  limitsize = FALSE,
-  dpi = 1200
-)
-
-plot_result_full <- (average_marginal_effect + plot_grand_mean)/
- (plot_grand_mean_ame + plot_grand_mean_ord) + plot_annotation(tag_levels = "a")
-plot_result_full
-
-ggsave(
-  plot_result_full,
-  path = here::here("figs"),
-  width = 10,
-  height =10,
-  units = "in",
-  filename = "plot_result_full.jpg",
-  device = 'jpeg',
-  limitsize = FALSE,
-  dpi = 1000
-)
-
-# Obtain evalues
-
-# marginal mean 
-effect_draws_or %>% mean_hdi()
-
-#SE = (upper limit – lower limit) / 3.92.
-
-oa <- effect_draws_or %>% mean_hdi() 
-(oa$.upper - oa$.lower)/3.92 #
-
-# SE = .075
-
-library(EValue)
-evalues.OLS(-0.187 , se = .06733407, sd = 1, delta = 1)
-
-#             point     lower     upper
-# RR       0.8435214 0.7482451 0.9509295
-# E-values 1.6544616        NA 1.2845523
-
-#SE = (upper limit – lower limit) / 3.92.
-
-ta <- effect_draws_cn %>% mean_hdi() 
-ta
-(ta$.upper - ta$.lower)/3.92 #
-
-evalues.OLS(-0.147  , se =  0.05607611, sd = 1, delta = 1.1)
-#              point     lower     upper
-# RR       0.8631671 0.7734115 0.9633389
-# E-values 1.5870732        NA 1.2368137
-```
-
-### Test priors ordinal model
-
-```{r}
-##### TeST PRIORS
-b_ord_test_weak_priors <- brm(
-  anx_s|weights(ipwCH) ~ mo(ch_ord) + (1 + mo(ch_ord) | country),
-  data = dp,
-  prior = c(
-    prior(student_t(3, 0, 2.5), class = Intercept),
-    prior(normal(0,1), class = "b"),
-    prior(dirichlet(1,1,1,1,1,1), class = "simo", coef = moch_ord1)),
-  seed = 1234,
-  warmup = 1000,
-  iter = 2000,
-  chains = 4,
-  sample_prior = "only",
-  backend = "cmdstanr"
-)
-
-plot_ord_wk_priors<-brms::conditional_effects(b_ord_test_weak_priors, "ch_ord", 
-                              spaghetti = T,
-                              ndraws = 1000,
-                              categorical = F,
-                              prob = .95,
-                              method = 'fitted')
-
-test_or_wk<- plot(plot_ord_wk_priors, plot = FALSE)[[1]]   +  xlab("Service attendance (monotonic)") + ylab ("Anxiety (standardised)") + labs(title = "Check Priors: Baysian Mixed Effects Marginal Model", subtitle = "Sampling from weak priors only (ordinal) ") + scale_y_continuous(limits = c(-20,15))+
-  theme_classic()
-test_or_wk
-
-d1p<-readRDS(here::here("mods","d1p"))
-dp<-readRDS(here::here("mods","dp"))
-
-b_cont_test_st_priors <- brm(
-  anx_s|weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
-  data = d1p,
-    prior = c(
-    prior(student_t(3, 0,.25), class = Intercept),
-    prior(normal(-.05, .1), class = b)),
-  seed = 1234,
-  warmup = 1000,
-  iter = 2000,
-  chains = 4,
-  sample_prior = "only",
-  backend = "cmdstanr"#,
- # file = here::here("mods","b_cont_test_st_priors")
-)
-
-con_st_priors<-brms::conditional_effects(b_cont_test_st_priors, "ch_s", 
-                              spaghetti = T,
-                              ndraws = 1000,
-                              categorical = F,
-                              prob = .95,
-                              method = 'fitted')
-
-test_con_st<- plot(con_st_priors, plot = FALSE)[[1]]   +  xlab("Service attendance (sd)") + ylab ("Anxiety (standardised)") + labs(title = "Check Priors: Baysian Mixed Effects Marginal Model", subtitle = "Sampling from weak priors only (continuous) ") +   scale_y_continuous(limits = c(-10,10))+
-  theme_classic()
-
-b_cont_test_wk_priors <- brm(
-  anx_s|weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
-  data = d1p,
-   prior = c(
-    prior(normal(0, 2.5), class = Intercept),
-    prior(normal(0, 1), class = b)),
-  seed = 1234,
-  warmup = 1000,
-  iter = 2000,
-  chains = 4,
-  sample_prior = "only",
-  backend = "cmdstanr"#
-)
-
-con_wk_priors<-brms::conditional_effects(b_cont_test_wk_priors, "ch_s", 
-                              spaghetti = T,
-                              ndraws = 1000,
-                              categorical = F,
-                              prob = .95,
-                              method = 'fitted')
-
-test_con_wk_priors<- plot(con_wk_priors, plot = FALSE)[[1]]   +  xlab("Service attendance (sd)") + ylab ("Anxiety (standardised)") + labs(title = "Check Priors: Baysian Mixed Effects Marginal Model", subtitle = "Sampling from weak priors only (continuous) ")   + scale_y_continuous(limits = c(-10,10))+
-  theme_classic()
-test_con_wk_priors
-
-## THIS IS IT.  TO MAKE WEAKER INCREASE INTERCEPT 
-b_or_test_st_priors4 <- brm(
-  anx_s|weights(ipwCH) ~ mo(ch_ord) + (1 + mo(ch_ord) | country),
-  data = dp,
-  seed = 1234,
-  prior = c(
-    prior(student_t(3, 0, .5), class = Intercept),
-    prior(normal(-.05,.1), class = "b"),
-    prior(dirichlet(1,1,1,1,1,1), class = "simo", coef = moch_ord1)),
-  warmup = 1000,
-  iter = 2000,
-  chains = 4,
-  sample_prior = "only",
-  backend = "cmdstanr"#,
- # file = here::here("mods","b_or_test_st_priors4")
-)
-
-prior_summary(b_or_test_st_priors4)
-
-b_or_test_st_priors4<-brms::conditional_effects(b_or_test_st_priors4, "ch_ord", 
-                              spaghetti = T,
-                              ndraws = 1000,
-                              categorical = F,
-                              prob = .95,
-                              method = 'fitted')
-
-or_strong<- plot(b_or_test_st_priors4, plot = FALSE)[[1]]   +  xlab("Service attendance (monotonic)") + ylab ("Anxiety (standardised)") + labs(title = "Check Priors: Baysian Mixed Effects Marginal Model", subtitle = "Sampling from strong priors only (ordinal) ") + scale_y_continuous(limits = c(-20,15))+
-  theme_classic()
-or_strong
-
-#saveRDS(modmc, here::here("mods", "modmc.rds"))
-
-library(patchwork)
-
-test_priors <-(test_or_wk + or_strong )/(test_con_wk_priors+test_con_st) + plot_annotation(tag_levels = "a")
-
-test_priors
-
-ggsave(
-  test_priors,
-  path = here::here("figs"),
-  width = 12,
-  height = 12,
-  units = "in",
-  filename = "test_priors.jpg",
-  device = 'jpeg',
-  limitsize = FALSE,
-  dpi = 1200
-)
-
-```
-
-```{r}
 library(brms)
@@
 
+## WEAK PRIOR
+modmc_wt_prior_weak <- brm(
+  anx_s|weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
+  data = d1p,
+  prior = c(prior(normal(0, 1), class = b)),
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  sample_prior = "only",
+  backend = "cmdstanr"#,
+ # file = here::here("mods","modmc_wt_prior")
+)
+```
+
+```{r, cache = TRUE}
+library(viridis)
+
+# make new data
+minchur<- min(d1p$ch_s)
+maxchur<- max(d1p$ch_s)
+
+library(emmeans)
+grand_mean
+
+p0<-brms::conditional_effects(modmc_wt_priorStrong, "ch_s", 
+                              spaghetti = T,
+                              ndraws = 1000,
+                              categorical = F,
+                              prob = .95,
+                              method = 'fitted')
+
+church_plot <- plot(p0, plot = FALSE)[[1]]   +  xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + labs(title = "Baysian Multi-level Model",
+subtitle = "Inverse Probability Weighting") + scale_y_continuous(limits = c(-.5,.5))+ 
+  theme_classic()
+church_plot
+
+p0priorTEST<-brms::conditional_effects(modmc_wt_priorStrong_TEST, "ch_s", 
+                              spaghetti = T,
+                              ndraws = 1000,
+                              categorical = F,
+                              prob = .95,
+                              method = 'fitted')
+
+church_plot_test_strong <- plot(p0priorTEST, plot = FALSE)[[1]]   +  xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + labs(title = "Prior Sanity Check: Baysian Multi-level Model", subtitle = "Sampling from STRONG Priors Only") + scale_y_continuous(limits = c(-.5,.25))+
+  theme_classic()
+church_plot_test_strong
+
+pred_brms
+
+p0priorTEST_weak<-brms::conditional_effects(modmc_wt_prior_weak, "ch_s", 
+                              spaghetti = T,
+                              ndraws = 1000,
+                              categorical = F,
+                              prob = .95,
+                              method = 'fitted')
+
+church_plot_test_weak <- plot(p0priorTEST_weak, plot = FALSE)[[1]]   +  xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + labs(title = "Prior Sanity Check: Baysian Multi-level Model", subtitle = "Sampling from WEAK Priors Only") + scale_y_continuous(limits = c(-.5,.5))+
+  theme_classic()
+church_plot_test_weak
+
+ipw_gee <- plot(pred,
+    #add.data=T,dot.alpha=.1, 
+     facets = F)  + 
+  scale_y_continuous(limits=c(-.1,.1)) + xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + 
+  theme_classic() +labs(title = "Generalized Estimating Equations", 
+  subtitle = "Inverse Probability Weighting") + scale_y_continuous(limits = c(-.5,.5))
+ipw_gee
+church_plot
+model_plot <- ipw_gee + church_plot + church_plot_test_weak + church_plot_test_strong
+model_plot <- model_plot + plot_annotation(tag_levels = "a")
+
+model_plot
+
+ggsave(
+  model_plot,
+  path = here::here("figs"),
+  width = 10,
+  height = 5,
+  units = "in",
+  filename = "model_plot.jpg",
+  device = 'jpeg',
+  limitsize = FALSE,
+  dpi = 1200
+)
+
+library(brmsmargins)
+modmc_wt_priorStrong
+
+max(d1p$ch_s)#-0.8321022,2.23193
+
+h <-.001
+ames <- brmsmargins(
+  object = modmc_wt_priorStrong,
+            at = data.frame(ch_s= c(-0.8321022,2.23193)),
+            contrasts = cbind("AME x" = c(0.8321022/ h, 2.23193/ h)),
+            effects = "integrateoutRE")
+                              )
+ames$Summary
+
+ames_st <- brmsmargins(
+  object = modmc_wt_priorStrong, 
+            at = data.frame(ch_s = c(-0.8321022,2.23193)),
+            contrasts = cbind("AME x" = c(-0.8321022/ h, 2.23193/ h)),
+            effects = "integrateoutRE", CI = 0.95)
+ames_st$ContrastSummary
+knitr::kable(ames$Summary, digits = 3, "latex")
+
+ames_ln <- brmsmargins(
+  object = modmc_wt_linear, 
+            at = data.frame(ch_s = c(-0.8321022,2.23193)),
+            contrasts = cbind("AME x" = c(-0.8321022/ h, 2.23193/ h)),
+            effects = "integrateoutRE", CI=.95)
+ames_ln$Summary
+knitr::kable(ames$Summary, digits = 3, "latex")
+
+### EEMEANS
+ modmc_wt_priorStrong %>% 
+  emmeans(~ ch_s,
+          at = list(ch_s = c(-0.8321022,2.23193),
+                    country = NA),
+          epred = TRUE) %>% 
+  contrast(method = "revpairwise")
+
+effect_draws <-  modmc_wt_priorStrong %>% 
+  emmeans(~ ch_s,
+          at = list(ch_s = c(-0.8321022,2.23193),
+                    country = NA),
+          epred = TRUE) %>% 
+  contrast(method = "revpairwise") %>% 
+  gather_emmeans_draws()
+
+# marginal mean 
+effect_draws %>% mean_hdi()
+
+# linear model, for EVALUES
+effect_draws_ln <-  modmc_wt_linear %>% 
+  emmeans(~ ch_s,
+          at = list(ch_s = c(-0.8321022,2.23193),
+                    country = NA),
+          epred = TRUE) %>% 
+  contrast(method = "revpairwise") %>% 
+  gather_emmeans_draws()
+
+# marginal mean 
+effect_draws_ln %>% mean_hdi()
+
+average_marginal_effect <- ggplot(autonomy_effect_draws, aes(x = .value)) +
+  stat_halfeye(fill = "dodgerblue") +
+  labs(x = "Average marginal effect of church attendance", y = "Density") +
+  theme_clean() +
+  theme(legend.position = "top")
+
+average_marginal_effect
+
+grand_mean <- modmc_wt_priorStrong %>% 
+  epred_draws(newdata = expand_grid(ch_s = seq(minchur, maxchur, by = 0.05)), 
+              re_formula = NA)
+
+  plot_grand_mean <- ggplot(grand_mean, 
+                                 aes(x = ch_s, y = .epred)) +
+  stat_lineribbon() +
+  scale_fill_brewer(palette = "Blues") +
+ # scale_fill_viridis_d(palette = "Blues") +
+  labs(x = "Frequency of religious service attendance (standard deviation units)", 
+       y = "Predicted anxiety (standard deviation units)",
+       fill = "95 HPD interval") +
+  theme_clean() +
+  theme(legend.position = "top")
+
+ plot_result <- average_marginal_effect + plot_grand_mean +  plot_annotation(tag_levels = "a")
+plot_result
+
+ggsave(
+  plot_result,
+  path = here::here("figs"),
+  width = 10,
+  height = 5,
+  units = "in",
+  filename = "plot_result.jpg",
+  device = 'jpeg',
+  limitsize = FALSE,
+  dpi = 1200
+)
+
+summary(modmc_wt_priorStrong)
+summary(modmc_wt_linear)
+
+#
+
+library(EValue)
+summary(ipb_ch_ln)
+summary(ipb_ch)
+summary(modmc_wt_linear)
+
+# sensitivity = Church
+evalues.OLS(-0.1093  , se = 0.0172, sd = 1, delta = 1.4)
+
+evalues.OLS(-0.2649 , se = 0.0648, sd = 1, delta = 1)
+
+evalues.OLS(-0.17 , se = 0.05, sd = 1, delta = 1)
+#1.61
+
+# multibias
+multi_evalue(est = RR(0.981), biases = selection())
+
+#
+bias_plot(rr = 0.857, xmax=5)
+
+multi_bound(multi_bias(confounding()),
+                RRAUc = 2.2, RRUcY = 1.7)
+
+biases <- multi_bias(confounding(), selection("S = U"),
+                         misclassification("exposure",
+                         rare_outcome = TRUE, rare_exposure = FALSE))
+
+```
+
+## Bayesian weighted regression correct sample
+
+```{r}
+## Bayesian wieghted regression
+
+```{r}
+b_ln <- brm(
+  anx_s|weights(ipwCH) ~ ch_s + (1 + ch | country),
+  data = dr,
+    prior = c(
+    prior(normal(0, .1), class = Intercept),
+    prior(normal(-.1,.1), class = b)),
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  backend = "cmdstanr",
+  file = here::here("mods","b_ln")
+)
+
+summary(b_ln)
+
+prior_summary(b_ln)
+summary(b_ln)
+
+library(brms)
+library(splines)
+
+listout <- readRDS(here::here("mods","listout"))
+modmc_wt_priorStrong <- brm(
+  anx_s|weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
+  data = d1p,
+  prior = c(
+    prior(normal(0, .1), class = Intercept),
+    prior(normal(-.1,.1), class = b)),
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  backend = "cmdstanr",
+  file = here::here("mods","modmc_wt_priorStrong")
+)
+
+##### TeST PRIORS
+modmc_wt_priorStrong_TEST <- brm(
+  anx_s|weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
+  data = d1p,
+  prior = c(
+    prior(normal(0, .1), class = Intercept),
+    prior(normal(-.1,.1), class = b)),
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  sample_prior = "only",
+  backend = "cmdstanr"#,
+ # file = here::here("mods","modmc_wt_priorStrong")
+)
+#saveRDS(modmc, here::here("mods", "modmc.rds"))
+summary(modmc_wt_prior)
+prior_summary(modmc_wt_prior)
+summary(modmc_wt_priorStrong)
+
+## WEAK PRIOR
+modmc_wt_prior_weak <- brm(
+  anx_s|weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
+  data = d1p,
+  prior = c(prior(normal(0, 1), class = b)),
+  seed = 1234,
+  warmup = 1000,
+  iter = 2000,
+  chains = 4,
+  sample_prior = "only",
+  backend = "cmdstanr"#,
+ # file = here::here("mods","modmc_wt_prior")
+)
+```
+
 ```

### 2022-03-12 History/analysis.Rmd
order files
create history foldrer

--- 2022-02-08 analysis.Rmd
@@
 ## https://cran.r-project.org/web/packages/parameters/vignettes/demean.html
-d1 <-
-  cbind(d1,
-        parameters::demean(d1, select = c("anx", "ch", "bg"), group = "country"))
-
-hist(d1$anx)
+# d1 <-
+#   cbind(d1,
+#         parameters::demean(d1, select = c("anx", "ch", "bg"), group = "country"))
+
+#hist(d1$anx)
 ```
@@
 
--   religion + ses + male + country + a different unmeasured latent variable1 (e.g. privilege) affects distress.
+-   religion + ses + male + country + ses
 
@@
 
-## Dag graph
+## DAG graph
 

### 2022-03-12 RBB-Publication-Files/bulbulia-RBB-2021-0063.tex
add rbb production files

% interactapasample.tex
% v1.05 - August 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
%\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled

\usepackage{hyperref}
\usepackage{listings}
\usepackage{setspace}

%\usepackage[longnamesfirst,sort]{natbib}% Citation support using natbib.sty
%\bibpunct[, ]{(}{)}{;}{a}{,}{,}% Citation support using natbib.sty
%\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% To set the list of references in 10 point font using natbib.sty

\usepackage[natbibapa,nodoi]{apacite}% Citation support using apacite.sty. Commands using natbib.sty MUST be deactivated first!
\setlength\bibhang{12pt}% To set the indentation in the list of references using apacite.sty. Commands using natbib.sty MUST be deactivated first!
\renewcommand\bibliographytypesize{\fontsize{10}{12}\selectfont}% To set the list of references in 10 point font using apacite.sty. Commands using natbib.sty MUST be deactivated first!

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

% JB's command for conditional independence
\newcommand{\indep}{\perp \!\!\! \perp}

\begin{document}

\articletype{Commentary MARP project}% Specify the article type or omit as appropriate

\title{A workflow for causal inference in cross-cultural psychology}

\author{
\name{Joseph A. Bulbulia \thanks{CONTACT Joseph A. Bulbulia. Email: joseph.bulbulia@vuw.ac.nz}}
\affil{Victoria University of Wellington, New Zealand// \href{https://orcid.org/0000-0002-5861-2056}{https://orcid.org/0000-0002-5861-2056}}
}

\maketitle

\begin{abstract}
The causal interpretation of a statistical association requires assumptions. Where the data are cross-sectional or cross-cultural these assumptions are even stronger. Here, I leverage a rigorous potential outcomes framework from contemporary epidemiology to (1) sharpen the causal question of whether religious service attendance reduces anxiety, and (2) develop a workflow for addressing this causal question using data from the Multiple Analysts of Religion Project (MARP, $N = 10,535$; 24 countries). This workflow clarifies how we may obtain a counterfactual contrast necessary to infer an average causal effect that is subject to (very) strong assumptions that causal inference requires in this setting. A sensitivity analysis helps to assess the robustness of the causal effect estimate to unmeasured confounding. This study is interesting because it clarifies the conditions under which the statistical association between religious service attendance and lower anxiety in the MARP dataset may have a causal interpretation. More generally, this study presents the potential outcomes framework as a rigorous method for investigating causal questions. I hope the potential outcomes framework will interest psychological scientists who seek better causal hygiene in their own research. 
\end{abstract}

\begin{keywords}
Causality, Causality Crisis, Culture, DAG, Distress, E-values, Marginal Structural Model, Potential outcomes, Religion, Sensitivity analysis, Target Trial, Well-being
\end{keywords}

\section{Introduction}

\subsection{Statistical associations do not quantify causal effects}

Statistical associations may be useful for predicting the future. Despite the prevalence of the term ``predict" in psychology journals, however, psychological scientists are rarely interested in prediction. Rather, we are interested in how cognition and behaviour works. We have causal interests. Controlled experiments are the gold standard for causal identification. However, many fundamental questions can only be addressed using observational data.

Psychological scientists typically use a framework of ``prediction" to interpret statistical associations in observational data. The terms ``predict" and "prediction" are meant to convey caution about causation. Absent a causal framework, however, the statistical associations upon which observational researchers base their predictions have little meaning. For this reason, reporting “predictions" in the absence of a causal framework is likely to mislead \citep{westreich_table_2013,mcelreath_statistical_2020,pearl_causal_2009,hernan_c-word_2018,bulbulia_causal_2021,rohrer_thinking_2018,vanderweele_explanation_2015}.

Here, I use the Many Analysts of Religion Project (MARP) data to develop a workflow for investigating the causal effects of religion on health in a cross-sectional, cross-cultural data-set \href{https://osf.io/v48gc/}{https://osf.io/v48gc/}. This workflow offers strategies for clarifying the causal assumptions required to obtain the counterfactual contrast that is necessary for quantifying a causal effect. Each causal inferential strategy must be developed bespoke for the scientific setting to which it is applied. Casual inference belongs to Savile Row not Sears Roebuck. However, because there are general features of any causal inference strategy -- such as obtaining credible counterfactual contrasts that are informed by previous science -- I hope that general features of my workflow will be helpful to researchers with different questions, data, and explanatory interests.

\subsection{The fundamental problem of causal inference}

Suppose we want to learn whether a dichotomous exposure has a causal effect. We must answer three questions: (1) What would happen with exposure? (2) What would happen with no exposure? (3) Do these potential outcomes differ?

Imagine a dichotomous exposure: exposed, $A = 1$; not-exposed, $A = 0$. $Y$ indicates the outcome for an individual who receives one of the two treatments: $Y|(A=1)$ or $Y|(A=0)$. We can observe only one exposure. We denote the potential outcome of $Y$ under the exposure $Y|(A=1)$ as $Y^{a=1}$. We denote the potential outcome of $Y$ under exposure $Y|(A=0)$ as $Y^{a=0}$. By physics, if we observe $Y^{a=1}$ we cannot observe $Y^{a=0}$ and vice versa. For this reason, we say that $Y^{a=1}$ and $Y^{a=0}$ are counterfactual or potential outcomes, which we write as $Y^{a}$, noting $(Y|(A=a) \neq Y^{a}$. This constraint in which we cannot observe the individuals outcomes that we require to evaluate a response under different exposures is called ``the fundamental problem of causal inference" \cite{rubin_inference_1976,holland_statistics_1986, gelman_bayesian_2020}.

\subsection{Assumptions for estimating an average causal effect in a population}

Despite the limitations of quantifying a causal effect of two or more exposures in an individual, we may be able to quantify an average causal effect of two or more exposures in a population:

$$
\begin{aligned}
 E[Y^{a=1}]-E[Y^{a=0}]  
\end{aligned}
$$

We read this as the difference in expected mean outcome when the population receives the exposure $E[Y^{a=1}]$ and the expected mean outcome when the population is withheld from the exposure $E[Y^{a=0}]$. Again, we can only obtain the statistical expectation under actual exposures $E[Y|A]$: we do not directly obtain the counterfactual expectation for the entire population. We cannot cheat physics. However, we may consistently estimate the counterfactual average response $E[Y^a]$ from statistical average response $E[Y|A]$ if the following conditions are satisfied:

\begin{enumerate}
    \item {\it Positivity}: there is a non-zero probability of the exposure among the non-exposed and of non-exposure among the exposed. That is, the exposure cannot be deterministic. 
    \item {\it Consistency}: an individual with observed exposure $A$ has observed outcome $Y$ equal to their counterfactual outcome $Y^{A}=Y^a$.
    \item {\it Exchangeability}: individuals are exchangeable when counterfactual outcomes are independent of the actual exposures. To describe conditional independence, we use the symbol: $\indep$. When  $Y^a\indep A$,  exchangeability is satisfied. In a randomised experiment, random allocation of participants to exposures will typically ensure exchangeability. In observational studies, however, exchangeability typically requires conditioning on a set of measured differences ($L=l$) such that $Y^a\indep A|L=l$. We say the counterfactual outcome of Y under exposure $A=a$ is conditionally independent of exposure assignment given a set of measured confounders $L=l$.
    \item {\it Correct model specification}: which includes that the outcome and the exposure are measured without error. The assumption is strong, unverifiable, and unlikely to be satisfied in observational settings. Failure of this assumption does not imply that causal inference in observational settings is impossible. It rather implies caution about claiming too much.
\end{enumerate}

I have stated the assumptions that we require for causal inference. We are now ready to proceed to the next step: asking a causal question \citep[for introductions to the key concepts in causal inference, see:][]{gelman_arm_2020, hernan_causal_2020,vanderweele_explanation_2015}.

\subsection{What is our causal question?}

In the abstract, I present the causal question: does religious service reduce anxiety? This question is motivated by previous theory and evidence that ritual behaviour reduces anxiety \citep[see:][]{malinowski_magic_2014,sosis_psalms_2007,lang_effects_2015, ejova_church_2020}. 
I am unaware of any evidence that religious service attendance, on average, causes a decrease in anxiety. However, it is unclear whether the buffering effect of religious service attendance for anxiety is cross-culturally robust.

If the assumptions for unbiased causal inference that I just considered were met, the MARP dataset could help us to compute the average causal effect of religious service attendance and, therefore, help to assess whether the average effect of religious service attendance for anxiety reduction is cross-culturally robust. However, my causal question has yet to be clarified. The definitions both of the exposure and the outcome remain vague. First, consider the exposure. Which specific interventions in religious service attendance do we hope to counterfactually contrast? Are we interested in contrasting a counterfactual intervention in which all people who did not attend religious service were made to attend? Or are we interested in contrasting a counterfactual intervention in which all people who participated in religious service were prevented from doing so? Or do we imagine that interventions by increments of attendance, up or down, constitute the interventions of interest? If so, what are the units of these increments? Causal inference requires satisfaction of the positivity assumption, that is, of a non-deterministic exposure assignment. The positivity assumption is violated if there is no chance that individuals could receive different exposures from the actual exposures that they received. VanderWeele raises this question for religious service attendance and argues for restriction of analysis to religiously-identified sub-samples \cite{vanderweele_causal_2017}
(For scepticism about comparisons of church attendance between secular and religious samples, see: \citet{vanderweele_religion_2017}, 
and for a related discussion focusing on the causal effects of “race" see: \citet{vanderweele_causal_2014}).
This objection is intuitively plausible. We cannot administer religious service attendance to an atheist like we might administer a medication.

Next, consider the outcome. Suppose we were able to measure anxiety without error. How long after initiating a change in behaviour would we expect the effects of religious service attendance on anxiety to manifest? One week? One month? Many months? A year? I have not specified when the effect of service attendance on anxiety is hypothesised to occur. Imagine that religious service attendance were to cause anxiety, which in turn leads to the attenuation of service attendance among those susceptible to service-induced anxiety. In such a circumstance, we would be left with an observed sample of individuals robust to the anxiety-provoking effect. Such a bias is called “immortality bias" \citep[see:][]{hernan_using_2016, hernan_specifying_2016}. The sample that will observe may not be the treatment sample.

In short, we cannot consider a counterfactual contrast without understanding the potential intervention and the potential outcome that we are contrasting and without ensuring that the four conditions necessary for causal inference have been satisfied. I have yet to provide this mission-critical clarity.

\subsection{Emulating a “target trial" to answer a causal question}

Miguel Hern\'{a}n and Jamie Robins suggest that observational researchers clarify their causal question by imagining an idealised experiment, or “target trial," in which researchers might ideally evaluate the effect of an intervention \citep{hernan_causal_2020, hernan_specifying_2016, hernan_using_2016}. 
Such an idealised experiment is typically implausible. If it were not implausible, there would be no point in using observational data rather than running the experiment. However, H\`{e}rnan and Robins argue that if researchers cannot state an idealised experiment for addressing a causal question -- one that clearly defines the outcome, the exposure, the contrasts, and the study design, then researchers might not know which causal question their analysis is answering. Additionally, H\`{e}rnan and Robins argue that by emulating an idealised target trial, observational researchers are better able to clarify their causal assumptions. In the present setting, where data collection is cross-sectional, we cannot remotely emulate a target trial. However, by clarifying how we fall short of the experiment we might want to perform to infer a causal effect our assumptions become clearer.

\paragraph*{What are the eligibility criteria for participation?}

All people who responded to the MARP study are eligible for inclusion in this hypothetical target trial. This consists of ($N=10,595$) individuals across 24 countries. Both religious service non-attenders and religious service attenders are included in the sample. I present the breakdown of individuals by their level of religious identification in Table \ref{tab:table1}. Notably, among those who state that they are “atheists", only $41.7\%$ report never going to religious service. Among those who state that they are “not religious," only $47.5\%$ report “never" going to religious service (see Table \ref{tab:table1}). Positivity is actual, and therefore possible, so the analysis I present here uses the full sample. However, I also perform a separate analysis restricted to those who identify as religious ($N = 3,712$) (The inference remains the same, see: Appendix \ref{app:relonly}).

\begin{table}[h!bt]
\caption{Cross tabulation of religious service attendance frequency and religious identification}
\scalebox{.61}{
\begin{tabular}{lllllllll}
\toprule
  & Never & Rarely & 1 $\times$ per year & Holy Days & 1 $\times$ month & 1 $\times$ week & \textgreater1 $\times$ per week & Overall\\
\midrule
 & (N=5061) & (N=1236) & (N=798) & (N=1371) & (N=618) & (N=957) & (N=494) & (N=10535)\\
Religious &  &  &  &  &  &  &  & \\
  Atheist & 2110 (41.7\%) & 172 (13.9\%) & 97 (12.2\%) & 101 (7.4\%) & 12 (1.9\%) & 8 (0.8\%) & 5 (1.0\%) & 2505 (23.8\%)\\
  Not-Religious & 2404 (47.5\%) & 686 (55.5\%) & 406 (50.9\%) & 544 (39.7\%) & 134 (21.7\%) & 107 (11.2\%) & 37 (7.5\%) & 4318 (41.0\%)\\
  Religious & 547 (10.8\%) & 378 (30.6\%) & 295 (37.0\%) & 726 (53.0\%) & 472 (76.4\%) & 842 (88.0\%) & 452 (91.5\%) & 3712 (35.2\%)\\
\bottomrule
\end{tabular}
}
\label{tab:table1}
\end{table}

\paragraph*{Treatment strategies} The MARP data are cross-sectional. We can only condition on the assumption that, after accounting for all measured confounding and performing a sensitivity analysis for unmeasured confounding, the co-variance of religious service attendance and anxiety quantifies a causal effect.

At first glance, this assumption might cast doubt about whether causal inference is sensible. How can we evaluate a causal effect that we assume? However, I have described experimental and national longitudinal evidence that religious service attendance buffers psychological distress. By conditioning my analysis on the causal assumption of potential buffering, cross-sectional data may help to evaluate whether the buffering effect of religious service on anxiety observed in previous research is robust to cultural variation.

\paragraph*{Randomisation} I generate a pseudo-randomised population using IPTW weighting of the exposure on measured confounders \citep{cole_constructing_2008,wal_ipw_2011}. 
IPTW weighting allows us to estimate a marginal (or population-level mean) effect that is balanced across the pseudo-population. IPTW weights are less efficient than stratified regression, however, standard stratification methods fail in longitudinal settings \citep{robins_new_1986}. 
Because the estimates from the stratified regression models and the IPTW  models are nearly identical and do not affect causal inference, I report results from the IPTW model.

I include in the set of all measured confounders $L$, any variable that might affect both the exposure and the outcome. I must ensure that this set does not include any variable that blocks the path between exposure and the outcome (a mediator) and that is not an effect of both exposure and the outcome (a ``collider" confounder) \citep{mcelreath_statistical_2020,bulbulia_causal_2021}
[For discussion see: Appendix \ref{app:dags}].

\paragraph*{Start and follow up} The data are cross-sectional: there can be neither an initiation of a start nor a termination by the end of follow up. Thus, inference requires the assumption that the causal effects of the exposure are immediate and that prior levels of religious service and prior levels of anxiety do not affect present levels of religious service attendance or current levels of anxiety. Again, my causal estimand will be biased if there are time-dependent dynamics, such as attrition of those for whom religious service is stressful.

\paragraph*{Outcomes} I assess anxiety using self-reported measures as described in the MARP protocols \href{https://osf.io/vy8z7/}{https://osf.io/vy8z7/}. The MARP anxiety measure asks: “How often do you have negative feelings such as blue mood, despair, anxiety, depression?" This item was measured on a Likert scale: (1) “Never" (2) “Seldom" (3) “Quite often" (4) “Very Often" (5) “Always." I standardise and scale this outcome at the sample average. I assume that anxiety is measured consistently and without error. The no-measurement error assumption is strong. However, failure of this assumption will not necessarily lead to bias against the causal null hypothesis. Non differential and non-directed measurement error in which measurement of the exposure and outcome are not correlated will attenuate the estimate of a true causal effect. By contrast, differential or directed measurement error risks overstating the true causal effect \citep{hernan_causal_2020}. 
I cannot test the no-measurement error assumption in the MARP dataset. However, I can perform a sensitivity analysis as part of my workflow. Sensitivity analysis is important for causal inference in observational research because we cannot generally know whether the assumptions required for causal inference have been met.

\paragraph*{Causal contrasts} The original measure of religious service attendance asks: “Apart from weddings and funerals, about how often do you attend religious services these days?" Answers are: (1) “More than once a week" (2) “Once a week" (3) “Once a month" (4) “Only on special holy days" (5) “Once a year" (6) “Less often" (7) “Never, or practically never." The MARP team reverse-scored these responses and scaled them to 0-1 \href{https://osf.io/vy8z7/}{https://osf.io/vy8z7/}. Given that responses on this scale are ordinal, I model religious service attendance as a monotonic effect \citep{burkner_bayesian_2021}. 
My focal contrast is “no attendance" versus “attendance more than once per week."  (My choice of statistical model here, as elsewhere, makes no practical difference to inference, see: Appendix \ref{tab:bayesmonotonic})

I also estimate counterfactual outcomes using a non-linear continuous model of religious service attendance exposure. These models were implemented using regression splines \citep{r_core_team_r_2021}. 
This second approach clarifies the robustness of the inference to model choice, see Appendix \ref{app:monotonic}.

\paragraph*{Analysis plan} I examine missing data using the Naniar package \citep{tierney_naniar_2021} 
in R \citep{r_core_team_r_2021}. 
High levels of missingness in the denomination variable prevents confounder adjustment for denominational affiliations ($n$ missing = $5,676$, or $53\%$ of the responses). To address selection bias from missingness in other variables, I multiply-impute ten missing data sets using both the Amelia package \citep{honaker_amelia_2011} 
and the Mice package \citep{buuren_mice_2011}
in R \citep{r_core_team_r_2021}.

I estimate a (quasi)-Bayesian mixed effect model using \citep{burkner_bayesian_2021}, 
see Appendix \ref{app:margmod}. Note my statistical model is not fully Bayesian because I obtain IPTW estimates using the IPW package in R \citep{wal_ipw_2011}. 
I also estimate a marginal structural model using a Generalised Estimating Equations (GEE) using weighted (and stratified) confounder adjustment strategies to ensure that estimation \citep{halekoh_r_2006}. 
GEE models also enable adjustment for country-level dependencies. My inference proved to be practically equivalent (see online materials). However, my preferred method for estimating the marginal effect of religious service attendance on anxiety is a (quasi)-Bayesian mixed-effect model with IP weights. In this model, I marginalise over the country-level variances by integrating out group level effects. Bayesian estimation has many valuable features, including clarity about model assumptions and the recovery of intuitive posterior probabilities.\footnote{Note that for time-series models with time-varying confounding Bayesian estimation remains experimental. Although this topic does relate to the questions at hand, I want to caution readers against any simple adaptation of regression based models to panel data.}

Following \citet{gelman_bayesian_2020} 
and \citet{mcelreath_statistical_2020}, I employ strong priors in our Bayesian models, and compare these to weaker priors. (Results were robust to the choice of priors, see Appendix \ref{app:priors}).

To recover an estimate of a marginal causal effect I must integrate out the country-level variation in intercepts and slopes. For this purpose, I used the emmeans package in R \citep{lenth_emmeans_2022}. 
The analysis can be found at \href{https://github.com/go-bayes/many_analysts}{https://github.com/go-bayes/many\_analysts}.

\section{Results}

\subsection{Causal Graph}

Causal inference requires causal assumptions \citep{vanderweele_explanation_2015,pearl_causal_2009, pearl_seven_2019,mcelreath_statistical_2020,vanderweele_outcome-wide_2020, bulbulia_causal_2021}. 
Following \citet{vanderweele_explanation_2015}, I include in my list of confounders the full set of demographic variables in the MARP dataset: these were \{age, country, education, male, SES\}. High levels of missing data prevented me from including ethnicity ($n  = 219$ missing) or denomination ($n=5,446$ missing). I then identify the minimal adjustment set required to quantify our causal associations. I do this to avoid introducing confounding by over-conditioning (see Figure \ref{fig:Figure1}b/c/d).

Figure \ref{fig:Figure1}.a presents a Direct Acyclic Graph that clarifies my causal assumptions. Following H\`{e}rnan, I present a simplified graph that shows as few nodes as are necessary to convey my causal assumptions. I identify the adjustment set of confounders as: $L$ \{age, country, education, male, SES\}. As indicated in Figure \ref{fig:Figure1}.b, conditioning on a confounder that is a collider of two unmeasured confounders will introduce bias. Figure \ref{fig:Figure1}.c, describes a special case, in which an unmeasured confounder of the mediator and outcome can artificially increase the effect of exposure on an outcome can introduce bias. I do not “control" for variables that religious service attendance might affect, such as religiosity. To do so would be to condition on a post-treatment variable or mediator and would partially block the effect of religious service attendance. In Appendix \ref{app:simulation}, I simulate post-treatment confounding. On the problem of controlling for post-treatment variables, \citep[see:][]{westreich_table_2013}).

\begin{figure}[h!bt]
\centering
    \caption{Panel (a) presents my strategy for confounder control. I assume that adjustment for set of measured confounders $L$ \{age,country, education, ses\} is sufficient to close all backdoor paths between exposure $A$ (service attendance) and the outcome $Y$ (anxiety). This assumption is almost certainly too strong, hence I also perform sensitivity analysis. Panel (b) presents the risk of over-conditioning by adjusting for a measured confounder that opens a closed path between unmeasured confounders liking $A$ and $Y$. Panel (c) presents the risk of conditioning on a measured confounder that is jointly caused by the exposure and and an unmeasured confounder. For example, imagine distressing experiences were to increase belief in God, and that church attendance also causes belief in God. Suppose belief in God does not affect anxiety, but church attendance does. In this case, conditioning on church attendance may artificially increase the negative statistical association between church attendance and distress. Indeed including belief in God in the regression model has this effect. This DAG illustrates another peril of including variables in a regression model that are not required to ensure exchangeability. I include a simulation in Appendix \ref{app:simulation} that replicates such confounding. Panel (d) presents the threat of confounding by selection bias, a form of collider confounding, in which an association between $A$ and $Y$ is opened because sampling is biased. Such confounding is likely from MARP data, however sensitivity analysis clarifies how much confounding would be required to negate an inferred causal association.} 
    \includegraphics[width=0.99\columnwidth]{figs/daggraph.jpg}
    \label{fig:Figure1}
\end{figure}

\subsection{Marginal Effects}

I consider the marginal predictions of the Bayesian estimator to give the most accessible and straightforward estimate of our uncertainty for the marginal effect of church attendance on anxiety. Figure \ref{fig:Figure2}.a presents the marginal effect estimate contrasting zero religious service attendance with those who participate more frequently than once a week: (ACE = -0.186, 95\% HDP(-0.323, -0.0586)). Figure \ref{fig:Figure2}.a graphs average causal effect of the exposure. Figure \ref{fig:Figure2}.b graphs the predicted marginal effects of religious service attendance at the extremes.

\begin{figure}
\centering
    \caption{Panels (a) average causal effect of religious service on anxiety. Full church attendance may cause a .2 SD average reduction in anxiety; (b) Predicted marginal means for attending “more than once per week" versus attending “never."}
    \includegraphics[width=0.9\columnwidth]{figs/plot_result_ord2.jpg}
    \label{fig:Figure2}
\end{figure}

Conditional on the strong assumptions described above, we find an average causal effect of religious service attendance on anxiety across the populations represented in the MARP dataset.

\subsection{Sensitivity Analysis}

The EValues package in R provides a user-friendly tool for investigating the sensitivity of a causal estimate to unmeasured confounding \citep{vanderweele_sensitivity_2017, mathur_website_2018,r_core_team_r_2021}.
An E-value for unmeasured confounding is defined as ``the minimal strength of association on the risk ratio scale that an unmeasured confounder (or confounders) would require in its association with both the exposure and the outcome to explain away a causal effect" \citep{mathur_website_2018}. 
I obtain an E-value for the marginal effect estimate in the monotonic mixed-effects model value (E-value = 1.65, upper limit = 1.28). I infer that an unmeasured confounder associated with both the exposure and the outcome by risk ratios of between 1.65-fold to 1.28-fold each would be sufficient to nullify the causal association obtained from the statistical model. In the context of a cross-sectional comparative study in which we might expect strong demand characteristics, measuring error, and sampling bias, I characterise the evidence for the causal effect of religious service attendance on anxiety reduction as modest. Note that it is possible that the actual causal effect is more substantial than what I have estimated here. A stronger true causal relationship is possible if there is measurement error both in the exposure and in the outcome and if these sources of error are independent.

\section{Discussion}

Conditional on the strong assumptions I have described, the MARP dataset offers modest evidence for a moderate average causal effect of religious service attendance for reducing anxiety. This effect is robust both to cultural variation and a small amount of unmeasured confounding. What else have we learned?

First, although causal inference resembles ordinary regression, the assumptions required for causal inference must be considered with care. Absent a potential outcomes causal framework, estimating associations from a regression model is misleading for causal inferential tasks. {\bf Do not report regression coefficients as causal effects unless you have clarified your causal assumptions.}

A corollary: \textbf{all regression tables may be safely moved to the supplements. You need only report the coefficients that pertain to your causal interests: typically that is one coefficient.}

Second, causal estimation requires stating one's assumptions about the data and one's model. Formulating such assumptions require expert knowledge about the topic at hand. The data do not generally contain sufficient information to warrant these assumptions. Here, I have written a statistical model whose features rely on previous theory and research. Based on this research I have assumed that habitual religious service attendance is causally antecedent to religious identification and other indicators of religiosity. For this reason I have not included these indicators because to do so would introduce post-treatment confounding.  I have also assumed that demographic indicators such as age, education, gender, SES, and country of residence may affect both church attendance and anxiety. All causal assumptions cannot generally be derived entirely from the data alone, it is important to remember that expert knowledge is constantly undergoing revision. Therefore: {\bf state your assumptions explicitly and graphically} and {\bf when expert knowledge is in doubt, perform multiple analyses} (see \citet{bulbulia_causal_2021}).

Third, I have presented a workflow based on Hern'{a}n  and Robins's strategy of specifying and emulating a “target trial" \citep{hernan_specifying_2016}. 
In this study, the target trial that I have specified falls well short of any experimental standard. For example, no experiment would enlist people after they have received their treatment. However, by imagining a hypothetical intervention that could test a causal hypothesis we can better appreciate both the strengths and the shortcomings of an observational study. It is easy to focus on the limitations of observational data such as the MARP dataset. However, cultural data need not be discarded because they were not obtained from experiments or because repeated measures are lacking. Rather, what we need is clarity about our causal assumptions. {\bf Clarify your casual question by stating and emulating a target trial. How your target trial fails will clarify your assumptions. Remain open to the possibility your question cannot be answered, either because the question is too vague or because the data cannot resolve it.}

Fourth, in observational research although we cannot generally guarantee there is no unmeasured confounding we can usually perform a sensitivity analysis to investigate robustness of a causal estimate to unmeasured confounding. The EValues package in R is freely available. Moreover there is an online calculator at: \href{https://www.evalue-calculator.com/}{https:\/\/www.evalue-calculator.com\/}. {\bf Perform a sensitivity analysis to evaluate the robustness of your causal effect estimate.}

Fifth, causal inference is the holy grail of scientific research. However in the psychological sciences, there has been relatively little discussion about improving workflows for causal inference \citep[although see:][]{rohrer_thinking_2018}. 
I have argued that the use of non-causal language such as “prediction" and “predicts" may be misleading. A prediction is not a conservative causal estimate. Absent a causal framework, a prediction has no causal implication at all. For this reason, researchers who limit observational inference to “prediction" invite misleading interpretations of their research. They also obscure their theoretical interests, which are generally causal \citep{hernan_c-word_2018}. Help is available. {\bf Do not settle for reporting a “prediction" from a statistical ``association." Instead, report your attempt to estimate an average causal effect, and perform a sensitivity analysis.}

A final comment. The development of powerful frameworks for causal inference during the past several decades has resulting in “causal revolution" across many sciences \citep{pearl_book_2018}. During this same period, however, most observational research in psychology has operated in the absence of any explicit causal framework. The widespread adoption of causality-naive practices in observational psychology has resulted in what might be deemed a ``causality crisis." The causality crisis in observational psychology has parallels to the replication crisis in experimental psychology -- causal naivety casts nearly all previous findings into doubt. Unlike the replication crisis in experimental psychology, however, the magnitude of the causality crisis remains to be fully appreciated. For those unfamiliar with the potential outcomes framework, I hope this study will stimulate your interest in developing a more systematic and rigorous approach to causal inference.

\section*{Acknowledgements}
I am grateful to the MARP team and their funders for obtaining the data used in this project, and to Templeton Religion Trust grant TRT0196 for support of my time. I am grateful to Dr. Usman Afzali, Professor John Shaver, and an anonymous reviewer for feedback.

\section*{Disclosure statement}

The author declares that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

\section*{Funding}
This research was supported by Templeton Religion Trust Grant 0196 and Royal Society of New Zealand Grants 19-UOO-090 and Marsden 3721245. A complete description of the data (including missing data) and statistical analysis has been posted to \url{https://github.com/go-bayes/many_analysts}.

\section*{Notes on contributor}
 Joseph A. Bulbulia is a professor of Psychology and Director of the Centre for Applied Cross-Cultural Research at Victoria University in Wellington, New Zealand. His research focuses on the evolutionary and longitudinal study of religion, cooperation, and psychological health. He is one of four senior managers for the New Zealand Attitudes and Values Study, a researcher for the Max Plank Institute for the Science of Human History.%\\
%\href{https://josephbulbulia.netlify.app}{https://josephbulbulia.netlify.app}\\
%\href{https://orcid.org/0000-0002-5861-2056}{https://orcid.org/0000-0002-5861-2056}

\bibliographystyle{apacite}
\bibliography{references2}

\newpage
\section*{Appendices}

\appendix

\section{Priors for Bayesian estimation}
\label{app:priors}

\subsection{Priors for monotonic predictor model}

In an ordinal predictor model where there are $R$ indicator responses, there are $C = R-1, c = 1\dots C$ thresholds to estimate the ordered responses predictors. In this model, $b^c$ locates the direction and size of the effect. That is $b^c$ locates the expected average difference between two adjacent categories of the ordinal predictor. Such models also require the parameter $\zeta^c$, which locates the normalised distances between consecutive predictor categories, producing the shape of each monotonic effect:

$$
\begin{aligned}
g(y_i) = \alpha + b^c \zeta^c x_i
\end{aligned}
$$

For the monotonic model I used the following strong priors:

$$
\begin{aligned}
y_{ij} \sim \text{Normal}(\mu_{ij}) \nonumber\\
\text{Normal}(\mu_{ij})\sim \boldsymbol{\alpha} + \boldsymbol{b} \boldsymbol{\zeta} \nonumber\\
\boldsymbol{\alpha} \sim \alpha_{0} +\alpha_j \nonumber\\
\alpha_{0} \sim \text{Student~T}(3, 0,.25)\nonumber\\
%\alpha_j \sim \text{StudentT}^+(3,0,10)\nonumber\\
\boldsymbol{\beta}\sim \boldsymbol{b} + \beta_j \nonumber\\
\boldsymbol{b}\sim \sim \text{Normal}(-.05,.1) \nonumber\\
\boldsymbol{\zeta}\sim \text{Dirichlet}(1)
\nonumber\\
\begin{bmatrix}
\alpha_j \\
\beta_j
\end{bmatrix}
\sim 
\text{MVNormal}
\begin{pmatrix}
\begin{bmatrix}
0\\
0
\end{bmatrix}
,\boldsymbol{S} \nonumber
\end{pmatrix}\\
\boldsymbol{S} = 
\begin{pmatrix}
\sigma_{\alpha_j} & 0 \\
0 & \sigma_{\beta_j} \nonumber
\end{pmatrix} \boldsymbol{R} \begin{pmatrix}
\sigma_{\alpha_j} & 0 \\
0 & \sigma_{\beta_j} \nonumber
\end{pmatrix}\\
\sigma_{\alpha_j} \sim 
\text{Student~T}(3,0,2.5) \nonumber\\
\sigma_{\beta_j} \sim 
\text{Student~T}(3,0,2.5) \nonumber\\
\boldsymbol{R}\sim \text{LKJcorr}(1) \nonumber
\end{aligned}
$$

Here, the vector $\boldsymbol{\zeta}$ contains six coefficients for the $c = 1\dots 6$ simplex parameters estimated for the thresholds.

\newpage
\subsection{Priors for continuous predictor model}

The strong priors for the continuous predictor model were as follows:

$$
\begin{aligned}
y_{ij} \sim \text{Normal}(\mu_{ij}) \nonumber\\
\text{Normal}(\mu_{ij})\sim \boldsymbol{\alpha +\beta} \nonumber\\
\boldsymbol{\alpha} \sim \alpha_{0} +\alpha_j \nonumber\\
\alpha_{0} \sim \text{Student~t}(3, 0,.25)\nonumber\\
%\alpha_j \sim \text{StudentT}^+(3,0,10)\nonumber\\
\boldsymbol{\beta}\sim \boldsymbol{b} + \beta_j \nonumber\\
\boldsymbol{b}\sim\text{Normal}(-.05,.1) \nonumber\\
\begin{bmatrix}
\alpha_j \\
\beta_j
\end{bmatrix}
\sim 
\text{MVNormal}
\begin{pmatrix}
\begin{bmatrix}
0\\
0
\end{bmatrix}
,\boldsymbol{S} \nonumber
\end{pmatrix}\\
\boldsymbol{S} = 
\begin{pmatrix}
\sigma_{\alpha_j} & 0 \\
0 & \sigma_{\beta_j} \nonumber
\end{pmatrix} \boldsymbol{R} \begin{pmatrix}
\sigma_{\alpha_j} & 0 \\
0 & \sigma_{\beta_j} \nonumber
\end{pmatrix}\\
\sigma_{\alpha_j} \sim
\text{Student~T}(3,0,2.5) \nonumber\\
\sigma_{\beta_j} \sim 
\text{Student~T}(3,0,2.5)  \nonumber\\
\boldsymbol{R}\sim \text{LKJcorr}(1) \nonumber
\end{aligned}
$$

As indicated in Figure \ref{fig:Figure3}, choice of priors did not affect the location estimates in either the monotonic or continuous model of religious service attendance on anxiety.

\begin{figure}
\centering
    \caption{Panel (a) presents posterior predictions sampling only from the weakly default priors for ordinal church attendance model. These priors allow location estimates for each monotonic location with up to 10 standard deviations width to either side of zero. Panel (b) presents posterior predictions sampling only from the stronger priors I used for the ordinal religious attendance model. Panel (c) presents posterior predictions sampling only from the weakly default priors for the continuous religious attendance model. Here too, the default priors allow location estimates of up to 10 standard deviations to either side of zero across the range of church attendance responses. Panel (b) presents posterior predictions sampling only from the stronger priors I used for the continuous church attendance model. Estimates from weakly and strongly informative models in both the monotonic and continuous models converged to within MCMC error. This result indicates the posteriors parameter estimates were informed by the data, not by the priors.}
    \includegraphics[width=0.9\columnwidth]{figs/test_priors.jpg}
    \label{fig:Figure3}
\end{figure}

\newpage
\section{Monotonic and continuous models of religious service: no difference.}
\label{app:monotonic}

Estimates for the monotonic model were:

\begin{table}[h!bt]
\caption{Parameter estimates for the monotonic effect model of religious service on anxiety}
\centering
\begin{tabular}[t]{llllllll}
\toprule
Parameter & Component & Median & CI & CI\_low & CI\_high & pd & Rhat\\
\midrule
Intercept & conditional & 0.044 & 0.95 & -0.043 & 0.139 & 0.834 & 1.006\\
bsp\_ord & conditional & -0.031 & 0.95 & -0.054 & -0.01 & 0.997 & 1.004\\
simo\_ord1[1] & simplex & 0.27 & 0.95 & 0.133 & 0.411 & 1 & 1\\
simo\_ord1[2] & simplex & 0.139 & 0.95 & 0.04 & 0.262 & 1 & 1\\
simo\_ord1[3] & simplex & 0.104 & 0.95 & 0.028 & 0.206 & 1 & 1.001\\
\addlinespace
simo\_ord1[4] & simplex & 0.11 & 0.95 & 0.036 & 0.206 & 1 & 0.999\\
simo\_ord1[5] & simplex & 0.159 & 0.95 & 0.053 & 0.278 & 1 & 0.999\\
simo\_ord1[6] & simplex & 0.185 & 0.95 & 0.06 & 0.329 & 1 & 1\\
sigma & sigma & 0.971 & 0.95 & 0.958 & 0.983 & 1 & 0.999\\
\bottomrule
\end{tabular}
\label{tab:bayesmonotonic}
\end{table}

Estimates for the continuous model were:

\begin{table}[h!bt]

\caption{Parameter estimates for continuous (spline) model of religious service on anxiety}
\centering
\begin{tabular}[t]{llllllll}
\toprule
Parameter & Component & Median & CI & CI\_low & CI\_high & pd & Rhat\\
\midrule
Intercept & conditional & 0.036 & 0.95 & -0.053 & 0.12 & 0.791 & 1.011\\
ch\_s1 & conditional & -0.104 & 0.95 & -0.218 & 0.011 & 0.958 & 1\\
ch\_s2 & conditional & -0.065 & 0.95 & -0.202 & 0.061 & 0.842 & 1.002\\
ch\_s3 & conditional & -0.147 & 0.95 & -0.261 & -0.041 & 0.996 & 1.004\\
sigma & sigma & 0.971 & 0.95 & 0.958 & 0.984 & 1 & 1\\
\bottomrule
\end{tabular}
\label{tab:bayescontinuous}
\end{table}

As indicated in Figure\ref{fig:Figure4}, marginal effects estimates for the monotonic religious service attendance model and the continuous religious service attendance model do not lead to practical difference in the causal inference.

\begin{figure}
\centering
    \caption{Panels (a) and (b) graph the marginal effect estimates and marginal predictions for a continuous model of church attendance, estimated with splines. In Panel (a) the marginal contrast is taken from the extremes of the range attendance range: more than once per week vs never. Panels (c) and (d) graph the marginal effect estimates and marginal predictions for a monotonic predictor model of church attendance, where again the contrast is taken from the extremes of the range. There is no practical difference between these model choices. I select the monotonic model because the categories of religious service responses are ordinal choices.}
    \includegraphics[width=0.9\columnwidth]{figs/plot_result_full.jpg}
    \label{fig:Figure4}
\end{figure}

\newpage
\section{Religious-only sub-sample}
\label{app:relonly}
Vanderweele observes that positivity might not hold among those who do not identify as religious. I have estimated causal effect estimates over the entire sample because I observe that over half of the non-religious sample attend religious service. However, I do not believe that causal analysis should be restricted to a single option. Some may find it useful to quantify the causal effects in the religious-only sample.

Using the identical procedure to that of the full sample estimate (Bayesian mononotic mixed effect model) we obtain an average marginal effect of (ATE = -0.21, HPD[0.34, -0.08]. we estimate an E-value of 1.71, with a lower bound of 1.35, indicating a stronger causal effect of religious service attendance in diminishing anxiety in the only-religious sample (see: Figure \ref{fig:graphR}).

\begin{figure}[h!bt]
\centering
    \caption{Causal effect of religious service on anxiety in the religious-only sample}
    \includegraphics[width=0.95\columnwidth]{figs/graph_r.jpg}
    \label{fig:graphR}
\end{figure}

\newpage
\section{Marginal Structural Model}
\label{app:margmod}

Our counterfactual contrast of interest is the difference in the expected average response in the (entire) population under exposure and the expected average response in the (entire) population under no exposure. Here, we compare the non-attending exposure with the attending-more-than-once-per-week exposure. %\footnote{It can be shown that this contrast is equivalent to the average of the expected differences $E[Y^{a=1} -Y^{a=0}]$ (cite Hernan).}

It is essential to realise that estimating a statistical difference in the observed associations in our data:

$$
\begin{aligned}
E[Y|A=1]-E[Y|A=0]
\end{aligned}
$$

does not provide the contrast that we require for causal inference, in which the whole population does, or does not, receive the exposure:

$$
\begin{aligned}
E[Y^{a=1}] - E[Y^{a=0}]
\end{aligned}
$$

Rather than estimating the expectation of the observed exposures, as in standard regression analysis, a marginal structural model estimates:

$$
\begin{aligned}
E[Y^a] = \beta_0 + \beta_1a 
\end{aligned}
$$

where:

$$
\begin{aligned}
E[Y^{a=0}] = \beta_0
\end{aligned}
$$

and:

$$
\begin{aligned}
E[Y^{a=0}] - E[Y^{a=1}] = \beta_1a
\end{aligned}
$$

A marginal structural model, then, allows us to estimate a difference in average counterfactual expectations (or equivalently, an average difference in counterfactual expectations). The method requires satisfaction of the assumption that those who received an exposure are similar to those who did not. To estimate a marginal structural model, we weight responses by the inverse probability of exposure; that is, we use IPTW weighting. In the simulated population that we obtain from IPTW weighting: $E[Y|A] =\theta_0 + \theta_1A$, where $\theta_1A$ is equivalent to $\beta_1a$ in the counterfactual model \citep{hernan_causal_2020}.

\newpage
\section{Graphical methods for confounder control}
\label{app:dags}
There are four primary conditions for confounder control:  (1) There is no unopened or “backdoor" path leading from exposure to outcomes: we block a backdoor path by conditioning on a variable along the path; (2) There is no intervening mediator between an exposure and an outcome has been conditioned on. Unless we are interested in mediation, we open a path by omitting this mediator from analysis; (3) There is no common outcome of the exposure and outcome that has been conditioned on. Conditioning on a “collider" will induce a spurious association that is removed by omitting the collider from the analysis; (4) No descendants of a collider have been conditioned on, which we address by omitting any descendants from the analysis.

\newpage
\section{Simulation of mediation confounding}
\label{app:simulation}

The simulation (in R) reported below clarifies the bias of conditioning on a mediator where there is unmeasured mediator-outcome confounding. Specifically, A mediator affected by an exposure that is only associated with the outcome by a hidden confounder may attenuate the estimate of an actual causal effect for the exposure on the outcome. I present a DAG for such confounding in Figure \ref{fig:Figure1}.c. The solution to the problem is straightforward. Do not condition on a mediator unless your causal question pertains to mediation. And if your causal question pertains to mediation, it is essential to assess the robustness of the indirect effect to mediator outcome confounding. For advice on mediation see: \citet{vanderweele_explanation_2015}.

\singlespace
\begin{lstlisting}[language=R]
set.seed(123)

```{r}
sim_fun_A = function() {
  n <- 1000
  U <- rnorm(n, 1) # unmeasured distress 
  A <- rnorm(n, 1) # religious service attendance,
  M <- rnorm(n , A + .2 * U)# religious identification
  Y <- rnorm(n , - A *.2 + U *.4) #  distress reduced by attendance.

# Religious id, caused by attendance + distress

# simulate data
simdat_A <- data.frame(
  M = M, # rel identification
  A = A, # rel service
  U = U, # distressing events (unmeasured)
  Y = Y )# outcome

sim_A <- lm(Y ~ A + M, data = simdat_A)
sim_A
}

# Replicate including the mediator
r_lm_A <- NA
r_lm_A = replicate(100, sim_fun_A(), simplify = FALSE )

# mediator attenuates true effect ~ .1 instead of ~.1
parameters::pool_parameters(r_lm_A)

## Repeat but with no mediator in model
set.seed(123)
sim_fun_B = function() {
  n <- 1000
  U <- rnorm(n, 1) # unmeasured distress 
  A <- rnorm(n, 1) # religious service attendance,
  M <- rnorm(n , A + .2 * U)# religious identification
  Y <- rnorm(n , - A *.2 + .4 * U) #  distress reduced by attendance.

# Simulate data
simdat_B <- data.frame(
  M = M, # rel identification
  A = A, # rel service attendance
  U = U, # distressing events (unmeasured)
  Y = Y ) # outcome

# Only model exposure
sim_B <- lm(Y ~ A, data = simdat_B)
sim_B
}

# Replication 100xs
r_lm_B <- NA
r_lm_B = replicate(100, sim_fun_B(), simplify = FALSE )

# diminished recover true effect ~ .2
parameters::pool_parameters(r_lm_B)

\end{lstlisting}

\end{document}

### 2022-03-12 analysis-religious-final.Rmd
updated for publication

@@
 ```{r global_options, echo=FALSE, include = FALSE}
+# setup
 knitr::opts_chunk$set(#
@@
 # source("libs.R") # libraries 
-# source("funs.R") # custom functions
-library("tidyverse")
+# source("funs.R") # custom functions, not used
 library("easystats")
@@
 library("splines")
+library("magick")
+library("tidyverse")
+library("Amelia")
+library("mice")
+
+# Bayes options
 rstan_options(auto_write=TRUE)
@@
 
-citation("daggity")
-toBibtex(citation("ggdag"))
-print(citation("EValue") , bibtex=TRUE)
-```
-
-## Workflow and analysis
+# citations
+# citation("daggity")
+# toBibtex(citation("ggdag"))
+# print(citation("EValue") , bibtex=TRUE)
+```
 
 First, we import the data (see functions). [@tchetgen2012]
+
+## Data import and set up
 
@@
 ```{r cache = TRUE}
-# clean data:use informative names !
-# note from Suzanne: 
-# I do think it's actually a negative relation, as the anxiety item was reverse-coded (following the documentation from the WHO quality of life scale, from which the items were taken).
-This is indicated in the data documentation, but I can totally understand this got buried under all information that was provided.
-We tried to make the dataset as clean as possible since most teams just aggregated all well-being items (and the religiosity items).
-But in retrospect, this recoding did cause some confusion for a couple of teams that used only specific items (and prefered the ordinal scale).
-This is definitely something to take into account for future projects like this one.
-
+# clean data:use informative names
 df <- d %>%
@@
   )
-max(df$rel_EXTENTBELIEVEGOD)
-max(df$rel_CHURCHFREQ)
-length(unique(df$country))
-table(df$country)
-```
-
-```{r}
-# get data into shape
-
+```
+
+```{r}
+# get data into shap
 d1 <- df %>%
   dplyr::filter(
-   #  !is.na(wb_psychANXIETY),
-   #  !is.na(age),
-   #  !is.na(rel_CHURCHFREQ),
-   # # !is.na(wb_psychMEANING),
-   #  !is.na(male),
-   #  !is.na(rel_EXTENTBELIEVEGOD),
   ) %>%
@@
     ch = rel_CHURCHFREQ,
-  #  mn  = wb_psychMEANING,
     edu = education,
     bg =  rel_EXTENTBELIEVEGOD#,
-  #  ch_s = scale(rel_CHURCHFREQ),
-  #  mn_s = scale(wb_psychMEANING),
-   # bg_s = scale(rel_EXTENTBELIEVEGOD)
   ) %>%
@@
                 ses,
-                #    mn,
                 edu,
                 bg,
-                # anx_s,
-                # age_s,
-                #  ch_s,
-                #  mn_s,
                 edu,
@@
                 religious,
-                #   bg_s,
                 denomination,
@@
 lt <-table1::table1(~  denomination, data = d1, transpose=F)
-lt
 lt <-table1::table1(~ ch + religious + age + male + edu + ses + denomination |country, data = d1, transpose=F)
 lt
+# Table for LaTeX
 kable(as.data.frame(lt), booktabs=TRUE, "latex")
-
-nrow(dr)
 ```
@@
 print( adjustmentSets( dg1a,  "belief","anx" ) )
-
-# 
-# d0a <-ggdag::ggdag(dg1a, text = FALSE, use_labels = "label")  + theme_dag() + ggtitle("What if we were to assume distressing events increase belief in God?")
-# d0a
-# # graph adjustment sets
-# 
-# d1a<-ggdag::ggdag_adjustment_set(dg1a, node_size = 14, text = FALSE, use_labels = "label") +
-#   theme(legend.position = "bottom") +
-#   theme_dag() + ggtitle("If we assume distressing events affect belief and service we cannot estimate causal effects ")
-# 
-# d1a 
-# 
-# 
-# # 
-# dp0 <-ggdag::ggdag(dg1, text = FALSE, use_labels = "label")  + theme_dag() + ggtitle("Assumed unmeasured causal effects of distressing experiences")
-# 
-# dp0
-# # # graph adjustment sets
-# # 
-# #dp1<-ggdag::ggdag_adjustment_set(dg1, node_size = 14, text = FALSE, use_labels = "label") +
-# #   theme(legend.position = "bottom") +
-# #   theme_dag() + ggtitle("Adjustment sets for Belief in God")
-# #dp1
-# dp0
 ```
@@
 
-# 
-# ggdag::ggdag(dg2) + theme_dag()
-# 
-# nocause <- ggdag::ggdag(dg2, text = FALSE, use_labels = "label")  + theme_dag() + labs(title = "Assume unmeasured distressing experience does not affect Belief or Service")
-# nocause
-# 
-# dg0<-ggdag::ggdag_adjustment_set(dg2, node_size = 14, text = FALSE, use_labels = "label") +
-#   theme(legend.position = "bottom") +
-#   theme_dag() + ggtitle("Adjustment set for religious service
-# assuming service does not change from distress")
-# 
-# dp0
-#
+## Dag
 
@@
 
-ggdag::ggdag(dpz) + theme_dag()
-# 
-# 
 # # graph adjustment sets
@@
 
-<!-- ```{r dg1, eval=FALSE} -->
-<!-- # save graph -->
-<!-- ggsave( -->
-<!--   dg1, -->
-<!--   path = here::here("Figs"), -->
-<!--   width = 10, -->
-<!--   height = 5, -->
-<!--   units = "in", -->
-<!--   filename = "dag_belief.jpg", -->
-<!--   device = 'jpeg', -->
-<!--   limitsize = FALSE, -->
-<!--   dpi = 1200 -->
-<!-- ) -->
-
-<!-- ``` -->
-
-<!-- ```{r dp0, eval=FALSE} -->
-<!-- # save graph -->
-<!-- ggsave( -->
-<!--   dg1, -->
-<!--   path = here::here("Figs"), -->
-<!--   width = 10, -->
-<!--   height = 5, -->
-<!--   units = "in", -->
-<!--   filename = "dag_church.jpg", -->
-<!--   device = 'jpeg', -->
-<!--   limitsize = FALSE, -->
-<!--   dpi = 1200 -->
-<!-- ) -->
-
-<!-- ``` -->
-
 ## Multiple imputation
 
-### Visualise missing data
-
-Next we need to examine patterns of missing-ness, which might spoil inference.
+Missing data can introduce selection bias.
+
+We examine patterns of missingness and find substantial missingness in denominations
 
@@
 library("naniar")
-library(magick)
 vis_miss(df) # substantial pattern of missingness in a few variables,
-df
-df$wb_soc_3
-dff<-df%>%
-  select(-wb_soc_3)
-
-miss_graph<- gg_miss_upset(dff)
-miss_graphR <-image_ggplot(miss_graph)
-# save graph
-miss_graph
-
-ggsave(miss_graph,
-  path = here::here("figs"),
-  width = 16,
-  height = 9,
-  units = "in",
-  filename = "miss_graph.jpg",
-  device = 'jpeg',
-  limitsize = FALSE,
-  dpi = 1200
-)
-
-library(Amelia)
-
-  # x is the longitudional dataframe (here 'd_3')
+
+# save missing file -- this does not work
+# ggsave(miss_graph,
+#   path = here::here("figs"),
+#   width = 16,
+#   height = 9,
+#   units = "in",
+#   filename = "miss_graph.jpg",
+#   device = 'jpeg',
+#   limitsize = FALSE,
+#   dpi = 1200
+# )
+```
+
+## Impute missing data
+
+```{r cache = TRUE}
 set.seed(1234)
@@
 
-table1::table1(~ ch |religious, data = d1)
-
-```
-
-We find there is substantial missingness in denomination. There is also missingness in ethnicity. Overall missingness is low in other indicators. However, 
-Because missingness can introduce bias, we multiply impute responses.
-
-## Model for belief
-
-### Marginal structural model Church
+# check
+#table1::table1(~ ch |religious, data = d1)
+```
+
+## Impute missingness
 
@@
 #d3 <-readRDS(here::here("mods","d3"))
+
 d4<-d3%>%
@@
          country=as.factor(country))
-
+```
+
+Imputation of the religion only dataset
+```{r}
 ## MICE ONLY RELIGIOUS
-table(d1$religious)
-
-# Check
 d1r<- d1 %>%
@@
 
+# save
+# saveRDS(d1p, here::here(
+#   "mods","d1p"
+# ))
+
 d2r<-mice(d1r)
 
-#check religious rumbers (1= religious, 2= not religious, 3=atheist) [R] So note that 3 = religious!!
-
 d3r<-complete(d2r)
 
+# Save data 
 #saveRDS(d3,here::here("mods","d3r"))
@@
 
+# read data
 d3 <-readRDS(here::here("mods","d3"))
@@
 
+# Select variables 
 d4<-d3%>%
@@
          country=as.factor(country))
-
+```
+
+Create inverse probability weights
+
+```{r}
 library(ipw)
@@
 
-max(d1p$ipwCH)
+# check max value
+#max(d1p$ipwCH)
 
@@
 # fine, well below 10
-max(dr$ipwCH)
+#max(dr$ipwCH)
 ```
@@
 lt <-table1::table1(~ religious|ch_ord, data = dp)
-
+lt
 kable(as.data.frame(lt), booktabs=TRUE, "latex")
-
-nrow(dr)
-```
-
-### Simulation
-
-```{r}
-#\begin{lstlisting}[language=R]
-
-set.seed(123)
-sim_fun_A = function() {
-  n <- 1000
-  U <- rnorm(n, 1) # Unmeasured distress 
-  A <- rnorm(n, 1) # religious service attendance,
-  M <- rnorm(n , A + U)# rel id
-  Y <- rnorm(n , - A*.2 + U + M*.1) #  distress reduced by attendance.
-
-# Religious id, caused by attendance + distress
-
-# simulate data
-simdat_A <- data.frame(
-  M = L, # rel id
-  A = A, # rel service
-  U = U, # distressing events (unmeasured)
-  Y = Y ) # outcome
-
-sim_A <- lm(Y ~ A + M, data = simdat_A)
-sim_A
-}
-
-# Replicate including the mediator
-r_lm_A <- NA
-r_lm_A = replicate(100, sim_fun_A(), simplify = FALSE )
-
-# average over outcomes: effect of religion is attenuated
-parameters::pool_parameters(r_lm_A)
-
-## Repeat but with no mediator in model
-set.seed(123)
-sim_fun_B = function() {
-  n <- 1000
-  U <- rnorm(n, 1) # Unmeasured distress 
-  A <- rnorm(n, 1) # religious service attendance,
-  M <- rnorm(n , A + U)# rel id + distress
-  Y <- rnorm(n , - A *.2 + U + M*.1) #  distress reduced by attendance.
-
-# simulate data
-simdat_B <- data.frame(
-  M = L, # rel id
-  A = A, # rel service
-  U = U, # distressing events (unmeasured)
-  Y = Y ) # outcome
-
-# Only exposure
-sim_B <- lm(Y ~ A, data = simdat_A)
-sim_B
-}
-
-# Replication 1000xs
-r_lm_B <- NA
-r_lm_B = replicate(100, sim_fun_B(), simplify = FALSE )
-
-# average over outcomes: effect of religion is attenuated
-parameters::pool_parameters(r_lm_B)
-
-#\end{lstlisting}
-```
+```
+
+Relable variables
 
@@
 
+# save
 #saveRDS(dr, here::here("mods","dr"))
-
-# saveRDS(d1p, here::here(
-#   "mods","d1p"
-# ))
-
+```
+
+Bayesian model with strong priors
+
+```{r}
 b_orR <- brm(
@@
 church_plot_R
-
+```
+
+Graphs
+
+```{r}
 ### EEMEANS
@@
 
-## include religion
+##  Gee models of various sorts
 
@@
 library(splines)
-
 ipb_ch_ln_no_mediation <- geeglm(
@@
 
-## Rel only
-
+# splines
 library(splines)
@@
 
-# model_parameters(ipb_bg)
-
 abs(min(d1p$ch_s) - max(d1p$ch_s))
@@
 evalues.OLS(0.04, se = 0.01, sd = 1, delta = 1.4)
-
 bias_plot(1.3009531, xmax = 6)
 
-### Religous Only
+### Religious Only
 
@@
 
-Adjusted GEE no weights
-
-```{r}
-# 
-# ipb_bg_noweights <- geeglm(
-#   anx_s ~ bg_s + ch_s + age_s + edu_s + male_2 + ses_s,
-#   data = d1p,  
-#   id = country,
-#   corstr = "exchangeable"
-# )
-# saveRDS(ipb_bg_noweights, here::here("mods","ipb_bg_noweights"))
-#
-
-ipb_ch_noweights <- geeglm(
-  anx_s ~ bs(ch_s) + age_s + edu_s + male_2 + ses_s,
-  data = d1p,  
-  id = country,
-  corstr = "exchangeable"
-)
-saveRDS(ipb_ch_noweights, here::here("mods","ipb_ch_noweights"))
-
-msm_lmer <- lmer(
-  anx_s ~ bs(ch_s) + age_s + edu_s + male_2 + ses_s +  (1+ ch_s|country),
-  data = d1p, 
-)
-saveRDS(msm_lmer, here::here("mods","msm_lmer"))
-
-msm_lmer_wg <- lmer(
-  anx_s ~ bs(ch_s) + (1 + ch_s|country),
-  weights = ipwCH,
-  data = d1p
-)
-saveRDS(msm_lmer_wg, here::here("mods","msm_lmer_wg"))
-
-# model_parameters(ipb_bg_noweights)
-# model_parameters(ipb_bg_noweights)
-# model_parameters(ipb_bg)
-# model_parameters(ipb_bg)
-
-model_parameters(ipb_ch_noweights)
-model_parameters(ipb_ch)
-model_parameters(ipb_ch)
-model_parameters(msm_lmer)
-model_parameters(msm_lmer_wg)
-
-pred <- ggeffects::ggemmeans(msm_lmer_wg, terms = c("ch_s[all]"))
-pred
-
-plot(pred,
-    # add.data=T,dot.alpha=.1, 
-     facets = F)  + 
-  scale_y_continuous(limits=c(-.7,.7)) + 
-  theme_classic() + ggtitle("") # title to be suppled
-
-```
-
-We note that the multi-level models give different estimates.
-
-### Bayesian Church Attendaance
+### Bayesian Stratified Regression
+
+Not used
 
@@
 
-#saveRDS(modmc, here::here("mods", "modmc.rds"))
 summary(modmc)
-
-```
+```
+
 ## Bayesian wieghted regression
@@
 summary(modmc_wt_linear)
-
+```
+
+Priors used
+
+```{r}
 prior_summary(modmc_wt_linear)
-summary(modmc_wt_linear)
-
+```
+
+Use of strong prior and splines
+
+```{r}
 library(brms)
@@
 prior_summary(modmc_wt_priorStrong)
-
+```
+
+Test strong priors on non-linear model
+
+```{r}
 ##### TeST PRIORS
@@
 )
-
-## WEAK PRIOR
+```
+
+Test a weaker prior.
+
+```{r}
+## Weak prior
 modmc_wt_prior_weak <- brm(
@@
 
+## Graph
+
 ```{r, cache = TRUE}
@@
 abs(maxchur-minchur)
-
-library(emmeans)
-grand_mean
 
@@
 church_plot
-
-ipw_gee <- plot(pred,
-    #add.data=T,dot.alpha=.1, 
-     facets = F)  + 
-  scale_y_continuous(limits=c(-.1,.1)) + xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + 
-  theme_classic() +labs(title = "Generalized Estimating Equations", 
-  subtitle = "Inverse Probability Weighting") + scale_y_continuous(limits = c(-.5,.5))
-ipw_gee
-church_plot
-model_plot <- ipw_gee + church_plot + church_plot_test_weak + church_plot_test_strong
-model_plot <- model_plot + plot_annotation(tag_levels = "a")
-
-model_plot
-
-ggsave(
-  model_plot,
-  path = here::here("figs"),
-  width = 10,
-  height = 5,
-  units = "in",
-  filename = "model_plot.jpg",
-  device = 'jpeg',
-  limitsize = FALSE,
-  dpi = 1200
-)
-
-library(brmsmargins)
-modmc_wt_priorStrong
-
-max(d1p$ch_s)#-0.8321022,2.23193
-
-h <-.001
-ames <- brmsmargins(
-  object = modmc_wt_priorStrong,
-            at = data.frame(ch_s= c(-0.8321022,2.23193)),
-            contrasts = cbind("AME x" = c(0.8321022/ h, 2.23193/ h)),
-            effects = "integrateoutRE")
-                              )
-ames$Summary
-
-ames_st <- brmsmargins(
-  object = modmc_wt_priorStrong, 
-            at = data.frame(ch_s = c(-0.8321022,2.23193)),
-            contrasts = cbind("AME x" = c(-0.8321022/ h, 2.23193/ h)),
-            effects = "integrateoutRE", CI = 0.95)
-ames_st$ContrastSummary
-knitr::kable(ames$Summary, digits = 3, "latex")
-
-ames_ln <- brmsmargins(
-  object = modmc_wt_linear, 
-            at = data.frame(ch_s = c(-0.8321022,2.23193)),
-            contrasts = cbind("AME x" = c(-0.8321022/ h, 2.23193/ h)),
-            effects = "integrateoutRE", CI=.95)
-ames_ln$Summary
-knitr::kable(ames$Summary, digits = 3, "latex")
-
+```
+
+### More graphs
+
+```{r, cache = TRUE}
 ### EEMEANS
@@
   scale_fill_brewer(palette = "Blues") +
- # scale_fill_viridis_d(palette = "Blues") +
   labs(x = "Frequency of religious service attendance (SD)", 
@@
 plot_result
+
+## Save
 
@@
 summary(modmc_wt_linear)
-
 #
-
+```
+
+Test evalues
+
+```{r, cache = TRUE}
 library(EValue)
@@
 # sensitivity = Church
-#yola
-
+
+# Obtain standard error 
 #SE = (upper limit – lower limit) / 3.92.
@@
 ta <- effect_draws_cn %>% mean_hdi()
-ta
+
 (ta$.upper - ta$.lower)/3.92 #
-
+```
+
+Evalue
+
+```{r, cache = TRUE}
 evalues.OLS(-0.147 , se = 0.05607611, sd = 1, delta = 1.1)
@@
 # E-values 1.5870732        NA 1.2368137
-
-# multibias
-multi_evalue(est = RR( 0.8631671), biases = selection())
-
+```
+
+Bias plot
+
+```{r, cache = TRUE}
 #
@@
 ```
-
-## Scaled 0-1
 
@@
 dp <- d1p %>% 
-  mutate(ipwCH = dchr_sc$weights.trunc)
+  mutate(ipwCH = dchr_sc$weights.trunc
 
@@
 
-# Montonic effects
+# Montonic effects model
+
+This is the model reported in the study
 
@@
 # bayesian monotonic effects
-
 dp$ch_ord <- factor(dp$ch, ordered=TRUE)
@@
 summary(b_or)
-
+```
+
+## Graph ( not used)
+
+```{r}
 church_plot <-plot(ggeffects::ggemmeans(b_or, terms = "ch_ord")) +   xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + labs(title = "Baysian Multi-level Model",
@@
 
-```{r}
-# tab_or <- model_parameters(b_or,   verbose = F, ci_method = "hdi",test = c("pd"))
-# kbl(tab_or, "latex",booktabs=T)
-
-# sjstats::tidy_stan(tab_or)
-
-# library(broom.mixed)
-# tidy(
-#   b_or,
-#  # parameters = NA,
-#   effects = c("fixed", "ran_pars"),
-#   robust = FALSE,
-#   conf.int = TRUE,
-#   conf.level = 0.95,
-#   conf.method = c("HPDinterval"),
-#   fix.intercept = TRUE
-# )
-
+Table usd in the publication
+
+```{r}
 parameters::parameters(
@@
   group_level = FALSE,
-  diagnostic = "Rhat") %>%
+  diagnostic = "Rhat"
+) %>%
   mutate_if(is.numeric, ~ round(., 3)) %>%
@@
 
-```{r}
-# pord<-brms::conditional_effects(b_or, "ch_ord", 
-#                               spaghetti = T,
-#                               ndraws = 1000,
-#                               categorical = F,
-#                               prob = .9,
-#                               method = 'fitted')
-# 
-# 
-# church_plot <- plot(pord, plot = FALSE)[[1]]   +  xlab("Service attendance (standardised)") + ylab ("Anxiety (standardised)") + labs(title = "Baysian Multi-level Model",
-# subtitle = "Inverse Probability Weighting") + scale_y_continuous(limits = c(-.5,.5))+ 
-#   theme_classic()
-# church_plot
+## Graphs used in the study
+
+```{r}
 library(tidybayes)
@@
 b_or %>%
-  emmeans(~ ch_ord,
-          at = list(ch_ord = c("never",">1xpw"),
+  emmeans( ~ ch_ord,
+           at = list(ch_ord = c("never", ">1xpw"),
                      country = NA),
@@
 effect_draws <-  b_or %>%
-  emmeans(~ ch_ord,
-         at = list(ch_ord = c("never",">1xpw"),
+  emmeans( ~ ch_ord,
+           at = list(ch_ord = c("never", ">1xpw"),
                      country = NA),
@@
 effect_draws_or <-  b_or %>%
-  emmeans(~ ch_ord,
-         at = list(ch_ord = c("never",">1xpw"),
+  emmeans( ~ ch_ord,
+           at = list(ch_ord = c("never", ">1xpw"),
                      country = NA),
@@
 effect_draws_or
+
 # marginal mean 
@@
 # se(effect_draws_ln$.value)
+```
+
+## Graph used in the study
+
+```{r}
 library(tidybayes)
@@
 
-# plot_grand_mean <- ggplot(grand_mean, 
-#                                  aes( x = .epred, fill=(ch_ord))) +
-#   scale_fill_brewer(palette = "Blues") +
-#     stat_halfeye(slab_alpha = 0.8) +
-# 
-#   labs(x = "Church attendence frequency", 
-#        y = "Marginal Density",
-#        fill = "") +
-#   theme_classic() +
-#   theme(legend.position = "top")
-# 
-# plot_grand_mean
-
 grand_mean_autonomy_dist <- b_or %>% 
@@
   stat_halfeye(slab_alpha = 0.7) +
-    scale_x_continuous(limits=c(-.6,.4)) +
-  labs(x = "Predicted anxiety (standardised)", y = "Density",
+  scale_x_continuous(limits = c(-.6, .4)) +
+  labs(
+    x = "Predicted anxiety (standardised)",
+    y = "Density",
     fill = "Church frequency",
-       subtitle = "Posterior predictions: monotonic model") +
+    subtitle = "Posterior predictions: monotonic model"
+  ) +
   theme_classic() +
-     theme(legend.position = "bottom",
+  theme(
+    legend.position = "bottom",
     legend.title = element_text(size = 10),
-          legend.text = element_text(size = 8))
+    legend.text = element_text(size = 8)
+  )
 
@@
   stat_halfeye(slab_alpha = 0.7) +
-    scale_x_continuous(limits=c(-.6,.4)) +
-  labs(x = "Predicted anxiety (standardised)", y = "Density",
+  scale_x_continuous(limits = c(-.6, .4)) +
+  labs(
+    x = "Predicted anxiety (standardised)",
+    y = "Density",
     fill = "Church frequency",
-       subtitle = "Posterior predictions: monotonic model") +
+    subtitle = "Posterior predictions: monotonic model"
+  ) +
   theme_classic() +
-     theme(legend.position = "bottom",
+  theme(
+    legend.position = "bottom",
     legend.title = element_text(size = 12),
-          legend.text = element_text(size = 8))
+    legend.text = element_text(size = 8)
+  )
 
@@
 
+# Save plot
 ggsave(
@@
 )
-
+```
+
+More evalues
+
+```{r}
 # Obtain evalues
@@
 
-plot_ord_wk_priors<-brms::conditional_effects(b_ord_test_weak_priors, "ch_ord", 
+plot_ord_wk_priors <-
+  brms::conditional_effects(
+    b_ord_test_weak_priors,
+    "ch_ord",
     spaghetti = T,
@@
     prob = .95,
-                              method = 'fitted')
-
-test_or_wk<- plot(plot_ord_wk_priors, plot = FALSE)[[1]]   +  xlab("Service attendance (monotonic)") + ylab ("Anxiety (standardised)") + labs(title = "Check Priors: Baysian Mixed Effects Marginal Model", subtitle = "Sampling from weak priors only (ordinal) ") + scale_y_continuous(limits = c(-20,15))+
+    method = 'fitted'
+  )
+
+test_or_wk <-
+  plot(plot_ord_wk_priors, plot = FALSE)[[1]]   +  xlab("Service attendance (monotonic)") + ylab ("Anxiety (standardised)") + labs(title = "Check Priors: Baysian Mixed Effects Marginal Model", subtitle = "Sampling from weak priors only (ordinal) ") + scale_y_continuous(limits = c(-20, 15)) +
   theme_classic()
@@
 
-con_st_priors<-brms::conditional_effects(b_cont_test_st_priors, "ch_s", 
+con_st_priors <-
+  brms::conditional_effects(
+    b_cont_test_st_priors,
+    "ch_s",
     spaghetti = T,
@@
     prob = .95,
-                              method = 'fitted')
-
-test_con_st<- plot(con_st_priors, plot = FALSE)[[1]]   +  xlab("Service attendance (sd)") + ylab ("Anxiety (standardised)") + labs(title = "Check Priors: Baysian Mixed Effects Marginal Model", subtitle = "Sampling from weak priors only (continuous) ") +   scale_y_continuous(limits = c(-10,10))+
+    method = 'fitted'
+  )
+
+test_con_st <-
+  plot(con_st_priors, plot = FALSE)[[1]]   +  xlab("Service attendance (sd)") + ylab ("Anxiety (standardised)") + labs(title = "Check Priors: Baysian Mixed Effects Marginal Model", subtitle = "Sampling from weak priors only (continuous) ") +   scale_y_continuous(limits = c(-10, 10)) +
   theme_classic()
@@
 b_cont_test_wk_priors <- brm(
-  anx_s|weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
+  anx_s | weights(ipwCH) ~ bs(ch_s) + (1 + ch_s | country),
   data = d1p,
-   prior = c(
-    prior(normal(0, 2.5), class = Intercept),
+  prior = c(prior(normal(0, 2.5), class = Intercept),
             prior(normal(0, 1), class = b)),
@@
 
-con_wk_priors<-brms::conditional_effects(b_cont_test_wk_priors, "ch_s", 
+con_wk_priors <-
+  brms::conditional_effects(
+    b_cont_test_wk_priors,
+    "ch_s",
     spaghetti = T,
@@
     prob = .95,
-                              method = 'fitted')
-
-test_con_wk_priors<- plot(con_wk_priors, plot = FALSE)[[1]]   +  xlab("Service attendance (sd)") + ylab ("Anxiety (standardised)") + labs(title = "Check Priors: Baysian Mixed Effects Marginal Model", subtitle = "Sampling from weak priors only (continuous) ")   + scale_y_continuous(limits = c(-10,10))+
+    method = 'fitted'
+  )
+
+test_con_wk_priors <-
+  plot(con_wk_priors, plot = FALSE)[[1]]   +  xlab("Service attendance (sd)") + ylab ("Anxiety (standardised)") + labs(title = "Check Priors: Baysian Mixed Effects Marginal Model", subtitle = "Sampling from weak priors only (continuous) ")   + scale_y_continuous(limits = c(-10, 10)) +
   theme_classic()
@@
 b_or_test_st_priors4 <- brm(
-  anx_s|weights(ipwCH) ~ mo(ch_ord) + (1 + mo(ch_ord) | country),
+  anx_s | weights(ipwCH) ~ mo(ch_ord) + (1 + mo(ch_ord) | country),
   data = dp,
@@
     prior(student_t(3, 0, .5), class = Intercept),
-    prior(normal(-.05,.1), class = "b"),
-    prior(dirichlet(1,1,1,1,1,1), class = "simo", coef = moch_ord1)),
+    prior(normal(-.05, .1), class = "b"),
+    prior(dirichlet(1, 1, 1, 1, 1, 1), class = "simo", coef = moch_ord1)
+  ),
   warmup = 1000,
@@
 
-b_or_test_st_priors4<-brms::conditional_effects(b_or_test_st_priors4, "ch_ord", 
+b_or_test_st_priors4 <-
+  brms::conditional_effects(
+    b_or_test_st_priors4,
+    "ch_ord",
     spaghetti = T,
@@
     prob = .95,
-                              method = 'fitted')
-
-or_strong<- plot(b_or_test_st_priors4, plot = FALSE)[[1]]   +  xlab("Service attendance (monotonic)") + ylab ("Anxiety (standardised)") + labs(title = "Check Priors: Baysian Mixed Effects Marginal Model", subtitle = "Sampling from strong priors only (ordinal) ") + scale_y_continuous(limits = c(-20,15))+
+    method = 'fitted'
+  )
+
+or_strong <-
+  plot(b_or_test_st_priors4, plot = FALSE)[[1]]   +  xlab("Service attendance (monotonic)") + ylab ("Anxiety (standardised)") + labs(title = "Check Priors: Baysian Mixed Effects Marginal Model", subtitle = "Sampling from strong priors only (ordinal) ") + scale_y_continuous(limits = c(-20, 15)) +
   theme_classic()
@@
 
+# Graph
 library(patchwork)
@@
 ```
+
+## More priors tested
 
@@
 
-### Summary of selection bias
-
-```{r}
-set.seed(1234)
-n <- 10000 # n
-a <- rnorm(n, 0, 1) # rel service attendance
-y <- rnorm(n, -a*.2, 1) # rel service affects anxiety 
-bg <- rnorm(n, a, 1) # rel service affects belief 
-s <- rnorm(n, .2*y + .2*bg,1) # selection bias
-
-dbind <-data.frame(y,a,s, bg)
-
-summary(lm(s ~  a + bg, data = dbind))
-summary(lm(s ~  a , data = dbind))
-
-##
-
-set.seed(123)
-sim_fun = function() {
-n <- 10000 # n
-a <- rnorm(n, 0, 1) # rel service attendance
-y <- rnorm(n, -a*.2, 1) # rel service affects anxiety 
-l <- rnorm(n, a, 1) # rel service affects belief 
-s <- rnorm(n, .2*l - 5*y,1) # selection bias
-
-# simulate data
-simdat <-data.frame(y, a, s, l)
-
-  sim_s <- lm(s ~ a + l, data = simdat) # belief + church
-  sim_s  # confounded 
-  summary(sim_s)
-}
-
-# Replication 1000xs
-r_lm = replicate(1000, sim_fun(), simplify = FALSE )
-
-# extract simulations
-tab_sim<-purrr::map_dfr(r_lm, broom::tidy)
-
-# extract belief coefficients
-mout<-tab_sim %>%
-  dplyr::filter(term =="l")%>%
-  dplyr::mutate_if(is.numeric, round, 5)
-
-# calculate proportion of estimates greater than 0
-# all positive
-sum(mout$estimate > 0) / length(mout$estimate)
-
-# proportion of belief coefficient p.values < .001
-sum(mout$p.value <.001) / length(mout$p.value)
-
-# extract religious service coefficients
-tout<-tab_sim %>%
-  dplyr::filter(term =="a")%>%
-  dplyr::mutate_if(is.numeric, round, 5)
-
-# calculate proportion of estimates less than 0
-# all negative
-sum(tout$estimate < 0) / length(tout$estimate)
-
-# proportion of service coefficient p.values < .001
-sum(tout$p.value <.001) / length(tout$p.value)
-```
-
-```{r}
-set.seed(123)
-sim_fun = function() {
-n <- 10000 # n
-a <- rnorm(n, 0, 1) # rel service attendance
-y <- rnorm(n, -a*.2, 1) # rel service affects anxiety 
-bg <- rnorm(n, a, 1) # rel service affects belief 
-s <- rnorm(n, .2*y + .2*bg,1) # selection bias
-
-# simulate data
-simdat <-data.frame(y, a, s, bg)
-
-  sim_s <- lm(s ~  a, data = dbind) # belief + church
-  sim_s  # confounded 
-}
-
-```
+### Images for DAGS in the study
+
+Made in TikZ, rendered in Magick and ggplot2
 
@@
 
-Coefficient plots
-
-Believe in God predicts more anxiety
-
-```{r}
-parF1 <- parameters::model_parameters(m1 ) 
-parF1
-plot(parF1,  sort = TRUE) +theme_lucid() #option -->
-```
-
-Variability in country intercepts, slope interactions not reliable.
-
-India, however, is clearly an outlier.
-
-```{r}
-parR1<-parameters::model_parameters(m1,effects = "random" ) 
-belief_random <- plot(parR1,  sort = TRUE) +theme_lucid()  + labs(title = "Country-level variability in anxiety intercepts are evident; country-level anxiety-on-belief co-variances
-are not reliable: India is an outlier.") 
-belief_random
-```
-
-No country level differences trending at the high end of the anxiety scale for slope/intercept covariance, but not at the low end of anxiety this pattern is reversed, leading to a canceling out ...
-such that overall the correlation isn't reliable [cor(Intercept,ch_s) 0.24 [-0.38, 0.77].
-But worth investigating...and perhaps a more complex mixture model.
-
-```{r}
-plot(ggeffects::ggpredict(
-  m1,
-  terms = c("country[all]", "bg_s[minmax]"),
-  type = "re",
-  ci.lvl = 0.89
-)) + scale_color_viridis_d() +  scale_y_continuous(limits = c(0,1)) + 
-  theme_bw()
-```
-
-Clear relationship between lower distress and greater religious service attendance
-
-```{r, cache = TRUE}
-library(viridis)
-p0<-brms::conditional_effects(m1, "bg_s", 
-                              spaghetti = T,
-                              ndraws = 500,
-                              categorical = F,
-                              prob = .89,
-                              method = 'fitted')
-
-belief_fixed <- plot(p0, plot = FALSE)[[1]]   +  xlab("belief (standardised)") + ylab ("anxiety (ordinal 1-5)") + labs(title = "Increasing belief is associated with increasing anxiety")
-belief_fixed
-```
-
-### Checks:
-
-```{r}
-# YES
-pp_check(m1)
-```
-
-Trace plots:
-
-```{r, cache = TRUE}
-mcmc_plot(m1,
-          type = "trace")
-```
-
-## Results summary: Anxiety \~ church
-
-Does church attendance predict anxiety?
-
-### Model
-
-```{r echo}
-glimpse(d1)
-m2 <- brm(
-    anx ~ ch_s +
-    age_s  +
-    male +
-    (1 + ch_s | country),
-  data = d1,
-  family = cumulative(link ="logit"),
-  seed = 1234,
-  warmup = 1000,
-  iter = 3000,
-  chains = 4,
-  file = here::here("models", "m1o_real")
-)
-
-parameters::parameters(
-  m2,
-  test = "pd",
-  priors = FALSE, 
-  effects = "all",
-  drop = TRUE,
-  group_level = FALSE,
-  diagnostic = "Rhat") %>%
-    mutate_if(is.numeric, ~ round(., 2)) %>%
-    mutate_all(funs(str_replace(., "b_", ""))) %>%
-    kable(booktabs = T, "latex", caption =  "Parameter estimates for model of Religious Service on Anxiety") %>%
-    print()
-```
-
-### Results
-
-Coefficient plots
-
-More church attendance predicts less anxiety
-
-```{r}
-parc<-parameters::model_parameters(m2 ) 
-plot(parc,  sort = TRUE) + theme_lucid() #option -->
-```
-
-Variability in country intercepts, slope interactions not reliable.
-
-India, however, is clearly an outlier.
-
-```{r}
-parcR<-parameters::model_parameters(m2,effects = "random" ) 
-church_random <- plot(parcR,  sort = TRUE) +theme_lucid() + labs(title = "Country-level variability in anxiety intercepts are evident; country-level anxiety-on-service co-variances
-are not reliable: India is again an outlier.")
-
-church_random
-```
-
-No country level differences trending at the high end of the anxiety scale for slope/intercept covariance, but not at the low end of anxiety this pattern is reversed, leading to a canceling out ...
-such that overall the correlation isn't reliable [cor(Intercept,ch_s) 0.24 [-0.38, 0.77].
-But worth investigating...and perhaps a more complex mixture model.
-
-```{r}
-plot(ggeffects::ggpredict(
-  m2,
-  terms = c("country[all]", "ch_s[minmax]"),
-  type = "re",
-  ci.lvl = 0.89
-)) + scale_color_viridis_d() + 
-  theme_classic()
-```
-
-Clear relationship between lower distress and greater religious service attendance
-
-```{r, cache = TRUE}
-library(viridis)
-p0b<-brms::conditional_effects(m2, "ch_s", 
-                              spaghetti = T,
-                              ndraws = 500,
-                              categorical =T,
-                              prob = .89,
-                              method = 'fitted')
-
-church_fixed <-plot(p0b, plot = FALSE)[[1]]  + xlab("religious service attendance (standardised)") + ylab ("anxiety (ordinal 1-5)") +
-  labs(title = "Increasing religious service attendence is associated with lower anxiety")
-
-church_fixed
-```
-
-## Results summary Belief \~ Anxiety
-
-```{r}
-m3<- brm(
-  bg_s ~
-    mo(anx)+
-    ch_s +
-    age_s  +
-    male +
-    (1 + mo(anx)| country),
-  data = d1,
-  seed = 1234,
-  warmup = 1000,
-  iter = 3000,
-  chains = 4,
-  file = here::here("models", "m3o_real")
-)
-
-parameters::model_parameters(m3, test = "p_direction")
-brms::ranef(m3)
-
-parameters::parameters(
-  m3,
-  test = "pd",
-  priors = FALSE, 
-  effects = "all",
-  drop = TRUE,
-  group_level = FALSE,
-  diagnostic = "Rhat") %>%
-    mutate_if(is.numeric, ~ round(., 2)) %>%
-    mutate_all(funs(str_replace(., "b_", ""))) %>%
-    kable(booktabs = T, "latex", caption =  "Parameter estimates for model of Anxiety on Belief") %>%
-    print()
-```
-
-<!-- ```{r} -->
-
-<!-- this is no better -->
-
-<!-- m3a<- brm( -->
-
-<!--   bg_s ~ -->
-
-<!--     anx_s+ -->
-
-<!--     ch_s + -->
-
-<!--     age_s  + -->
-
-<!--     male + -->
-
-<!--     (1 + anx_s | country), -->
-
-<!--   data = d1, -->
-
-<!--   seed = 1234, -->
-
-<!--   warmup = 1000, -->
-
-<!--   iter = 3000, -->
-
-<!--   chains = 4, -->
-
-<!--   file = here::here("models", "m3a") -->
-
-<!-- ) -->
-
-<!-- parameters::model_parameters(m3a, test = "p_direction") -->
-
-<!-- brms::ranef(m3a) -->
-
-<!-- summary(m3a) -->
-
-<!-- hist(d1$bg_s) -->
-
-<!-- # YES -->
-
-<!-- pp_check(m3a) -->
-
-<!-- ``` -->
-
-Coefficient plots
-
-Anxiety predicts belief
-
-```{r}
-parF3 <- parameters::model_parameters(m3 ) 
-parF3
-plot(parF3,  sort = TRUE) +theme_lucid() #option -->
-```
-
-Variability in country intercepts, slope interactions not reliable.
-
-India, however, is clearly an outlier.
-
-```{r}
-parR3<-parameters::model_parameters(m3,effects = "random" ) 
-anx_belief_random <- plot(parR3,  sort = TRUE) +theme_lucid()  + labs(title = "Country-level variabilty in belief intercepts are evident; country-level belief-on-anxiety co-variances
-are not reliable.")
-
-anx_belief_random
-```
-
-No country level differences trending at the high end of the anxiety scale for slope/intercept covariance, but not at the low end of anxiety this pattern is reversed, leading to a canceling out ...
-such that overall the correlation isn't reliable [cor(Intercept,ch_s) 0.24 [-0.38, 0.77].
-But worth investigating...and perhaps a more complex mixture model.
-
-```{r}
-plot(ggeffects::ggpredict(
-  m3,
-  terms = c("country[all]", "anx"),
-  type = "re",
-  ci.lvl = 0.89
-)) + scale_color_viridis_d() +  scale_y_continuous(limits = c(0,1)) + 
-  theme_bw()
-```
-
-```{r, cache = TRUE}
-library(viridis)
-p3<-brms::conditional_effects(m3, "anx", 
-                              spaghetti = T,
-                              ndraws = 500,
-                              categorical = F,
-                              prob = .89,
-                              method = 'fitted')
-
-belief_anxiety_random_fixed <- plot(p3, plot = FALSE)[[1]]   + xlab("Anxiety (monotonic ordinal 1-5)") + ylab ("Belief in God( standardised))") +
-  labs(title = "Increasing anxiety is associated with increasing belief in God")
-
-belief_anxiety_random_fixed
-```
-
-### Checks:
-
-```{r}
-# YES
-pp_check(m3)
-```
-
-Trace plots:
-
-```{r, cache = TRUE}
-mcmc_plot(m3,
-          type = "trace")
-```
-
-## Results summary Church \~ Anxiety
-
-```{r}
-m4<- brm(
-  ch_s ~
-    mo(anx)+
-    age_s  +
-    male +
-    (1 + mo(anx)| country),
-  data = d1,
-  seed = 1234,
-  warmup = 1000,
-  iter = 3000,
-  chains = 4,
-  file = here::here("models", "m4o_real")
-)
-
-parameters::model_parameters(m4, test = "p_direction")
-brms::ranef(m4)
-
-parameters::parameters(
-  m4,
-  test = "pd",
-  priors = FALSE, 
-  effects = "all",
-  drop = TRUE,
-  group_level = FALSE,
-  diagnostic = "Rhat") %>%
-    mutate_if(is.numeric, ~ round(., 2)) %>%
-    mutate_all(funs(str_replace(., "b_", ""))) %>%
-    kable(booktabs = T, "latex", caption =  "Parameter estimates for model of Anxiety on Religious Service") %>%
-    print()
-```
-
-<!-- ```{r} -->
-
-<!-- this is no better -->
-
-<!-- m3a<- brm( -->
-
-<!--   bg_s ~ -->
-
-<!--     anx_s+ -->
-
-<!--     ch_s + -->
-
-<!--     age_s  + -->
-
-<!--     male + -->
-
-<!--     (1 + anx_s | country), -->
-
-<!--   data = d1, -->
-
-<!--   seed = 1234, -->
-
-<!--   warmup = 1000, -->
-
-<!--   iter = 3000, -->
-
-<!--   chains = 4, -->
-
-<!--   file = here::here("models", "m3a") -->
-
-<!-- ) -->
-
-<!-- parameters::model_parameters(m3a, test = "p_direction") -->
-
-<!-- brms::ranef(m3a) -->
-
-<!-- summary(m3a) -->
-
-<!-- hist(d1$bg_s) -->
-
-<!-- # YES -->
-
-<!-- pp_check(m3a) -->
-
-<!-- ``` -->
-
-Coefficient plots
-
-Anxiety predicts belief
-
-```{r}
-parF4 <- parameters::model_parameters(m4 ) 
-parF4
-plot(parF4,  sort = TRUE) +theme_lucid() #option -->
-```
-
-Variability in country intercepts, slope interactions not reliable.
-
-India, however, is clearly an outlier.
-
-```{r}
-par4<-parameters::model_parameters(m4,effects = "random" ) 
-anx_church_random <- plot(par4,  sort = TRUE) + theme_lucid()  + labs(title = "Country-level variabilty in religious service intercepts are evident; country-level service-on-anxiety co-variances
-are not reliable.")
-
-anx_church_random
-```
-
-```{r}
-plot(ggeffects::ggpredict(
-  m4,
-  terms = c("country[all]", "anx"),
-  type = "re",
-  ci.lvl = 0.89
-)) + scale_color_viridis_d() +  scale_y_continuous(limits = c(0,1)) + 
-  theme_bw()
-```
-
-```{r, cache = TRUE}
-library(viridis)
-p4<-brms::conditional_effects(m4, "anx", 
-                              spaghetti = T,
-                              ndraws = 500,
-                              categorical = F,
-                              prob = .89,
-                              method = 'fitted')
-
-church_anxiety_random_fixed <- plot(p4, plot = FALSE)[[1]]   + xlab("Anxiety (monotonic ordinal 1-5)") + ylab ("Religious Service(standardised))") +
-  labs(title = "Increasing anxiety is associated with decreasing church attendance.")
-
-church_anxiety_random_fixed
-```
-
-### Checks:
-
-```{r}
-# YES
-pp_check(m4)
-```
-
-Trace plots:
-
-```{r, cache = TRUE}
-mcmc_plot(m4,
-          type = "trace")
-```
-
-## Combined plot
-
-```{r}
-library(patchwork)
-
-(belief_fixed +  belief_random)/(church_fixed+ church_random  ) / 
-(belief_anxiety_random_fixed + anx_belief_random) / (church_anxiety_random_fixed + anx_church_random) +
-  plot_annotation(tag_levels = "a")# + plot_layout(guides = "collect")
-
-# coef_graph 15 x 25
-```
-
-### Checks:
-
-```{r}
-# YES
-pp_check(m2)
-```
-
-Trace plots:
-
-```{r, cache = TRUE}
-mcmc_plot(m2,
-          type = "trace")
-```
-
-### Overall summary
-
-Belief in God predicts anxiety. This is consistent with theories that anxiety leads to greater demand for belif in God. However, this question awaits future research.
-
-Church attendance positively predicts lower anxiety.
-
-### Group level-predictions
-
-There is substantial variation in country-level intercepts for anxiety. And substantial cultural level variation in religious practice and belief. All this needs to be explained. India is an outlier.
-
-## Simulation
-
-```{r}
-set.seed(123)
-n <- 1000
-# distressing events (unmeasured)
-U <- rnorm( n ) 
-# religious service attendance increases with distress
-C <- rnorm( n, U ) 
-# anxiety caused by distress experience and
-# buffered by religious service
-Y <- rnorm( n , U - C )
-
-# belief increases with service, anxiety,
-# and experience
-B <- rnorm( n , Y + C + U )
-
-# simulate data
-simdat <- data.frame(
-  Belief = B, # belief
-  ReligiousService = C, # religious service
-  U = U, # distressing events (unmeasured)
-  Anxiety = Y  # outcome
-  )
-
-# summarise results
-summary(
-  mod1 <- lm(Anxiety ~ Belief + ReligiousService, data = simdat)
-)
-
-# summarise results
-summary(
-  mod1b <- lm(Anxiety ~ ReligiousService, data = simdat)
-)
-
-summary(
-  mod2 <- lm(Belief  ~ Anxiety + ReligiousService, data = simdat)
-)
-
-summary(
-  mod3 <- lm(ReligiousService  ~ Anxiety, data = simdat)
-)
-
-belief_on_anxiety_sim <- plot( ggeffects::ggpredict(mod1, terms = c("Belief")), add.data = TRUE, dot.alpha =.5 ) + labs(title = "Simulation recovers positive relationship between belief on anxiety") + scale_y_continuous(limits = c(-7,7))+ scale_x_continuous(limits = c(-10,10))
-
-service_on_anxiety_sim <- plot( ggeffects::ggpredict(mod1b, terms = c("ReligiousService")), add.data = TRUE, dot.alpha =.5 ) + labs(title = "Simulation recovers negative relationship between religious service on anxiety")+ scale_y_continuous(limits = c(-7,7))+ scale_y_continuous(limits = c(-7,7))+ scale_x_continuous(limits = c(-10,10))
-service_on_anxiety_sim
-
-service_on_anxiety_sim <- plot( ggeffects::ggpredict(mod1, terms = c("ReligiousService")), add.data = TRUE, dot.alpha =.5 ) + labs(title = "Simulation recovers negative relationship between religious service on anxiety")+ scale_y_continuous(limits = c(-7,7))+ scale_x_continuous(limits = c(-10,10))
-service_on_anxiety_sim
-
-anxiety_on_belief_sim <- plot( ggeffects::ggpredict(mod2, terms = c("Anxiety")), add.data = TRUE, dot.alpha =.5 ) + labs(title = "Simulation recovers positive relationship between anxiety on belief")+ scale_y_continuous(limits = c(-7,7))+ scale_x_continuous(limits = c(-10,10))
-anxiety_on_belief_sim
-
-anxiety_on_church_sim <- plot( ggeffects::ggpredict(mod3, terms = c("Anxiety")), add.data = TRUE, dot.alpha =.5 ) + labs(title = "Simulation recovers negative relationship between anxiety on religious service")+ scale_y_continuous(limits = c(-7,7))
-anxiety_on_church_sim
-
-library(patchwork)
-
-(belief_on_anxiety_sim +  service_on_anxiety_sim)/ (anxiety_on_belief_sim + anxiety_on_church_sim)
-
-# sim_graph 15 x 15
-
-```
-
-We can convince ourselves of this result by replicating it many times
-
-```{r}
-sim_fun = function() {
-n <- 1000
-U <- rnorm( n ) # distressing events (unmeasured)
-C <- rnorm( n, U ) # religious service attendance, increase with distress
-Y <- rnorm( n , U - C ) # measured distress caused by distress, reduced by church attendance. 
-B <- rnorm( n , Y + C + U )# belief increase with service + distressing events + anxiety
-
-# simulate data
-simdat <- data.frame(
-  Belief = B, # belief
-  ReligiousService = C, # religious service
-  U = U, # distressing events (unmeasured)
-  Anxiety = Y  # outcome
-  )
-
-  sim_s <- lm(Anxiety ~ Belief + ReligiousService, data = simdat)
-  sim_s
-}
-
-# Replication 1000xs
-r_lm = replicate(1000, sim_fun(), simplify = FALSE )
-
-# extract simulations
-tab_sim<-purrr::map_dfr(r_lm, broom::tidy)
-
-# extract belief coefficients
-mout<-tab_sim %>%
-  dplyr::filter(term =="Belief")%>%
-  dplyr::mutate_if(is.numeric, round, 5)
-
-# calculate proportion of estimates greater than 0
-# all positive
-sum(mout$estimate > 0) / length(mout$estimate)
-
-# proportion of belief coefficient p.values < .001
-sum(mout$p.value <.001) / length(mout$p.value)
-
-# extract religious service coefficients
-tout<-tab_sim %>%
-  dplyr::filter(term =="ReligiousService")%>%
-  dplyr::mutate_if(is.numeric, round, 5)
-
-# calculate proportion of estimates less than 0
-# all negative
-sum(tout$estimate < 0) / length(tout$estimate)
-
-# proportion of service coefficient p.values < .001
-sum(tout$p.value <.001) / length(tout$p.value)
-```
-
-˝## Simulation tables
-
-```{r}
-parameters::parameters(
-  mod1,
-  test = "pd",
-  priors = FALSE, 
-  effects = "all",
-  drop = TRUE,
-  group_level = FALSE,
-  diagnostic = "Rhat") %>%
-    mutate_if(is.numeric, ~ round(., 2)) %>%
-    mutate_all(funs(str_replace(., "b_", ""))) %>%
-    kable(booktabs = T, "latex", caption =  "Simulation reveals how church attendance and belief can deviate where distress increases religious service attendance and belief in God") %>%
-    print()
-
-parameters::parameters(
-  mod2,
-  test = "pd",
-  priors = FALSE, 
-  effects = "all",
-  drop = TRUE,
-  group_level = FALSE,
-  diagnostic = "Rhat") %>%
-    mutate_if(is.numeric, ~ round(., 2)) %>%
-    mutate_all(funs(str_replace(., "b_", ""))) %>%
-    kable(booktabs = T, "latex", caption =  "Same simulation tracks anxiety as cause of religious belief") %>%
-    print()
-
-parameters::parameters(
-  mod3,
-  test = "pd",
-  priors = FALSE, 
-  effects = "all",
-  drop = TRUE,
-  group_level = FALSE,
-  diagnostic = "Rhat") %>%
-    mutate_if(is.numeric, ~ round(., 2)) %>%
-    mutate_all(funs(str_replace(., "b_", ""))) %>%
-    kable(booktabs = T, "latex", caption =  "Same simulation tracks anxiety as cause of decling religious service") %>%
-    print()
-```
-
-## DETAILS OF ANLAYSIS FOR REPLICATION
+## DETAILS OF ANLAYSIS FROM PREVIOUS ITERATIONS OF THE STUDY
 
@@
 
-#### The group level effects of country
+#### The group level variances of country
 
@@
 
-Which we interpret as follows:
-
--   Countries have different levels of anxiety;
-
--   Within a country, those whose who participate more frequently in church can be expected to have higher anxiety as those who do not participate as frequently in church.
-
--   Between_country predictions trend in the opposite direction, though not reliably at the high end of the anxiety spectrum this indicates.
-
-Let's look at the country-level variances more closely
-
-```{r}
-par<-parameters::model_parameters(m1o,effects = "random" ) 
-plot(par,  sort = TRUE) +theme_lucid() #option -->
-```
-
-Countries who differ in their average levels of church attendance differ present differences in their expected anxiety levels, with an unreliable trend in the opposite direction.
-
-## Belief in God
-
-Lets consider the non-trivial theory that belief in God (bg) affects psychological anxiety (wb) by affording a sense of meaning in life.
-
-We begin by drawing the dag.
-
-And again, we face a collider bias. Our causal model shows us that if we were to include our meaning-making indicator in our statistical we will compromise inference!
-
-```{r}
-ggdag::ggdag_collider(dg1,
-                      from = "church",
-                      to = "wb",
-                      controlling_for(c("male", "support", "country", "age")))
-```
-
-We find weak effect for belief in God. No correlation of intercept and slope but India is an outlier.
-
 ```{r cache = TRUE}
